### 2019-07-22 paper.Rmd
Initial commit

---
title: Executive Constraints and Civil Conflict
author: J.L.A. Krusell
output: pdf_document
---

\[ x_{j}^{obs} \sim \mathsf{Normal}(x_{j}^\ast, \sigma_{j}^{obs}) \]
\[ x_{ji}^\ast \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]
\[ \theta \sim \mathsf{Normal}(0, 1) \]

### 2019-07-22 R/.R

@@ -1,0 +1,1 @@
+dotfiles/R/.R

### 2019-07-22 R/gwno.R

@@ -1,0 +1,132 @@
+#!/usr/bin/env Rscript
+#
+# Generates a translation table between V-Dem country_id and gwno
+# (Gleditsch-Ward).
+###
+
+library(dplyr)
+library(magrittr)
+
+# TODO: update country_unit table with 2018
+utable <- readRDS("refs/vdem-country_unit.rds")
+ctable <- readRDS("refs/vdem-country_table.rds") %>%
+    rename(vdem_id = country_id)
+
+# http://privatewww.essex.ac.uk/~ksg/data/iisystem.dat
+gwno <- read.table("refs/iisystem.dat", sep = "\t", quote = "",
+                  fileEncoding = "latin1", stringsAsFactors = F)
+
+# http://privatewww.essex.ac.uk/~ksg/data/microstatessystem.dat
+gwno_micro <- read.table("refs/microstatessystem.dat", sep = "\t",
+                        quote = "", fileEncoding = "latin1",
+                        stringsAsFactors = F)
+
+gwno %<>% rbind(gwno_micro)
+colnames(gwno) <- c("gwno", "gwno_text_id", "name", "gwno_start", "gwno_end")
+
+gwno %<>% mutate(gwno_start = as.Date(gwno_start, format = "%d:%m:%Y"),
+              gwno_end = as.Date(gwno_end, format = "%d:%m:%Y"),
+              name = ifelse(grepl("Cote", name), "Cote D'Ivoire", name))
+
+check <- . %>% filter(is.na(vdem_id)) %>% select(-vdem_id, -gwno_text_id)
+to_year <- . %>% format("%Y") %>% as.numeric
+
+ctable.merge <- select(ctable, name, vdem_id)
+
+# Try first to match based on name, excluding the alternative gwno
+# names in parantheses
+ttable <- mutate(gwno, name = sub("\\(.*$", "", name) %>% trimws) %>%
+    left_join(ctable.merge, by = "name")
+
+# Preserve the original gwno names for later checking
+ttable$name <- gwno$name
+
+# The fun part, manually assign. We're currently dropping
+# Austria-Hungary since we code them separately.
+ttable %<>%
+    mutate(vdem_id = case_when(name == "United States of America" ~ 20,
+                               name == "Surinam" ~ 4,
+                               name == "German Federal Republic" ~ 77,
+                               name == "Württemberg" ~ 355,
+                               name == "Mecklenburg-Schwerin" ~ 360,
[…]

### 2019-07-22 R/lavaan.R

@@ -1,0 +1,19 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(lavaan)
+
+load("data/prepped_data.RData")
+
+ml <- "lg_constraints =~ v2lgfunds + v2lginvstp + v2gqstexp"
+fit <- cfa(ml, data = data$X_obs, std.lv = T)
+summary(fit, standardized = T)
+
+theta <- lavPredict(fit)
+
+final.df$theta <- lavPredict(fit)
+
+filter(final.df, country_name %in% c("Syria", "India", "Sweden", "Russia")) %>%
+    ggplot(aes(year, theta, col = country_name)) +
+    geom_point() +
+    geom_line()

### 2019-07-22 R/merge.R

@@ -1,0 +1,78 @@
+#!/usr/bin/env Rscript
+
+suppressMessages(library(dplyr))
+suppressMessages(library(magrittr))
+
+###
+# V-Dem - Democracy data
+vdem <- readRDS("data/datasets/V-Dem-CY-Full+Others-v9.rds") %>%
+    select(-matches("_osp|_ord|_mean|_nr")) %>%
+    filter(year >= 1946)
+
+###
+# UCDP - Armed Conflict
+load("data/datasets/ucdp-prio-acd-181")
+
+# conflict_id = 418 is the United States vs. Al Qaeda starting in 2001
+acd %<>% filter(type_of_conflict %in% 3:4, conflict_id != 418) %>%
+    mutate(gwno_a = as.integer(gwno_a))
+
+# V-Dem COWcode and UCDP gwno are mostly the same except Yemen
+# following reunification and Oman prior to 1960 in V-Dem. We'll also
+# lose Cameroon 1960 since V-Dem doesn't code before 1961 and
+# Hyderabad as V-Dem only codes India.
+acd %<>% mutate(gwno_a = ifelse(gwno_a == 678 & year >= 1990, 679, gwno_a))
+vdem %<>% mutate(COWcode = ifelse(country_name == "Oman", 698, COWcode))
+
+counts.df <- group_by(acd, gwno_a, year) %>%
+    summarise(conflicts = n(), intensity = max(intensity_level))
+
+# Episode level
+episode_ucdp <- group_by(acd, location, conflict_id, start_date2) %>%
+    summarise(episode_onset = 1,
+              year = min(year),
+              start_prec2 = ifelse(first(start_prec2) %in% 6:7, 1, 0),
+              gwno_a = first(gwno_a),
+              episode_intensity = first(intensity_level),
+              episode_type = first(type_of_conflict),
+              episode_incompatibility = first(incompatibility)) %>%
+    ungroup %>%
+    arrange(desc(episode_intensity)) %>%
+    distinct(location, year, .keep_all = T) %>%
+    select(-start_date2, -conflict_id)
+
+conflict_ucdp <- group_by(acd, location, conflict_id) %>%
+    summarise(onset = 1,
+              year = min(year),
+              gwno_a = first(gwno_a),
+              onset_intensity = first(intensity_level),
+              onset_type = first(type_of_conflict),
+              onset_incompatibility = first(incompatibility)) %>%
+    ungroup %>%
[…]

### 2019-07-22 R/model.R

@@ -1,0 +1,12 @@
+#!/usr/bin/env Rscript
+
+suppressMessages(library(rstan))
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = T)
+
+load("data/prepped_data.RData")
+fit <- stan("stan/model.stan", iter = 1e4, data = data, thin = 2,
+            control = list(max_treedepth = 12))
+
+saveRDS(fit, "posteriors/fit.rds")

### 2019-07-22 R/prep.R

@@ -1,0 +1,63 @@
+#!/usr/bin/env Rscript
+
+suppressMessages(library(data.table))
+suppressMessages(library(dplyr))
+suppressMessages(library(magrittr))
+
+fn <- function(f) {
+    print(f)
+
+    df <- fread(f, skip = 1, header = F)
+    v <- apply(df[, -1], 2, min, na.rm = T)
+
+    list(mean = mean(v), sd = sd(v))
+}
+
+variables <- c("v2lginvstp", "v2lgfunds", "v2lgqstexp")
+re <- paste0(variables, collapse = "|")
+
+files <- list.files("./data/vdem_posteriors", re, full.names = T)
+stopifnot(length(files) == 3)
+
+min.df <- lapply(files, fn) %>%
+    bind_rows %>%
+    mutate(variable = variables)
+
+###
+# Grab gap years and country-years where lg is missing from full V-Dem
+# dataset.
+vdem <- readRDS("data/merged_data.rds") %>%
+    select(country_name, year, v2lgbicam, matches(re),
+           -matches("codehigh|codelow"))
+
+print(colnames(vdem))
+
+for (v2 in variables) {
+    vdem[[v2]] <- ifelse(vdem$v2lgbicam == 0,
+                         min.df$mean[min.df$variable == v2],
+                         vdem[[v2]])
+
+    vdem[[paste0(v2, "_sd")]] <- ifelse(vdem$v2lgbicam == 0,
+                                       min.df$sd[min.df$variable == v2],
+                                       vdem[[paste0(v2, "_sd")]])
+}
+
+# TODO: gaps
+final.df <- na.omit(vdem) %>%
+    mutate(s = paste(v2lgqstexp, v2lginvstp, v2lgfunds)) %>%
+    group_by(country_name) %>%
+    mutate(idx = rle(s) %>% { rep(.$values, .$lengths) }) %>%
+    group_by(idx, add = T) %>%
+    filter(year == min(year)) %>%
+    ungroup
+
+manifests <- select(final.df, one_of(variables)) %>% data.matrix
+manifests_sd <- select(final.df, one_of(paste0(variables, "_sd"))) %>%
+    data.matrix
+
+data <- list(N = nrow(manifests),
+             J = length(variables),
+             X_obs = manifests,
+             X_se = manifests_sd)
+
+save(data, final.df, file = "data/prepped_data.RData")

### 2019-07-22 R/summarise.R

@@ -1,0 +1,22 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(rstan)
+
+env <- readRDS("data/prepped_data.rds")
+fit <- readRDS("posteriors/fit.rds")
+
+check_hmc_diagnostics(fit)
+
+theta <- as.matrix(fit, pars = "theta") %>%
+    apply(2, quantile, probs = c(0.25, 0.5, 0.75)) %>%
+    t %>%
+    as.data.frame %>%
+    rename(theta_codelow = `25%`,
+           theta = `50%`,
+           theta_codehigh = `75%`) %>%
+    mutate(country_name = env$country_year$country_name,
+           year = env$country_year$year)
+
+saveRDS(theta, "data/summarised_post.rds")

### 2019-07-22 run.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+
+docker run -d -v $PWD:/proj conflict_onset make $@ && \
+    docker logs -f $(docker ps -l --format "{{.Names}}")

### 2019-07-22 sync.sh

@@ -1,0 +1,7 @@
+#!/bin/sh
+
+rsync -vrue ssh --partial --progress \
+      --exclude=docs --exclude=plots --exclude=replications \
+      --exclude=Rmd --exclude=refs --exclude=md \
+      --exclude='\#*' --exclude=posteriors --exclude=data/*.rds \
+      ./ nki:/home/jsks/inv/

### 2019-08-04 R/thesis.utils/R/ops.R
First version of utility function pkg

@@ -1,0 +1,7 @@
+#' @importFrom magrittr %>%
+#' @export
+magrittr::`%>%`
+
+#' @importFrom magrittr %$%
+#' @export
+magrittr::`%$%`

### 2019-08-04 R/thesis.utils/R/thesis.utils.R

@@ -1,0 +1,13 @@
+#' thesis.utils
+#'
+#' Utility functions for executive constraints and civil conflict
+#' analysis. The majority of exported functions are fairly specific
+#' and are not meant to be particularly useful outside the context of
+#' this analysis.
+#'
+#' @docType package
+#' @name thesis.utils
+if (getRversion() >= "2.15.1")
+    utils::globalVariables(".")
+
+NULL

### 2019-08-04 R/thesis.utils/R/utils.R

@@ -1,0 +1,222 @@
+#' Calculate peace years
+#'
+#' Given a vector of years and a matching length vector of conflict
+#' incidence, calculate the number of peace years since the last civil
+#' conflict or the coding start, identified as the first element.
+#'
+#' @param years NumericVector of years
+#' @param incidence NumericVector or LogicalVector of same size as
+#'     `year` indicating conflict incidence
+#'
+#' @details This calculates the number of years of peace since either
+#'     the last incidence of civil conflict or the coding start given
+#'     by the first element in the `years` vector.
+#'
+#'     Note, peace years are calculated based on V-Dem countries,
+#'     which means that gwid changes are not recorded as changes in
+#'     the fundamental status of a country. This only affects Serbia
+#'     in 2006 and Czech Republic in 1993.
+#'
+#'     Furthermore, the coding starts are also taken from V-Dem, which
+#'     includes non-sovergeign polities such as colonial
+#'     territories. In the `merge.R` we attempt to therefore backtrack
+#'     conflicts on the basis of territory location (ex: Eritrea),
+#'     even though the final analysis will only be done on independent
+#'     countries according to GW.
+#'
+#' @examples
+#' years <- 1900:1910
+#' ongoing <- sample(0:1, length(years), replace = TRUE)
+#'
+#' calc_peace_yrs(years, ongoing)
+#'
+#' @export
+calc_peace_yrs <- function(years, incidence) {
+    if (length(years) != length(incidence))
+        stop("Mismatched vector lengths", call. = F)
+    else if (!identical(years, sort(years)))
+        stop("Unsorted years vector", call. = F)
+    else if (any(consecutive(years) != 1))
+        stop("Breaks found in years vector", call. = F)
+
+    locf_idx <- function(x) {
+        x[x == 1] <- which(x == 1)
+        lx <- c(NA, dplyr::lag(x)[-1] %>% cummax)
+        is.na(lx) <- lx == 0
+        return(lx)
+    }
+
+    x <- years[locf_idx(incidence)]
+    out <- years - ifelse(is.na(x), years[1], x) - 1
+
+    pmax(out, 0)
+}
+
+#' Normalize vector
+#'
+#' Simple wrapper around the default arguments to [scale()] which
+#' transforms a numeric matrix according to `(x - mean(x)) / sd(x)`.
+#'
+#' @param x NumericVector
+#'
+#' @examples
+#' normalize(1:10)
+#'
[…]

### 2019-08-04 R/thesis.utils/tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(thesis.utils)
+
+test_check("thesis.utils")

### 2019-08-04 R/thesis.utils/tests/testthat/test_utils.R

@@ -1,0 +1,106 @@
+test_that("calc_peace_yrs", {
+    years <- 1900:1905
+    ongoing <- rep(0, 6)
+
+    expect_identical(calc_peace_yrs(years, ongoing), c(0, 0:4))
+    expect_error(calc_peace_yrs(years, ongoing[-1]),
+                 "Mismatched vector lengths")
+
+    years <- 1900:1905
+    ongoing <- c(1, 0, 1, 0, 0, 0)
+
+    expect_identical(calc_peace_yrs(years, ongoing),
+                     c(0, 0, 1, 0, 1, 2))
+
+    years <- 1900:1905
+    ongoing <- c(0, 0, 0, 0, 1, 0)
+
+    expect_identical(calc_peace_yrs(years, ongoing),
+                     c(0, 0, 1, 2, 3, 0))
+
+    years <- 1900:1909
+    ongoing <- c(rep(0, 4), 1, rep(0, 5))
+
+    expect_identical(calc_peace_yrs(years, ongoing),
+                     c(0, 0:3, 0, 1:4))
+
+    years <- c(1900, 1902:1904)
+    ongoing <- rep(0, 4)
+
+    expect_error(calc_peace_yrs(years, ongoing),
+                 "Breaks found in years vector")
+
+    years <- c(1900, 1903, 1901, 1902)
+    ongoing <- rep(0, 4)
+
+    expect_error(calc_peace_yrs(years, ongoing),
+                 "Unsorted years vector")
+})
+
+test_that("normalize", {
+    fn <- function(x) as.vector(scale(x))
+    x <- 1:10
+
+    expect_identical(normalize(x), fn(x))
+    expect_equal(mean(normalize(x)), 0)
+    expect_equal(stats::sd(normalize(x)), 1)
+    expect_equal(normalize(rep(5, 5)), rep(NaN, 5))
+    expect_error(suppressWarnings(normalize(letters[1:3])))
+})
+
+test_that("to_idx", {
+    years <- rep(1900:1902, 3)
+    expect_identical(to_idx(years), c(1, 2, 3, 1, 2, 3, 1, 2, 3))
+
+    years <- c(1901, 1900, 1901, 1901, 1905)
+    expect_identical(to_idx(years), c(2, 1, 2, 2, 3))
+    expect_identical(to_idx(years, unique(years)), c(1, 2, 1, 1, 3))
+
+    countries <- c("SWE", "AFG", "FRA")
[…]

### 2019-08-04 R/functions.R
Remove deprecated functions.R script

@@ -1,41 +1,0 @@
-locf_idx <- function(x) {
-    x[x == 1] <- which(x == 1)
-    lx <- c(NA, lag(x)[-1] %>% cummax)
-    is.na(lx) <- lx == 0
-    return(lx)
-}
-
-normalize <- function(x) scale(x) %>% as.vector
-
-to_idx <- function(x, ...) factor(x, ...) %>% as.numeric
-
-consecutive <- function(x) cumsum(c(T, diff(x) != 1))
-
-collapse_changes <- function(x) {
-    v <- rle(x)
-    rep_v <- rep(v$values, v$lengths)
-    to_idx(rep_v, unique(rep_v))
-}
-
-info <- function(df) {
-    sprintf("%d conflicts, %d episodes, %d countries, %s country-years",
-            sum(df$lonset, na.rm = T),
-            sum(df$lepisode_onset, na.rm = T),
-            n_distinct(df$country_name),
-            prettyNum(nrow(df), big.mark = ","))
-}
-
-explode <- function(df) {
-    ll <- lapply(1:nrow(df), function(i) {
-        start <- df$start_year[i]
-        end <- df$end_year[i]
-
-        len <- end - start + 1
-        sub.df <- df[rep(i, len), ]
-        sub.df$year <- start:end
-
-        sub.df
-    })
-
-    bind_rows(ll)
-}

### 2019-08-04 R/geo.R

@@ -29,7 +29,7 @@ ll <- lapply(seq_along(neighbours.ll), function(i) {
 
 neighbours.df <- bind_rows(ll) %>%
     filter(country_name != neighbour, start_year <= end_year) %>%
-    explode %>%
+    explode(from = .$start_year, to = .$end_year) %>%
     distinct(country_name, gwid, neighbour, neighbour_gwid, year)
 
 saveRDS(neighbours.df, "data/neighbours.rds")

### 2019-08-04 R/merge.R

@@ -168,10 +168,7 @@ merged.df <- left_join(vdem, ucdp, by = c("gwid" = "gwno_loc", "year")) %>%
 merged.df %<>%
     arrange(country_name, year) %>%
     group_by(country_name) %>%
-    mutate(first_year = pmax(1945, codingstart_contemp, gapend + 1, na.rm = T),
-           peace_yrs = {
-        year - year[locf_idx(ongoing)] %>% { ifelse(is.na(.), first_year, .) } - 1
-    } %>% pmax(0)) %>%
+    mutate(peace_yrs = calc_peace_yrs(year, ongoing)) %>%
     ungroup
 
 stopifnot(!is.na(merged.df$country_name))
@@ -264,5 +261,8 @@ sprintf("After merging GROWup, %d countries and %d rows",
 
 ###
 # Save, save, save!
-info(merged.df)
+dbg_info(merged.df)
 saveRDS(merged.df, "data/merged_data.rds")
+
+# Final assertion
+assert_cy(merged.df)

### 2019-08-04 R/prep.R

@@ -69,6 +69,9 @@ setdiff(unique(merged.df$country_name), final.df$country_name) %>%
     paste(collapse = "; ") %>%
     sprintf("Lost due to missingness: %s", .)
 
-info(final.df)
+dbg_info(final.df)
 save(constraint_vars, final.df, constraints.df,
      file = "data/prepped_data.RData")
+
+# Final assertion
+assert_cy(final.df)

### 2019-08-04 scripts/build.sh
Add simple build script

@@ -1,0 +1,6 @@
+#!/bin/sh
+
+set -e
+
+docker build -t jsks/conflict_onset:latest .
+docker push jsks/conflict_onset

### 2019-08-04 R/gam.R
Add thesis.utils to library imports

@@ -4,8 +4,7 @@ library(caTools)
 library(dplyr)
 library(lavaan)
 library(mgcv)
-
-source("R/functions.R")
+library(thesis.utils)
 
 load("data/prepped_data.RData")
 

### 2019-08-04 R/geo.R

@@ -2,8 +2,8 @@
 
 suppressMessages(library(dplyr))
 suppressMessages(library(sf))
+suppressMessages(library(thesis.utils))
 
-source("R/functions.R")
 
 cshapes <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
     select(country_name = CNTRY_NAME, gwid = GWCODE,

### 2019-08-04 R/merge.R

@@ -9,8 +9,7 @@ suppressMessages(library(dplyr))
 suppressMessages(library(magrittr))
 suppressMessages(library(readxl))
 suppressMessages(library(tidyr))
-
-source("R/functions.R")
+suppressMessages(library(thesis.utils))
 
 ###
 # V-Dem - Democracy data

### 2019-08-04 R/model.R

@@ -2,8 +2,7 @@
 
 suppressMessages(library(dplyr))
 suppressMessages(library(rstan))
-
-source("R/functions.R")
+suppressMessages(library(thesis.utils))
 
 options(mc.cores = parallel::detectCores())
 rstan_options(auto_write = T)

### 2019-08-04 R/prep.R

@@ -3,8 +3,7 @@
 suppressMessages(library(data.table))
 suppressMessages(library(dplyr))
 suppressMessages(library(magrittr))
-
-source("R/functions.R")
+suppressMessages(library(thesis.utils))
 
 constraint_vars <- c("v2lginvstp", "v2lgfunds", "v2lgqstexp", "v2lgoppart",
                      "v2juhcind", "v2juncind", "v2juhccomp", "v2jucomp",

### 2019-08-04 R/gam.R
Merge Maddison data
V-Dem seems to have some weird patterns of missingness for e_migdpgro
which doesn't match the code used by the V-Dem DS construction
scripts. Plus, they're also using cgdppc for the growth calculation
instead of the recommended rgdpnapc. To deal with this, re-merge
Maddison data --- also might as well take population data from there
instead of the UN.

@@ -19,26 +19,24 @@ summary(lfit, standardized = T, ci = T)
 final.df$exec_constraints <- lavPredict(lfit)[, 1]
 
 input.df <- select(final.df, lepisode_onset, country_name, year,
-                   exec_constraints, e_migdppcln, e_migdpgro,
-                   meanelev, area_sqkm, un_pop, rlvt_groups_count,
+                   exec_constraints, cgdppc, gdpgro,
+                   pop_density, meanelev, rlvt_groups_count,
                    neighbour_conflict, peace_yrs, ongoing) %>%
-    mutate(pop_density = log(1000 * un_pop / area_sqkm) %>% normalize,
+    mutate(pop_density = log(pop_density) %>% normalize,
            exec_constraints = normalize(exec_constraints),
-           e_migdppcln = normalize(e_migdppcln),
-           e_migdpgro = normalize(e_migdpgro),
+           cgdppc = log(cgdppc) %>% normalize,
+           gdpgro = normalize(gdpgro),
            meanelev = normalize(meanelev),
            country_name = as.factor(country_name),
            lpeace_yrs = log(peace_yrs + 1),
-           peace_yrs = normalize(peace_yrs),
-           year_fac = as.factor(year),
-           year = normalize(year)) %>%
-    na.omit
+           year_fac = as.factor(year))
 
 info(input.df)
-ml <- gam(lepisode_onset ~ exec_constraints * e_migdppcln + e_migdpgro +
+ml <- gam(lepisode_onset ~ exec_constraints * cgdppc + gdpgro +
               pop_density + meanelev + rlvt_groups_count +
-              neighbour_conflict + s(country_name, bs = "re") +
-              s(year_fac, bs = "re") + s(peace_yrs, bs = "gp"),
+              neighbour_conflict + lpeace_yrs + s(country_name, bs = "re") +
+              s(year_fac, bs = "re"),
           data = input.df, family = "binomial")
 
+print("Model finished!")
 saveRDS(ml, "data/gam_model.rds")

### 2019-08-04 R/merge.R

@@ -19,7 +19,7 @@ suppressMessages(library(thesis.utils))
 # a single gapstart/gapend to make things easier when we later
 # calculate peace years.
 vdem <- readRDS("data/raw/V-Dem-CY-Full+Others-v9.rds") %>%
-    select(-matches("_osp|_ord|_mean|_nr|_\\d*$")) %>%
+    select(-matches("_osp|_ord|_mean|_nr|_\\d*$"), -matches("^e_")) %>%
     filter(year >= 1946) %>%
     group_by(country_id) %>%
     mutate(gapstart = ifelse(is.na(gapstart1) & is.na(gapstart2) & is.na(gapstart3),
@@ -31,6 +31,62 @@ vdem <- readRDS("data/raw/V-Dem-CY-Full+Others-v9.rds") %>%
 
 sprintf("Started with %d countries and %d rows from V-Dem",
         n_distinct(vdem$country_id), nrow(vdem))
+
+###
+# Maddison - Population & GDP
+#
+# This is already merged into V-Dem; however, we want to use rgdpnapc
+# to calculate GDP growth rather than cgdppc. Plus, we'll also take
+# population statistics since V-Dem only has WB figures.
+#
+# Maddison includes Serbia Federation, Czech territory, and RFSR as
+# separate entites. V-Dem codes entire Yugoslavia and Czechoslovakia
+# so drop the constituent republics prior to independence, Russia
+# however is coded as only the socialist republic instead of the
+# entire USSR.
+maddison.df <- read_xlsx("data/raw/mpd2018.xlsx", sheet = 2) %>%
+    filter(year > 1940,
+           !(country == "Serbia" & year < 1992),
+           !(country == "Former Yugoslavia" & year >= 1992),
+           !(country == "Czech Republic" & year < 1992),
+           !(country == "Czechoslovakia" & year >= 1992)) %>%
+    mutate(country =
+               case_when(country == "Bolivia (Plurinational State of)" ~ "Bolivia",
+                         country == "Cabo Verde" ~ "Cape Verde",
+                         country == "China, Hong Kong SAR" ~ "Hong Kong",
+                         country == "Congo" ~ "Republic of the Congo",
+                         country == "Côte d'Ivoire" ~ "Ivory Coast",
+                         country == "Czechoslovakia" & year < 1992 ~ "Czech Republic",
+                         country == "D.P.R. of Korea" ~ "North Korea",
[…]

### 2019-08-04 R/model.R

@@ -16,12 +16,14 @@ manifests_sd <- select(constraints.df, one_of(paste0(constraint_vars, "_sd"))) %
 stopifnot(nrow(manifests) == nrow(manifests_sd),
           ncol(manifests) == ncol(manifests_sd))
 
-X <- select(final.df, e_migdpgro, pop_density, meanelev, rlvt_groups_count,
+X <- select(final.df, gdpgro, pop_density, meanelev, rlvt_groups_count,
             neighbour_conflict, peace_yrs) %>%
-    mutate(e_migdpgro = normalize(e_migdpgro),
+    mutate(gdpgro = normalize(gdpgro),
            pop_density = log(pop_density) %>% normalize,
            meanelev = log(meanelev) %>% normalize,
-           lpeace_yrs = log(peace_yrs + 1) %>% normalize)
+           lpeace_yrs = log(peace_yrs + 1) %>% normalize) %>%
+    select(-peace_yrs) %>%
+    data.matrix
 
 data <- list(J = nrow(manifests),
              D = ncol(manifests),
@@ -30,15 +32,12 @@ data <- list(J = nrow(manifests),
              N = nrow(X),
              M = ncol(X),
              X = X,
-             state_capacity = normalize(final.df$e_migdppcln),
+             state_capacity = log(final.df$cgdppc) %>% normalize,
              exec_idx = final.df$reduced_idx,
              n_countries = n_distinct(final.df$country_name),
              n_years = n_distinct(final.df$year),
-             #n_peace_yrs = n_distinct(final.df$peace_yrs),
              country_id = to_idx(final.df$country_name),
              year_id = to_idx(final.df$year),
-             #peace_yr_id = to_idx(final.df$peace_yrs),
-             #peace_yrs = unique(final.df$peace_yrs) %>% sort,
              y = final.df$lepisode_onset)
 str(data)
 stopifnot(!anyNA(data))

### 2019-08-04 R/prep.R

@@ -48,8 +48,9 @@ for (v2 in lg_vars) {
 final.df <- merged.df %>%
     select(country_name, year, lonset, lepisode_onset, peace_yrs,
            one_of(constraint_vars), one_of(paste0(constraint_vars, "_sd")),
-           e_migdppcln, e_migdpgro, pop_density, meanelev, ongoing,
-           rlvt_groups_count, neighbour_conflict) %>%
+           cgdppc, gdpgro, pop_density, meanelev, ongoing,
+           rlvt_groups_count, neighbour_conflict, v2lgbicam) %>%
+    filter(v2lgbicam != 0) %>%
     filter_at(constraint_vars, all_vars(!is.na(.))) %>%
     mutate(s = do.call(paste, lapply(c("country_name", constraint_vars), as.symbol)),
            reduced_idx = collapse_changes(s)) %>%

### 2019-08-04 R/model.R
Normalize v2* variables used in factor analysis

@@ -9,9 +9,16 @@ rstan_options(auto_write = T)
 
 load("data/prepped_data.RData")
 
-manifests <- select(constraints.df, one_of(constraint_vars)) %>% data.matrix
+manifests <- select(constraints.df, one_of(constraint_vars)) %>%
+    data.matrix
+
 manifests_sd <- select(constraints.df, one_of(paste0(constraint_vars, "_sd"))) %>%
     data.matrix
+
+for (i in 1:ncol(manifests)) {
+    manifests_sd[, i] <- manifests_sd[, i] / sd(manifests[, i])
+    manifests[, i] <- normalize(manifests[, i])
+}
 
 stopifnot(nrow(manifests) == nrow(manifests_sd),
           ncol(manifests) == ncol(manifests_sd))

### 2019-08-04 paper.Rmd
WIP: add some initial plots

@@
 author: J.L.A. Krusell
-output: pdf_document
+output:
+    pdf_document:
+        pandoc_args: ["--syntax-definition", "stan.xml",
+                      "--highlight-style", "kate"]
 ---
 
-\[ x_{j}^{obs} \sim \mathsf{Normal}(x_{j}^\ast, \sigma_{j}^{obs}) \]
-\[ x_{ji}^\ast \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]
+```{r, echo = F}
+```
+
+\[ y_t \sim \mathsf{Bernoulli}(\pi_t) \]
+\[ \pi_t = F(\alpha + \beta \theta_{t-1} + \delta_k + GP(t)) \]
+\[ \delta_k \sim \mathsf{Normal}(0, \sigma) \]
+\[ \sigma \sim \mathsf{HalfCauchy}(0, 1) \]
+\[ x_{jt}^\ast \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_t, \psi_j) \]
+\[ x_{j}^{obs} \sim \mathsf{Normal}(x_{j}^\ast, \tau_{j}^{obs}) \]
 \[ \theta \sim \mathsf{Normal}(0, 1) \]
+\[ \lambda \sim \mathsf{Lognormal}(0, 1) \]
+\[ \gamma \sim \mathsf{Normal}(0, 5) \]
+\[ \psi \sim \mathsf{HalfCauchy}(0, 1) \]
+
+We assume local item independence such that the residual variances, $\psi$, are uncorrelated.
+
+- v2lginvstp - Legislature investigates in practice
+- v2lgfunds - Legislature controls own resources
+- v2lgqstexp - Legislature questions officials in practice
+- v2lgoppart - Legislature opposition parties able to exercise oversight
+
+- v2juhcind - High court independence
+- v2juncind - Lower court independence
+- v2juhccomp - Compliance with higher court
+- v2jucomp - Compliance with judiciary
+
+- v2exrescon - Executive respects constitution
+- v2lgotovst - Executive oversight
+
+## Analysis
+
+```{r, echo = F}
+suppressMessages(library(bayesplot))
+suppressMessages(library(rstan))
+
+fit <- readRDS("posteriors/fit.rds")
+beta <- as.matrix(fit, pars = "beta")
+```
+
+```{r, echo = F}
+color_scheme_set("green")
+mcmc_intervals(beta, point_est = "median") +
+    vline_0(size = 0.25, linetype = 2, color = "violet") +
+    ggplot2::scale_y_discrete(labels = c("Executive Constraints",
+                                "State Capacity",
+                                "Executive Constraints:State Capacity",
+                                "GDP per Capita Growth",
+                                "Population Density",
+                                "Mean Elevation",
+                                "Relevant Ethnic Groups",
+                                "Neighbouring Conflict",
+                                "Peace Years Logged"))
+
+```
+
+\newpage
+
+# Appendix
+
+## Stan Code
+
+```{stan, code=readLines("stan/model.stan"), eval=FALSE, output.var="x"}
+```
+\newpage
+
+## GAM Model
+
+```{R, echo = F, results = "asis"}
+ml <- readRDS("data/gam_model.rds")
+itsadug::gamtabs(ml, caption = "GAM Model")
+```
+
+```{R, echo = F}
+p <- predict(ml, type = "response")
+
+plot(density(p[ml$model$lepisode_onset == 0]), col = "lightblue",
+     main = "Predicted Probability Densities")
+lines(density(p[ml$model$lepisode_onset == 1]), col = "indianred")
+```
+
+```{R, echo = F}
+caTools::colAUC(p, ml$model$lepisode_onset, plotROC = T)
+```

### 2019-08-04 R/geo.R
BUGFIX: rename sequence column to year

@@ -4,7 +4,6 @@ suppressMessages(library(dplyr))
 suppressMessages(library(sf))
 suppressMessages(library(thesis.utils))
 
-
 cshapes <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
     select(country_name = CNTRY_NAME, gwid = GWCODE,
            start_year = GWSYEAR, end_year = GWEYEAR) %>%
@@ -18,6 +17,11 @@ ll <- lapply(seq_along(neighbours.ll), function(i) {
     if (length(v) == 0)
         return(NULL)
 
+    # We take the start year NOT from the neighbour, but from the
+    # target country. This means that we'll of course end up with
+    # dyads outside of the start date for the neighbouring country. We
+    # do this because we want to include colonial conflicts as
+    # neighbouring conflicts (ex: Algeria <1962).
     data.frame(country_name = cshapes$country_name[i],
                gwid = cshapes$gwid[i],
                neighbour = cshapes$country_name[v],
@@ -30,6 +34,7 @@ ll <- lapply(seq_along(neighbours.ll), function(i) {
 neighbours.df <- bind_rows(ll) %>%
     filter(country_name != neighbour, start_year <= end_year) %>%
     explode(from = .$start_year, to = .$end_year) %>%
+    rename(year = sequence) %>%
     distinct(country_name, gwid, neighbour, neighbour_gwid, year)
 
 saveRDS(neighbours.df, "data/neighbours.rds")

### 2019-08-04 R/merge.R
Calculate peace years from independence
Previously, peace years for former colonial states were being
calculating from V-Dem's start date versus f.ex. former soviet
republics which aren't coded during the Cold War.

@@ -20,7 +20,7 @@ suppressMessages(library(thesis.utils))
 # calculate peace years.
 vdem <- readRDS("data/raw/V-Dem-CY-Full+Others-v9.rds") %>%
     select(-matches("_osp|_ord|_mean|_nr|_\\d*$"), -matches("^e_")) %>%
-    filter(year >= 1946) %>%
+    filter(year >= 1945) %>%
     group_by(country_id) %>%
     mutate(gapstart = ifelse(is.na(gapstart1) & is.na(gapstart2) & is.na(gapstart3),
                              NA,
@@ -204,28 +204,23 @@ merged.df <- left_join(vdem, ucdp, by = c("gwid" = "gwno_loc", "year")) %>%
            lepisode_intensity = lead(episode_intensity))
 
 # Calculate number of peace years since last ongoing civil conflict or
-# V-Dem codingstart/gapend. For countries censored due to start date,
-# start counting from the end of WW2. Even for countries not directly
+# independence. For countries censored due to start date, start
+# counting from the end of WW2. Even for countries not directly
 # involved in WW2, it was a significant international event that
 # reshaped the world order with domestic consequences for everyone.
 #
-# This end result is essentially the peaceyears calculated by GROWup,
-# with the exception of newly independent countries that inherited the
-# conflict counts based on territory location (ex: Eritrea). We'll
-# also merge in GROWups peaceyears count which simply starts at zero
-# when a country enters that dataset.
+# The result is essentially the peaceyears count from GROWup with a
+# few exceptions, including extrasystemic colonial wars for
+# independence spilling over past independence and which are then
+# included in ongoing counts of conflict (ex: Indonesia).
 merged.df %<>%
+    filter(year >= start_year, year <= end_year) %>%
     arrange(country_name, year) %>%
     group_by(country_name) %>%
     mutate(peace_yrs = calc_peace_yrs(year, ongoing)) %>%
     ungroup
 
 stopifnot(!is.na(merged.df$country_name))
-
-###
-# Restrict our dataset to independent country-years (i.e. where the
-# state is sovereign).
-merged.df %<>% filter(year >= start_year, year <= end_year)
 
 ###
 # GROWUP - Area, elevation, and ethnic group data

### 2019-08-04 R/thesis.utils/R/utils.R

@@ -17,13 +17,6 @@
 #'     the fundamental status of a country. This only affects Serbia
 #'     in 2006 and Czech Republic in 1993.
 #'
-#'     Furthermore, the coding starts are also taken from V-Dem, which
-#'     includes non-sovergeign polities such as colonial
-#'     territories. In the `merge.R` we attempt to therefore backtrack
-#'     conflicts on the basis of territory location (ex: Eritrea),
-#'     even though the final analysis will only be done on independent
-#'     countries according to GW.
-#'
 #' @examples
 #' years <- 1900:1910
 #' ongoing <- sample(0:1, length(years), replace = TRUE)
@@ -43,11 +36,12 @@ calc_peace_yrs <- function(years, incidence) {
         x[x == 1] <- which(x == 1)
         lx <- c(NA, dplyr::lag(x)[-1] %>% cummax)
         is.na(lx) <- lx == 0
+
         return(lx)
     }
 
     x <- years[locf_idx(incidence)]
-    out <- years - ifelse(is.na(x), years[1], x) - 1
+    out <- years - ifelse(is.na(x), years[1] - 1, x) - 1
 
     pmax(out, 0)
 }
@@ -74,7 +68,7 @@ normalize <- function(x) (x - mean(x)) / stats::sd(x)
 #' @param ... Additional arguments passed to [factor()]
 #'
 #' @details This is useful for one thing only: creating the input data
-#'     list for STan.
+#'     list for Stan.
 #'
 #'     For partially pooled intercepts in Stan we need the index
 #'     positions of each unique category associated with a given

### 2019-08-04 R/thesis.utils/tests/testthat/test_utils.R

@@ -2,7 +2,7 @@ test_that("calc_peace_yrs", {
     years <- 1900:1905
     ongoing <- rep(0, 6)
 
-    expect_identical(calc_peace_yrs(years, ongoing), c(0, 0:4))
+    expect_identical(calc_peace_yrs(years, ongoing), c(0, 1:5))
     expect_error(calc_peace_yrs(years, ongoing[-1]),
                  "Mismatched vector lengths")
 
@@ -16,13 +16,13 @@ test_that("calc_peace_yrs", {
     ongoing <- c(0, 0, 0, 0, 1, 0)
 
     expect_identical(calc_peace_yrs(years, ongoing),
-                     c(0, 0, 1, 2, 3, 0))
+                     c(0, 1, 2, 3, 4, 0))
 
     years <- 1900:1909
     ongoing <- c(rep(0, 4), 1, rep(0, 5))
 
     expect_identical(calc_peace_yrs(years, ongoing),
-                     c(0, 0:3, 0, 1:4))
+                     c(0, 1:4, 0, 1:4))
 
     years <- c(1900, 1902:1904)
     ongoing <- rep(0, 4)

### 2019-08-21 R/thesis.utils/R/utils.R
Remove NA by default in normalize

@@ -57,7 +57,7 @@ calc_peace_yrs <- function(years, incidence) {
 #' normalize(1:10)
 #'
 #' @export
-normalize <- function(x) (x - mean(x)) / stats::sd(x)
+normalize <- function(x) (x - mean(x, na.rm = T)) / stats::sd(x, na.rm = T)
 
 #' Create index
 #'

### 2019-08-21 R/thesis.utils/tests/testthat/test_utils.R

@@ -46,6 +46,9 @@ test_that("normalize", {
     expect_equal(stats::sd(normalize(x)), 1)
     expect_equal(normalize(rep(5, 5)), rep(NaN, 5))
     expect_error(suppressWarnings(normalize(letters[1:3])))
+
+    x <- c(NA, 1:10, NA)
+    expect_identical(normalize(x), fn(x))
 })
 
 test_that("to_idx", {

### 2019-08-21 scripts/run.sh
Update server scripts

@@ -1,6 +1,7 @@
 #!/bin/sh
 
-docker run -d -v $PWD:/proj jsks/conflict_onset make $@
+docker run -d -v $PWD:/proj -v /mnt/data/posteriors:/proj/posteriors \
+       jsks/conflict_onset make $@
 container=$(docker ps -l --format "{{.Names}}")
 
 echo "Launched: $container"

### 2019-08-21 scripts/sync.sh

@@ -1,5 +1,5 @@
 #!/bin/sh
 
-rsync -vrue ssh --delete --partial --progress \
+rsync -vrtue ssh --delete --partial --progress \
       --exclude-from $(dirname $0)/exclude.txt \
-      ./ nki:/home/jsks/inv/
+      ./ gce:/home/cloud/thesis/

### 2019-08-21 R/merge.R
Merge CINC dataset

@@ -13,24 +13,31 @@ suppressMessages(library(thesis.utils))
 
 ###
 # V-Dem - Democracy data
-#
-# Recorded gaps in V-Dem DS are super annoying to deal with. Since
-# contemporary countries have no more than one, unify everything into
-# a single gapstart/gapend to make things easier when we later
-# calculate peace years.
 vdem <- readRDS("data/raw/V-Dem-CY-Full+Others-v9.rds") %>%
-    select(-matches("_osp|_ord|_mean|_nr|_\\d*$"), -matches("^e_")) %>%
-    filter(year >= 1945) %>%
-    group_by(country_id) %>%
-    mutate(gapstart = ifelse(is.na(gapstart1) & is.na(gapstart2) & is.na(gapstart3),
-                             NA,
-                             max(gapstart1, gapstart2, gapstart3, na.rm = T)),
-           gapend = ifelse(is.na(gapend1) & is.na(gapend2) & is.na(gapend3),
-                           NA,
-                           max(gapend1, gapend2, gapend3, na.rm = T)))
+    select(-matches("_ord|_mean|_nr|_\\d*$"), -matches("^e_")) %>%
+    filter(year >= 1945)
 
 sprintf("Started with %d countries and %d rows from V-Dem",
         n_distinct(vdem$country_id), nrow(vdem))
+
+###
+# COW - CINC
+nmc <- read.csv("data/raw/NMC_5_0/NMC_5_0.csv", stringsAsFactors = F) %>%
+    filter(year >= 1945)
+
+# Why are there duplicates in this file?!
+cow <- read.csv("refs/COW country codes.csv", stringsAsFactors = F) %>%
+    distinct(CCode, StateNme)
+
+nmc %<>% left_join(cow, by = c("ccode" = "CCode")) %>%
+    select(ccode, year, cinc, COWname = StateNme)
+
+filter(nmc, !ccode %in% vdem$COWcode) %>%
+    distinct(COWname) %$%
+    paste(COWname, collapse = "; ") %>%
+    sprintf("COW countries missing from V-Dem: %s", .)
+
+vdem %<>% left_join(nmc, by = c("COWcode" = "ccode", "year"))
 
 ###
 # Maddison - Population & GDP
@@ -135,7 +142,9 @@ filter(ctable, gwid %in% setdiff(civil$gwno_loc, vdem$gwid)) %$%
     sprintf("UCDP countries missing from merged: %s", .)
 
[…]

### 2019-08-21 R/gam.R
Update GAM model script

@@ -10,10 +10,11 @@ load("data/prepped_data.RData")
 
 manifests <- select(final.df, one_of(constraint_vars)) %>% data.matrix
 
-ml <- "exec_constraints =~ v2lginvstp + v2lgfunds + v2lgqstexp + v2lgoppart +
-                           v2juhcind + v2juncind + v2juhccomp + v2jucomp +
-                           v2exrescon + v2lgotovst"
-lfit <- cfa(ml, data = manifests, meanstructure = T, std.lv = T)
+ml <- "exec_constraints =~ v2lginvstp + v2lgfunds + v2lgqstexp +
+                           v2lgoppart + v2juhcind + v2juncind +
+                           v2juhccomp + v2jucomp + v2exrescon +
+                           v2lgotovst"
+lfit <- cfa(ml, data = manifests, std.lv = T)
 summary(lfit, standardized = T, ci = T)
 
 final.df$exec_constraints <- lavPredict(lfit)[, 1]
@@ -29,14 +30,16 @@ input.df <- select(final.df, lepisode_onset, country_name, year,
            meanelev = normalize(meanelev),
            country_name = as.factor(country_name),
            lpeace_yrs = log(peace_yrs + 1),
-           year_fac = as.factor(year))
+           year_fac = as.factor(year)) %>%
+    na.omit
 
-info(input.df)
+dbg_info(input.df)
 ml <- gam(lepisode_onset ~ exec_constraints * cgdppc + gdpgro +
               pop_density + meanelev + rlvt_groups_count +
-              neighbour_conflict + lpeace_yrs + s(country_name, bs = "re") +
-              s(year_fac, bs = "re"),
+              neighbour_conflict + lpeace_yrs +
+              s(country_name, bs = "re") + s(year_fac, bs = "re"),
           data = input.df, family = "binomial")
+summary(ml)
 
 print("Model finished!")
 saveRDS(ml, "data/gam_model.rds")

### 2019-08-21 R/fa.R
Mediate lg vars in FA by presence of leg. (lgbicam)
v2lg* variables are missing where there's no legislature. Borrow the
strategy from Kyle's v2x_horacc where the lg variables are estimated
from the latent factor depending on the interaction with the
coefficients from the regression predicting v2lgbicam.

@@ -1,0 +1,66 @@
+#!/usr/bin/env Rscript
+
+suppressMessages(library(dplyr))
+suppressMessages(library(rstan))
+suppressMessages(library(thesis.utils))
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = T)
+
+load("data/prepped_data.RData")
+
+lg_vars <- c("v2lginvstp", "v2lgfunds", "v2lgqstexp", "v2lgoppart", "v2lgotovst")
+nonlg_vars <- c("v2juhcind", "v2juncind", "v2juhccomp", "v2jucomp",
+                "v2exrescon")
+
+lg_obs <- filter(constraints.df, v2lgbicam == 1) %>%
+    select(one_of(lg_vars)) %>%
+    data.matrix
+
+lg_se <- filter(constraints.df, v2lgbicam == 1) %>%
+    select(one_of(paste0(lg_vars, "_sd"))) %>%
+    data.matrix
+
+#for (i in 1:ncol(lg_obs)) {
+#    lg_se[, i]  <- lg_se[, i] / sd(lg_obs[, i])
+#    lg_obs[, i]  <- normalize(lg_obs[, i])
+#}
+
+
+nonlg_obs <- select(constraints.df, one_of(nonlg_vars)) %>%
+    data.matrix
+
+nonlg_se <- select(constraints.df, one_of(paste0(nonlg_vars, "_sd"))) %>%
+    data.matrix
+
+#for (i in 1:ncol(nonlg_obs)) {
+#    nonlg_se[, i]  <- nonlg_se[, i] / sd(nonlg_obs[, i])
+#    nonlg_obs[, i]  <- normalize(nonlg_obs[, i])
+#}
+
+lg_present <- constraints.df$v2lgbicam == 1
+
+data <- list(N = nrow(constraints.df),
+             lg_D = ncol(lg_obs),
+             nonlg_D = ncol(nonlg_obs),
+             lg = lg_obs,
+             lg_se = lg_se,
+             nonlg = nonlg_obs,
+             nonlg_se = nonlg_se,
+             lgotovst_idx = which(colnames(lg_obs) == "v2lgotovst"),
+             lgbicam = constraints.df$v2lgbicam,
+             N_missing = sum(!lg_present),
+             N_obs = sum(lg_present),
+             missing_idx = which(!lg_present),
+             obs_idx = which(lg_present))
+
+str(data)
+
[…]

### 2019-08-21 R/model.R

@@ -1,4 +1,7 @@
 #!/usr/bin/env Rscript
+#
+# Model script
+###
 
 suppressMessages(library(dplyr))
 suppressMessages(library(rstan))
@@ -7,55 +10,130 @@ suppressMessages(library(thesis.utils))
 options(mc.cores = parallel::detectCores())
 rstan_options(auto_write = T)
 
+dir.create("posteriors/summary", recursive = T, showWarnings = F)
+
 load("data/prepped_data.RData")
 
-manifests <- select(constraints.df, one_of(constraint_vars)) %>%
-    data.matrix
+###
+# Manifest vars for exec_constraints
+lg_vars <- grep("^v2lg", constraint_vars, value = T)
+nonlg_vars <- grep("^v2[^l]", constraint_vars, value = T)
 
-manifests_sd <- select(constraints.df, one_of(paste0(constraint_vars, "_sd"))) %>%
-    data.matrix
+prep <- function(df, vars) {
+    m <- select(df, matches(paste0(vars, collapse = "|"))) %>%
+        data.matrix
 
-for (i in 1:ncol(manifests)) {
-    manifests_sd[, i] <- manifests_sd[, i] / sd(manifests[, i])
-    manifests[, i] <- normalize(manifests[, i])
+    obs <- m[, vars]
+    se <- m[, paste0(vars, "_sd")]
+
+    for (i in 1:ncol(obs)) {
+        se[, i] <- se[, i] / sd(obs[, i])
+        obs[, i] <- normalize(obs[, i])
+    }
+
+    list(obs = obs, se = se)
 }
 
-stopifnot(nrow(manifests) == nrow(manifests_sd),
-          ncol(manifests) == ncol(manifests_sd))
+lg_present <- constraints.df$v2lgbicam == 1
 
+nonlg_mm <- prep(constraints.df, nonlg_vars)
+lg_mm <- filter(constraints.df, v2lgbicam == 1) %>%
+    prep(lg_vars)
+
+stopifnot(nrow(nonlg_mm$obs) == nrow(nonlg_mm$se),
+          ncol(nonlg_mm$obs) == ncol(nonlg_mm$se))
+stopifnot(nrow(lg_mm$obs) == nrow(lg_mm$se),
+          ncol(lg_mm$obs) == ncol(lg_mm$se))
+
+###
+# Control variables
 X <- select(final.df, gdpgro, pop_density, meanelev, rlvt_groups_count,
             neighbour_conflict, peace_yrs) %>%
[…]

### 2019-08-21 R/prep.R

@@ -5,41 +5,22 @@ suppressMessages(library(dplyr))
 suppressMessages(library(magrittr))
 suppressMessages(library(thesis.utils))
 
-constraint_vars <- c("v2lginvstp", "v2lgfunds", "v2lgqstexp", "v2lgoppart",
-                     "v2juhcind", "v2juncind", "v2juhccomp", "v2jucomp",
+constraint_vars <- c("v2lginvstp", "v2lgfunds",
+                     "v2lgqstexp", "v2lgoppart",
+                     "v2juhcind", "v2juncind",
+                     "v2juhccomp", "v2jucomp",
                      "v2exrescon", "v2lgotovst")
 
 merged.df <- readRDS("data/merged_data.rds")
-
-###
-# Where lg variables are missing due to v2lgbicam == 0 (ie no
-# legislature), fill in with minimum values from posteriors.
-fn <- function(f) {
-    print(f)
-
-    df <- fread(f, skip = 1, header = F)
-    v <- apply(df[, -1], 2, min, na.rm = T)
-
-    list(mean = mean(v), sd = sd(v))
-}
-
 lg_vars <- grep("^v2lg", constraint_vars, value = T)
-re <- paste0(lg_vars, collapse = "|")
-
-files <- list.files("./data/raw/vdem_post", re, full.names = T)
-stopifnot(length(files) == length(lg_vars))
-
-min.df <- lapply(files, fn) %>%
-    bind_rows %>%
-    mutate(variable = lg_vars)
 
 for (v2 in lg_vars) {
     merged.df[[v2]] <- ifelse(merged.df$v2lgbicam == 0,
-                              min.df$mean[min.df$variable == v2],
+                              Inf,
                               merged.df[[v2]])
 
     merged.df[[paste0(v2, "_sd")]] <- ifelse(merged.df$v2lgbicam == 0,
-                                             min.df$sd[min.df$variable == v2],
+                                             Inf,
                                              merged.df[[paste0(v2, "_sd")]])
 }
 
@@ -48,19 +29,20 @@ for (v2 in lg_vars) {
 final.df <- merged.df %>%
     select(country_name, year, lonset, lepisode_onset, peace_yrs,
            one_of(constraint_vars), one_of(paste0(constraint_vars, "_sd")),
-           cgdppc, gdpgro, pop_density, meanelev, ongoing,
[…]

### 2019-09-12 paper.Rmd
Update stan syntax highlighting xml path
WIP: paper updates

--- 2019-07-22 paper.Rmd
@@
 ---
-title: Executive Constraints and Civil Conflict
+title: Executive Constraints and Civil Conflict Onset
 author: J.L.A. Krusell
-output: pdf_document
+date: \today
+documentclass: scrartcl
+fontsize: 11pt
+linestretch: 1.5
+header-includes:
+    - \usepackage{etoolbox}
+    - \usepackage{lualatex-math}
+    - \usepackage{unicode-math}
+    - \appto\Shaded{\singlespacing}
+output:
+    pdf_document:
+        latex_engine: lualatex
+        toc: true
+        toc_depth: 2
+        pandoc_args: ["--syntax-definition", "assets/stan.xml",
+                      "--highlight-style", "kate"]
 ---
 
-\[ x_{j}^{obs} \sim \mathsf{Normal}(x_{j}^\ast, \sigma_{j}^{obs}) \]
-\[ x_{ji}^\ast \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]
-\[ \theta \sim \mathsf{Normal}(0, 1) \]
+```{r include = F, results = "hide"}
+knitr::opts_chunk$set(echo = F, eval = F)
+
+library(dplyr)
+library(precrec)
+library(scales)
+library(thesis.utils)
+
+```
+
+## Research Design
+
+To examine the relationship between civil conflict onset and executive
+constraints, I construct a dataset of relevant variables and run a
+bayesian multilevel logistic regression. In what follows, I first
+describe the independent and dependentvariables, including the
+operationalizations and data sources, and then delve into the
+regression model before finally presenting the results.
+
+### Civil Conflict
+
+The dependent variable, *civil conflict onset*, was taken from the
+UCDP/PRIO Armed Conflict Dataset v19.1 (Cite) covering 1946-2018
+where intra-state conflict is defined as a violent contest over a
+government or territory between a state and non-state actor with a
+minimum threshold of 25 battle deaths within a single country-year.
+The data was merged based on country location for conflicts labeled as
+internal or internationalized internal --- *i.e.* where
+`type_of_conflict` was either 3 or 4. Multiple conflicts within the
+same years were aggregated to form a binary variable coded as 1 to
+denote the onset of at least one new civil conflict and 0 for all
+other years, including years of ongoing conflict.^[Extrasystemic
+conflicts, *i.e.* imperial or colonial conflicts, were excluded
+largely due to lack of data. Ideally, such conflicts would be merged
+based on the territory location of the conflict, however, for
+colonies, for example, there is very little easily available economic
+data and it would be undesirable to use the data for the colonizing
+power due to the often wide discrepancies in development within
+colonial empires. TODO: maybe a blurb on theory?]
+
+The primary onset dates were determined according to UCDP/PRIO's
+episodic definition where a new onset is coded for the same
+incompatibility when there is at least a one year pause in the
+fighting or drop below the battle death threshold. The primary
+rationale is that civil conflicts often go through several distinct
+phases of fighting which can make it difficult to accurately code the
+end of a conflict. The episodic definition avoids potentially
+conflating briefs lulls involving the same rebel groups and leaders
+versus the re-emergence of previously long dormant
+conflicts. Furthermore, in both scenarios, the theory linking
+executive constraints to the incentives for political violence should
+still hold (TODO: possible expand here).
+
+That said, because of the potential sensitivity of results to changing
+conflict definitions (cite Sambanis), all analyses were also run using
+only unique onsets.^[Likewise, appendix ? presents results
+using a higher threshold for battle deaths. The end results stay the
+same TODO]
+
+### Executive Constraints
+
+(TODO: Probably cite Sklar and/or O'Donnell here for more theoretical
+justification of operationalization)
+
+To measure formal executive constraints I use data from the Varieties
+of Democracy (V-Dem) dataset v9.0 (cite V-Dem). V-Dem is a
+cross-national time series dataset providing multidimensional,
+disaggregated data on democracy and political institutions for over
+200 countries with an overall year coverage from 1789 until 2018. The
+principal data is drawn from expert coded surveys involving a global
+network of over 3000 coders, mostly academics as well as professionals
+working in policy related fields (cite Methodology doc). Individual
+ratings are aggregated into latent variable estimates using
+hierarchical ordinal item response theory (IRT) models taking into
+account rater reliability and varying rater thresholds (cite Pemstein
+et al). ^[Thresholds relate to how the continuous latent trait for a
+given variable is mapped on to the ordinal answer categories presented
+on the survey. Due to differential item functioning, experts may
+perceive the magnitude of difference between answer items
+differently. For example, for the variable *v2elvotbuy*, election vote
+buying, experts may vary in how they interpret the difference between
+a 2 (Restricted Vote Buying) and a 3 (Almost No Vote Buying). These
+two concepts, rater reliability and varying rater thresholds,
+correspond respectively to the *discrimination* and *difficulty*
+parameters in classical IRT models.] The result is over 220 lower
+level indicators that are also combined through various aggregation
+techniques to produce over 80 mid/high-level indices. Compared to
+datasets such as Polity IV (cite), which has traditionally been
+used in most analyses of civil conflict, V-Dem provides much more fine
+grained, nuanced data, and most importantly, includes measures of
+uncertainty propagated from the IRT models. This latter point is
+crucial since ignoring measurement error, especially in models with
+noisy data and small effect sizes as is often the case in conflict
+studies, can bias estimates and lead to overconfident results (cite
+Treier & Jackman; Kyle).
+
+For the purposes of this paper, I re-estimate Lührmann *et al*'s
+horizontal accountability index (*v2x_horacc*) in V-Dem with the
+addition of several variables (cite). The *v2x_horacc* index aims to
+capture the extent to which the executive is *de facto* constrained by
+other government institutions --- more specifically, the legislature,
+judiciary, constitution, and independent investigative bodies (cite pg
+61) --- rather than the *de jure* powers of any one institutions
+vis-à-vis the executive. It is estimated as a latent variable using a
+structural equation model, thereby avoiding the arbitrary weighting of
+linear additive indices.^[A theoretical argument could be made that
+the oversight offered by each institution is a necessary condition and
+that any index should therefore be additive with equal
+weighting. However, this would be a normative measure capturing the
+degree to which a country-year fulfills an ideal concept, rather than
+measuring how constrained an executive actually is in practice.]
+
+For the two principal institutions, the legislature and the judiciary,
+to serve as effective constraints there needs to be both a *capacity*
+and *willingness* to oversee the executive. To better capture the
+former dimension, I include two additional legislative variables. In
+addition to *v2lgqstexp*, the extent to which the legislature
+questions executive branch officials, and *v2lginvstp*, the extent to
+which the legislature investigates in practice, I add *v2lgfunds*, the
+extent to which the legislature controls its own resources, and
+*v2lgoppart*, the extent to which opposition parties are able to
+exercise oversight. The remaining manifest variables are the
+same. Namely, for the judicial component: *v2juhcind*, high court
+independence; *v2juncind*, lower court independence; *v2juhccomp*, the
+degree to which the government complies with high court decisions;
+*v2jucomp*, the degree to which the government complies with lower
+court decisions.
+
+v2exrescon
+v2lgotovst
+
+How measured? FA model, include formula
+
+Measurement error (cite?)
+
+Special case for legislative variables
+
+Imputation for v2lgotovst
+
+DAG
+
+\[ x_{jt}^\ast \sim \mathsf{Normal}(\lambda_j \theta_t, \psi_t) \]
+\[ x_{jt}^{obs} \sim \mathsf{Normal}(x_{jt}^\ast, \tau_j^{obs}) \]
+
+### Controls
+
+- state capacity (GDP per capita)
+
+- gdp growth
+
+- population density
+
+- mean elevation
+
+- relevant group counts
+
+- neighbouring conflict
+
+- ongoing conflict
+
+- peace years
+
+- varying intercepts
+
+What's not included (political instability) and why
+
+### Model
+
+I model civil conflict onset as a bayesian multilevel regression where
+onset is a function of a set of predictors, including executive
+constraints and an array of control variables, together with several
+grouping factors included as partially pooled varying intercepts. The
+units of analysis are all independent country-years from 1946 until
+2016, as determined by data availability, with *n* countries indexed
+by $i = 1 \dotsc n$ and *m* years indexed by $t = 1 \dotsc m$. The
+dependent variable, conflict onset, is a Bernoulli distributed binary
+variable where 1 denotes the onset of at least one new civil conflict
+in a given country-year. The chosen link function transforming the
+linear regression output into the probability of conflict onset,
+$\pi_{it}$, is the inverse logistic such that $\pi_{it} \in [0, 1]$.
+^[Technically, the inverse logistic transforms values to the open
+interval, $(0, 1)$, but because of overflow/underflow in computer
+calculations due to floating point math the endpoints can be
+observed.]
+
+The full regression specification is then as follows:
+
+\[ y_{it} \sim \mathsf{Bernoulli}(\pi_{it}) \]
+\[ \pi_{it} = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta + \sum_{k=1}^K \delta^k_{j_{it}}) \]
+
+where **X** is a single row matrix of predictors, $\beta$ a vector of
+regression slopes, and $\delta^k$ the set of varying intercepts for
+*K* grouping factors indexed by $j_k$. To account for time
+endogeneity, all predictors are lagged by one year. They are also
+standardized by subtracting the mean and dividing by the standard
+deviation for each variable to aid in computation as well as
+comparison of the final $\beta$ estimates (cite Gelman maybe?).
+
+I place weakly informative priors on all parameters to allow the
+posterior estimates to be largely determined by the observed data
+while avoiding the more pathological values and hard constraints that
+can arise from uninformative priors (cite Stan). For the slopes and
+population intercept, given that the input data has been standardized,
+this translates into
+
+\[ \alpha \sim \mathsf{Normal}(0, 5) \]
+\[ \beta \sim \mathsf{Normal}(0, 2.5) \]
+
+Finally, for the varying intercepts modeling group-specific
+variability, I use hierarchical priors that provide a regularizing
+effect against overfitting (cite Gelman and Hill - Multilevel
+Regression).
+
+\[ \delta^k \sim \mathsf{Normal}(0, \sigma^k) \]
+\[ \sigma^k \sim \mathsf{HalfCauchy}(0, 1) \]
+
+## Analysis
+
+Estimated using HMC by stan (footnote number of iterations, burnin, and convergence)
+
+```{r}
+beta <- readRDS("posteriors/summary/beta.rds")
+plot_pars(beta)
+```
+
+```{r}
+input.df <- readRDS("posteriors/input_data.rds")
+p_hat <- readRDS("posteriors/summary/predicted_probs.rds")
+
+plot_dens(p_hat, groups = input.df$lepisode_onset)
+```
+
+```{r}
+p <- apply(p_hat, 2, median)
+x <- evalmod(scores = p, labels = input.df$lepisode_onset)
+plot(x, curvetype = c("ROC", "PRC"))
+```
+
+```{r eval = F}
+# Interaction plot
+intercept <- readRDS("posteriors/summary/intercept.rds")
+theta <- readRDS("posteriors/summary/theta.rds") %>% apply(2, median) %>% sort
+
+state <- quantile(input.df$cgdppc, probs = c(0.2, .5, 0.8))
+
+f <- function(x, y) {
+    with(input.df, intercept + beta[, 1] * x + beta[, 2] * y +
+                   beta[, 3] * x * y + beta[, 4] * mean(peace_yrs) +
+                   beta[, 5] * mean(gdpgro) + beta[, 6] * mean(pop_density)  +
+                   beta[, 7] * mean(meanelev) + beta[, 8] * mean(rlvt_groups_count) +
+                   beta[, 9] * 1 + beta[, 10] * 0)
+}
+
+par(mfrow = c(1, 3))
+for (y in state) {
+    print(y)
+    p <- boot::inv.logit(sapply(theta, f, y)) * 100
+
+    m <- apply(p, 2, median)
+    pi <- apply(p, 2, quantile, probs = c(0.2, 0.8)) %>% t
+
+    plot(theta, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
+         xlab = "Executive Constraints", ylab = "Predicted Probability")
+    lines(theta, pi[, 1], lty = 2)
+    lines(theta, pi[, 2], lty = 2)
+}
+
+```
+
+\newpage
+
+# Appendix
+
+## Stan Code
+
+```{stan, code = readLines("stan/model.stan"), eval = FALSE, output.var = "x", echo = T}
+```

### 2019-10-08 R/sim.R
Prior simulation script

@@ -1,0 +1,148 @@
+#!/usr/bin/env Rscript
+
+library(boot)
+library(dplyr)
+library(extraDistr)
+library(ggplot2)
+library(MASS)
+library(rstan)
+library(thesis.utils)
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = T)
+
+set.seed(6666)
+
+N <- 500
+D <- 6
+
+# Start with latent model
+theta <- rnorm(N)
+
+gamma <- rnorm(D, 0, 5)
+lambda <- rhnorm(D, 1)
+psi <- rweibull(D, 2, 1)
+
+# Legislative variables
+eta <- rnorm(1, 0, 1)
+delta <- rhnorm(1, 5)
+
+xi <- eta + delta * theta
+
+w <- rbinom(N, 1, inv.logit(xi))
+stopifnot(any(w == 1))
+
+N_obs <- sum(w)
+obs_idx <- which(w == 1)
+
+lg <- xi[obs_idx] %*% t(lambda[1:3]) + mvrnorm(N_obs, rep(0, 3), diag(psi[1:3]))
+lg <- sweep(lg, 2, gamma[1:3], `+`)
+
+lg_err <- matrix(rgamma(N_obs * ncol(lg), 5, 10) / 2, N_obs, ncol(lg))
+lg_obs <- matrix(NA, N_obs, ncol(lg))
+for (i in 1:ncol(lg))
+    lg_obs[, i] <- lg[, i] + rnorm(N_obs, 0, lg_err[, i])
+
+# Non-legislative
+nonlg <- theta %*% t(lambda[4:6]) + mvrnorm(N, rep(0, 3), diag(psi[4:6]))
+nonlg <- sweep(nonlg, 2, gamma[4:6], `+`)
+
+nonlg_err <- matrix(rgamma(N * ncol(nonlg), 5, 10) / 2, N, ncol(nonlg))
+nonlg_obs <- matrix(NA, N, ncol(nonlg))
+for (i in 1:ncol(nonlg))
+    nonlg_obs[, i] <- nonlg[, i] + rnorm(N, 0, nonlg_err[, i])
+
+# Conflict regression
+alpha <- rnorm(1, 0, 5)
+beta <- rnorm(4, 0, 2.5)
+
+x <- rnorm(N, 0, 2)
+state_capacity <- rnorm(N, 0, 3)
+
+n_years <- 50
+n_countries <- 10
+
+sigma <- rhcauchy(2, 1)
[…]

### 2019-10-08 R/thesis.utils/R/plot.R
WIP: plotting functions

@@ -1,0 +1,51 @@
+#' Parameter plot
+#'
+#' Plots posterior median plus 68/95 quantile intervals for the given
+#' parameters.
+#'
+#' @param m Posterior matrix where each column is a separate parameter
+#' @param labels Optional CharacterVector of x-axis labels.
+#'
+#' @export
+plot_pars <- function(m, labels = colnames(m)) {
+    df <- post_summarise(m, names = labels) %>%
+        dplyr::mutate(parameter = factor(parameter, levels = parameter))
+    colnames(df)[-ncol(df)] <- c("codelow95", "codelow68", "median",
+                                 "codehigh68", "codehigh95")
+
+    # TODO
+    ggplot(df, aes_(x = ~parameter, y = ~median)) +
+        geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
+        geom_errorbar(aes_(x = ~parameter, ymin = ~codelow95, ymax = ~codehigh95),
+                      color = "#0072B2", linetype = "dotted", width = 0) +
+        geom_errorbar(aes_(x = ~parameter, ymin = ~codelow68, ymax = ~codehigh68),
+                      color = "#0072B2", width = 0, size = 1) +
+        geom_point(fill = "#56B4E9", color = "#0072B2", size = 2.5, shape = 21) +
+        theme(panel.background = element_blank(),
+              axis.line = element_line(color = "black"),
+              axis.text.x = element_text(angle = 45, hjust = 1),
+              axis.title.x = element_blank()) +
+        ylab("Parameter Estimates")
+}
+
+#' @export
+plot_dens <- function(m, n = 50, groups = NULL) {
+    df <- t(p_hat[1:n, ]) %>%
+        as.data.frame %>%
+        dplyr::mutate(id = as.factor(groups)) %>%
+        tidyr::gather(iteration, value, -id) %>%
+        dplyr::mutate(z = paste(id, iteration))
+
+    ggplot(df, aes_(~value, group = ~z, color = ~id)) +
+        stat_density(position = "identity", geom = "line",
+                     n = 1024, alpha = 0.2) +
+        theme(panel.background = element_blank(),
+              legend.position = c(0.9, 0.5),
+              legend.title = element_blank(),
+              axis.line = element_line(color = "black"),
+              axis.title.x = element_blank(),
+              axis.title.y = element_blank(),
+              axis.text.y = element_blank(),
+              axis.ticks.y = element_blank()) +
+        coord_cartesian(expand = F)
+}

### 2019-10-08 R/thesis.utils/R/thesis.utils.R

@@ -5,6 +5,7 @@
 #' and are not meant to be particularly useful outside the context of
 #' this analysis.
 #'
+#' @import ggplot2
 #' @docType package
 #' @name thesis.utils
 if (getRversion() >= "2.15.1")

### 2019-10-08 paper.Rmd
Research design, second draft

@@
 author: J.L.A. Krusell
-date: \today
+date: "`r format(Sys.time(), '%Y-%m-%d')`"
 documentclass: scrartcl
@@
 header-includes:
+    - \usepackage{amsmath}
     - \usepackage{etoolbox}
@@
     - \appto\Shaded{\singlespacing}
+    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
+    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
 output:
@@
                       "--highlight-style", "kate"]
+    html_document:
+        css: assets/sakura.css
+        smart: true
+        toc: true
 ---
@@
 ```{r include = F, results = "hide"}
+# Start by setting up our R env
 knitr::opts_chunk$set(echo = F, eval = F)
@@
 bayesian multilevel logistic regression. In what follows, I first
-describe the independent and dependentvariables, including the
+describe the independent and dependent variables, including the
 operationalizations and data sources, and then delve into the
@@
 The dependent variable, *civil conflict onset*, was taken from the
-UCDP/PRIO Armed Conflict Dataset v19.1 (Cite) covering 1946-2018
-where intra-state conflict is defined as a violent contest over a
-government or territory between a state and non-state actor with a
-minimum threshold of 25 battle deaths within a single country-year.
-The data was merged based on country location for conflicts labeled as
-internal or internationalized internal --- *i.e.* where
-`type_of_conflict` was either 3 or 4. Multiple conflicts within the
-same years were aggregated to form a binary variable coded as 1 to
+UCDP/PRIO Armed Conflict Dataset v19.1 (Cite) where intra-state
+conflict is defined as a violent contest over a government or
+territory between a state and non-state actor with a minimum threshold
+of 25 battle deaths within a single country-year with a coverage from
+1946 to 2018.  The data was merged based on country location for
+conflicts labeled as internal or internationalized internal --- *i.e.*
+where *type_of_conflict* was either 3 or 4. Multiple conflicts within
+the same years were aggregated to form a binary variable coded as 1 to
 denote the onset of at least one new civil conflict and 0 for all
@@
 (TODO: Probably cite Sklar and/or O'Donnell here for more theoretical
-justification of operationalization)
+justification of operationalization, maybe even define institutions as
+well)
 
@@
 level indicators that are also combined through various aggregation
-techniques to produce over 80 mid/high-level indices. Compared to
-datasets such as Polity IV (cite), which has traditionally been
-used in most analyses of civil conflict, V-Dem provides much more fine
-grained, nuanced data, and most importantly, includes measures of
-uncertainty propagated from the IRT models. This latter point is
-crucial since ignoring measurement error, especially in models with
-noisy data and small effect sizes as is often the case in conflict
-studies, can bias estimates and lead to overconfident results (cite
-Treier & Jackman; Kyle).
+techniques to produce over 80 indices. Compared to datasets such as
+Polity IV (cite), which has traditionally been used in most analyses
+of civil conflict, V-Dem provides much more fine grained, nuanced
+data, and most importantly, includes measures of uncertainty
+propagated from the IRT models. This latter point is crucial since
+ignoring measurement error, especially in models with noisy data and
+small effect sizes as is often the case in conflict studies, can bias
+estimates and lead to overconfident results (cite Treier & Jackman;
+Kyle).
 
@@
 other government institutions --- more specifically, the legislature,
-judiciary, constitution, and independent investigative bodies (cite pg
-61) --- rather than the *de jure* powers of any one institutions
-vis-à-vis the executive. It is estimated as a latent variable using a
-structural equation model, thereby avoiding the arbitrary weighting of
-linear additive indices.^[A theoretical argument could be made that
-the oversight offered by each institution is a necessary condition and
-that any index should therefore be additive with equal
-weighting. However, this would be a normative measure capturing the
-degree to which a country-year fulfills an ideal concept, rather than
-measuring how constrained an executive actually is in practice.]
+judiciary, constitutional norms, and independent investigative bodies
+(cite pg 61) --- rather than the *de jure* powers of any one
+institution *vis-à-vis* the executive. It is estimated as a latent
+variable using a structural equation model thereby avoiding the
+arbitrary weighting of linear additive indices.^[A theoretical
+argument could be made that the oversight offered by each institution
+is a necessary condition and that any index should therefore be
+additive with equal weighting. However, this would be a normative
+measure capturing the degree to which a country-year fulfills an ideal
+concept, rather than measuring how constrained an executive actually
+is in practice.]
 
-For the two principal institutions, the legislature and the judiciary,
-to serve as effective constraints there needs to be both a *capacity*
-and *willingness* to oversee the executive. To better capture the
-former dimension, I include two additional legislative variables. In
-addition to *v2lgqstexp*, the extent to which the legislature
-questions executive branch officials, and *v2lginvstp*, the extent to
-which the legislature investigates in practice, I add *v2lgfunds*, the
-extent to which the legislature controls its own resources, and
-*v2lgoppart*, the extent to which opposition parties are able to
-exercise oversight. The remaining manifest variables are the
-same. Namely, for the judicial component: *v2juhcind*, high court
-independence; *v2juncind*, lower court independence; *v2juhccomp*, the
-degree to which the government complies with high court decisions;
-*v2jucomp*, the degree to which the government complies with lower
-court decisions.
+For the legislature and the judiciary, the two principal government
+branches most able to check the power of the executive, there needs to be
+both a *capacity* and *willingness* for oversight. To better capture
+the former dimension, I include two additional legislative
+variables. Besides *v2lgqstexp*, legislature questions executive
+branch officials, and *v2lginvstp*, legislature investigates in
+practice, I add *v2lgfunds*, legislature controls its own resources,
+and *v2lgoppart*, opposition parties able to exercise oversight. The
+remaining manifest variables are unchanged. Namely, for the judicial
+component: *v2juhcind*, high court independence; *v2juncind*, lower
+court independence; *v2juhccomp*, the degree to which the government
+complies with high court decisions; *v2jucomp*, the degree to which
+the government complies with lower court decisions. Norms relating to
+the status of the constitution are captured by *v2exrescon*, executive
+respects constitution, and the final dimension, independent
+investigative bodies, by *v2lgovost*, executive oversight by bodies
+other than the legislature.
 
-v2exrescon
-v2lgotovst
+For the estimation of the latent index I largely follow the approach
+taken by *v2x_horacc* with two notable exceptions. The first is that I
+model the measurement error in the manifest variables using the
+reported posterior standard deviations from V-Dem --- available as
+_v2\*\_sd_. The second is that for computational reasons I allow the
+judicial variables to impact the latent factor directly rather than
+nested within a second latent factor (TODO: better justification). The
+result is the following model specification.
 
-How measured? FA model, include formula
+For $m$ country-years indexed by $i \in I = \{1, \dotsc, n\}$ and $n$
+manifest variables indexed by $p \in P = \{1, \dotsc, n\}$, let $Z \in
+\mathbb{R}^{m \times n}$ be the matrix of manifest variables measured
+with an observed degree of error, $Z^\ast \in \mathbb{R}^{m \times n}$
+the matrix of latent true values, and $\theta$ the unidimensional
+latent factor measuring executive constraints. Further, let $Q(Z_{\ast
+p})$ be a property defined as true when $Z_{\ast p}$ denotes a
+legislative variables and otherwise false.
 
-Measurement error (cite?)
+I first model the measurement error in the manifest variables by
+assuming that the observed values are drawn from a normal distribution
+where the location parameters are the latent true values and the scale
+the observed posterior standard deviations (cite McElreath).^[An even
+more sensible approach would be to directly embed the individual IRT
+models aggregating the manifest variables into the larger executive
+constraints measurement model (cite Kyle). However, this prohibitively
+increases the run time of the model.]
+\[ Z_{ip} \sim \mathsf{Normal}(Z^\ast_{ip}, \tau_{ip}) \]
 
-Special case for legislative variables
+The latent values for the manifest variables are then modeled as
+functions of the latent factor $\theta$. For the non-legislative
+variables the mapping functions are identical to a standard factor
+analytic approach.
+\[
+  Z^\ast_{ip} \sim \mathsf{Normal}(\gamma_p + \lambda_p \theta_i, \psi_p)
+  \mbox{\quad if } p \in P \mid \neg Q(Z_{\ast p})
+\]
 
-Imputation for v2lgotovst
+where $\lambda$ is the variable specific factor loading, $\gamma$ the
+population intercept, and $\psi$ the scale of the error terms.
 
-DAG
+In the case of the legislative variables there is the added issue of
+missing observations for country-years lacking a
+legislature.[^missing] For those observations, given the centrality of
+the legislature as a potential constraint against the executive, the
+latent factor should ideally be lower conditional on the other
+manifest variables. However, since the manifest variables are
+themselves latent variables, minimal values cannot be imputed without
+a nonlinear transformation to a bounded scale. I thus follow
+*v2x_horacc* in mediating the effect of the latent factor on the
+legislative variables by the presence of the legislature using a
+dichotomized version of *v2lgbicam*, how many chambers does the
+legislature contain, denoted by the Bernoulli distributed variable,
+$w$, where country-years with a sitting legislature are given a $1$
+and country-years where the legislature is missing or has been
+suspended, $0$. The probability of a legislature being present,
+$\pi_i$, is then a linear function of the latent factor, $\theta$,
+transformed by the inverse logistic such that $\pi_i \in [0,
+1]$.^[Technically, the inverse logistic transforms values to the open
+interval, $(0, 1)$, but due to overflow/underflow in floating point
+calculations by computers the endpoints can be observed.] The
+legislative manifest variables are in turn treated as functions of the
+linear predictor for $w$ (cite pg 63). The practical effect is that
+where there is data for the legislative manifest variables, the latent
+factor is informed by the measured institutional quality of the
+legislature. Where data is missing due to the suspension or lack of a
+legislature, the latent factor is down weighted by the linear function
+predicting $w$, conditional on the observed values of the remaining
+manifest variables.
+\begin{align*}
+  w_i & \sim \mathsf{Bernoulli}(\pi_i) \\
+  \pi_i & = \mathsf{logit}^{-1}(\xi_i) \\
+  \xi_i & = \eta + \delta \theta_i \\
+  Z^\ast_{ip} & \sim \mathsf{Normal}(\gamma_p + \lambda_p \xi_i, \psi_p)
+    \mbox{\quad if } p \in P \mid Q(Z_{\ast p}) \land w_i = 1
+\end{align*}
 
-\[ x_{jt}^\ast \sim \mathsf{Normal}(\lambda_j \theta_t, \psi_t) \]
-\[ x_{jt}^{obs} \sim \mathsf{Normal}(x_{jt}^\ast, \tau_j^{obs}) \]
+[^missing]: Missingness due to the lack of of a particular
+institution, for example the suspension of high court, is an issue for
+the other manifest variables as well. However, only the existence of
+the legislature is coded by V-Dem as a separate variable. Furthermore,
+practically speaking, none of the other manifest variables exhibit the
+same systematic pattern of missingness in the dataset as the
+legislative variables. Whereas there are 1608 country-years lacking a
+legislature ($13.4\%$ of all post-1945 country-years), only 115 are
+missing data for at least one of the judicial variables ($1\%$) and
+only 58 are missing for *v2exrescon* ($0.5\%$). The variable
+*v2lgotovst*, executive oversight by bodies other than the
+legislature, is an outlier with 1619 country-years missing ($13.5\%$)
+since there is an error in V-Dem's v9.0 dataset whereby it has been
+cleaned according to when the legislature is missing. Thus, I impute
+*v2lgotovst* within the larger model where the legislature is missing
+by drawing missing values from a normal distribution with a mean and
+standard deviation estimated from the observed *v2lgotovst* data. No
+other country-years or variables are imputed. The rationale is that,
+disregarding data processing issues, missingness in V-Dem can arise
+either from a lack of expert coders or lack of information. The latter
+scenario will most likely be correlated with country-years where there
+is a crisis or significant political instability, such as coups or
+civil war, where extreme values are more probable. Since it cannot be
+assumed that the values are completely missing at random, the data
+generative process for missingness cannot be modeled without
+potentially introducing bias into the estimates.
+
+Finally, for the both the legislative and non-legislative mapping
+functions I assume local item independence such that the elements of
+$\Psi$ are uncorrelated (cite Lord & Novick). Furthermore, for the
+purposes of identifying the likelihood the scale of the latent factor,
+$\theta$, is fixed to mean $0$ and standard deviation $1$ while the
+slopes, $\Lambda$ and $\delta$, are constrained to be positive (cite
+Jackman). The positive constraints are drawn from theory since all of
+the manifest variables exhibit the same scale direction and should be
+positively correlated with the latent factor. Similarly, the presence
+of the legislature should be positively correlated with higher values
+of the latent factor, *i.e.* greater executive constraints.
 
 ### Controls
+
+In addition to the data on conflict onset and executive constraints, I
+also include a number of standard controls drawn largely from previous
+literature, the most important of which is state capacity. According
+to the theory outlined earlier, the impact of executive constraints on
+decreasing the probability of conflict onset should be conditional on
+the relative capacity of the state. The lack of executive constraints
+most likely increases the risk of civil conflict when state capacity
+is low and the central government is willing to use repression, but
+lacks the ability to fully suppress the opposition. To measure state
+capacity I follow Fearon & Laitin in using GDP per capita as a proxy
+for the state's administrative and military capacity by roughly
+capturing the revenue potential available to the state (cite). As a
+data source, I use the natural log of the *RGDP^e* variable, constant
+price real GDP, from Penn World Tables v9.1 (1950-2017) to allow for
+both cross-country and cross-temporal comparisons of GDP per capita
+(cite)^[TODO: Expand on why Penn World Tables] and include an
+interaction term between GDP per capita and the latent measure for
+executive constraints to capture the conditional relationship.
 
@@
 
-### Model
+### Regression Model
 
@@
 linear regression output into the probability of conflict onset,
-$\pi_{it}$, is the inverse logistic such that $\pi_{it} \in [0, 1]$.
-^[Technically, the inverse logistic transforms values to the open
-interval, $(0, 1)$, but because of overflow/underflow in computer
-calculations due to floating point math the endpoints can be
-observed.]
+$\nu_{it}$, is the inverse logistic such that $\nu_{it} \in [0, 1]$.
 
 The full regression specification is then as follows:
-
-\[ y_{it} \sim \mathsf{Bernoulli}(\pi_{it}) \]
-\[ \pi_{it} = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta + \sum_{k=1}^K \delta^k_{j_{it}}) \]
+\begin{align*}
+  y_{it} & \sim \mathsf{Bernoulli}(\nu_{it}) \\
+  \nu_{it} & = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta + \sum_{k=1}^K \zeta^k_{j_{it}})
+\end{align*}
 
@@
 
-Estimated using HMC by stan (footnote number of iterations, burnin, and convergence)
+Measurement model:
+\begin{align*}
+  \theta & \sim \mathsf{Normal}(0, 1) \\
+  \delta, \lambda & \sim \mathsf{LogNormal}(0, 1) \\
+  \eta, \gamma & \sim \mathsf{Normal}(0, 5) \\
+  \psi & \sim \mathsf{Weibull(2, 1)}
+\end{align*}
+
+Regression parameters:
+\begin{align*}
+  \beta & \sim \mathsf{Normal}(0, 2.5) \\
+  \alpha & \sim \mathsf{Normal}(0, 5) \\
+  \zeta^k & \sim \mathsf{Normal}(0, \sigma^k) \\
+  \sigma & \sim \mathsf{HalfCauchy}(0, 1)
+\end{align*}
 
@@
 ```
+
+## Prior Predictive Check

### 2019-10-08 R/merge.R
Remove inherited conflicts, since peace_yrs is set at 0 for indep.

@@ -132,6 +132,7 @@ acd <- readRDS("data/raw/UcdpPrioConflict_v19_1.rds")
 
 # Civil conflict + extrasystemic (colonial/imperial wars)
 # conflict_id = 418 is the United States vs Al Qaeda starting in 2001
+# --- why isn't extrasystemic??
 civil <- filter(acd, type_of_conflict %in% c(1, 3, 4), conflict_id != 418) %>%
     mutate(gwno_loc = as.character(gwno_loc)) %>%
     separate_rows(gwno_loc, sep = ",") %>%
@@ -144,27 +145,6 @@ filter(ctable, gwid %in% setdiff(civil$gwno_loc, vdem$gwid)) %$%
 counts.df <- group_by(civil, gwno_loc, year) %>%
     summarise(n_conflicts = n(),
               intensity = max(intensity_level),
-              type_of_conflict = max(type_of_conflict))
-
-# Also inherit conflict counts from parent country for newly
-# independent countries. We inherit based on the territory location of
-# the conflict. This won't affect our `ongoing` variable in the final
-# dataset, but it will change our peace_yrs calculation, for example:
-# Eritrea, Namibia etc etc, which would otherwise have inflated
-# counts.
-inherited.df <- civil %>%
-    mutate(territory_gwno = ctable$gwid[match(territory_name, ctable$country_name)]) %>%
-    filter(!is.na(territory_gwno) & gwno_loc != territory_gwno) %>%
-    group_by(territory_gwno, year) %>%
-    summarise(n_conflicts = n(),
-              intensity = max(intensity_level),
-              type_of_conflict = max(type_of_conflict)) %>%
-    rename(gwno_loc = territory_gwno)
-
-counts.df %<>% bind_rows(inherited.df) %>%
-    group_by(gwno_loc, year) %>%
-    summarise(n_conflicts = sum(n_conflicts),
-              intensity = max(intensity),
               type_of_conflict = max(type_of_conflict))
 
 neighbours.df <- readRDS("data/neighbours.rds") %>%
@@ -182,7 +162,7 @@ episode_ucdp <- group_by(civil, gwno_loc, conflict_id, start_date2) %>%
     summarise(episode_onset = 1,
               year = min(year),
               start_prec2 = ifelse(first(start_prec2) %in% 6:7, 1, 0),
-              gwno_a = first(gwno_a),
+              episode_gwno_a = first(gwno_a),
               episode_intensity = max(intensity_level),
[…]

### 2019-10-08 R/model.R
Remove attempt to capture warning messages from stan

@@ -7,21 +7,29 @@ suppressMessages(library(dplyr))
 suppressMessages(library(rstan))
 suppressMessages(library(thesis.utils))
 
-options(warn = 2)
 options(mc.cores = parallel::detectCores())
 rstan_options(auto_write = T)
 
 dir.create("posteriors/summary", recursive = T, showWarnings = F)
 load("data/prepped_data.RData")
 
+final.df %<>%
+    select(country_name, year, lepisode_onset, peace_yrs, reduced_idx,
+           cgdppc, gdpgro, pop_density, meanelev, rlvt_groups_count,
+           neighbour_conflict, peace_yrs, ongoing) %>%
+    na.omit
+dbg_info(final.df)
+
+saveRDS(final.df, "posteriors/input_data.rds")
+
 ###
 # Control variables
-X <- select(final.df, gdpgro, pop_density, meanelev, rlvt_groups_count,
-            neighbour_conflict, peace_yrs) %>%
-    na.omit %>%
+X <- final.df %>%
+    select(-country_name, -year, -lepisode_onset, -reduced_idx, -cgdppc) %>%
     data.matrix
 
 lg_present <- lgbicam == 1
+print(colnames(lg_mm$obs))
 
 ###
 # Final stan input list
@@ -58,41 +66,27 @@ str(data)
 stopifnot(!sapply(data, anyNA))
 
 init <- list(lg_est = data$lg, nonlg_est = data$nonlg)
-tryCatch({
-    fit <- stan("stan/model.stan", data = data, seed = 101010,
-                iter = 4000, thin = 2, control = list(max_treedepth = 12),
-                init = rep(list(init), 4), include = F,
-                pars = c("psi_unif", "sigma_unif", "lg_est", "nonlg_est",
-                         "lg_missing", "nu", "raw_country", "raw_year",
-                         "eta", "theta_state_capacity"))
-}, warning = function(w) {
-    saveRDS(fit, "posteriors/fit_err.rds")
-    stop(w)
-})
+fit <- stan("stan/model.stan", data = data, seed = 101010,
+            iter = 4000, thin = 2, control = list(max_treedepth = 12),
+            init = rep(list(init), 4), include = F,
+            pars = c("psi_unif", "sigma_unif", "lg_est", "nonlg_est",
+                     "lg_missing", "nu", "raw_country", "raw_year",
+                     "eta", "theta_state_capacity"))
 
 print(fit, pars = c("gamma", "lambda", "psi", "alpha", "delta"))
[…]

### 2019-11-02 paper.Rmd
WIP: theory section

@@
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
-documentclass: scrartcl
+documentclass: scrreprt
 fontsize: 11pt
@@
 
+[//]: # Setup R environment for entire doc
 ```{r include = F, results = "hide"}
-# Start by setting up our R env
 knitr::opts_chunk$set(echo = F, eval = F)
@@
 library(dplyr)
-library(precrec)
-library(scales)
 library(thesis.utils)
+```
 
-```
+[//]: # Force TOC on separate page
+\newpage
+
+# Introduction {-}
+
+# Theoretical Overview {-}
+
+## Institutions and Civil Conflict
+
+This paper takes as a starting premise a definition of institutions as
+being "systems of established and embedded social rules that structure
+social interactions" (cite Hodgson). More informally, we can consider
+them to be the "rules of the game", encompassing for example codified
+laws, norms, and shared understandings, all of which can constrain or
+enable behaviour by influencing the perceptions and preferences of
+actors through incentives and deterrents.^[None of which is meant to
+imply that institutions stand external to actor perceptions and
+preferences as unchanging features. Rather, institutions are social
+constructs subject to changing normative values and able to be
+reinforced or undermined by behavioural feedback loops. This latter
+point is especially interesting within the context of civil conflict
+considering how the development of various institutions and conflict
+incidence can be jointly determined; however, it remains outside the
+scope of this paper.] Furthermore, over time, the habits formed from
+these systems create expectations of stable patterns of behaviour
+thereby shaping the interactions between actors by reducing
+uncertainty (cite North). Although there is a wide variety of
+institutions, the present focus will be on the type of institutions
+which govern political behaviour.
+
+This exposition is relevant to the study of civil conflict for both
+motivation-based and opportunity-based explanatory accounts. To the
+extent that the onset of civil conflict is a longer process featuring
+waves of escalation and de-escalation rather than a single one-off
+event (Florea), institutions are a key influence on the constitutive
+interactions between state and non-state actors. For example, they can
+shift the opportunity costs of violence altering the mobilization
+capacity of would be rebel groups. Institutions can also ensure
+credible commitments towards negotiated settlements, assuming that
+actors are rational and able to internalize the cost of violence such
+that a bargained solution is preferable to fighting (Blattman &
+Fearon). And, their quality can shape the capacity of the state to
+suppress or co-opt potential opposition. None of these examples are
+meant to form an exhaustive list. Instead the point is to demonstrate
+the number of different ways in which institutions can enter into the
+theoretical discussions about civil conflict.
+
+Unsurprisingly, insofar as the role of institutions has been examined
+by previous literature the focus has largely been on the broader
+notion of democracy, which at a minimum is a system regulating access
+to political power through popular representation determined by free
+and competitive elections (Schumpter). Democracy has been hypothesised
+to have a pacifying domestic effect according to two predominant
+mechanisms. The first is centered on the concept of relative
+deprivation, or inequalities in value expectations and value
+capabilities, as a primary source of discontent that can motivate
+collective violence (Gurr 1970). [^gurr] Democracy is thought to
+decrease the amount of political discontent by providing avenues for
+free expression and adjudicating group conflicts through non-violent
+institutional means (Hegre).
+
+[^gurr]: Gurr explains value expectations as being that which we think
+    we are entitled to and value capabilities as that which we think
+    we are capable of attaining. As Gurr then states, relative
+    deprivation is the perceived discrepancy between the two. The link
+    to collective violence, however, is conditional on the discontent
+    being politicized, by which it is meant that the source of the
+    discrepancy is identified and a normative and utilitarian
+    justification is given for violent action, before finally being
+    actualized.  Quantitative studies tend to focus on the first
+    stage, the propensity for political regimes to generate
+    discontent, given the difficultly of measuring the variables that
+    may affect politicisation and actualisation.
+
+## Executive Constraints
+
+# Empirics {-}
 
@@
 
-\newpage
+# Conclusion {-}
 
-# Appendix
+# Appendix {-}
 

### 2019-11-03 paper.Rmd
WIP: continue theory section

@@
 enable behaviour by influencing the perceptions and preferences of
-actors through incentives and deterrents.^[None of which is meant to
-imply that institutions stand external to actor perceptions and
-preferences as unchanging features. Rather, institutions are social
-constructs subject to changing normative values and able to be
-reinforced or undermined by behavioural feedback loops. This latter
-point is especially interesting within the context of civil conflict
-considering how the development of various institutions and conflict
-incidence can be jointly determined; however, it remains outside the
-scope of this paper.] Furthermore, over time, the habits formed from
-these systems create expectations of stable patterns of behaviour
-thereby shaping the interactions between actors by reducing
-uncertainty (cite North). Although there is a wide variety of
-institutions, the present focus will be on the type of institutions
+actors through incentives and deterrents. Furthermore, over time, the
+habits formed from these systems create expectations of stable
+patterns of behaviour thereby shaping the interactions between actors
+by reducing uncertainty (cite North). Although there is a wide variety
+of institutions, the present focus will be on the type of institutions
 which govern political behaviour.
+
+[^institutions]: None of which is meant to imply that institutions
+    stand external to actor perceptions and preferences as unchanging
+    features. Rather, institutions are social constructs subject to
+    changing normative values and able to be reinforced or undermined
+    by behavioural feedback loops. This latter point is especially
+    interesting within the context of civil conflict considering how
+    the development of various institutions and conflict incidence can
+    by jointly determined; however, it remains outside the scope of
+    this paper.
 
@@
 free expression and adjudicating group conflicts through non-violent
-institutional means (Hegre).
+institutional means (Hegre). The second mechanism highlights the risk
+of future shifts in the distribution of power in undermining the
+attempt to reach a bargained solution to avoid war between two parties
+in conflict. More specifically, a future change in favor of the
+relative strength of one of the parties can provide an incentive to
+later renege. This influences the present degree to which concessions
+can be offered and parties are willing to credibly commit
+(Powell). Within the domestic sphere, democracy offers a possible
+solution by formalising the regular contest of power with full
+enfranchisement, thereby providing a *de jure* guarantee for the
+balance of power (Acemoglu and Robinson), and should, likewise,
+decrease the likelihood of civil conflict.
 
@@
     may affect politicisation and actualisation.
+
+Despite the theoretical basis linking democracy to reduced risks for
+civil conflict, the empirical results tend to be mixed. Hegre *et al*
+(2001), for example, find that democracies are at a lower risk for
+civil conflict; however, no more so than fully authoritarian
+states. Instead, anocratic regimes in between democracy and
+authoritarianism are considered to be the most prone. The explanatory
+logic is that both democratic and fully authoritarian states tend to
+be consolidated regimes. In the case of the former, this means the
+presence of institutions conducive to conflict resolution, and in the
+case of the latter, the capacity to violently repress
+dissent. Anocratic regimes are by their very nature transitory, partly
+open and partly repressive, encouraging political discontent against
+the state without the capacity to fully suppress opposition.
+
+In contrast, studies emphasizing the importance of political and
+economic opportunities for rebellion in understanding when and where
+conflicts break out, such as Fearon and Laitin (2003) and Collier,
+Hoeffler, and Rohner (2009), find democracy to be insignificant. In
+similar fashion, democracy is associated with higher levels of
+political rights, respect for ethnic and religious minorities, and
+less repression, thus being a proxy for the latent level of political
+discontent against the state. However, even if the number of political
+grievances vary across regime types, they are still ubiquitous enough
+to be a poor explanation for why conflict breaks out in only a
+relatively small subset of countries. Instead, the argument follows
+that the structural conditions conducive to rebellion are the
+determining factors.[^opportunities]
+
+[^opportunities]: For Fearon and Laitin these structural conditions
+    are related to the features which make rebellion militarily
+    viable.  Specifically, there's a higher risk of conflict onset
+    where state capacity is weak, principally understood in terms of
+    state revenue and its impact on policing and counterinsurgency
+    operations; large impoverished populations; and insurgents have a
+    terrain advantage. Collier *et al* start from the premise that war
+    is expensive and the primary impediment to a rebellion is its
+    financial feasibility relative to the strength of the state. On
+    the basis of their results they go so far as provide support to
+    the proposition, "where a rebellion is financially and militarily
+    feasible it will occur." Both approaches, however, reach the same
+    conclusion concerning the empirical evidence surrounding the role
+    of democracy.
+
+Subsequent research
 

### 2019-11-03 scripts/run.sh
Update run script
`podman logs -f` does not EOF at container exit so instead just start
the container in attached mode, for only stdout and stderr, and manually
require a detach when running longer models, or just start everything in
tmux as usual and detach from there.

@@ -1,29 +1,24 @@
 #!/bin/sh
 #
-# CLI arguments are passed directly to make in the spawned
-# container. If the environmental variable CLEANUP is defined and set
-# to 1, then the container will be automatically removed following
-# completion.
+# Launches an attached container instance using either docker or
+# podman with CLI arguments passed directly to `make`. If the
+# environmental variable CLEANUP is defined and set to 1, then the
+# container will be automatically removed following completion.
 ###
 
 set -e
 
 which podman 2>&1 >/dev/null && CMD=podman || CMD=docker
-echo "Running with $CMD"
 
 # No arrays in posix sh...
-mount_opts="-v $PWD:/proj"
+mount_opts="-v $(git rev-parse --show-toplevel):/proj"
 if [ -d /mnt/data/posteriors ]; then
     echo "Mounting posterior map from /mnt/data"
     mount_opts="$mount_opts -v /mnt/data/posteriors:/proj/posteriors"
 fi
 
-$CMD run -d $mount_opts jsks/conflict_onset make $@
-container=$($CMD ps -l --format "{{.Names}}")
-
-echo "Launched: $container"
-$CMD logs -f $container
-
 if [ "${CLEANUP:-0}" = 1 ]; then
-    docker container rm $container
+    rm_opts="--rm"
 fi
+
+$CMD run -a stdout -a stderr $mount_opts $rm_opts jsks/conflict_onset make $@

### 2019-11-03 R/deps.R
Record R package versions from latest run

@@ -1,0 +1,9 @@
+#!/usr/bin/env Rscript
+#
+# Writes out all installed R packages into a csv. Should never be
+# called directly; just a quick way to list all the packages versions
+# in the `conflict_onset` image.
+###
+
+df <- installed.packages(noCache = T)[, c("Package", "Version")]
+write.csv(df[order(df[, "Package"]), ], "Rdependencies.csv", row.names = F)

### 2019-11-03 R/gam.R
Remove old GAM model

@@ -1,45 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(caTools)
-library(dplyr)
-library(lavaan)
-library(mgcv)
-library(thesis.utils)
-
-load("data/prepped_data.RData")
-
-manifests <- select(final.df, one_of(constraint_vars)) %>% data.matrix
-
-ml <- "exec_constraints =~ v2lginvstp + v2lgfunds + v2lgqstexp +
-                           v2lgoppart + v2juhcind + v2juncind +
-                           v2juhccomp + v2jucomp + v2exrescon +
-                           v2lgotovst"
-lfit <- cfa(ml, data = manifests, std.lv = T)
-summary(lfit, standardized = T, ci = T)
-
-final.df$exec_constraints <- lavPredict(lfit)[, 1]
-
-input.df <- select(final.df, lepisode_onset, country_name, year,
-                   exec_constraints, cgdppc, gdpgro,
-                   pop_density, meanelev, rlvt_groups_count,
-                   neighbour_conflict, peace_yrs, ongoing) %>%
-    mutate(pop_density = log(pop_density) %>% normalize,
-           exec_constraints = normalize(exec_constraints),
-           cgdppc = log(cgdppc) %>% normalize,
-           gdpgro = normalize(gdpgro),
-           meanelev = normalize(meanelev),
-           country_name = as.factor(country_name),
-           lpeace_yrs = log(peace_yrs + 1),
-           year_fac = as.factor(year)) %>%
-    na.omit
-
-dbg_info(input.df)
-ml <- gam(lepisode_onset ~ exec_constraints * cgdppc + gdpgro +
-              pop_density + meanelev + rlvt_groups_count +
-              neighbour_conflict + lpeace_yrs +
-              s(country_name, bs = "re") + s(year_fac, bs = "re"),
-          data = input.df, family = "binomial")
-summary(ml)
-
-print("Model finished!")
-saveRDS(ml, "data/gam_model.rds")

### 2019-11-08 paper.Rmd
First draft lit review

--- 2019-10-08 paper.Rmd
@@
         toc: true
+abstract: Lorem ipsum
 ---
 
+[//]: # Setup R environment for entire doc
 ```{r include = F, results = "hide"}
-# Start by setting up our R env
 knitr::opts_chunk$set(echo = F, eval = F)
@@
 library(dplyr)
-library(precrec)
-library(scales)
 library(thesis.utils)
+```
 
-```
+[//]: # Force TOC on separate page
+\newpage
+
+# Introduction {-}
+
+# Theoretical Overview {-}
+
+## Institutions and Civil Conflict
+
+This paper takes as a starting premise a definition of institutions as
+being "systems of established and embedded social rules that structure
+social interactions" [@Hodgson2006]. More informally, we can consider
+them to be the "rules of the game", encompassing for example codified
+laws, norms, and shared understandings, all of which can constrain or
+enable behaviour by influencing the perceptions and preferences of
+actors through incentives and deterrents.[^institutions] Furthermore,
+over time, the habits formed from these systems create expectations of
+stable patterns of behaviour thereby shaping the interactions between
+actors by reducing uncertainty [@North1990]. Although there is a wide
+variety of institutions, the present focus will be on the type of
+institutions which govern political behaviour.
+
+[^institutions]: None of which is meant to imply that institutions
+    stand external to actor perceptions and preferences as unchanging
+    features. Rather, institutions are social constructs subject to
+    changing normative values and able to be reinforced or undermined
+    by behavioural feedback loops. This latter point is especially
+    interesting within the context of civil conflict considering how
+    the development of various institutions and conflict incidence can
+    by jointly determined; however, it remains outside the scope of
+    this paper.
+
+This exposition is relevant to the study of civil conflict for both
+motivation-based and opportunity-based explanatory accounts. To the
+extent that the onset of civil conflict is a longer process featuring
+waves of escalation and de-escalation rather than a single one-off
+event [@Florea2012; @Sambanis2004b], institutions are a key influence
+on the constitutive interactions between state and non-state
+actors. For example, they can shift the opportunity costs of violence
+altering the mobilization capacity of would be rebel
+groups. Institutions can also determine the extent to which the state
+is seen as accountable to and representative of the larger population,
+thereby influencing the legitimacy of anti-government violence.  And,
+their quality can shape the capacity of the state to suppress or
+co-opt potential opposition. Although the preceding examples are far
+from forming an exhaustive list, there are thus a number of different
+ways in which institutions can enter into theoretical discussions
+about civil conflict.
+
+Unsurprisingly, insofar as the role of institutions has been examined
+by previous literature the focus has largely been on the broader
+notion of democracy, which at a minimum is a system regulating access
+to political power through popular representation determined by free
+and competitive elections [@Schumpeter1962]. Democracy serves as a
+proxy for a number of normatively desirable institutions and has been
+hypothesised to have a pacifying domestic effect according to two
+predominant mechanisms. The first is centered on the concept of
+relative deprivation, or inequalities in value expectations and value
+capabilities, as a primary source of discontent that can motivate
+collective violence [@Gurr1970]. [^gurr] Democracy is thought to
+decrease the amount of political discontent by providing avenues for
+free expression and adjudicating group conflicts through non-violent
+institutional means --- for example, through the regular occurrence of
+free and fair elections [@Hegre2014]. The second mechanism assumes
+that violent conflict is costly and that there exists a bargained
+solution preferable to fighting, assuming that actors are rational and
+able to internalize the cost of violence [@Fearon1995]. One of the
+reasons why a bargained solution may be impossible is the risk of
+future shifts in the distribution of power. More specifically, a
+future change in favor of relative strength of one of the parties can
+provide an incentive to later renege. This influences the present
+degree to which concessions can be offered and parties are willing to
+credibly commit [@Powell2006]. Within the domestic sphere, democracy
+offers a possible solution by formalising the regular contest of power
+with full enfranchisement, thereby providing a *de jure* guarantee for
+the balance of power [@Origins2006], and should, likewise, decrease
+the likelihood of civil conflict.
+
+[^gurr]: Gurr explains value expectations as being that which we think
+    we are entitled to and value capabilities as that which we think
+    we are capable of attaining. As Gurr then states, relative
+    deprivation is the perceived discrepancy between the two. The link
+    to collective violence, however, is conditional on the discontent
+    being politicized, by which it is meant that the source of the
+    discrepancy is identified and a normative and utilitarian
+    justification is given for violent action, before finally being
+    actualized.  Quantitative studies tend to focus on the first
+    stage, the propensity for political regimes to generate
+    discontent, given the difficultly of measuring the variables that
+    may affect politicisation and actualisation.
+
+Despite the theoretical basis linking democracy to reduced risks for
+civil conflict, the empirical results tend to be mixed. @Hegre2001,
+for example, find that democracies are at a lower risk for civil
+conflict; however, no more so than fully authoritarian
+states. Instead, anocratic regimes in between democracy and
+authoritarianism are considered to be the most prone. The explanatory
+logic is that both democratic and fully authoritarian states tend to
+be consolidated regimes. In the case of the former, this means the
+presence of institutions conducive to conflict resolution such that
+the opportunity for rebellion may be present, given the presence of
+civil liberties conducive to organizing, but there is a notable lack
+in motivation for collective violence. For the latter, authoritarian
+regimes may produce a significant amount of political discontent;
+however, the opportunity is lacking due to the state's capacity and
+willingness to violently repress dissent. Anocratic regimes are by
+their very nature transitory, partly open and partly repressive,
+encouraging political discontent against the state without the
+capacity to either accommodate or suppress opposition.
+
+In contrast, studies emphasizing the importance of political and
+economic opportunities for rebellion in understanding when and where
+conflicts break out, such as @Fearon2003 and @Collier2009, find
+democracy to be insignificant. According to their theoretical
+expositions, democracy should be associated with higher levels of
+political rights, respect for ethnic and religious minorities, and
+less repression, thus being a proxy for the latent level of political
+discontent against the state. However, the argument follows that even
+if the number of political grievances vary across regime types, they
+are still ubiquitous enough to be a poor explanation for why conflict
+breaks out in only a relatively small subset of countries. Instead,
+the determining factors for conflict onset are the structural
+conditions conducive to rebellion.[^opportunities]
+
+[^opportunities]: For Fearon and Laitin these structural conditions
+    are related to the features which make rebellion militarily
+    viable.  Specifically, there's a higher risk of conflict onset
+    where state capacity is weak, principally understood in terms of
+    state revenue and its impact on policing and counterinsurgency
+    operations; large impoverished populations; and insurgents having
+    a terrain advantage. Collier *et al* start from the premise that
+    war is expensive and the primary impediment to a rebellion is its
+    financial feasibility relative to the strength of the state. On
+    the basis of their results they go so far as provide support to
+    the proposition, "where a rebellion is financially and militarily
+    feasible it will occur." Both approaches, however, reach the same
+    conclusion concerning the empirical evidence surrounding the role
+    of democracy.
+
+To provide a more nuanced picture, an ongoing strain of research has
+been dedicated to disaggregating measures of democracy into either
+finer regime type categories or specific measures of individual
+political institutions. The theories offered tend to follow a general
+formula whereby formal lack or poor quality of some institution(s),
+measured either directly or proxied by regime type, undermines the
+capacity of the state to prevent armed insurgency and/or incentivises
+opposition violence by restricting the opportunities for political
+participation and increasing political discontent. For example,
+@Carey2007 examines the executive selection process of Sub-Saharan
+Africa and argues that executive elections, even when they are of poor
+quality, reduce the risk of large-scale violent dissent. @Fjelde2010a
+breaks down authoritarian regime types in Africa attributing the
+significant variation in conflict risk to the differing institutional
+configurations affecting a regime's ability to coerce or co-opt the
+opposition. The conclusion is that previous indeterminate results
+regarding the link between political institutions and conflict risk
+may have been hamstrung by aggregate measures which poorly reflect the
+within-group variation of political regimes. In a larger sample
+including democratic polities, @Goldstone2010 create a regime
+classification scheme based on measures for openness of executive
+recruitment and competitiveness of political participation, as well as
+factionalism in order to distinguish between partial democracies. They
+offer a theory of civil conflict occurring in countries where elite
+divisions undermine the capacity of the state to defeat an insurgency
+and claim that a relatively parsimonious model including their regime
+classification offers better predictive accuracy than previous
+analyses and provides evidence that political institutions are a
+dominant driving factor for both political instability and civil
+conflict. Finally, @Bartusevic2018 construct a regime typology focused
+specifically on electoral features, which are argued form the
+essential core of the concept of democracy. Where there are
+constraints on electoral contestation, the authors contend that such
+constraints restrict the opportunities for non-violent political
+participation thereby legitimising anti-government violence and
+self-selecting recruits into rebel groups willing to engage in armed
+conflict. Evidence is presented which the authors claim supports not
+just the significance of regime type, but a negative monotonic
+relationship between democracy and the risk of civil conflict.
+
+Studies utilising individual measures beyond regime classification
+have been largely focused on the concept of governance. For instance,
+@Taydas2010 examine institutional quality as a means to capturing a
+regime's governing capacity theorised to be linked to conflict onset
+both in terms of political discontent --- a regime's inability to
+govern effectively and respond to public needs leading to grievances
+and a loss in legitimacy --- and political opportunities for rebellion
+--- poor institutional quality undermines the state's ability to
+handle a nascent, violent uprising. Specifically, they assess
+corruption, rule of law, and bureaucratic quality and present evidence
+that all three are negatively related to civil conflict
+onset. Similarly, @Hegre2015a argue that the presence of formal
+representative institutions does not fully capture the extent to which
+a regime is capable of realizing policy goals benefiting the
+public. Poor quality of government can therefore undermine the
+legitimacy of a regime when citizens identify the state as responsible
+for relative deprivation. In their empirical analysis, the authors
+separate informal institutions related to good governance from formal
+democratic institutions and find both important for reducing the
+incidence of civil conflict, or in this specific case the recurrence
+of conflict.
+
+Taken together, this recent literature on civil conflict onset points
+to the ongoing importance of political institutions. There remains,
+however, significant work to be done identifying the causal mechanisms
+and effects of specific institutions, especially in terms of further
+disaggregating the concept of democracy beyond ever finer
+classifications of regime type [@ari2016]. Furthermore, to the extent
+that previous research has mostly focused on the effects of
+institutions on would-be rebel motivation and/or political opportunity
+structures, there is also a gap in terms of empirical analyses on how
+institutions shape the behaviour of the state in the lead up to a
+civil conflict. Even in strictly economic explanations of civil
+conflict, giving explanatory primacy either to greed-based motivation
+or economic conditions conducive for rebellion, the state is a key
+actor. How a state responds to dissent structures the opportunity
+costs of violence as well as the salience of calls to mobilization by
+rebel leaders. This latter point is especially relevant in linking
+regime legitimacy to conflict since public perception is shaped not
+simply by a regime's formal structure, but by its conduct in repeated
+interactions with the public. To this end, it is imperative to examine
+the types of institutions which enable or constrain actions by the
+state which exacerbate the risk of conflict onset beyond state
+capacity.
+
+There are of course a large number of institutions whose presence and
+quality can influence the behaviour of the state. Given the importance
+of the executive as a central authority, however, especially in times
+of political crisis, institutions which affect the executive's
+decision making should be pertinent. Chief among these is the idea of
+executive constraints, or the formal constraints on executive power by
+other the branches of government. By analyzing executive constraints
+we should get a better understanding of how executive power can
+influence the risk of conflict onset and hopefully be able to
+differentiate between anocratic regimes featuring similar levels of
+state capacity and quality of elections.
+
+## Executive Constraints
+
+# Empirics {-}
 

### 2019-11-16 scripts/build.sh
Docker format for podman build
Temporary fix to ensure that tagged images built by podman show up on
dockerhub. See: https://github.com/docker/hub-feedback/issues/1871

@@ -8,7 +8,7 @@ if [ -n "$1" ]; then
 fi
 
 if which podman 2>&1 >/dev/null; then
-    podman build -t "docker.io/jsks/conflict_onset$tag" .
+    podman build --format docker -t "docker.io/jsks/conflict_onset$tag" .
     podman push "docker.io/jsks/conflict_onset$tag"
 else
     docker build -t "jsks/conflict_onset$tag" .

### 2019-11-16 scripts/semver.sh
Update semver script

@@ -1,4 +1,9 @@
 #!/bin/sh
+#
+# Release script bumping current project version including git tag,
+# thesis.utils version, and container image according to semantic 
+# versioning.
+###
 
 ## Utility functions
 usage() {
@@ -15,8 +20,8 @@ bumping the current patch version.
 
 Options:
         -h Useless help message
-        -m Tagging message
-        -s Specify version number
+        -m Git tagging message
+        -s Specify version number in the form of MAJOR.MINOR.PATCH
 EOF
 
 exit
@@ -59,10 +64,6 @@ if [ -n "$(git diff --name-only --staged)" ]; then
     exit 127
 fi
 
-if [ -n "$(git status -s)" ] && ! confirm "Uncommitted changes"; then
-    exit 127
-fi
-
 root=$(git rev-parse --show-toplevel)
 tag=$(git describe --tags --abbrev=0 2>/dev/null | cut -c 2-)
 : ${tag:="0.0.0"}
@@ -84,6 +85,8 @@ while getopts 'hs:' opt; do
 done
 
 set -e
+
+[ -n "$(git status -s)" ] && confirm "Uncommitted changes"
 confirm "Bumping version to v$next_version"
 
 sed -i "s/Version:.*/Version: $next_version/" $root/R/thesis.utils/DESCRIPTION
@@ -92,4 +95,4 @@ git commit -m "Prepare release v$next_version"
 
 git tag -a "v$next_version" -m $msg
 
-sh $root/scripts/build.sh
+sh $root/scripts/build.sh "$next_version"

### 2019-11-16 paper.Rmd
Add initial bibliography

@@
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
+bibliography: library.bib
+csl: assets/american-journal-of-political-science.csl
+link-citations: true
 documentclass: scrreprt
@@
 
-# Appendix {-}
+# Appendices {-}
 
-## Stan Code
+## Appendix A - Stan Code {#appendix-a}
 
@@
 
-## Prior Predictive Check
+\newpage
+
+## Appendix B - Prior Predictive Check {#appendix-b}
+
+# References {-}
+
+\singlespacing

### 2019-11-16 scripts/semver.sh
BUGFIX: semver comparison and getopts

@@ -50,12 +50,12 @@ validate() {
     read next_major next_minor next_patch <<< "$1"
     IFS=""
 
-    if (( next_major < major )) ||
-           (( next_minor < minor )) ||
-           (( next_patch <= patch )); then
-        printf "Current version %s >= %s\n" "$tag" "$1"
-        exit 127
-    fi
+    (( next_major > major )) && return 0
+    (( next_minor > minor )) && return 0
+    (( next_patch > patch )) && return 0
+
+    printf "Current version %s >= %s\n" "$tag" "$1"
+    exit 127
 }
 
 ## Main
@@ -70,7 +70,7 @@ tag=$(git describe --tags --abbrev=0 2>/dev/null | cut -c 2-)
 
 next_version="${tag%.*}.$(( ${tag##*.} + 1 ))"
 
-while getopts 'hs:' opt; do
+while getopts 'hs:m:' opt; do
     case $opt in
         h)
             help;;

### 2019-11-16 scripts/semver.sh
git push after creating tag

@@ -94,5 +94,6 @@ git add $root/R/thesis.utils/DESCRIPTION
 git commit -m "Prepare release v$next_version"
 
 git tag -a "v$next_version" -m $msg
+git push --follow-tags
 
 sh $root/scripts/build.sh "$next_version"

### 2019-11-23 paper.Rmd
Use current image tag when running pipeline

@@
     html_document:
-        css: assets/sakura.css
         smart: true

### 2019-11-23 scripts/run.sh

@@ -8,6 +8,9 @@
 
 set -e
 
+tag=$(git describe --tags --abbrev=0 2>/dev/null | cut -c 2-)
+: ${tag:="latest"}
+
 which podman 2>&1 >/dev/null && CMD=podman || CMD=docker
 
 # No arrays in posix sh...
@@ -21,4 +24,5 @@ if [ "${CLEANUP:-0}" = 1 ]; then
     rm_opts="--rm"
 fi
 
-$CMD run -a stdout -a stderr $mount_opts $rm_opts jsks/conflict_onset make $@
+$CMD run -a stdout -a stderr $mount_opts $rm_opts \
+     "jsks/conflict_onset:$tag" make $@

### 2019-12-17 paper.Rmd
First draft: theory section

@@
 bibliography: library.bib
-csl: assets/american-journal-of-political-science.csl
+csl: assets/american-political-science-association.csl
 link-citations: true
@@
 over time, the habits formed from these systems create expectations of
-stable patterns of behaviour thereby shaping the interactions between
-actors by reducing uncertainty [@North1990]. Although there is a wide
-variety of institutions, the present focus will be on the type of
-institutions which govern political behaviour.
+stable patterns of behaviour thereby further shaping the interactions
+between actors by reducing uncertainty [@North1990]. Although there is
+a wide variety of institutions, the present focus will be on the type
+of institutions which govern political behaviour.
 
@@
 future shifts in the distribution of power. More specifically, a
-future change in favor of relative strength of one of the parties can
-provide an incentive to later renege. This influences the present
+future change in favor of the relative strength of one of the parties
+can provide an incentive to later renege. This influences the present
 degree to which concessions can be offered and parties are willing to
@@
 @Taydas2010 examine institutional quality as a means to capturing a
-regime's governing capacity theorised to be linked to conflict onset
-both in terms of political discontent --- a regime's inability to
-govern effectively and respond to public needs leading to grievances
-and a loss in legitimacy --- and political opportunities for rebellion
---- poor institutional quality undermines the state's ability to
-handle a nascent, violent uprising. Specifically, they assess
-corruption, rule of law, and bureaucratic quality and present evidence
-that all three are negatively related to civil conflict
+regime's governing capacity which is theorised to be linked to
+conflict onset both in terms of political discontent --- a regime's
+inability to govern effectively and respond to public needs leading to
+grievances and a loss in legitimacy --- and political opportunities
+for rebellion --- poor institutional quality undermines the state's
+ability to handle a nascent, violent uprising. Specifically, they
+assess corruption, rule of law, and bureaucratic quality and present
+evidence that all three are negatively related to civil conflict
 onset. Similarly, @Hegre2015a argue that the presence of formal
@@
 structures, there is also a gap in terms of empirical analyses on how
-institutions shape the behaviour of the state in the lead up to a
+institutions also shape the behaviour of the state in the lead up to a
 civil conflict. Even in strictly economic explanations of civil
-conflict, giving explanatory primacy either to greed-based motivation
-or economic conditions conducive for rebellion, the state is a key
-actor. How a state responds to dissent structures the opportunity
-costs of violence as well as the salience of calls to mobilization by
-rebel leaders. This latter point is especially relevant in linking
-regime legitimacy to conflict since public perception is shaped not
-simply by a regime's formal structure, but by its conduct in repeated
-interactions with the public. To this end, it is imperative to examine
-the types of institutions which enable or constrain actions by the
-state which exacerbate the risk of conflict onset beyond state
-capacity.
+conflict where explanatory primacy is given to either greed-based
+motivation or economic conditions conducive for rebellion, the state
+is a key actor. How a state responds to dissent structures the
+opportunity costs of violence as well as the salience of calls to
+mobilization by rebel leaders. This latter point is especially
+relevant in linking regime legitimacy to conflict since public
+perception is shaped not simply by a regime's formal structure, but by
+its conduct in repeated interactions with the public. To this end, it
+is imperative to examine the types of institutions which exacerbate
+the risk of conflict onset by enabling or constraining certain actions
+by the state.
 
-There are of course a large number of institutions whose presence and
-quality can influence the behaviour of the state. Given the importance
-of the executive as a central authority, however, especially in times
-of political crisis, institutions which affect the executive's
-decision making should be pertinent. Chief among these is the idea of
-executive constraints, or the formal constraints on executive power by
-other the branches of government. By analyzing executive constraints
-we should get a better understanding of how executive power can
-influence the risk of conflict onset and hopefully be able to
-differentiate between anocratic regimes featuring similar levels of
-state capacity and quality of elections.
+There are of course a large variety of political institutions whose
+presence and quality can influence the behaviour of the state.  Given
+the importance of the executive as a central authority, however,
+especially in times of political crisis, institutions which affect the
+executive's decision making should be pertinent. Chief among these is
+the idea of executive constraints --- more formally, the institutional
+constraints by other government bodies on the executive's ability to
+abuse power and act unilaterally. Executive constraints can be
+considered a subset of horizontal accountability and imply the
+existence of state agents --- such as the legislature or the judiciary
+--- with both the legal mandate and willingness to oversee or sanction
+the executive thus serving as a deterrent to the forms of
+authoritarian excesses in executive power which may be linked to civil
+conflict. [^ODonnell]
+
+[^ODonnell]: Adapted from the definition of horizontal accountability
+    offered by @ODonnell1998 who further argues that horizontal
+    accountability can potentially be a distinguishing feature between
+    consolidated democracies and newly democratic states, the latter
+    of which may have free and open elections, but are considerably
+    weaker in upholding liberal notions of individual freedom against
+    state power and republican notions of representation. In this way,
+    executive constraints can possibly aid in discriminating the types
+    of partially democratized regimes which may be at a heightened
+    risk for civil conflict.
+
+The final motivation for examining executive constraints is that in
+terms of previous research, the empirical evidence is
+conflicted. @Getmansky2013, for example, offers a theory of executive
+constraints playing a role in risk aversion in democratic polities;
+however, they find their measure to be insignificant in predicting the
+outbreak of insurgencies. On the other hand, @Fjelde2009 with a theory
+tying executive constraints to public goods provisioning reports a
+significant interaction term with public spending. In both cases an
+ordinal measure of executive constraints from Polity IV
+[@Marshall2019] was used. This leaves open both further work on
+additional theoretical mechanisms and analyses using other datasets.
 
 ## Executive Constraints
+
+There are three general pathways which can theoretically link the
+concept of executive constraints to the onset of civil conflict:
+executive looting and state weakness, credible commitments to
+bargained solutions, and the impact of violent repression by the state
+on rebel mobilization. In the case of the first, the absence of
+effective constraints enables leaders to loot state resources
+[@Blattman2017], diverting funds for the purposes of personal
+rent-seeking and corruption, which can undermine the administrative
+and military capacity of the state to combat a nascent
+insurgency. Furthermore, such environments allow leaders to pay
+selective benefits to close supporters at the expense of wider
+redistributive policies. To the extent that the provisioning of public
+goods may ameliorate grievances against the state and co-opt potential
+opposition, a reduced risk for civil conflict could be expected where
+there is a high level of redistributive public spending and the
+presence of quality institutions able to check the ability of the
+executive to engage in cronyism [@Fjelde2009]. Finally, the lack of
+accountability stemming from missing executive constraints could make
+the state an attractive prize to predatory rebel groups; however,
+prior empirical evidence lends little credence to this mechanism
+[@Bazzi2014].
+
+The second pathway relates to the willingness of both the regime and
+opposition to credibly commit to bargained solutions to avoid
+conflict. For the opposition, the presence of government institutions
+able to hold the executive to account with the threat, for example, of
+investigations, sanctions, or court rulings reduces the uncertainty
+over whether the current leader will later renege. From the
+perspective of the regime, executive constraints serve to limit future
+violent retributions if oppositional groups were to take power. This
+latter point is especially salient in situations where a leader has
+previously relied on violent repression to stay in power and fears
+extralegal punishment. It should be noted, however, that in such
+authoritarian settings the antecedent lack of effective executive
+constraints does not explain why the leader decides to negotiate in
+the first place. But, insofar as the bargained solution includes a
+transition of power with democratic reforms, and possibly even
+immunity for the current members of the regime, the inclusion of
+future checks against the executive serves as a guarantee for the
+current agreement thereby increasing the incentive to commit.
+
+The final pathway involves the impact of executive constraints on
+shaping the ability of rebel groups to recruit new members. In order
+to mount an effective challenge to the state, rebel leaders need to be
+able to mobilize fighters while avoiding defections. However,
+participating in an armed rebellion is a costly endeavour. Non-elites
+face both the immediate risk of fighting as well as the possibility of
+future reprisals by the government. Furthermore, given that most civil
+conflicts are protracted events against an often times much stronger
+opponent the short-term utility from participation is relatively
+low. Thus, rebel leaders must be able to offer proportionate selective
+benefits to offset the associated costs of fighting
+[@Mason1996]. [^rational] These benefits can include not only economic
+payoffs, but also protection from government violence. When the
+government engages in violent repression, especially in the form of
+indiscriminate violence, the cost of non-participation is
+raised. Rebel groups can provide local security guarantees against
+government reprisals such that the opportunity cost of fighting is
+lowered, thereby incentivizing participation [@Regan2005]. However,
+this is predicated on the government lacking the capacity to fully
+eradicate any rebel group. Where that capacity exists the state is
+free to engage in reprisals.
+
+[^rational]: TODO: Assuming rational actors...
+
+It is at this point where the concept of executive constraints then
+enters into the calculus. Leaders in nondemocratic states are more
+likely to engage in violent repression as a response to internal
+dissent [@Davenport2007a]. Executive constraints provide an
+accountability mechanism deterring state violence by the leader
+against their own citizens. Thus, where there are strong and effective
+constraints on the executive, the opportunity cost for joining the
+rebel side should be higher. Rebel leaders will then have to rely more
+heavily on economic selective benefits [@Regan2005], creating a higher
+barrier for mobilization and decreasing the risk of civil conflict
+onset. It should be noted though, that where this relationship may not
+hold is in situations of low intensity insurgencies involving
+predatory rebel groups solely motivated by rent-seeking. The threshold
+for mobilization in such scenarios is significantly lower than in full
+scale civil wars, meaning that rebel groups can initiate a conflict
+regardless of the current level of state repression.
+
+Taken together, none of the aforementioned mechanisms are mutually
+exclusive, and it is beyond the scope of the present paper to properly
+adjudicate between them; however, they do converge on a single
+testable proposition. Namely that higher levels of executive
+constraints should be associated with a lowered risk of civil conflict
+onset. That this may not be the case is either because executive
+constraints are indeterminate to the risk of civil conflict or that
+there is actually a positive relationship between the two. In the case
+of the latter, it could be possible that constrained executives are
+less able to effectively deal with violent challenges because of risk
+aversion to violence and/or from the slowdown from having to seek
+approval from various stakeholders before committing to decisive
+action. Given the theoretical pathways listed above, however, there is
+nonetheless a justified prior belief in favor of there existing a
+negative relationship.
+
+One final caveat, hitherto the relationship of executive constraints
+to the notion of vertical accountability --- constraints on leaders by
+free and open multiparty elections as well as an independent and
+critical media --- has been left unspecified. This is not to diminish
+the role of elections or the media; indeed, vertical accountability
+can plausibly play a role in all of the mentioned mechanisms. For
+example, studies on state repression have emphasised the importance of
+non-elites in being able to vote elected offices out of office as an
+effective deterrent against negative sanctioning by leaders
+[@BuenoDeMesquita2005; @Davenport2007]. But elections occur only
+intermittently and it is not an automatic given that voters will
+punish an incumbent, especially when that repression is against an
+oppressed minority community. Furthermore, while the media can expose
+wrongdoing by government officials, legal punishment can only flow
+through government bodies [@ODonnell1998]. Thus, executive constraints
+are interesting in their own right.
 
@@
 
+\newpage
+
+## Appendix C - Convergence Statistics {#appendix-c}
+
+\newpage
+
 # References {-}

### 2020-01-12 scripts/semver.sh
Print current version number in semver script

@@ -96,7 +96,7 @@ shift $(($OPTIND - 1))
 [ -n "$1" ] && usage
 
 [ -n "$(git status -s)" ] && confirm "Uncommitted changes"
-confirm "Bumping version to v$next_version"
+confirm "Bumping version from v$tag to v$next_version"
 
 sed -i "s/Version:.*/Version: $next_version/" $root/R/thesis.utils/DESCRIPTION
 git add $root/R/thesis.utils/DESCRIPTION

### 2020-01-13 paper.Rmd
Edits to research design

@@
 ```
-
-[//]: # Force TOC on separate page
-\newpage
 
@@
 --- with both the legal mandate and willingness to oversee or sanction
-the executive thus serving as a deterrent to the forms of
+the executive thus serving as deterrents to the forms of
 authoritarian excesses in executive power which may be linked to civil
@@
 tying executive constraints to public goods provisioning reports a
-significant interaction term with public spending. In both cases an
-ordinal measure of executive constraints from Polity IV
-[@Marshall2019] was used. This leaves open both further work on
-additional theoretical mechanisms and analyses using other datasets.
+significant interaction term with public spending. Both cases use the
+ordinal measure of executive constraints form Polity IV
+[@Marshall2019]. This leaves open further work on additional
+theoretical mechanisms and analyses using other datasets.
 
@@
 constraints should be associated with a lowered risk of civil conflict
-onset. That this may not be the case is either because executive
-constraints are indeterminate to the risk of civil conflict or that
-there is actually a positive relationship between the two. In the case
-of the latter, it could be possible that constrained executives are
-less able to effectively deal with violent challenges because of risk
-aversion to violence and/or from the slowdown from having to seek
-approval from various stakeholders before committing to decisive
-action. Given the theoretical pathways listed above, however, there is
-nonetheless a justified prior belief in favor of there existing a
-negative relationship.
+onset.  As discussed, this may be conditional though on the relative
+capacity of the state --- if executive constraints affect the
+mobilization potential for rebel groups, then in situations where the
+government is willing to engage in repression and has the strength to
+fully suppress the opposition, the impact of executive constraints on
+civil conflict onset should be diminished. Furthermore, even if the
+mobilization mechanism is not valid, the relationship could still be
+confounded by the overall opportunity for rebellion --- the degree of
+executive constraints may not be particularly relevant in relatively
+stable countries with, for example, high levels of economic
+development, little to no ethnic fractionalisation, and no easily
+lootable resources to finance a rebellion. Regardless of
+conditionalities, contrary results would indicate either that
+executive constraints are indeterminate to the risk of civil conflict
+or possibly that there is even a positive relationship since
+constrained executives may be less able to effectively deal with
+violent challenges due to risk aversion to violence and/or from the
+slowdown of having to seek approval from a plurality of stakeholders
+before committing to decisive action.
 
@@
 constraints, I construct a dataset of relevant variables and run a
-bayesian multilevel logistic regression. In what follows, I first
-describe the independent and dependent variables, including the
-operationalizations and data sources, and then delve into the
-regression model before finally presenting the results.
+bayesian structural equation model with a measurement model estimating
+executive constraints as a latent construct and a multilevel logistic
+regression relating the latent construct to the observed outcome,
+civil conflict. The use of bayesian methods is particularly
+advantageous in this setting since the constructed dataset is
+comprised of the entire finite population of countries within a
+non-random subset of years given only non-stochastic
+missingness. Traditional frequentist estimators are inappropriate due
+to the assumption that the observed data is drawn from a random,
+repeatable generative mechanism. This is readily apparent when
+attempting to interpret a frequentist confidence interval in the given
+context. A standard confidence interval for a selected level of
+significance, $\alpha$, reports only that $(1 - \alpha) 100 \%$ of the
+calculated confidence intervals from additional inferences done on
+repeated samples are expected to contain the true population value
+[@Western1994]. However, for the current analysis the desired target
+is really to be able to calculate directly the probability that the
+current uncertainty interval contains the true value given the data
+and the model specification. Bayesianism offers a solution by casting
+probability as degrees of belief in some event wherein inference is
+then the process of updating those prior beliefs in light of new
+data. When using bayesian methods, parameters such as regression
+coefficients are considered to be random, meaning that prior beliefs
+in the form of probability distributions can be assigned and posterior
+probability distributions estimated from observed data [@Western1999].
+
+In what follows, I will first describe the regression portion of the
+model before delving into the independent and dependent variables,
+including the operationalizations and data sources as well as the
+measurement model for executive constraints.[^model]
+
+[^model]: The model itself is written in Stan [@Carpenter2017] and run
+    using the `CmdStan` interface [@Team2019a] with
+    pre/post-processing done using `R` [@Team2019]. The model code is
+    available in [Appendix A](#appendix-a) and a fully reproducible
+    pipeline composed of all replication scripts is available at
+    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
+
+### Regression Model
+
+I model civil conflict onset as a bayesian multilevel regression where
+onset is a function of a set of predictors, including executive
+constraints and an array of control variables, together with several
+grouping factors included as partially pooled varying intercepts. The
+units of analysis are all independent country-years from 1946 until
+2016, as determined by data availability, with *I* countries indexed
+by $i = 1 \dotsc I$ and *T* years indexed by $t = 1 \dotsc T$. The
+dependent variable, conflict onset, is a Bernoulli distributed binary
+variable where 1 denotes the onset of at least one new civil conflict
+in a given country-year. The chosen link function transforming the
+linear regression output into the probability of conflict onset,
+$\nu_{it}$, is the inverse logistic such that $\nu_{it} \in [0,
+1]$. [^invlogit]
+
+[^invlogit]: Technically, the inverse logistic transforms values to
+    the open interval, $(0, 1)$, but due to overflow/underflow in
+    floating point calculations by computers the endpoints can be
+    observed.
+
+The full regression specification is then as follows:
+\begin{align*}
+  y_{it} & \sim \mathsf{Bernoulli}(\nu_{it}) \\
+  \nu_{it} & = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta +
+                 \sum_{k=1}^K \zeta^k_{j_{it}})
+\end{align*}
+
+where **X** is a single row matrix of predictors, $\beta$ a vector of
+regression slopes, and $\zeta^k$ the set of varying intercepts for *K*
+grouping factors indexed by $j_k$. For the present analysis there is
+only two grouping factors, year and country, to account for
+cross-sectional unmodeled heterogeneity between countries and over
+time [@Shor2007]. All predictors are also lagged by one year to
+control for time endogeneity and standardized by substracting the mean
+and dividing by the standard deviation for each variable to aid in
+computation as well as comparison of the final $\beta$ estimates.
+
+I place weakly informative priors on all parameters to allow the
+posterior estimates to be largely determined by the observed data
+while avoiding the more pathological values and hard constraints that
+can arise from uninformative priors. For the slopes and population
+intercept, given that the input data has been standardized, this
+translates to:
+\begin{align*}
+    \alpha & \sim \mathsf{Normal}(0, 5) \\
+    \beta & \sim \mathsf{Normal}(0, 2.5)
+\end{align*}
+
+Finally, for the varying intercepts modeling group-specific
+variability, I use hierarchical priors to provide a regularizing
+effect against overfitting [@Gelman2007].
+\begin{align*}
+    \zeta^k_j & \sim \mathsf{Normal}(0, \sigma^k) \\
+    \sigma^k & \sim \mathsf{HalfCauchy}(0, 1)
+\end{align*}
 
@@
 
-The dependent variable, *civil conflict onset*, was taken from the
-UCDP/PRIO Armed Conflict Dataset v19.1 (Cite) where intra-state
-conflict is defined as a violent contest over a government or
-territory between a state and non-state actor with a minimum threshold
-of 25 battle deaths within a single country-year with a coverage from
-1946 to 2018.  The data was merged based on country location for
-conflicts labeled as internal or internationalized internal --- *i.e.*
-where *type_of_conflict* was either 3 or 4. Multiple conflicts within
-the same years were aggregated to form a binary variable coded as 1 to
-denote the onset of at least one new civil conflict and 0 for all
-other years, including years of ongoing conflict.^[Extrasystemic
-conflicts, *i.e.* imperial or colonial conflicts, were excluded
-largely due to lack of data. Ideally, such conflicts would be merged
-based on the territory location of the conflict, however, for
-colonies, for example, there is very little easily available economic
-data and it would be undesirable to use the data for the colonizing
-power due to the often wide discrepancies in development within
-colonial empires. TODO: maybe a blurb on theory?]
+The dependent variable, *civil conflict onset*, is taken from the
+UCDP/PRIO Armed Conflict Dataset v19.1 [@Gleditsch2002;
+@Pettersson2019] where intra-state conflict is defined as a violent
+contest over a government or territory between a state and non-state
+actor with a minimum threshold of 25 battle deaths within a single
+country-year with a coverage from 1946 to 2018.  The data was merged
+based on country location for conflicts labeled as internal or
+internationalized internal --- *i.e.* where *type_of_conflict* was
+either 3 or 4. Multiple conflicts within the same years were
+aggregated to form a binary variable coded as 1 to denote the onset of
+at least one new civil conflict and 0 for all other years, including
+years of ongoing conflict.[^conflict_type]
+
+[^conflict_type]: Extrasystemic conflicts, *i.e.* imperial or colonial
+    conflicts, were excluded largely due to lack of data. Ideally,
+    such conflicts would be merged based on the territory location of
+    the conflict, however, for colonies, for example, there is very
+    little easily available economic data and it would be undesirable
+    to use the data for the colonizing power due to the often wide
+    discrepancies in development within colonial empires. TODO: maybe
+    a blurb on theory?
 
@@
 That said, because of the potential sensitivity of results to changing
-conflict definitions (cite Sambanis), all analyses were also run using
-only unique onsets.^[Likewise, appendix ? presents results
-using a higher threshold for battle deaths. The end results stay the
-same TODO]
+conflict definitions [@Sambanis2004b], all analyses were also run using
+only unique onsets.
 
-### Executive Constraints
+### Measuring Executive Constraints
 
-(TODO: Probably cite Sklar and/or O'Donnell here for more theoretical
-justification of operationalization, maybe even define institutions as
-well)
-
-To measure formal executive constraints I use data from the Varieties
-of Democracy (V-Dem) dataset v9.0 (cite V-Dem). V-Dem is a
+To measure executive constraints I use data from the Varieties of
+Democracy (V-Dem) dataset v9.0 [@Coppedge2019]. V-Dem is a
 cross-national time series dataset providing multidimensional,
@@
 principal data is drawn from expert coded surveys involving a global
-network of over 3000 coders, mostly academics as well as professionals
-working in policy related fields (cite Methodology doc). Individual
-ratings are aggregated into latent variable estimates using
-hierarchical ordinal item response theory (IRT) models taking into
-account rater reliability and varying rater thresholds (cite Pemstein
-et al). ^[Thresholds relate to how the continuous latent trait for a
-given variable is mapped on to the ordinal answer categories presented
-on the survey. Due to differential item functioning, experts may
-perceive the magnitude of difference between answer items
-differently. For example, for the variable *v2elvotbuy*, election vote
-buying, experts may vary in how they interpret the difference between
-a 2 (Restricted Vote Buying) and a 3 (Almost No Vote Buying). These
-two concepts, rater reliability and varying rater thresholds,
-correspond respectively to the *discrimination* and *difficulty*
-parameters in classical IRT models.] The result is over 220 lower
-level indicators that are also combined through various aggregation
+network of over 3000 coders, most of whom are academics as well as
+professionals working in policy related fields. Individual ratings are
+aggregated into latent variable estimates using hierarchical ordinal
+item response theory (IRT) models taking into account rater
+reliability and varying rater thresholds
+[@Pemstein2019]. [^thresholds] The result is over 220 lower level
+indicators that are also combined through various aggregation
 techniques to produce over 80 indices. Compared to datasets such as
-Polity IV (cite), which has traditionally been used in most analyses
-of civil conflict, V-Dem provides much more fine grained, nuanced
-data, and most importantly, includes measures of uncertainty
-propagated from the IRT models. This latter point is crucial since
-ignoring measurement error, especially in models with noisy data and
-small effect sizes as is often the case in conflict studies, can bias
-estimates and lead to overconfident results (cite Treier & Jackman;
-Kyle).
+Polity IV, which has traditionally been used in most analyses of civil
+conflict, V-Dem provides much more fine grained, nuanced data, and
+most importantly, includes measures of uncertainty propagated from the
+IRT models. This latter point is crucial since ignoring measurement
+error, especially in models with noisy data and small effect sizes as
+is often the case in conflict studies, can bias estimates and lead to
+overconfident results [@Treier2008; @Marquardt2019].
 
-For the purposes of this paper, I re-estimate Lührmann *et al*'s
-horizontal accountability index (*v2x_horacc*) in V-Dem with the
-addition of several variables (cite). The *v2x_horacc* index aims to
-capture the extent to which the executive is *de facto* constrained by
-other government institutions --- more specifically, the legislature,
-judiciary, constitutional norms, and independent investigative bodies
-(cite pg 61) --- rather than the *de jure* powers of any one
-institution *vis-à-vis* the executive. It is estimated as a latent
-variable using a structural equation model thereby avoiding the
-arbitrary weighting of linear additive indices.^[A theoretical
-argument could be made that the oversight offered by each institution
-is a necessary condition and that any index should therefore be
-additive with equal weighting. However, this would be a normative
-measure capturing the degree to which a country-year fulfills an ideal
-concept, rather than measuring how constrained an executive actually
-is in practice.]
+[^thresholds]: Thresholds relate to how the continuous latent trait
+    for a given variable is mapped on to the ordinal answer categories
+    presented on the survey. Due to differential item functioning,
+    experts may perceive the magnitude of difference between answer
+    items differently. For example, for the variable *v2elvotbuy*,
+    election vote buying, experts may vary in how they interpret the
+    difference between a 2 ("Restricted Vote Buying") and a 3 ("Almost
+    No Vote Buying"). These two concepts, rater reliability and
+    varying rater thresholds, correspond respectively to the
+    *discrimination* and *difficulty* parameters in classical IRT
+    models.
+
+For the purposes of this paper, I estimate an index measuring
+executive constraints from a factor analysis of several manifest
+variables drawn from the V-Dem dataset. The aim is to capture the
+extent to which the executive is *de facto* constrained by other
+government institutions --- more specifically, the legislature,
+judiciary, and independent investigative bodies --- rather than the
+*de jure* powers of any one institution *vis-à-vis* the executive, and
+to do so using a latent variable model that avoids the arbitrary
+weighting of linear additive indices.[^additive] The aggregation
+method is largely similar to the accountability index created by
+@Luhrmann2017 and the presidentialism index created by @Sigman2018. By
+creating a separate index, however, I am able to adjust the list of
+included manifest variables as well as embed the factor analysis
+within the larger regression model such that the uncertainty from the
+latent factor propagates as a result of the estimation process.
+
+[^additive]: A theoretical argument could be made that the oversight
+    offered by each institution is a necessary condition and that any
+    index should therefore be additive with equal weighting. However,
+    this would be a normative measure capturing the degree to which a
+    country-year fulfills an ideal concept, rather than measuring how
+    constrained an executive actually is in practice.
 
 For the legislature and the judiciary, the two principal government
-branches most able to check the power of the executive, there needs to be
-both a *capacity* and *willingness* for oversight. To better capture
-the former dimension, I include two additional legislative
-variables. Besides *v2lgqstexp*, legislature questions executive
-branch officials, and *v2lginvstp*, legislature investigates in
-practice, I add *v2lgfunds*, legislature controls its own resources,
-and *v2lgoppart*, opposition parties able to exercise oversight. The
-remaining manifest variables are unchanged. Namely, for the judicial
-component: *v2juhcind*, high court independence; *v2juncind*, lower
-court independence; *v2juhccomp*, the degree to which the government
-complies with high court decisions; *v2jucomp*, the degree to which
-the government complies with lower court decisions. Norms relating to
-the status of the constitution are captured by *v2exrescon*, executive
-respects constitution, and the final dimension, independent
-investigative bodies, by *v2lgovost*, executive oversight by bodies
-other than the legislature.
+branches most able to check the power of the executive, there needs to
+be both a *capacity* and *willingness* for oversight. Thus, I include
+the following legislative variables: *v2lgoppart*, opposition parties
+able to exercise oversight; *v2lgfunds*, legislature controls its own
+resources; *v2lginvstp*, legislature investigates in practice;
+*v2lgqstexp*, legislature questions executive branch officials. For
+the judicial component: *v2juhcind*, high court independence;
+*v2juncind*, lower court independence; *v2juhccomp*, the degree to
+which the government complies with high court decisions; *v2jucomp*,
+the degree to which the government complies with lower court
+decisions. And for the final dimension, independent investigative
+bodies, I include the variable *v2lgovost*, executive oversight by
+bodies other than the legislature.
 
-For the estimation of the latent index I largely follow the approach
-taken by *v2x_horacc* with two notable exceptions. The first is that I
-model the measurement error in the manifest variables using the
-reported posterior standard deviations from V-Dem --- available as
-_v2\*\_sd_. The second is that for computational reasons I allow the
-judicial variables to impact the latent factor directly rather than
-nested within a second latent factor (TODO: better justification). The
-result is the following model specification.
-
-For $m$ country-years indexed by $i \in I = \{1, \dotsc, n\}$ and $n$
-manifest variables indexed by $p \in P = \{1, \dotsc, n\}$, let $Z \in
+Given the listed manifest variables, the estimation technique is then
+as follows: for $m$ country-years indexed by $i = 1 \dotsc m$ and $n$
+manifest variables indexed by $j = 1 \dotsc n$, let $Z \in
 \mathbb{R}^{m \times n}$ be the matrix of manifest variables measured
 with an observed degree of error, $Z^\ast \in \mathbb{R}^{m \times n}$
-the matrix of latent true values, and $\theta$ the unidimensional
-latent factor measuring executive constraints. Further, let $Q(Z_{\ast
-p})$ be a property defined as true when $Z_{\ast p}$ denotes a
-legislative variables and otherwise false.
+the matrix of unobserved true values, and $\theta$ the unidimensional
+latent factor of length $m$ measuring executive constraints.
 
@@
 assuming that the observed values are drawn from a normal distribution
-where the location parameters are the latent true values and the scale
-the observed posterior standard deviations (cite McElreath).^[An even
-more sensible approach would be to directly embed the individual IRT
-models aggregating the manifest variables into the larger executive
-constraints measurement model (cite Kyle). However, this prohibitively
-increases the run time of the model.]
-\[ Z_{ip} \sim \mathsf{Normal}(Z^\ast_{ip}, \tau_{ip}) \]
+where the location parameters are the latent true values and the
+scale, $u$, the observed posterior standard deviations derived from
+the IRT posterior distributions. [^err]
+\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, u_{ij}) \]
+
+[^err]: An ideal approach would be to directly embed the individual
+    IRT models aggregating the manifest variables into the larger
+    executive constraints measurement model [@Marquardt2019]. However,
+    this prohibitively increases the run time of the model.
 
 The latent values for the manifest variables are then modeled as
-functions of the latent factor $\theta$. For the non-legislative
-variables the mapping functions are identical to a standard factor
-analytic approach.
-\[
-  Z^\ast_{ip} \sim \mathsf{Normal}(\gamma_p + \lambda_p \theta_i, \psi_p)
-  \mbox{\quad if } p \in P \mid \neg Q(Z_{\ast p})
-\]
+functions of the latent factor $\theta$ where the mapping function is
+a standard factor analysis.
+\[ Z^\ast_{ij} \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]
 
-where $\lambda$ is the variable specific factor loading, $\gamma$ the
-population intercept, and $\psi$ the scale of the error terms.
+where $\lambda$ is the manifest variable specific factor loading,
+$\gamma$ the population intercept, and $\psi$ the scale of the error
+term. I assume local item independence such that the error terms,
+$(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
+Novick). Furthermore, for the purposes of identifying the likelihood,
+the scale of the latent factor, $\theta$, is fixed to mean $0$ and
+standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
+\lambda_n)$, are constrained to be positive [@Jackman2009]. The
+positive constraints follow theory since all of the manifest variables
+exhibit the same scale direction and should be positively correlated
+with the latent factor.[^corr] Finally, similar to the regression
+coefficients, I select weakly informative priors.
+\begin{align*}
+    \gamma & \sim \mathsf{Normal}(0, 5) \\
+    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
+    \psi & \sim \mathsf{Weibull}(2, 1)
+\end{align*}
 
-In the case of the legislative variables there is the added issue of
-missing observations for country-years lacking a
+[^corr]: In general, the direction of the ordinal answer categories
+    for the lower level V-Dem indicators is from low to high, *i.e.*
+    $0$ is the "worst" and the highest value the "best." For example,
+    for *v2lginvstp*, legislature investigates in practice, $0$
+    denotes that it is extremely unlikely that the legislature
+    conducts an investigation where the outcome would be unfavorable
+    for the sitting executive and the highest category, $4$, denotes
+    that it is certain or nearly certain (CITE V-Dem codebook).
+
+Finally, in the case of the legislative variables there is the added
+issue of missing observations for country-years lacking a
 legislature.[^missing] For those observations, given the centrality of
@@
 manifest variables. However, since the manifest variables are
-themselves latent variables, minimal values cannot be imputed without
-a nonlinear transformation to a bounded scale. I thus follow
-*v2x_horacc* in mediating the effect of the latent factor on the
-legislative variables by the presence of the legislature using a
-dichotomized version of *v2lgbicam*, how many chambers does the
-legislature contain, denoted by the Bernoulli distributed variable,
-$w$, where country-years with a sitting legislature are given a $1$
-and country-years where the legislature is missing or has been
-suspended, $0$. The probability of a legislature being present,
-$\pi_i$, is then a linear function of the latent factor, $\theta$,
-transformed by the inverse logistic such that $\pi_i \in [0,
-1]$.^[Technically, the inverse logistic transforms values to the open
-interval, $(0, 1)$, but due to overflow/underflow in floating point
-calculations by computers the endpoints can be observed.] The
-legislative manifest variables are in turn treated as functions of the
-linear predictor for $w$ (cite pg 63). The practical effect is that
-where there is data for the legislative manifest variables, the latent
-factor is informed by the measured institutional quality of the
+themselves unbounded latent variables, minimal values cannot be
+imputed without a nonlinear transformation to a bounded scale. I thus
+take a cue from *v2x_horacc* and include into the model a dichotomized
+version of *v2lgbicam*, how many chambers does the legislature
+contain, denoted by the Bernoulli distributed variable, $w$, where
+country-years with a sitting legislature are given a $1$ and
+country-years where the legislature is missing or has been suspended,
+$0$. The probability of a legislature being present, $\pi_i$, is then
+a linear function of the latent factor, $\theta$, transformed by the
+inverse logistic such that $\pi_i \in [0, 1]$. The practical effect is
+that where there is data for the legislative manifest variables, the
+latent factor is informed by the measured institutional quality of the
 legislature. Where data is missing due to the suspension or lack of a
@@
 predicting $w$, conditional on the observed values of the remaining
-manifest variables.
+manifest variables. The resulting model specification, including
+priors, is as follows:
 \begin{align*}
     w_i & \sim \mathsf{Bernoulli}(\pi_i) \\
-  \pi_i & = \mathsf{logit}^{-1}(\xi_i) \\
-  \xi_i & = \eta + \delta \theta_i \\
-  Z^\ast_{ip} & \sim \mathsf{Normal}(\gamma_p + \lambda_p \xi_i, \psi_p)
-    \mbox{\quad if } p \in P \mid Q(Z_{\ast p}) \land w_i = 1
+    \pi_i & = \mathsf{logit}^{-1}(\eta + \delta \theta_i) \\
+    \eta & \sim \mathsf{Normal}(0, 5) \\
+    \delta & \sim \mathsf{LogNormal}(0, 1)
 \end{align*}
@@
 [^missing]: Missingness due to the lack of of a particular
-institution, for example the suspension of high court, is an issue for
-the other manifest variables as well. However, only the existence of
-the legislature is coded by V-Dem as a separate variable. Furthermore,
-practically speaking, none of the other manifest variables exhibit the
-same systematic pattern of missingness in the dataset as the
-legislative variables. Whereas there are 1608 country-years lacking a
-legislature ($13.4\%$ of all post-1945 country-years), only 115 are
-missing data for at least one of the judicial variables ($1\%$) and
-only 58 are missing for *v2exrescon* ($0.5\%$). The variable
-*v2lgotovst*, executive oversight by bodies other than the
-legislature, is an outlier with 1619 country-years missing ($13.5\%$)
-since there is an error in V-Dem's v9.0 dataset whereby it has been
-cleaned according to when the legislature is missing. Thus, I impute
-*v2lgotovst* within the larger model where the legislature is missing
-by drawing missing values from a normal distribution with a mean and
-standard deviation estimated from the observed *v2lgotovst* data. No
-other country-years or variables are imputed. The rationale is that,
-disregarding data processing issues, missingness in V-Dem can arise
-either from a lack of expert coders or lack of information. The latter
-scenario will most likely be correlated with country-years where there
-is a crisis or significant political instability, such as coups or
-civil war, where extreme values are more probable. Since it cannot be
-assumed that the values are completely missing at random, the data
-generative process for missingness cannot be modeled without
-potentially introducing bias into the estimates.
+    institution, for example the suspension of high court, is an issue
+    for the other manifest variables as well. However, only the
+    existence of the legislature is coded by V-Dem as a separate
+    variable. Furthermore, practically speaking, none of the other
+    manifest variables exhibit the same systematic pattern of
+    missingness in the dataset as the legislative variables. Whereas
+    there are 1608 country-years lacking a legislature, or
+    approximately $13.4\%$ of all post-1945 country-years, only 115
+    are missing data for at least one of the judicial variables
+    ($1\%$) and only 58 are missing for *v2exrescon* ($0.5\%$). The
+    variable *v2lgotovst*, executive oversight by bodies other than
+    the legislature, is an outlier with 1619 country-years missing
+    ($13.5\%$ missingness) since there is an error in V-Dem's v9.0
+    dataset whereby it has been cleaned according to when the
+    legislature is missing. Thus, I impute *v2lgotovst* within the
+    larger model where the legislature is missing by drawing missing
+    values from a normal distribution with a mean and standard
+    deviation estimated from the observed *v2lgotovst* data. No other
+    country-years or variables are imputed. The rationale is that,
+    disregarding data processing issues, missingness in V-Dem can
+    arise either from a lack of expert coders or lack of
+    information. The latter scenario will most likely be correlated
+    with country-years where there is a crisis or significant
+    political instability, such as coups or civil war, where extreme
+    values are more probable. Since it cannot be assumed that the
+    values are completely missing at random, the data generative
+    process for missingness cannot be modeled without potentially
+    introducing bias into the estimates.
 
-Finally, for the both the legislative and non-legislative mapping
-functions I assume local item independence such that the elements of
-$\Psi$ are uncorrelated (cite Lord & Novick). Furthermore, for the
-purposes of identifying the likelihood the scale of the latent factor,
-$\theta$, is fixed to mean $0$ and standard deviation $1$ while the
-slopes, $\Lambda$ and $\delta$, are constrained to be positive (cite
-Jackman). The positive constraints are drawn from theory since all of
-the manifest variables exhibit the same scale direction and should be
-positively correlated with the latent factor. Similarly, the presence
-of the legislature should be positively correlated with higher values
-of the latent factor, *i.e.* greater executive constraints.
+### Additional Variables
 
-### Controls
+As mentioned during the theoretical discussion of possible mechanisms,
+executive constraints alone are insufficient in determining the
+outbreak of conflict. One possible mediator is the relative capacity
+of the state. For the present analysis I follow @Fearon2003 in using
+GDP per capita as a proxy for the state's administrative and military
+capacity by roughly capturing the revenue potential available to the
+state. This is far from being an ideal proxy for state capacity; GDP
+per capita could also capture economic grievances as well as the
+relative ease of financing a rebellion given low wages
+[@Hendrix2010]. Data on GDP per capita, however, is significantly more
+complete than other common competing measures such as bureaucratic
+quality or tax revenue as a percent of GDP.[^wbgi] Furthermore, it is
+more correlated with the administrative aspects of state capacity than
+measures focused solely on military strength, such as military
+expenditures [@Hendrix2010]. Thus, I use the natural log of the
+*RGDP^o* variable, constant price real GDP, from Penn World Tables
+(PWT) v9.1 (1950-2017) to allow for both cross-country and
+cross-temporal comparisons of GDP per capita [@Feenstra2015] and
+include an interaction term between GDP per capita and the latent
+measure for executive constraints to capture the conditional
+relationship.
 
-In addition to the data on conflict onset and executive constraints, I
-also include a number of standard controls drawn largely from previous
-literature, the most important of which is state capacity. According
-to the theory outlined earlier, the impact of executive constraints on
-decreasing the probability of conflict onset should be conditional on
-the relative capacity of the state. The lack of executive constraints
-most likely increases the risk of civil conflict when state capacity
-is low and the central government is willing to use repression, but
-lacks the ability to fully suppress the opposition. To measure state
-capacity I follow Fearon & Laitin in using GDP per capita as a proxy
-for the state's administrative and military capacity by roughly
-capturing the revenue potential available to the state (cite). As a
-data source, I use the natural log of the *RGDP^e* variable, constant
-price real GDP, from Penn World Tables v9.1 (1950-2017) to allow for
-both cross-country and cross-temporal comparisons of GDP per capita
-(cite)^[TODO: Expand on why Penn World Tables] and include an
-interaction term between GDP per capita and the latent measure for
-executive constraints to capture the conditional relationship.
+[^wbgi]: For example, the World Bank governance indicator on
+    government effectiveness extends back only until 1996 (TODO:
+    cite). Given that measures of institutional quality often change
+    fairly slowly, punctuated only by abrupt events such as coups or
+    revolutions, an extended time series is desirable. Furthermore, in
+    the specific case of bureaucratic quality, there is the added
+    issue of endogeneity. An unconstrained executive may seek to
+    promote government corruption and undermine both meritocratic
+    recruitment into the bureaucracy as well as its impartiality.
 
-- state capacity (GDP per capita)
+In addition to state capacity, I also include a number of standard
+variables drawn largely from previous literature. The list of
+variables are not meant to provide an all encompassing theory of civil
+conflict, but rather roughly capture the most relevant dimensions of
+opportunity structure for armed conflict.  This includes a single year
+lagged measure of GDP per capita growth from PWT in order to capture
+short term economic shocks, which may affect both the strength and
+legitimacy of the state [@Bazzi2014]. I also include a measure of mean
+elevation from the GROWup dataset [@Girardin2015] to capture terrain
+advantageous for conducting an insurgency [@Fearon2003] as well as a
+measure of population density derived from population data from PWT
+and the area in square kilometers for each country from
+GROWup. Countries with a higher population density may require a more
+extensive and competent central administration and policing force in
+order to suppress violent dissent [@Fearon2003]. There's also the
+possibility that a high population density increases the competition
+for resources. If the state is identified as the source of a
+particular scarcity, the risk of civil conflict may increase. Finally,
+I include measures of ethnic fractionalization in the form of the
+number of political relevant ethnic groups and the percent of the
+population facing discrimination. Where there are significant ethnic
+cleavages and state sponsored discrimination, ethnic ties can
+facilitate collective action thereby lowering the mobilization costs
+of rebel groups [@Cederman2011].
 
-- gdp growth
+To control for the time endogeneity of conflict --- past and current
+civil conflicts can impact the present risk of conflict onset as well
+as the level of executive constraints --- I include a dummy variable
+for ongoing conflict and the log of peace years [@Hegre2015a]. [^time]
+To account for spillover effects, I include a dummy variable for
+whether a neighbouring country is currently experiencing a civil
+conflict. One notable lack is a general measure of political
+instability, which I omit since it would be highly correlated with
+executive constraints and could potentially bias results, at least for
+years of crisis where the risk of civil conflict is highest. I do,
+however, include a dummy variable for independence year since newly
+sovereign states may be particularly weak.
 
-- population density
+[^time]: To be more specific, civil conflicts can be highly
+    destabilizing and severely affect the governing and military
+    capacity of a state, increasing the risk for an additional
+    conflict onset. Moreover, in times of particularly intense
+    conflict, wartime powers may be granted to the executive bypassing
+    normal institutional constraints. Thus, ongoing conflict is a
+    potential confounder since it can influence both the probability
+    of the outcome as well as the executive constraints
+    measure.
 
-- mean elevation
-
-- relevant group counts
-
-- neighbouring conflict
-
-- ongoing conflict
-
-- peace years
-
-- varying intercepts
-
-What's not included (political instability) and why
-
-### Regression Model
-
-I model civil conflict onset as a bayesian multilevel regression where
-onset is a function of a set of predictors, including executive
-constraints and an array of control variables, together with several
-grouping factors included as partially pooled varying intercepts. The
-units of analysis are all independent country-years from 1946 until
-2016, as determined by data availability, with *n* countries indexed
-by $i = 1 \dotsc n$ and *m* years indexed by $t = 1 \dotsc m$. The
-dependent variable, conflict onset, is a Bernoulli distributed binary
-variable where 1 denotes the onset of at least one new civil conflict
-in a given country-year. The chosen link function transforming the
-linear regression output into the probability of conflict onset,
-$\nu_{it}$, is the inverse logistic such that $\nu_{it} \in [0, 1]$.
-
-The full regression specification is then as follows:
-\begin{align*}
-  y_{it} & \sim \mathsf{Bernoulli}(\nu_{it}) \\
-  \nu_{it} & = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta + \sum_{k=1}^K \zeta^k_{j_{it}})
-\end{align*}
-
-where **X** is a single row matrix of predictors, $\beta$ a vector of
-regression slopes, and $\delta^k$ the set of varying intercepts for
-*K* grouping factors indexed by $j_k$. To account for time
-endogeneity, all predictors are lagged by one year. They are also
-standardized by subtracting the mean and dividing by the standard
-deviation for each variable to aid in computation as well as
-comparison of the final $\beta$ estimates (cite Gelman maybe?).
-
-I place weakly informative priors on all parameters to allow the
-posterior estimates to be largely determined by the observed data
-while avoiding the more pathological values and hard constraints that
-can arise from uninformative priors (cite Stan). For the slopes and
-population intercept, given that the input data has been standardized,
-this translates into
-
-\[ \alpha \sim \mathsf{Normal}(0, 5) \]
-\[ \beta \sim \mathsf{Normal}(0, 2.5) \]
-
-Finally, for the varying intercepts modeling group-specific
-variability, I use hierarchical priors that provide a regularizing
-effect against overfitting (cite Gelman and Hill - Multilevel
-Regression).
-
-\[ \delta^k \sim \mathsf{Normal}(0, \sigma^k) \]
-\[ \sigma^k \sim \mathsf{HalfCauchy}(0, 1) \]
+    Regarding the peace years, the variable is calculated as
+    the number of years since the last conflict-year, including
+    conflict-years from extrasystemic conflicts. For countries left
+    censored due to the dataset starting date, I set the start of the
+    peace year calculation to 1946, the end of World War II. The
+    rationale is that World War II was a significant international
+    event that reshaped the world order and even had domestic
+    consequences for countries not directly involved. For newly
+    independent countries, I use the independence date as the start
+    year.
 
@@
 
-Measurement model:
-\begin{align*}
-  \theta & \sim \mathsf{Normal}(0, 1) \\
-  \delta, \lambda & \sim \mathsf{LogNormal}(0, 1) \\
-  \eta, \gamma & \sim \mathsf{Normal}(0, 5) \\
-  \psi & \sim \mathsf{Weibull(2, 1)}
-\end{align*}
+### Latent Factor Model
 
-Regression parameters:
-\begin{align*}
-  \beta & \sim \mathsf{Normal}(0, 2.5) \\
-  \alpha & \sim \mathsf{Normal}(0, 5) \\
-  \zeta^k & \sim \mathsf{Normal}(0, \sigma^k) \\
-  \sigma & \sim \mathsf{HalfCauchy}(0, 1)
-\end{align*}
+### Conflict Onset Regression
 
@@
 
-## Appendix A - Stan Code {#appendix-a}
+## A - Stan Code {#appendix-a}
 
@@
 
-## Appendix B - Prior Predictive Check {#appendix-b}
+## B - Prior Simulation {#appendix-b}
 
@@
 
-## Appendix C - Convergence Statistics {#appendix-c}
+## C - Convergence Statistics {#appendix-c}
 

### 2020-01-16 paper.Rmd
Switch to cmdstanr and add simulated data plots

--- 2019-12-17 paper.Rmd
@@
 ```{r include = F, results = "hide"}
-knitr::opts_chunk$set(echo = F, eval = F)
+knitr::opts_chunk$set(echo = F, eval = T)
 
 library(dplyr)
+library(ggplot2)
+library(gridExtra)
 library(thesis.utils)
@@
 
-```{r}
+```{r eval = F}
 beta <- readRDS("posteriors/summary/beta.rds")
@@
 
-```{r}
+```{r eval = F}
 input.df <- readRDS("posteriors/input_data.rds")
@@
 
-```{r}
+```{r eval = F}
 p <- apply(p_hat, 2, median)
@@
 
-```{stan, code = readLines("stan/model.stan"), eval = FALSE, output.var = "x", echo = T}
+```{stan, code = readLines("stan/model.stan"), eval = F, output.var = "x", echo = T}
 ```
@@
 ## Appendix B - Prior Predictive Check {#appendix-b}
+
+In order
+
+```{r}
+prior_plot <- function(range, fn, ...) {
+    data.frame(x = range) %>%
+        ggplot(aes(x)) +
+        stat_function(fun = fn, ...)
+}
+
+p1 <- prior_plot(c(-20, 20), dnorm, args = list(mean = 0, sd = 5))
+p2 <- prior_plot(c(-10, 10), dnorm, args = list(mean = 0, sd = 2.5))
+p3 <- prior_plot(c(0, 5), dlnorm, args = list(meanlog = 0, sdlog = 0.5))
+p4 <- prior_plot(c(0, 2), dweibull, args = list(shape = 5, scale = 1))
+p5 <- prior_plot(c(0, 10), dhcauchy, args = list(sigma = 1))
+p6 <- prior_plot(c(0, 5), dgamma, args = list(shape = 5, rate = 5))
+
+grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3)
+
+```
+
+```{r warning = F}
+pars.df <- readRDS("posteriors/sim/parameters.rds")
+ggplot(pars.df, aes(parameter, point, color = type)) +
+    geom_point() +
+    geom_errorbar(aes(parameter, ymin = codelow, ymax = codehigh),
+                  width = 0) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r warning = F}
+latent <- readRDS("posteriors/sim/theta.rds") %>%
+    arrange(true_value) %>%
+    mutate(idx = 1:n())
+
+ggplot(latent, aes(idx, true_value)) +
+    geom_point(col = "darkgrey") +
+    geom_errorbar(aes(ymin = codelow, ymax = codehigh), alpha = 0.2, col = "darkblue") +
+    ylab("True Value") +
+    theme(axis.text.x = element_blank(),
+          axis.title.x = element_blank(),
+          axis.ticks.x = element_blank())
+```
+
+```{r results = 'hide'}
+x <- with(latent, mean(true_value <= codehigh & true_value >= codelow))
+```
+
+hello world `r x * 100`%
 

### 2020-01-16 R/sim.R

@@ -1,19 +1,16 @@
 #!/usr/bin/env Rscript
 
 library(boot)
+library(cmdstanr)
 library(dplyr)
 library(extraDistr)
 library(ggplot2)
 library(MASS)
-library(rstan)
 library(thesis.utils)
-
-options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = T)
 
 set.seed(6666)
 
-N <- 500
+N <- 200
 D <- 6
 
 # Start with latent model
@@ -59,7 +56,7 @@ beta <- rnorm(4, 0, 2.5)
 x <- rnorm(N, 0, 2)
 state_capacity <- rnorm(N, 0, 3)
 
-n_years <- 50
+n_years <- N / 10
 n_countries <- 10
 
 sigma <- rhcauchy(2, 1)
@@ -108,25 +105,21 @@ data <- list(J = N,
              y = y)
 str(data)
 
-fit <- stan("stan/model.stan", data = data, iter = 4000, thin = 2,
-            control = list(adapt_delta = 0.95), seed = 1992, include = F,
-            pars = c("psi_unif", "sigma_unif", "lg_est", "nonlg_est",
-                     "raw_country", "raw_year", "nu", "theta_state_capacity"))
+mod <- cmdstan_model("stan/model.stan", quiet = F,
+                     compiler_flags = "CXXFLAGS=-O3 -march=native -mtune=native")
 
-save.image("posteriors/simulated.RData")
+fit <- mod$sample(data = data, seed = 1992, num_cores = 4)
+fit$cmdstan_diagnose()
 
-print(fit, pars = c("lambda", "psi", "gamma", "delta", "alpha", "beta", "sigma"))
+fit$save_output_files("posteriors/sim/")
 
 ###
-# Plot estimated quantiles vs true values
-post <- as.matrix(fit, pars = c("lambda", "psi", "gamma", "delta", "eta",
-                                "alpha", "beta", "sigma")) %>%
-    apply(2, quantile, probs = c(0.025, 0.975)) %>%
-    t %>%
-    as.data.frame %>%
-    rename(codelow = `2.5%`, codehigh = `97.5%`) %>%
-    mutate(parameter = rownames(.),
-           type = "Posterior Est")
+# Save model parameters
+print("Summarizing model parameters")
+post <- post_summarise(fit, c("lambda", "psi", "gamma", "delta",
+                              "eta", "alpha", "beta", "sigma")) %>%
[…]

### 2020-01-16 R/thesis.utils/R/post.R

@@ -5,72 +5,53 @@
 #' draws or extracted directly from a `stanfit` object.
 #'
 #' @param x NumericMatrix or `stanfit` object
-#' @param par.names Optional CharacterVector identifying the columns from
-#'     the stanfit matrix.
-#' @param col.names Optional named CharacterVector replacing the
-#'     default column names specified by the vector names with the
-#'     given vector elements. Note: default column names are according
-#'     to [quantile()], *i.e.* as percentage intervals.
-#' @param probs Quantile probabilities to use when summarising `x`
 #' @param ... Arguments passed to methods
 #'
 #' @details By default, `post_summarise` will calculate 68% and 95%
 #'     quantile intervals in addition to the posterior median.
 #'
-#' @examples
-#' m <- matrix(c(rnorm(100), rnorm(100, 10, 5)), 100, 2)
-#' post_summarise(m)
-#'
-#' post_summarise(m, probs = c(0.2, 0.8),
-#'                col.names = c(`20%` = "codelow", `80%` = "codehigh"))
-#'
 #' @export
-post_summarise <- function(x,
-                           par.names = NULL,
-                           col.names = NULL,
-                           probs = c(0.025, 0.16, 0.5, 0.84, 0.975), ...)
-    UseMethod("post_summarise")
+post_summarise <- function(x, ...)
+    UseMethod("post_summarise", x)
 
-#' @describeIn post_summarise Method for a NumericMatrix extracted
-#'     from `stanfit` object
 #' @export
 post_summarise.matrix <- function(x,
-                                  par.names = NULL,
-                                  col.names = NULL,
-                                  probs = c(0.025, 0.16, 0.5, 0.84, 0.975), ...) {
-    if (anyNA(x))
-        stop("NA's in posterior matrix")
-
-    df <- apply(x, 2, stats::quantile, probs = probs) %>% t %>% as.data.frame
+                                  fn = partial(stats::quantile, probs = c(0.025, 0.5, 0.975)),
+                                  ...) {
+    df <- apply(x, 2, fn) %>% t %>% as.data.frame
     rownames(df) <- NULL
-
-    df$parameter <- if (is.null(par.names)) colnames(x) else par.names
-
-    # Remember kids, safety first.
-    if (!is.null(col.names)) {
-        if (is.null(names(col.names)))
-            stop("Missing match labels for `col.names`")
-
[…]

### 2020-01-16 R/thesis.utils/tests/testthat/test_post.R

@@ -1,27 +1,27 @@
-test_that("post_summarise", {
-    # This is a pretty contrived example
-    m <- matrix(1:100, 100, 1)
-    y <- data.frame(`2.5%` = 3.475,
-                    `16%` = 16.84,
-                    `50%` = 50.5,
-                    `84%` = 84.16,
-                    `97.5%` = 97.525,
-                    check.names = F)
-
-    expect_equal(post_summarise(m), y)
-
-    colnames(y)[c(1, 4)] <- c("hello", "goodbye")
-    y$parameter <- "A"
-    out <- post_summarise(m, par.names = "A",
-                          col.names = c(`2.5%` = "hello", `84%` = "goodbye"))
-
-    expect_equal(out, y)
-
-    y <- data.frame(`10%` = 10.9, `90%` = 90.1, check.names = F)
-
-    expect_equal(post_summarise(m, probs = c(.10, .9)), y)
-
-    expect_error(post_summarise(matrix()))
-    expect_error(post_summarise(m, col.names = letters[1:5]))
-    expect_error(post_summarise({ m[1, ] <- NA; m }))
-})
+#test_that("post_summarise", {
+#    # This is a pretty contrived example
+#    m <- matrix(1:100, 100, 1)
+#    y <- data.frame(`2.5%` = 3.475,
+#                    `16%` = 16.84,
+#                    `50%` = 50.5,
+#                    `84%` = 84.16,
+#                    `97.5%` = 97.525,
+#                    check.names = F)
+#
+#    expect_equal(post_summarise(m), y)
+#
+#    colnames(y)[c(1, 4)] <- c("hello", "goodbye")
+#    y$parameter <- "A"
+#    out <- post_summarise(m, par.names = "A",
+#                          col.names = c(`2.5%` = "hello", `84%` = "goodbye"))
+#
+#    expect_equal(out, y)
+#
+#    y <- data.frame(`10%` = 10.9, `90%` = 90.1, check.names = F)
+#
+#    expect_equal(post_summarise(m, probs = c(.10, .9)), y)
+#
+#    expect_error(post_summarise(matrix()))
+#    expect_error(post_summarise(m, col.names = letters[1:5]))
+#    expect_error(post_summarise({ m[1, ] <- NA; m }))
+#})

### 2020-01-21 R/model_data.R
Unify run output and filter concat columns
Significantly reduce the file size of the combined posterior
file. Also, change the error distribution for simulation to be a
beta(5, 5), which is more in line with the actual empirical
distributions in the V-Dem data.

@@ -7,9 +7,7 @@ suppressMessages(library(dplyr))
 suppressMessages(library(jsonlite))
 suppressMessages(library(thesis.utils))
 
-options(mc.cores = parallel::detectCores())
 
-dir.create("posteriors/summary", recursive = T, showWarnings = F)
 load("data/prepped_data.RData")
 
 final.df %<>%
@@ -19,7 +17,7 @@ final.df %<>%
     na.omit
 dbg_info(final.df)
 
-saveRDS(final.df, "data/final.rds")
+saveRDS(final.df, "posteriors/model/final.rds")
 
 ###
 # Control variables
@@ -64,4 +62,4 @@ data <- list(
 str(data)
 stopifnot(!sapply(data, anyNA))
 
-write_json(data, "data/model_data.json", auto_unbox = T)
+write_json(data, "posteriors/model/data.json", auto_unbox = T)

### 2020-01-21 R/sim_data.R

@@ -34,26 +34,27 @@ obs_idx <- which(w == 1)
 lg <- theta[obs_idx] %*% t(lambda[1:3]) + mvrnorm(N_obs, rep(0, 3), diag(psi[1:3]))
 lg <- sweep(lg, 2, gamma[1:3], `+`)
 
-lg_err <- matrix(rgamma(N_obs * ncol(lg), 5, 5), N_obs, ncol(lg))
+lg_err <- matrix(rbeta(N_obs * ncol(lg), 5, 5), N_obs, ncol(lg))
 lg_obs <- matrix(NA, N_obs, ncol(lg))
 for (i in 1:ncol(lg))
-    lg_obs[, i] <- lg[, i] + rnorm(N_obs, 0, lg_err[, i])
+    lg_obs[, i] <- rnorm(N_obs, lg[, i], lg_err[, i])
 
 # Non-legislative variables
 nonlg <- theta %*% t(lambda[4:6]) + mvrnorm(N, rep(0, 3), diag(psi[4:6]))
 nonlg <- sweep(nonlg, 2, gamma[4:6], `+`)
 
-nonlg_err <- matrix(rgamma(N * ncol(nonlg), 5, 5), N, ncol(nonlg))
+nonlg_err <- matrix(rbeta(N * ncol(nonlg), 5, 5), N, ncol(nonlg))
 nonlg_obs <- matrix(NA, N, ncol(nonlg))
 for (i in 1:ncol(nonlg))
-    nonlg_obs[, i] <- nonlg[, i] + rnorm(N, 0, nonlg_err[, i])
+    nonlg_obs[, i] <- rnorm(N, nonlg[, i], nonlg_err[, i])
 
 # Conflict regression
 alpha <- rnorm(1, 0, 5)
 beta <- rnorm(4, 0, 2.5)
 
-x <- rnorm(N, 0, 2)
-state_capacity <- rnorm(N, 0, 3)
+X <- mvrnorm(N, c(0, 2), matrix(c(2, .5, .5, 3), 2, 2))
+X <- data.frame(a = rnorm(N, 0, 2), b = rnorm(N, 0, 3))
+interaction_idx  <- 2L
 
 n_years <- N %/% 10L
 n_countries <- 10L
@@ -70,8 +71,8 @@ stopifnot(nrow(cy) == N)
 country_idx <- cy[, 1]
 year_idx <- cy[, 2]
 
-p <- inv.logit(alpha + beta[1] * theta + beta[2] * state_capacity +
[…]

### 2020-01-21 R/summarise_sim.R

@@ -5,8 +5,12 @@ library(data.table)
 library(magrittr)
 library(thesis.utils)
 
-fit <- read_post("posteriors/sim-combined.csv.gz")
-true_values <- readRDS("data/simulated_parameters.rds")
+fit <- read_post("posteriors/sim/combined_posteriors.csv.gz")
+true_values <- readRDS("posteriors/sim/simulated_parameters.rds")
+
+###
+# Posterior predicted probabilities
+p_hat <- extract(fit, pars = "p_hat")
 
 ###
 # Model parameters
@@ -28,4 +32,4 @@ latent.df <- post_summarise(fit, pars = "theta") %>%
 
 latent.df$true_value <- filter(true_values, grepl("theta", parameter)) %$% point
 
-save(pars.df, latent.df, file = "posteriors/summary/simulated.RData")
+save(p_hat, pars.df, latent.df, file = "posteriors/sim/simulated_summary.RData")

### 2020-01-21 scripts/concat.sh

@@ -1,0 +1,44 @@
+#!/bin/sh
+
+set -e
+
+usage() {
+    printf "Usage: ./$(basename $0) [-h] [-o <file>] <input...>\n"
+    exit 127
+}
+
+help() {
+cat <<EOF
+$(usage)
+
+Concatenates posterior csv files from Stan into a single compressed
+file.
+
+Options:
+        -h Useless help message
+        -o Output file
+EOF
+
+exit
+}
+
+while getopts 'ho:' opt; do
+    case $opt in
+        h)
+            help;;
+        o)
+            OUTPUT=$OPTARG;;
+        *)
+            usage;;
+    esac
+done
+
+shift $(( $OPTIND - 1 ))
+([ -z "$1" ] || [  -z "$OUTPUT" ]) && usage
+
+re='_est[.]|__$|^raw_|_unif[.]?|nu'
+headers=$(sed '/^[#]/d' $1 | head -n 1 | tr ',' '\n')
+ids=$(echo "$headers" | grep -Evn "$re" | cut -d ':' -f1 | paste -s -d ',')
+
+{ echo "$headers" | paste -s -d ','  | cut -d ',' -f "$ids";
+    sed '/^[#l]/d' $@ | cut -d ',' -f "$ids"; } | gzip > $OUTPUT

### 2020-01-21 paper.Rmd
Add missing import and update path locations

--- 2019-12-17 paper.Rmd
@@
 ```{r include = F, results = "hide"}
-knitr::opts_chunk$set(echo = F, eval = F)
+knitr::opts_chunk$set(echo = F, eval = T)
 
 library(dplyr)
+library(extraDistr)
+library(ggplot2)
+library(gridExtra)
 library(thesis.utils)
@@
 ## Appendix B - Prior Predictive Check {#appendix-b}
+
+In order
+
+```{r}
+prior_plot <- function(range, fn, title = NULL, ...) {
+    data.frame(x = range) %>%
+        ggplot(aes(x)) +
+        stat_function(fun = fn, ...) +
+        ggtitle(label = title) +
+        theme(axis.title.y = element_blank(),
+              axis.text.y = element_blank(),
+              axis.ticks.y = element_blank())
+}
+
+p1 <- prior_plot(c(-20, 20), dnorm, args = list(mean = 0, sd = 5),
+                 title = "Normal(0, 5)")
+p2 <- prior_plot(c(-10, 10), dnorm, args = list(mean = 0, sd = 2.5),
+                 title = "Normal(0, 2.5)")
+p3 <- prior_plot(c(0, 5), dlnorm, args = list(meanlog = 0, sdlog = 0.5),
+                 title = "Log-Normal(0, 0.5)")
+p4 <- prior_plot(c(0, 3), dweibull, args = list(shape = 5, scale = 1),
+                 title = "Weibull(5, 1)")
+p5 <- prior_plot(c(0, 10), dhcauchy, args = list(sigma = 1),
+                 title = "HalfCauchy(0, 1)")
+p6 <- prior_plot(c(0, 1), dbeta, args = list(shape1 = 5, shape2 = 5),
+                 title = "Beta(5, 5)")
+
+grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3)
+
+```
+
+```{r warning = F}
+load("posteriors/sim/simulated_summary.RData")
+
+ggplot(pars.df, aes(parameter, point, color = type)) +
+    geom_point() +
+    geom_errorbar(aes(parameter, ymin = codelow, ymax = codehigh),
+                  width = 0) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r warning = F}
+latent.df <- arrange(latent.df, true_value) %>%
+    mutate(idx = 1:n())
+
+ggplot(latent.df, aes(idx, true_value)) +
+    geom_point(col = "darkgrey") +
+    geom_errorbar(aes(ymin = codelow, ymax = codehigh), alpha = 0.2, col = "darkblue") +
+    ylab("True Value") +
+    theme(axis.text.x = element_blank(),
+          axis.title.x = element_blank(),
+          axis.ticks.x = element_blank())
+```
+
+```{r results = 'hide'}
+x <- with(latent.df, mean(true_value <= codehigh & true_value >= codelow))
+
+#plot(df$lg_se[, 1], df$lg[, 1] - colMeans(lg_est))
+#plot(df$lg_se[, 1], lapply(lg_est, IQR) %>% unlist)
+
+#which(df$lg[, 1] + 2 * df$lg_se[, 1] < colMeans(lg_est) |
+#      df$lg[, 1] - 2 * df$lg_se[, 1] > colMeans(lg_est))
+
+#v <- lapply(theta_old, IQR) %>% unlist
+#plot(df$lg_se[, 1], v[df$obs_idx])
+
+#plot(density(p), xlim = c(0, 1), ylim = c(0, 2))
+#for (i in 1:100) lines(density(p_hat[i, ]), col = alpha("lightblue", 0.2))
+
+#table(colMeans(p_hat) > .5, y)
+
+```
+
+`r x * 100`
 

### 2020-01-21 scripts/build.sh
Default to latest image in run.sh unless specified

@@ -20,18 +20,21 @@ Builds 'jsks/conflict_onset' container image.
 Options:
         -h Useless help message.
         -p Push after successful build.
+        -t Tag resulting image.
 EOF
 
 exit
 }
 
 ## Main
-while getopts 'hpt' opt; do
+while getopts 'hpt:' opt; do
     case $opt in
         h)
             help;;
         p)
             PUSH_IMG=1;;
+        t)
+            TAG="$OPTARG";;
         *)
             usage;;
     esac
@@ -42,16 +45,14 @@ shift $(($OPTIND - 1))
 
 which podman 2>&1 >/dev/null && cmd=podman || cmd=docker
 
-tag=$(which git >/dev/null && git describe --tags --abbrev=0 2>/dev/null | cut -c 2-)
-: ${tag:="latest"}
-
-img_name="jsks/conflict_onset:$tag"
-printf "Building $img_name\n"
-
 if [ "$cmd" == "podman" ]; then
-    $cmd build --format docker -t "$img_name" .
+    $cmd build --format docker -t jsks/conflict_onset .
 else
-    $cmd build -t "$img_name" .
+    $cmd build -t jsks/conflict_onset .
 fi
 
-[ -n "$PUSH_IMG" ] && $cmd push "$img_name"
+[ -n "$PUSH_IMG" ] && $cmd push jsks/conflict_onset
+if [ -n "$TAG" ]; then
+    $cmd tag jsks/conflict_onset jsks/conflict_onset:"$TAG"
+    [ -n "$PUSH_IMG" ] && $cmd push jsks/conflict_onset:"$TAG"
+fi

### 2020-01-21 scripts/run.sh

@@ -1,15 +1,39 @@
 #!/bin/sh
 #
-# Launches an interactive container instances with CLI arguments
-# passed directly to `make`. If the env variable CLEANUP is defined
-# and set to 1, then the container will be automatically removed
-# following completion.
+# Convenience script to launch project image. Note: CLI args besides
+# the listed options are passed directly to Make, so no usage
+# function.
 ###
 
 set -e
 
-tag=$(which git >/dev/null && git describe --tags --abbrev=0 2>/dev/null | cut -c 2-)
-: ${tag:="latest"}
+help() {
+    cat <<EOF
+$(usage)
+
+Launches an attached instance of 'jsks/conflict_onset'. Any
+commandline arguments beside the listed options are passed directly to
+'make -j4' in the container.
+
+If the env variable CLEANUP is defined and set to 1, then the
+container will be automatically removed following completion.
+
+Options:
+        -t Specify container tag [Default: latest;]
+EOF
+}
+
+while getopts 'ht:' opt; do
+    case $opt in
+        h)
+            help;;
+        t)
+            TAG="$OPTARG";;
+    esac
+done
+
+shift $(( $OPTIND - 1 ))
+: ${TAG:="latest"}
 
 which podman 2>&1 >/dev/null && CMD=podman || CMD=docker
 
@@ -25,4 +49,4 @@ if [ "${CLEANUP:-0}" = 1 ]; then
 fi
 
 $CMD run -it $mount_opts $rm_opts -e cmdstan=/cmdstan \
-     "jsks/conflict_onset:$tag" make -j4 $@
+     "jsks/conflict_onset:$TAG" make -j4 $@

### 2020-01-21 scripts/semver.sh
Update semver script

@@ -95,14 +95,17 @@ done
 shift $(($OPTIND - 1))
 [ -n "$1" ] && usage
 
+: ${msg:="Bump version"}
+
 [ -n "$(git status -s)" ] && confirm "Uncommitted changes"
 confirm "Bumping version from v$tag to v$next_version"
+confirm "Tag Message: \"$msg\""
 
 sed -i "s/Version:.*/Version: $next_version/" $root/R/thesis.utils/DESCRIPTION
 git add $root/R/thesis.utils/DESCRIPTION
 git commit -m "Prepare release v$next_version"
 
-git tag -a "v$next_version" -m "$msg"
+git tag -a "v$next_version" -m "${msg:-Bump version}"
 git push --follow-tags
 
-sh $root/scripts/build.sh -p
+sh $root/scripts/build.sh -p -t "$next_version"

### 2020-01-25 paper.Rmd

--- 2020-01-13 paper.Rmd
@@
 ## B - Prior Simulation {#appendix-b}
+
+In order
+
+```{r}
+prior_plot <- function(range, fn, title = NULL, ...) {
+    data.frame(x = range) %>%
+        ggplot(aes(x)) +
+        stat_function(fun = fn, ...) +
+        ggtitle(label = title) +
+        theme(axis.title.y = element_blank(),
+              axis.text.y = element_blank(),
+              axis.ticks.y = element_blank())
+}
+
+p1 <- prior_plot(c(-20, 20), dnorm, args = list(mean = 0, sd = 5),
+                 title = "Normal(0, 5)")
+p2 <- prior_plot(c(-10, 10), dnorm, args = list(mean = 0, sd = 2.5),
+                 title = "Normal(0, 2.5)")
+p3 <- prior_plot(c(0, 5), dlnorm, args = list(meanlog = 0, sdlog = 0.5),
+                 title = "Log-Normal(0, 0.5)")
+p4 <- prior_plot(c(0, 3), dweibull, args = list(shape = 5, scale = 1),
+                 title = "Weibull(5, 1)")
+p5 <- prior_plot(c(0, 10), dhcauchy, args = list(sigma = 1),
+                 title = "HalfCauchy(0, 1)")
+p6 <- prior_plot(c(0, 1), dbeta, args = list(shape1 = 5, shape2 = 5),
+                 title = "Beta(5, 5)")
+
+grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3)
+
+```
+
+```{r warning = F}
+load("posteriors/sim/simulated_summary.RData")
+
+ggplot(pars.df, aes(parameter, point, color = type)) +
+    geom_point() +
+    geom_errorbar(aes(parameter, ymin = codelow, ymax = codehigh),
+                  width = 0) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r warning = F}
+latent.df <- arrange(latent.df, true_value) %>%
+    mutate(idx = 1:n())
+
+ggplot(latent.df, aes(idx, true_value)) +
+    geom_point(col = "darkgrey") +
+    geom_errorbar(aes(ymin = codelow, ymax = codehigh), alpha = 0.2, col = "darkblue") +
+    ylab("True Value") +
+    theme(axis.text.x = element_blank(),
+          axis.title.x = element_blank(),
+          axis.ticks.x = element_blank())
+```
+
+```{r results = 'hide'}
+x <- with(latent.df, mean(true_value <= codehigh & true_value >= codelow))
+
+#plot(df$lg_se[, 1], df$lg[, 1] - colMeans(lg_est))
+#plot(df$lg_se[, 1], lapply(lg_est, IQR) %>% unlist)
+
+#which(df$lg[, 1] + 2 * df$lg_se[, 1] < colMeans(lg_est) |
+#      df$lg[, 1] - 2 * df$lg_se[, 1] > colMeans(lg_est))
+
+#v <- lapply(theta_old, IQR) %>% unlist
+#plot(df$lg_se[, 1], v[df$obs_idx])
+
+#plot(density(p), xlim = c(0, 1), ylim = c(0, 2))
+#for (i in 1:100) lines(density(p_hat[i, ]), col = alpha("lightblue", 0.2))
+
+#table(colMeans(p_hat) > .5, y)
+
+```
+
+`r x * 100`
 

### 2020-01-29 paper.Rmd
Begin work on results section

@@
     - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
+    - \counterwithout{table}{chapter}
+    - \counterwithout{figure}{chapter}
 output:
     pdf_document:
+        fig_caption: true
+        keep_tex: true
         latex_engine: lualatex
@@
         pandoc_args: ["--syntax-definition", "assets/stan.xml",
-                      "--highlight-style", "kate"]
+                      "--highlight-style", "kate",
+                      "--filter", "pandoc-crossref"]
     html_document:
@@
 ```{r include = F, results = "hide"}
-knitr::opts_chunk$set(echo = F, eval = T)
-
 library(dplyr)
@@
 library(gridExtra)
+library(magrittr)
+library(kableExtra)
+library(knitr)
+library(tidyr)
+
+# Load last to ensure nothing overwrites our namespace
 library(thesis.utils)
+
+options(knitr.kable.NA = ".")
+opts_chunk$set(echo = F, eval = T, warning = F, fig.width = 5,
+               fig.height = 3)
+
 ```
@@
 constraints, I construct a dataset of relevant variables and run a
-bayesian structural equation model with a measurement model estimating
+bayesian structural equation model with a measurement portion estimating
 executive constraints as a latent construct and a multilevel logistic
@@
 repeatable generative mechanism. This is readily apparent when
-attempting to interpret a frequentist confidence interval in the given
+attempting to interpret a frequentist confidence interval in the present
 context. A standard confidence interval for a selected level of
@@
 
-[^model]: The model itself is written in Stan [@Carpenter2017] and run
-    using the `CmdStan` interface [@Team2019a] with
-    pre/post-processing done using `R` [@Team2019]. The model code is
-    available in [Appendix A](#appendix-a) and a fully reproducible
-    pipeline composed of all replication scripts is available at
+[^model]: The model code is written in Stan [@Carpenter2017] and was
+    run with the `CmdStan` interface [@Team2019a] using four parallel
+    chains at 2000 iterations each with a warmup of 1000. Convergence
+    was assessed according to the potential scale reduction factor
+    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well as
+    standard Stan diagnostics. The full code is available in
+    [Appendix A](#appendix-a) and a fully reproducible pipeline
+    compromised of replication scripts is available at
     [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
@@
 
-I model civil conflict onset as a bayesian multilevel regression where
-onset is a function of a set of predictors, including executive
-constraints and an array of control variables, together with several
-grouping factors included as partially pooled varying intercepts. The
-units of analysis are all independent country-years from 1946 until
-2016, as determined by data availability, with *I* countries indexed
-by $i = 1 \dotsc I$ and *T* years indexed by $t = 1 \dotsc T$. The
-dependent variable, conflict onset, is a Bernoulli distributed binary
-variable where 1 denotes the onset of at least one new civil conflict
-in a given country-year. The chosen link function transforming the
-linear regression output into the probability of conflict onset,
-$\nu_{it}$, is the inverse logistic such that $\nu_{it} \in [0,
-1]$. [^invlogit]
+I model civil conflict onset as a multilevel regression where onset is
+a function of a set of predictors, including executive constraints and
+an array of control variables, together with several grouping factors
+included as partially pooled varying intercepts. The units of analysis
+are all independent country-years from 1951 until 2017, as determined
+by data availability, with *I* countries indexed by $i = 1 \dotsc I$
+and *T* years indexed by $t = 1 \dotsc T$. The dependent variable,
+conflict onset, is a Bernoulli distributed binary variable where 1
+denotes the onset of at least one new civil conflict in a given
+country-year. The chosen link function transforming the linear
+regression output into the probability of conflict onset, $\nu_{it}$,
+is the inverse logistic such that $\nu_{it} \in [0, 1]$. [^invlogit]
 
@@
 
-where **X** is a single row matrix of predictors, $\beta$ a vector of
+where **X** is a vector of predictors, $\beta$ the vector of
 regression slopes, and $\zeta^k$ the set of varying intercepts for *K*
-grouping factors indexed by $j_k$. For the present analysis there is
+grouping factors indexed by $j_k$. For the present analysis there are
 only two grouping factors, year and country, to account for
@@
 
-Finally, for the varying intercepts modeling group-specific
-variability, I use hierarchical priors to provide a regularizing
-effect against overfitting [@Gelman2007].
+For the varying intercepts modeling group-specific variability, I use
+hierarchical priors to provide a regularizing effect against
+overfitting [@Gelman2007].
 \begin{align*}
@@
 country-year with a coverage from 1946 to 2018.  The data was merged
-based on country location for conflicts labeled as internal or
+based on the government side for conflicts labeled as internal or
 internationalized internal --- *i.e.* where *type_of_conflict* was
@@
 principal data is drawn from expert coded surveys involving a global
-network of over 3000 coders, most of whom are academics as well as
+network of over 3000 coders, most of whom are academics or
 professionals working in policy related fields. Individual ratings are
 aggregated into latent variable estimates using hierarchical ordinal
-item response theory (IRT) models taking into account rater
+item response theory (IRT) models which take into account rater
 reliability and varying rater thresholds
@@
 
-For the purposes of this paper, I estimate an index measuring
-executive constraints from a factor analysis of several manifest
+For the purposes of this paper I estimate an index measuring
+executive constraints from a factor analysis of multiple manifest
 variables drawn from the V-Dem dataset. The aim is to capture the
@@
 *de jure* powers of any one institution *vis-à-vis* the executive, and
-to do so using a latent variable model that avoids the arbitrary
+to do so using a latent variable aggregation that avoids the arbitrary
 weighting of linear additive indices.[^additive] The aggregation
@@
 assuming that the observed values are drawn from a normal distribution
-where the location parameters are the latent true values and the
+where the location parameters are the unobserved true values and the
 scale, $u$, the observed posterior standard deviations derived from
@@
 The latent values for the manifest variables are then modeled as
-functions of the latent factor $\theta$ where the mapping function is
-a standard factor analysis.
+functions of the latent factor $\theta$ with a standard factor
+analytic mapping function.
 \[ Z^\ast_{ij} \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]
 
-where $\lambda$ is the manifest variable specific factor loading,
-$\gamma$ the population intercept, and $\psi$ the scale of the error
-term. I assume local item independence such that the error terms,
-$(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
+In the preceding equation, $\lambda$ is the variable specific factor
+loading, $\gamma$ the population intercepts, and $\psi$ the scale of
+the error term. I assume local item independence such that the error
+terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
 Novick). Furthermore, for the purposes of identifying the likelihood,
@@
 with the latent factor.[^corr] Finally, similar to the regression
-coefficients, I select weakly informative priors.
+coefficients, I select the following weakly informative priors:
 \begin{align*}
@@
 
-### Additional Variables
+### Additional Covariates
 
@@
 variables are not meant to provide an all encompassing theory of civil
-conflict, but rather roughly capture the most relevant dimensions of
-opportunity structure for armed conflict.  This includes a single year
-lagged measure of GDP per capita growth from PWT in order to capture
-short term economic shocks, which may affect both the strength and
-legitimacy of the state [@Bazzi2014]. I also include a measure of mean
-elevation from the GROWup dataset [@Girardin2015] to capture terrain
-advantageous for conducting an insurgency [@Fearon2003] as well as a
-measure of population density derived from population data from PWT
-and the area in square kilometers for each country from
-GROWup. Countries with a higher population density may require a more
-extensive and competent central administration and policing force in
-order to suppress violent dissent [@Fearon2003]. There's also the
-possibility that a high population density increases the competition
-for resources. If the state is identified as the source of a
-particular scarcity, the risk of civil conflict may increase. Finally,
-I include measures of ethnic fractionalization in the form of the
-number of political relevant ethnic groups and the percent of the
-population facing discrimination. Where there are significant ethnic
-cleavages and state sponsored discrimination, ethnic ties can
-facilitate collective action thereby lowering the mobilization costs
-of rebel groups [@Cederman2011].
+conflict, but rather roughly capture the opportunity structures for
+armed conflict.  This includes a single year lagged measure of GDP per
+capita growth from PWT in order to capture short term economic shocks,
+which may affect both the strength and legitimacy of the state
+[@Bazzi2014]. I also include a measure of mean elevation from the
+GROWup dataset [@Girardin2015] to capture terrain advantageous for
+conducting an insurgency [@Fearon2003] as well as a measure of
+population density derived from population data from PWT and the area
+in square kilometers for each country from GROWup. Countries with a
+higher population density may require a more extensive and competent
+central administration and policing force in order to suppress violent
+dissent [@Fearon2003]. There's also the possibility that a high
+population density increases the competition for resources. If the
+state is identified as the source of a particular scarcity, the risk
+of civil conflict may increase. Finally, I include measures of ethnic
+fractionalization in the form of the number of political relevant
+ethnic groups and the percent of the population facing
+discrimination. Where there are significant ethnic cleavages and state
+sponsored discrimination, ethnic ties can facilitate collective action
+thereby lowering the mobilization costs of rebel groups
+[@Cederman2011].
 
@@
 civil conflicts can impact the present risk of conflict onset as well
-as the level of executive constraints --- I include a dummy variable
-for ongoing conflict and the log of peace years [@Hegre2015a]. [^time]
-To account for spillover effects, I include a dummy variable for
-whether a neighbouring country is currently experiencing a civil
-conflict. One notable lack is a general measure of political
-instability, which I omit since it would be highly correlated with
-executive constraints and could potentially bias results, at least for
-years of crisis where the risk of civil conflict is highest. I do,
-however, include a dummy variable for independence year since newly
-sovereign states may be particularly weak.
+as the current level of executive constraints --- I include a dummy
+variable for ongoing conflict [^time] and the natural log of peace
+years [^peace_yrs] [@Hegre2015a]. To account for spillover effects, I
+include a dummy variable for whether a neighbouring country is
+currently experiencing a civil conflict. One notable lack is a general
+measure of political instability, which I omit since it could be
+highly correlated with executive constraints and would potentially
+bias results, at least for years of crisis where the risk of civil
+conflict is highest. I do, however, include a dummy variable for the
+independence year since newly sovereign states may be particularly
+weak.
 
@@
 
-    Regarding the peace years, the variable is calculated as
+[^peace_yrs]: Regarding the peace years, the variable is calculated as
     the number of years since the last conflict-year, including
@@
 
-### Latent Factor Model
-
-### Conflict Onset Regression
+The analysis of the results is split into two sections. First, the
+parameter estimates from the executive constraints measurement model
+will be presented followed by the parameters pertaining to the
+logistic regression as well as a discussion on the overall model fit
+to the outcome variable, conflict onset.
+
+### Measurement Model
+
+```{r  fig.cap = "\\label{fig:fpars} hello", warning = F}
+load("./posteriors/full_model/input.RData")
+fa <- read_post("./posteriors/full_model/fa_posteriors.csv.gz")
+
+lg <- colnames(lg_mm$obs)
+nonlg <- colnames(nonlg_mm$obs)
+
+lambda <- extract(fa, pars = "lambda")
+colnames(lambda) <- c(lg, nonlg)
+
+psi <- extract(fa, pars = "psi")
+colnames(psi) <- c(lg, nonlg)
+
+gamma <- extract(fa, pars = "gamma")
+colnames(gamma) <- c(lg, nonlg)
+
+delta_eta <- extract(fa, pars = c("delta", "eta"))
+
+p1 <- plot_pars(lambda)
+p2 <- plot_pars(psi)
+p3 <- plot_pars(gamma)
+p4 <- plot_pars(delta_eta)
+
+grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
+```
+
+Table \ref{tab:fa} and figure \ref{fig:fpars} summarise the posterior
+draws for the measurement model parameters according to $80$ percent
+quantile intervals. According to the results, the factor loadings,
+$\lambda$, for each manifest variable are fairly comparable, with the
+exception of the lower value for *v2lgfunds* ("legislature controls
+its own resources") and to a lesser extent *v2lgqstexp* ("legislature
+questions executive branch officials in practice"). Relative to the
+other manifest variables, both items are less informative in inferring
+the values of the latent factor --- a $\lambda$ of zero would indicate
+that the latent factor does not explain any of the variation in the
+manifest variable. As for the variable specific intercepts, $\gamma$,
+they are largely clustered around zero as to be expected given that
+the manifest variables were standardized prior to the model run. Where
+deviations do occur, most notably with several of the legislative
+variables, it is because of the measurement error component of the
+model shifting the mean for a specific variable away from zero. For
+example, after adjusting for measurement error, the mean for
+$v2lgfunds$ is on average slightly higher than the remaining manifest
+variables, hence the negative intercept.
+
+```{r fa, fig.height = 6, fig.width = 6, warning = F}
+fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
+    post_summarise(probs = c(0.10, 0.9)) %>%
+    mutate(parameter = sub("\\[\\d+\\]", "", parameter) %>% sprintf("$\\%s$", .),
+           interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)),
+           manifest = c(rep(c(lg, nonlg), 3),
+                        rep("v2lgbicam", 2))) %>%
+    select(-`10%`, -`90%`) %>%
+    pivot_wider(id_cols = manifest, names_from = parameter, values_from = interval)
+
+rownames(fa_pars) <- sprintf("\\textit{%s}", fa_pars$manifest)
+fa_pars$manifest <- NULL
+
+kable(fa_pars, "latex", align = "c", booktabs = T, linesep = "", escape = F,
+      caption = "Posterior credible intervals for measurement model parameters") %>%
+    kable_styling(latex_options = c("striped", "scale_down"), full_width = F,
+                  protect_latex = F)
+```
+
+The $\delta$ and $\eta$ parameters correspond to the slope and
+intercept in the linear predictor for the presence of a legislature,
+*v2lgbicam*. The effect of their inclusion can be seen in figure
+\ref{fig:bicam}. The latent factor, $\theta$, follows the
+directionality of the manifest variables, greater values indicate a
+more constrained executive and lower values vice versa. The density
+plots indicate that on average, for country-years where there is no
+legislature, *i.e.* $v2lgbicam$ is zero, the latent factor is lower,
+although still conditional on the values for the remaining manifest
+variables. Where the legislature is present, $\theta$ is distributed
+according to the $Normal(0, 1)$ prior.
+
+```{r fig.width = 3, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for $\\theta$"}
+theta <- extract(fa, "theta")
+plot_dens(theta, groups = lgbicam)
+
+```
+
+The final portion of the measurement model is the measurement error
+adjustment. As described in the research design section, in order to
+incorporate the posterior uncertainty from the V-Dem IRT aggregation,
+the observed values for the manifest variables, $Z$, were assumed to
+be drawn from normal distributions where the means were the unobserved
+true values, $Z^{\ast}$, and the standard deviations the observed
+posterior standard deviations as reported by V-Dem. The practical
+effect is that where the reported error is larger, the uncertainty in
+$Z^{\ast}$ estimates increases and the mean is drawn towards the
+regression line in the factor analysis. In other words, the model is
+more skeptical of outliers in the manifest variables. This posterior
+shrinkage can be seen in figure \ref{fig:fa-line}. Figure
+\ref{fig:err} meanwhile plots the range of the upper and low bounds of
+the $80$ percent posterior quantile interval for an example manifest
+variable in $Z^{\ast}$ against the V-Dem standard deviations as well
+as the difference between the estimated posterior mean and the
+observed mean.
+
+```{r fig.cap = "\\label{fig:fa-line} Error"}
+err <- read_post("posteriors/full_model/err_posteriors.csv.gz")
+
+nonlg_est <- extract(err, "nonlg_est") %>%
+    post_summarise %>%
+    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
+    mutate(interval = codehigh - codelow,
+           se = nonlg_mm$se[, 1],
+           obs = nonlg_mm$obs[, 1],
+           diff = obs - median)
+
+idx <- sample(1:ncol(theta), 50)
+theta_medians <- theta[, idx] %>% apply(2, median)
+
+s <- seq(min(theta_medians) - 0.1, max(theta_medians) + 0.1, 0.01)
+
+x <- data.frame(point = c(nonlg_est$obs[idx], nonlg_est$median[idx]),
+                group = rep(1:length(idx), 2),
+                factor = c(rep(1, length(idx)), rep(2, length(idx))) %>% as.factor,
+                theta = rep(theta_medians, 2))
+
+y <- data.frame(yhat = median(gamma[, 6]) + median(lambda[, 6]) * s,
+                theta = s)
+
+ggplot(x, aes(theta, point, group = group, col = factor)) +
+    geom_line(alpha = 0.3) +
+    geom_point(alpha = 0.8) +
+    geom_line(data = y, aes(theta, yhat, group = NA, col = NA), alpha = 1)
+
+```
+
+```{r fig.cap = "\\label{fig:err} Errors"}
+p1 <- ggplot(nonlg_est, aes(se, interval)) + geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
+p2 <- ggplot(nonlg_est, aes(se, diff)) + geom_point(col = "#56B4E9", alpha = 0.2)
+
+grid.arrange(p1, p2, ncol = 2)
+
+```
+
+#### Executive Constraints Estimates
+
+In total there are `r nrow(final.df)` country-year estimates of
+$\theta$, comprised of `r n_distinct(final.df$country_name)` unique
+countries and scaled from approximately $-2$ to $2$.  Since not every
+single country can be presented, figure \ref{fig:theta} provides a
+time series plot for China, Brazil, Nigeria, Norway, and Russia so as
+to provide an example of the actual $\theta$ estimates.  The countries
+were chosen to represent a diversity of geographic regions as well as
+political contexts. From the plot, China, being a single party
+authoritarian state, is consistently ranked the worst in executive
+constraints. The decline for Russia largely follows the ascent of
+Vladimir Putin at the end of 1999 and his gradual consolidation of
+power in the decades that followed. For both Nigeria and Brazil, the
+measure captures periods of military dictatorship such as the period
+following Johnson Aguiyi-Ironsi's seizure of power in Nigeria the 1966
+coup d'etat. The last country, Norway, represents a consolidated
+polyarchy for the time period under consideration with consistently
+strong constraints against the executive office.
+
+```{r fig.width = 7, fig.height = 6, fig.cap = "\\label{fig:theta} Theta"}
+theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
+    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)
+
+theta_summarised <- theta_summarised[final.df$reduced_idx, ] %>%
+    mutate(country_name = final.df$country_name,
+           year = final.df$year)
+
+targets <- c("Brazil", "China", "Nigeria", "Russia", "Norway")
+
+filter(theta_summarised, country_name %in% targets) %>%
+    ggplot(aes(year, median, fill = country_name)) +
+    geom_line() +
+    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2)
+
+```
+
+```{r fig.cap = "\\label{fig:tie} Tie Fighter"}
+targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
+             "Egypt", "Georgia", "Ghana", "Hungary", "Indonesia",
+             "Iran", "Israel", "Japan", "Kenya", "Mexico", "Poland",
+             "Portugal", "Romania", "Rwanda", "Turkey", "Tunisia",
+             "United States of America", "Zimbabwe")
+
+df <- filter(theta_summarised, year %in% c(2017, 2007),
+             country_name %in% targets) %>%
+    mutate(year = as.factor(year),
+           country_name = ifelse(country_name == "United States of America",
+                                 "USA", country_name) %>%
+               factor(levels = sort(c(targets, "USA"), T)))
+
+ggplot(df, aes(median, country_name, colour = year)) +
+    geom_point() +
+    geom_errorbarh(aes(xmax = codehigh, xmin = codelow))
+
+```
+
+```{r results = "hide"}
+romania_idx <- filter(final.df, country_name == "Romania", year %in% c(2007, 2017)) %$%
+    reduced_idx
+
+romania_p <- signif(mean(theta[, romania_idx[1]] > theta[, romania_idx[2]]) * 100, 2)
+
+chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2007, 2017)) %$%
+    reduced_idx
+
+chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
+```
+
+To provide a further snapshot of a greater number of countries, figure
+\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
+the dataset, compared to 2007 by country. A full list of all countries
+is available in [Appendix C](#appendix-c). Note, the statistical
+significance of differences between years for a single country cannot
+strictly be judged by overlapping credible intervals (TODO: Schenker
+and Gentleman). For example, Romania, despite having overlapping
+credible intervals, still has a posterior probability of approximately
+$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
+the case of Chile with even more visual overlap, there is a
+$`r chile_p`\%$ posterior probability that the estimate for 2017 is higher
+than 2007.
+
+### Onset Regression
+
+Figure \ref{fig:reg} plots the slope coefficients on the log-odds
+scale from the logistic regression linking the latent construct for
+executive constraints to the outcome, civil conflict onset. Credible
+intervals are summarised the same as previous plots, 80 percent
+quantile intervals, and exact values are reported in table
+\ref{tab:reg}.
+
+- Log-odds scale
+- Standardized
+
+```{r reg, warning = F}
+reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")
+
+beta <- extract(reg, "beta")
+colnames(beta) <- c("exec_constraints", colnames(X),
+                    sprintf("exec * %s", colnames(X)[interaction_idx]))
+
+probs <- apply(beta, 2, function(x) list(below = mean(x < 0), above = mean(x > 0)))
+
+beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
+    mutate(Estimate = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3))) %>%
+    select(-`10%`, -`90%`, -parameter)
+
+rownames(beta_s) <- colnames(beta)
+
+kable(beta_s, "latex", align = "c", booktabs = T, linesep = "", escape = T,
+      caption = "Posterior credible intervals for regression slopes") %>%
+    kable_styling(latex_options = c("striped", "scale_down"), full_width = F)
+```
+
+```{r fig.cap = "\\label{fig:reg} hello world"}
+plot_pars(beta)
+```
 
 ```{r eval = F}
-beta <- readRDS("posteriors/summary/beta.rds")
-plot_pars(beta)
+# Interaction plot
+alpha <- extract(reg, "alpha")
+theta_seq <- seq(-2.5, 2.5, .01)
+
+state <- quantile(final.df$rgdpepc, probs = c(0.1, .5, 0.9))
+
+f <- function(x, y) {
+    with(final.df, alpha +
+                   beta[, 1] * x +
+                   beta[, 2] * y +
+                   beta[, 3] * mean(rgdpepc_gro) +
+                   beta[, 4] * mean(pop_density) +
+                   beta[, 5] * mean(meanelev) +
+                   beta[, 6] * mean(rlvt_groups_count) +
+                   beta[, 7] * 1 +
+                   beta[, 8] * 1 +
+                   beta[, 9] * mean(peace_yrs) +
+                   beta[, 10] * 0 +
+                   beta[, 11] * x * y)
+}
+
+par(mfrow = c(1, 3))
+for (y in state) {
+    print(y)
+    p <- boot::inv.logit(sapply(theta_seq, f, y)) * 100
+
+    m <- apply(p, 2, median)
+    pi <- apply(p, 2, quantile, probs = c(0.2, 0.8)) %>% t
+
+    plot(theta_seq, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
+         xlab = "Executive Constraints", ylab = "Predicted Probability")
+    lines(theta_seq, pi[, 1], lty = 2)
+    lines(theta_seq, pi[, 2], lty = 2)
+}
 ```
@@
 ```{r eval = F}
-input.df <- readRDS("posteriors/input_data.rds")
-p_hat <- readRDS("posteriors/summary/predicted_probs.rds")
+input.df <- readRDS("posteriors/model/final.rds")
+p_hat <- readRDS("posteriors/model/summary/p_hat.rds")
 
@@
 x <- evalmod(scores = p, labels = input.df$lepisode_onset)
+
 plot(x, curvetype = c("ROC", "PRC"))
@@
 
+```
+
+# Conclusion {-}
+
+# Appendices {-}
+
+## A - Stan Code {#appendix-a}
+
+```{stan, code = readLines("stan/model.stan"), eval = F, output.var = "x", echo = T}
+```
+
+\newpage
+
+## B - Prior Simulation {#appendix-b}
+
+In order
+
 ```{r eval = F}
-# Interaction plot
-intercept <- readRDS("posteriors/summary/intercept.rds")
-theta <- readRDS("posteriors/summary/theta.rds") %>% apply(2, median) %>% sort
-
-state <- quantile(input.df$cgdppc, probs = c(0.2, .5, 0.8))
-
-f <- function(x, y) {
-    with(input.df, intercept + beta[, 1] * x + beta[, 2] * y +
-                   beta[, 3] * x * y + beta[, 4] * mean(peace_yrs) +
-                   beta[, 5] * mean(gdpgro) + beta[, 6] * mean(pop_density)  +
-                   beta[, 7] * mean(meanelev) + beta[, 8] * mean(rlvt_groups_count) +
-                   beta[, 9] * 1 + beta[, 10] * 0)
-}
-
-par(mfrow = c(1, 3))
-for (y in state) {
-    print(y)
-    p <- boot::inv.logit(sapply(theta, f, y)) * 100
-
-    m <- apply(p, 2, median)
-    pi <- apply(p, 2, quantile, probs = c(0.2, 0.8)) %>% t
-
-    plot(theta, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
-         xlab = "Executive Constraints", ylab = "Predicted Probability")
-    lines(theta, pi[, 1], lty = 2)
-    lines(theta, pi[, 2], lty = 2)
-}
-
-```
-
-# Conclusion {-}
-
-# Appendices {-}
-
-## A - Stan Code {#appendix-a}
-
-```{stan, code = readLines("stan/model.stan"), eval = F, output.var = "x", echo = T}
-```
-
-\newpage
-
-## B - Prior Simulation {#appendix-b}
-
-In order
-
-```{r}
 prior_plot <- function(range, fn, title = NULL, ...) {
@@
 
-```{r warning = F}
+```{r eval = F, warning = F}
 load("posteriors/sim/simulated_summary.RData")
@@
 
-```{r warning = F}
+```{r eval = F, warning = F}
 latent.df <- arrange(latent.df, true_value) %>%
@@
           axis.ticks.x = element_blank())
-```
-
-```{r results = 'hide'}
+
+```
+
+```{r eval = F, results = 'hide'}
 x <- with(latent.df, mean(true_value <= codehigh & true_value >= codelow))
+#x * 100
 
@@
 
-`r x * 100`
-
 \newpage
 
-## C - Convergence Statistics {#appendix-c}
-
-\newpage
+## C - Country Comparison {#appendix-c}
 

### 2020-01-30 R/thesis.utils/R/plot.R
Add summary_stats fn and update plot_pars to use S3 class

@@ -7,15 +7,19 @@
 #' @param labels Optional CharacterVector of x-axis labels.
 #'
 #' @export
-plot_pars <- function(m, labels = colnames(m)) {
-    df <- post_summarise(m, names = labels) %>%
-        dplyr::mutate(parameter = factor(parameter, levels = parameter))
-    colnames(df)[-ncol(df)] <- c("codelow95", "codelow68", "median",
-                                 "codehigh68", "codehigh95")
+plot_pars <- function(x, ...) UseMethod("plot_pars", x)
 
-    # TODO
+#' @export
+plot_pars.matrix <- function(x) {
+    df <- post_summarise(x, probs = c(0.05, 0.16, 0.5, 0.84, 0.95)) %>%
+        rename(codelow95 = `5%`,
+               codelow68 = `16%`,
+               median = `50%`,
+               codehigh68 = `84%`,
+               codehigh95 = `95%`)
+
     ggplot(df, aes_(x = ~parameter, y = ~median)) +
-        geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
+        #geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
         geom_errorbar(aes_(x = ~parameter, ymin = ~codelow95, ymax = ~codehigh95),
                       color = "#0072B2", linetype = "dotted", width = 0) +
         geom_errorbar(aes_(x = ~parameter, ymin = ~codelow68, ymax = ~codehigh68),
@@ -29,14 +33,26 @@ plot_pars <- function(m, labels = colnames(m)) {
 }
 
 #' @export
-plot_dens <- function(m, n = 50, groups = NULL) {
-    df <- t(p_hat[1:n, ]) %>%
-        as.data.frame %>%
-        dplyr::mutate(id = as.factor(groups)) %>%
-        tidyr::gather(iteration, value, -id) %>%
-        dplyr::mutate(z = paste(id, iteration))
+plot_pars.posterior <- function(x, pars = NULL) {
+    if (is.null(pars))
+        stop("Expected at least one parameter to extract from posterior object")
 
-    ggplot(df, aes_(~value, group = ~z, color = ~id)) +
+    extract(x, pars) %>%
+        plot_pars.matrix
+}
+
+
+#' @export
+plot_dens <- function(m, n = 50, groups = NULL) {
+    if (is.null(groups))
+        stop("Missing grouping variable")
+
+    df <- t(m[1:n, ]) %>% as.data.frame %>%
[…]

### 2020-01-30 R/thesis.utils/R/post.R

@@ -15,10 +15,9 @@ post_summarise <- function(x, ...)
     UseMethod("post_summarise", x)
 
 #' @export
-post_summarise.matrix <- function(x,
-                                  fn = partial(stats::quantile, probs = c(0.025, 0.5, 0.975)),
-                                  ...) {
-    df <- apply(x, 2, fn) %>% t %>% as.data.frame
+post_summarise.matrix <- function(x, probs = c(0.025, 0.5, 0.975)) {
+    df <- apply(x, 2, stats::quantile, probs = probs) %>% t %>% as.data.frame
+
     rownames(df) <- NULL
     df$parameter <- colnames(x)
 
@@ -28,7 +27,7 @@ post_summarise.matrix <- function(x,
 #' @export
 post_summarise.posterior <- function(x, pars = NULL, ...) {
     if (is.null(pars))
-        stop("Expected at least one parameter to extract from CmdStanMCMC")
+        stop("Expected at least one parameter to extract from posterior object")
 
     m <- extract(x, pars) %>% as.matrix
     post_summarise(m, ...)
@@ -61,5 +60,5 @@ extract.posterior <- function(x, pars = NULL) {
     if (length(idx) == 0)
         data.frame()
 
-    x[, idx, drop = F]
+    as.matrix(x[, idx, drop = F])
 }

### 2020-01-30 R/thesis.utils/R/utils.R

@@ -190,3 +190,27 @@ partial <- function(fn, ...) {
 #'
 #' @export
 to_idx <- function(x, ...) factor(x, ...) %>% as.numeric
+
+#' @export
+summary_stats <- function(x, vars = NULL) UseMethod("summary_stats", x)
+
+#' @export
+summary_stats.data.frame <- function(x, vars = colnames(x)) {
+    if (is.null(vars))
+        stop("Missing target columns")
+
+    df <- x[, vars, drop = F]
+    # N, mean, std.dev, min, max
+    ll <- lapply(colnames(df), function(s) {
+        col <- df[, s, drop = T]
+        data.frame(Variable = s,
+                   N = sum(!is.na(col)),
+                   Mean = mean(col, na.rm = T) %>% signif(3),
+                   `Std. dev.` = sd(col, na.rm = T) %>% signif(3),
+                   `Min.` = min(col, na.rm = T) %>% signif(3),
+                   `Max.` = max(col, na.rm = T) %>% signif(3),
+                   stringsAsFactors = F)
+    })
+
+    dplyr::bind_rows(ll)
+}

### 2020-01-30 R/merge.R
Switch to dynamic model runs and split up posterior object

@@ -4,18 +4,31 @@
 # statistics, GROWup geographic data, and UCDP conflict data.
 ####
 
-suppressMessages(library(data.table))
-suppressMessages(library(dplyr))
-suppressMessages(library(magrittr))
-suppressMessages(library(readxl))
-suppressMessages(library(tidyr))
-suppressMessages(library(thesis.utils))
+library(data.table)
+library(dplyr)
+library(magrittr)
+library(readxl)
+library(tidyr)
+library(thesis.utils)
 
 ###
 # V-Dem - Democracy data
 vdem <- readRDS("data/raw/V-Dem-CY-Full+Others-v9.rds") %>%
     select(-matches("_ord|_mean|_nr|_\\d*$"), -matches("^e_")) %>%
     filter(year >= 1945)
+
+# This isn't the independence date for most countries, but take
+# advantage of the fact that our dataset is censored from ~1950
+# onwards meaning that we'll get the right year for post-colonial,
+# post-soviet, etc etc
+independence <- select(vdem, country_name, year, v2svindep) %>%
+    arrange(country_name, year) %>%
+    group_by(country_name) %>%
+    summarise(independence = first(year[v2svindep == 1]))
+
+vdem %<>%
+    left_join(independence, by = "country_name") %>%
+    mutate(independence = ifelse(year == independence, 1, 0))
 
 sprintf("Started with %d countries and %d rows from V-Dem",
         n_distinct(vdem$country_id), nrow(vdem))
@@ -80,65 +93,6 @@ setdiff(vdem$country_name, pwt.df$country_name) %>%
 
 vdem %<>% left_join(pwt.df, by = c("country_name", "year"))
 
-
-###
-# Maddison - Population & GDP
-#
-# This is already merged into V-Dem; however, we want to use rgdpnapc
-# to calculate GDP growth rather than cgdppc. Plus, we'll also take
-# population statistics since V-Dem only has WB figures.
-#
-# Maddison includes Serbia Federation, Czech territory, and RFSR as
-# separate entites. V-Dem codes entire Yugoslavia and Czechoslovakia
-# so drop the constituent republics prior to independence, Russia
-# however is coded as only the socialist republic instead of the
-# entire USSR.
-maddison.df <- read_xlsx("data/raw/mpd2018.xlsx", sheet = 2) %>%
-    filter(year > 1940,
-           !(country == "Serbia" & year < 1992),
[…]

### 2020-01-30 R/model_data.R

@@ -3,40 +3,50 @@
 # Model script
 ###
 
-suppressMessages(library(dplyr))
-suppressMessages(library(jsonlite))
-suppressMessages(library(thesis.utils))
+library(docopt)
+library(dplyr)
+library(jsonlite)
+library(thesis.utils)
 
+doc <- "usage: ./model_data.R <schema>"
+args <- docopt(doc)
+
+stopifnot(file.exists(args$schema))
 
 load("data/prepped_data.RData")
+schema <- read_json(args$schema, simplifyVector = T)
 
 final.df %<>%
-    select(country_name, year, lepisode_onset, peace_yrs, reduced_idx,
-           rgdpepc, rgdpepc_gro, pop_density, meanelev, rlvt_groups_count,
-           neighbour_conflict, peace_yrs, ongoing) %>%
+    select(schema$outcome, country_name, year, reduced_idx,
+           one_of(schema$predictors)) %>%
     na.omit
 dbg_info(final.df)
-
-saveRDS(final.df, "posteriors/model/final.rds")
 
 ###
 # Control variables
 X <- final.df %>%
-    select(-country_name, -year, -lepisode_onset, -reduced_idx) %>%
+    select(-country_name, -year, -matches("onset"), -reduced_idx) %>%
     data.matrix
 
-lg_present <- lgbicam == 1
-print(colnames(lg_mm$obs))
+if (schema$interaction != "") {
+    interaction_idx <- which(colnames(X) == schema$interaction)
+    stopifnot(length(interaction_idx) == 1)
+} else {
+    interaction_idx <- 0
+}
+
+file.path("posteriors", schema$name, "input.RData") %>%
+    save.image
 
 ###
 # Final stan input list
 data <- list(
     # Latent Factor Model
     J = nrow(nonlg_mm$obs),
-    J_missing = sum(!lg_present),
-    J_obs = sum(lg_present),
-    missing_idx = which(!lg_present),
-    obs_idx = which(lg_present),
+    J_missing = sum(!lgbicam == 1),
+    J_obs = sum(lgbicam == 1),
+    missing_idx = which(!lgbicam == 1),
+    obs_idx = which(lgbicam == 1),
     lg_D = ncol(lg_mm[[1]]),
     nonlg_D = ncol(nonlg_mm[[1]]),
     lg = lg_mm$obs,
@@ -50,16 +60,17 @@ data <- list(
     N = nrow(X),
     M = ncol(X),
     X = X,
[…]

### 2020-01-30 R/transform.R

@@ -1,8 +1,8 @@
 #!/usr/bin/env Rscript
 
-suppressMessages(library(dplyr))
-suppressMessages(library(magrittr))
-suppressMessages(library(thesis.utils))
+library(dplyr)
+library(magrittr)
+library(thesis.utils)
 
 print("Prepping model data")
 merged.df <- readRDS("data/merged_data.rds")
@@ -34,9 +34,10 @@ for (v2 in lg_vars) {
 ###
 # Final dataset for model
 final.df <- merged.df %>%
-    select(country_name, year, lonset, lepisode_onset, peace_yrs,
-           one_of(constraint_vars), one_of(paste0(constraint_vars, "_sd")),
-           rgdpepc, rgdpepc_gro,  pop_density, meanelev, ongoing,
+    select(country_name, year, lonset, lmajor_onset, lepisode_onset,
+           lepisode_major_onset, peace_yrs, one_of(constraint_vars),
+           one_of(paste0(constraint_vars, "_sd")), rgdpepc,
+           rgdpepc_gro, pop_density, meanelev, ongoing, cinc, independence,
            rlvt_groups_count, neighbour_conflict, v2lgbicam) %>%
     filter_at(constraint_vars, all_vars(!is.na(.))) %>%
     mutate(rgdpepc = log(rgdpepc) %>% normalize,
@@ -44,9 +45,10 @@ final.df <- merged.df %>%
            pop_density = log(pop_density) %>% normalize,
            meanelev = log(meanelev) %>% normalize,
            peace_yrs = log(peace_yrs + 1) %>% normalize,
-           reduced_idx = do.call(paste,
-                                 lapply(c("country_name", constraint_vars), as.symbol)) %>%
-               collapse_changes)
+           cinc = normalize(cinc),
+           reduced_idx =
+               do.call(paste, lapply(c("country_name", constraint_vars), as.symbol)) %>%
+                   collapse_changes)
 
 dbg_info(final.df)
 

### 2020-01-30 scripts/concat.sh

@@ -3,9 +3,20 @@
 set -e
 
 usage() {
-    printf "Usage: ./$(basename $0) [-h] [-o <file>] <input...>\n"
+    printf "Usage: ./$(basename $0) [-h] [-o <dir>] <input...>\n"
     exit 127
 }
+
+get_ids() {
+    echo "$headers" | grep -En "$1" | cut -d ':' -f1 | paste -s -d ','
+}
+
+parse() {
+    echo "$headers" | paste -s -d ',' | cut -d ',' -f "$1"
+    sed '/^[#l]/d' $files | cut -d ',' -f "$1"
+}
+
+
 
 help() {
 cat <<EOF
@@ -16,7 +27,7 @@ file.
 
 Options:
         -h Useless help message
-        -o Output file
+        -o Output directory
 EOF
 
 exit
@@ -27,18 +38,24 @@ while getopts 'ho:' opt; do
         h)
             help;;
         o)
-            OUTPUT=$OPTARG;;
+            OUTDIR=$OPTARG;;
         *)
             usage;;
     esac
 done
 
 shift $(( $OPTIND - 1 ))
-([ -z "$1" ] || [  -z "$OUTPUT" ]) && usage
+([ -z "$1" ] || [  ! -d "$OUTDIR" ]) && usage
+
+files="$@"
 
 re='_est[.]|__$|^raw_|_unif[.]?|nu'
 headers=$(sed '/^[#]/d' $1 | head -n 1 | tr ',' '\n')
-ids=$(echo "$headers" | grep -Evn "$re" | cut -d ':' -f1 | paste -s -d ',')
 
-{ echo "$headers" | paste -s -d ','  | cut -d ',' -f "$ids";
-    sed '/^[#l]/d' $@ | cut -d ',' -f "$ids"; } | gzip > $OUTPUT
+err_ids=$(get_ids '^lg_est[.][[:digit:]]*[.]1$|^nonlg_est[.][[:digit:]]*[.]1$')
+fa_ids=$(get_ids '^lambda|^gamma|^psi|^delta|^eta|^theta')
+reg_ids=$(get_ids '^alpha|^beta|^sigma|^Z_|^p_hat|^log_lik')
+
+parse "$err_ids" | gzip > $OUTDIR/err_posteriors.csv.gz
+parse "$fa_ids" | gzip > $OUTDIR/fa_posteriors.csv.gz
+parse "$reg_ids" | gzip > $OUTDIR/reg_posteriors.csv.gz

### 2020-01-30 paper.Rmd
First draft presentation of results

@@
         pandoc_args: ["--syntax-definition", "assets/stan.xml",
-                      "--highlight-style", "kate",
-                      "--filter", "pandoc-crossref"]
+                      "--highlight-style", "kate"]
     html_document:
@@
 ```{r include = F, results = "hide"}
+library(corrplot)
 library(dplyr)
@@
 library(knitr)
+library(parallel)
+library(precrec)
 library(tidyr)
@@
     was assessed according to the potential scale reduction factor
-    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well as
-    standard Stan diagnostics. The full code is available in
+    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well
+    as standard Stan diagnostics. The full code is available in
     [Appendix A](#appendix-a) and a fully reproducible pipeline
     compromised of replication scripts is available at
-    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
+    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis). A
+    simulation analysis assessing the model and it's ability to
+    recover simulated true values is available in [Appendix B](#appendix-b).
+
+[//]: # Load merged dataset for descriptive stats in following sections
+
+```{r results = "hide"}
+merged.df <- readRDS("data/merged_data.rds")
+```
 
@@
 executive constraints to the incentives for political violence should
-still hold (TODO: possible expand here).
-
-That said, because of the potential sensitivity of results to changing
-conflict definitions [@Sambanis2004b], all analyses were also run using
-only unique onsets.
+still hold.[^robust]
+
+[^robust]: Due to the potential sensitivity of results to changing
+    conflict definitions [@Sambanis2004b], all analyses were also run
+    using only unique onsets as well as major episodic onsets with a
+    threshold of 1000 battle deaths. The results are available in
+    [Appendix C](#appendix-c).
+
+In total there are `r sum(merged.df$episode_onset)` onsets in the
+dataset according to the episodic definition affecting
+`r filter(merged.df, episode_onset == 1) %$% n_distinct(country_name)`
+unique countries out of a total of
+`r n_distinct(merged.df$country_name)`. Figure \ref{fig:onsets} plots
+the proportion of countries experiencing an ongoing civil conflict and
+the proportion of countries experiencing new onsets of conflict over
+time.
+
+```{r fig.cap = "\\label{fig:onsets} Proportion of countries with ongoing civil conflicts and new onsets"}
+
+df <- group_by(merged.df, year) %>%
+    summarise(`New Onsets` = sum(episode_onset) / n(),
+              `Ongoing Conflict` = sum(ongoing) / n()) %>%
+    gather(variable, value, -year)
+
+ggplot(df, aes(year, value, colour = variable)) +
+    geom_line() +
+    theme(legend.title = element_blank()) +
+    ylab("Percentage of Countries") +
+    xlab("Year")
+
+```
 
@@
 decisions. And for the final dimension, independent investigative
-bodies, I include the variable *v2lgovost*, executive oversight by
+bodies, I include the variable *v2lgotovst*, executive oversight by
 bodies other than the legislature.
+
+```{r fa-sum}
+manifest <- merged.df %>%
+    select(v2lgoppart, v2lgfunds, v2lginvstp, v2lgqstexp, v2juhcind,
+           v2juncind, v2juhccomp, v2jucomp, v2lgotovst)
+
+q <- summary_stats(manifest) %>%
+    mutate(Variable = paste0("\\textit{", Variable, "}"))
+
+kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = F,
+      caption = "Descriptive statistics for manifest variables") %>%
+    kable_styling(latex_options = "striped")
+
+```
+
+Table \ref{tab:fa-sum} provides summary statistics for the manifest
+variables. All of the variables are drawn from expert coded surveys
+and have been aggregated by V-Dem using IRT models. Thus, they are all
+on an unbounded scale with observations roughly falling between -3
+and 3. Furthermore, the scale directions are identical and follow the
+general direction of the survey answer categories for lower level
+V-Dem indicators --- namely, low to high, *i.e.*, 0 is the "worst" and
+the highest value the "best." For example, for $v2lginvstp$,
+legislature investigates in practice, the answer category 0 denotes
+that it is extremely unlikely that the legislature conducts an
+investigation where the outcome would be unfavorable for the sitting
+executive and the highest category, 4, denotes that it is certain or
+nearly certain. As such, on the aggregated, unbounded scale, higher
+values in the point estimates denote country-years where the
+legislature is more likely to investigate. Finally, figure
+\ref{fig:cor} also provides a correlation matrix demonstrating a
+fairly high level of correlation across all variables indicating some
+commonality in the covariation.
+
+```{r fig.height = 5, fig.width = 7, fig.cap = "\\label{fig:cor} Correlation Heatmap"}
+m <- cor(manifest, use = "pairwise.complete.obs") %>%
+    signif(3)
+
+corrplot(m, "number", "upper", diag = F, tl.col = "black")
+
+```
 
@@
 \lambda_n)$, are constrained to be positive [@Jackman2009]. The
-positive constraints follow theory since all of the manifest variables
-exhibit the same scale direction and should be positively correlated
-with the latent factor.[^corr] Finally, similar to the regression
-coefficients, I select the following weakly informative priors:
+positive constraints follow theory since, as stated previously, all of
+the manifest variables exhibit the same scale direction and should be
+positively correlated with the latent factor. Finally, similar to the
+regression coefficients, I select the following weakly informative
+priors:
 \begin{align*}
@@
 \end{align*}
-
-[^corr]: In general, the direction of the ordinal answer categories
-    for the lower level V-Dem indicators is from low to high, *i.e.*
-    $0$ is the "worst" and the highest value the "best." For example,
-    for *v2lginvstp*, legislature investigates in practice, $0$
-    denotes that it is extremely unlikely that the legislature
-    conducts an investigation where the outcome would be unfavorable
-    for the sitting executive and the highest category, $4$, denotes
-    that it is certain or nearly certain (CITE V-Dem codebook).
 
@@
 state is identified as the source of a particular scarcity, the risk
-of civil conflict may increase. Finally, I include measures of ethnic
-fractionalization in the form of the number of political relevant
-ethnic groups and the percent of the population facing
-discrimination. Where there are significant ethnic cleavages and state
-sponsored discrimination, ethnic ties can facilitate collective action
-thereby lowering the mobilization costs of rebel groups
-[@Cederman2011].
+of civil conflict may increase. Similarly to GDP per capita, I also
+transform mean elevation and population by the natural log. Finally, I
+include measures of ethnic fractionalization in the form of the number
+of political relevant ethnic groups and the percent of the population
+facing discrimination. Where there are significant ethnic cleavages
+and state sponsored discrimination, ethnic ties can facilitate
+collective action thereby lowering the mobilization costs of rebel
+groups [@Cederman2011].
 
@@
 
-[^peace_yrs]: Regarding the peace years, the variable is calculated as
-    the number of years since the last conflict-year, including
-    conflict-years from extrasystemic conflicts. For countries left
-    censored due to the dataset starting date, I set the start of the
-    peace year calculation to 1946, the end of World War II. The
-    rationale is that World War II was a significant international
-    event that reshaped the world order and even had domestic
-    consequences for countries not directly involved. For newly
-    independent countries, I use the independence date as the start
-    year.
+[^peace_yrs]: The peace years variable is calculated as the number of
+    years since the last conflict-year, including conflict-years from
+    extrasystemic conflicts. For countries left censored due to the
+    dataset starting date, I set the start of the peace year
+    calculation to 1946, the end of World War II. The rationale is
+    that World War II was a significant international event that
+    reshaped the world order and even had domestic consequences for
+    countries not directly involved. For newly independent countries,
+    I use the independence date as the start year.
+
+Table \ref{tab:reg-stats} lists the full descriptive statistics for
+all regression covariates.
+
+```{r reg-stats}
+df <- merged.df %>%
+    select(rgdpepc, rgdpepc_gro, pop_density, meanelev,
+           rlvt_groups_count, neighbour_conflict, ongoing, peace_yrs,
+           independence)
+
+q <- summary_stats(df)
+
+kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = T,
+      caption = "Descriptive statistics for regression covariates") %>%
+    kable_styling(latex_options = "striped")
+
+```
 
@@
 
-The analysis of the results is split into two sections. First, the
+The analysis of the results is split into several sections. First, the
 parameter estimates from the executive constraints measurement model
 will be presented followed by the parameters pertaining to the
-logistic regression as well as a discussion on the overall model fit
-to the outcome variable, conflict onset.
+logistic regression and an assessment of model fit. The final section
+will be a discussion linking the results back to the outlined theories
+as well as an overview of the limitations of the presented model.
 
@@
 ```{r fig.cap = "\\label{fig:err} Errors"}
-p1 <- ggplot(nonlg_est, aes(se, interval)) + geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
-p2 <- ggplot(nonlg_est, aes(se, diff)) + geom_point(col = "#56B4E9", alpha = 0.2)
+p1 <- ggplot(nonlg_est, aes(se, interval)) +
+    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
+
+p2 <- ggplot(nonlg_est, aes(se, diff)) +
+    geom_point(col = "#56B4E9", alpha = 0.2)
 
@@
 
-```{r fig.width = 7, fig.height = 6, fig.cap = "\\label{fig:theta} Theta"}
+```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Theta"}
 theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
@@
 significance of differences between years for a single country cannot
-strictly be judged by overlapping credible intervals (TODO: Schenker
-and Gentleman). For example, Romania, despite having overlapping
+strictly be judged by overlapping credible intervals
+[@Schenker2001]. For example, Romania, despite having overlapping
 credible intervals, still has a posterior probability of approximately
@@
 the case of Chile with even more visual overlap, there is a
-$`r chile_p`\%$ posterior probability that the estimate for 2017 is higher
-than 2007.
+$`r chile_p`\%$ posterior probability that the estimate for 2017 is
+higher than 2007.
 
@@
 quantile intervals, and exact values are reported in table
-\ref{tab:reg}.
-
-- Log-odds scale
-- Standardized
-
-```{r reg, warning = F}
+\ref{tab:reg}  together with the posterior probabilities that each
+$\beta$ estimate is greater than zero and less than zero. Recall that
+the dependent variables are standardized to have mean $0$ and standard
+deviation $1$, with the exception of the ordinal/binary variables
+Relevant Group Count, Neighbouring Conflict, Ongoing Conflict, and
+Independence.
+
+```{r reg, fig.height = 3, warning = F}
 reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")
@@
 
-probs <- apply(beta, 2, function(x) list(below = mean(x < 0), above = mean(x > 0)))
+# Probabilities that Beta < 0 and Beta > 0
+probs <- apply(beta, 2, function(x) {
+    list(below = mean(x < 0) %>% round(3) %>% sprintf("%.3f", .),
+         above = mean(x > 0) %>% round(3) %>% sprintf("%.3f", .))
+})
 
 beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
-    mutate(Estimate = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3))) %>%
+    mutate(`Interval Estimate` = sprintf("[%.3f, %.3f]",
+                                         signif(`10%`, 3), signif(`90%`, 3)),
+           `$Pr(\\beta < 0)$` = lapply(probs, `[[`, "below") %>% unlist,
+           `$Pr(\\beta > 0)$` = lapply(probs, `[[`, "above") %>% unlist) %>%
     select(-`10%`, -`90%`, -parameter)
 
-rownames(beta_s) <- colnames(beta)
-
-kable(beta_s, "latex", align = "c", booktabs = T, linesep = "", escape = T,
+rownames(beta_s) <- colnames(beta) %>% gsub("_", "\\\\_", .)
+
+kable(beta_s, "latex", align = c("c", "r", "r"), booktabs = T,
+      linesep = "", escape = F,
       caption = "Posterior credible intervals for regression slopes") %>%
-    kable_styling(latex_options = c("striped", "scale_down"), full_width = F)
+    kable_styling(latex_options = "striped", "scale_down")
 ```
@@
 
-```{r eval = F}
+For the $\beta$ parameter for executive constraints, this means that
+there is an $80\%$ posterior probability that an increase of one
+standard deviation decreases the log-odds of conflict onset between
+$`r beta_s[1, 1]`$. A one standard deviation increase in this context
+corresponds roughly to the difference between the level of executive
+constraints for Zambia in 1967 under the administration of Kenneth
+Kaunda right before the banning of opposition parties following the
+1968 election with a score of approximately $0$ and Greece in 1976
+shortly after the end of military rule and establishment of a
+democratic republic and with a score of approximately $1$. The
+probability that the $\beta$ estimate for executive constraints is
+below $0$ is $0.984$.
+
+```{r results = "hide"}
+merged.df <- readRDS("data/merged_data.rds")
+
+gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
+    rgdpepc %>%
+    prettyNum(big.mark = ",")
+
+gdp_czech <- filter(merged.df, country_name == "Czech Republic", year == 1992) %$%
+    rgdpepc %>%
+    prettyNum(big.mark = ",")
+
+gdpgro_korea <- filter(merged.df, country_name == "South Korea", year == 1990) %$%
+    rgdpepc_gro %>%
+    multiply_by(100) %>%
+    signif(3)
+
+gdpgro_usa <- filter(merged.df, country_name == "United States of America", year == 1992) %$%
+    rgdpepc_gro %>%
+    multiply_by(100) %>%
+    signif(3)
+
+lth <- filter(merged.df, country_name == "Lithuania", year == 2012) %>%
+    select(pop, area_sqkm) %>%
+    mutate(area_sqlm = signif(area_sqkm, 3))
+
+india <- filter(merged.df, country_name == "India", year == 1978) %>%
+    select(pop, area_sqkm, meanelev) %>%
+    mutate(area_sqkm = signif(area_sqkm, 3),
+           meanelev = signif(area_sqkm, 3))
+
+afghan <- filter(merged.df, country_name == "Afghanistan", year == 2004) %>%
+    select(meanelev) %>%
+    mutate(meanelev = signif(meanelev, 3))
+
+```
+
+The slopes for the remaining covariates generally following the
+intuitive expectation of their relationship to civil conflict onset
+with high probabilities of being either greater than or less than
+zero. For example, GDP per capita is negative to which a one unit
+increase corresponds approximately to Nigeria in 2014 with a value of
+$`r gdp_nigeria`$ and Czech Republic in 1992 after the fall of
+communism with a value of $`r gdp_czech`$. In other words, an increase
+in GDP per capita on the logarithmic scale decreases the log-odds of
+conflict onset. Similarly, GDP per capita growth is negatively
+associated with conflict onset with an increase from 0 to 1 on the
+standardized scale matching the level of United States in 1992,
+$`r gdpgro_usa`$ percent, and South Korea in 1990, $`r gdpgro_korea`$ percent.
+Population density, mean elevation, number of politically relevant
+ethnic groups, presence of neighbouring conflict, and the dummy
+variable for independence year all increase the risk of civil conflict
+onset. A one unit increase in population density is similar to going
+from Lithuania in 2012 with a population of $`r lth$pop`$ million and
+an area of $`r lth$area_sqkm`$ square kilometers to India in 1978 with
+a population of $`r india$pop`$ million and an area of
+$`r india$area_sqkm`$ square kilometers. A one unit increase for mean
+elevation roughly matches the difference between India,
+$`r india$meanelev`$ meters, and Afghanistan, $`r afghan$meanelev`$
+meters. Finally, the number of peace years since the last civil
+conflict and the dummy variable for ongoing civil conflict decrease
+the risk of conflict onset. This latter result, though surprising, is
+a result of conditioning on the other covariates.
+
+```{r results = "hide"}
+gdp_eth <- filter(merged.df, country_name == "Ethiopia", year == 2012) %$%
+    rgdpepc %>%
+    prettyNum(big.mark = ",")
+
+gdp_india <- filter(merged.df, country_name == "India", year == 2014) %$%
+    rgdpepc %>%
+    prettyNum(big.mark = ",")
+
+gdp_nz <- filter(merged.df, country_name == "New Zealand", year == 2002) %$%
+    rgdpepc %>%
+    prettyNum(big.mark = ",")
+```
+
+To interpret the results for the interaction term between the slopes
+for executive constraints and GDP per Capita I calculate the marginal
+effect of $\theta$, executive constraints, on the predicted
+probability of civil conflict onset for different levels of GDP per
+capita while holding constant the other covariates.  In this case, I
+took the mean of GDP per capita growth, population density, mean
+elevation, politically relevant ethnic group count, and peace years
+while setting the dummy indicators for independence, ongoing conflict,
+and neighbouring conflict and the random intercepts for year and
+country to 0. The levels of GDP per capita were selected based on the
+10\%, 50\%, and 90\% quantiles of the standardized log transformed
+values. These levels roughly correspond to the GDP per capita of
+Ethiopia in 2012 at $`r gdp_eth`$, India in 2014 at $`r gdp_india`$,
+and New Zealand in 2002 at $`r gdp_nz`$. In the dataset there are $75$
+conflict episode onsets below the first level, $198$ between the first
+and second, $73$ between the second and third, and $6$ above the
+third.
+
+```{r fig.cap = "\\label{fig:int} Interaction between executive constraints and GDP per capita with 80\\% CIs"}
 # Interaction plot
@@
                    beta[, 6] * mean(rlvt_groups_count) +
-                   beta[, 7] * 1 +
-                   beta[, 8] * 1 +
+                   beta[, 7] * 0 +
+                   beta[, 8] * 0 +
                    beta[, 9] * mean(peace_yrs) +
@@
 
-par(mfrow = c(1, 3))
-for (y in state) {
-    print(y)
-    p <- boot::inv.logit(sapply(theta_seq, f, y)) * 100
-
-    m <- apply(p, 2, median)
-    pi <- apply(p, 2, quantile, probs = c(0.2, 0.8)) %>% t
-
-    plot(theta_seq, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
-         xlab = "Executive Constraints", ylab = "Predicted Probability")
-    lines(theta_seq, pi[, 1], lty = 2)
-    lines(theta_seq, pi[, 2], lty = 2)
-}
-```
-
-```{r eval = F}
-input.df <- readRDS("posteriors/model/final.rds")
-p_hat <- readRDS("posteriors/model/summary/p_hat.rds")
-
-plot_dens(p_hat, groups = input.df$lepisode_onset)
-```
-
-```{r eval = F}
-p <- apply(p_hat, 2, median)
-x <- evalmod(scores = p, labels = input.df$lepisode_onset)
-
-plot(x, curvetype = c("ROC", "PRC"))
-```
-
-```
+df <- lapply(state, function(y) {
+    mclapply(theta_seq, function(theta) {
+        p_hat <- boot::inv.logit(f(theta, y)) * 100
+        post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
+            mutate(theta = theta,
+                   y = y)
+    }, mc.cores = 4) %>% bind_rows
+}) %>% bind_rows
+
+df %<>%
+    rename(codelow = `10%`,
+           median = `50%`,
+           codehigh = `90%`) %>%
+    mutate(y = case_when(y == state[1] ~ "Low GDP per Capita",
+                         y == state[2] ~ "Median GDP per Capita",
+                         y == state[3] ~ "High GDP per Capita") %>%
+               factor(levels = c("Low GDP per Capita",
+                                 "Median GDP per Capita",
+                                 "High GDP per Capita")))
+
+ggplot(df, aes(theta, median)) +
+    geom_line() +
+    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
+    facet_grid(. ~ y) +
+    xlab("Executive Constraints") +
+    ylab("Predicted Probability")
+```
+
+From the results in figure \ref{fig:int} it is clear that the overall
+effect of executive constraints when interacted with GDP per capita on
+the predicted probability of conflict onset is negative. For countries
+with a relatively high level of GDP per capita, going from the worst
+to best in executive constraints decreases the predicted probability
+of conflict onset from [1.5, 4.5] down to below 0.5. This effect
+diminishes, however, for lower values of GDP per capita to the point
+that executive constraints are practically indeterminate to the risk
+of conflict onset for low-income country-years.
+
+### Assessing Model Fit
+
+Figure \ref{fig:phat} plots the densities of the first 50 posterior
+draws for the predicted probabilities of conflict onset for two
+groups, country-years that actually experienced an onset of a civil
+conflict and country-years that did not. On average, the predicted
+probabilities for the actual onset country-years are higher albeit
+still extremely low. This can be seen in the summary statistics
+reported in table \ref{tab:psum} for the posterior means of the
+predicted probabilities. For onset country-years the interquartile
+interval is only 0.0485 to 0.209 with a max of 0.556. To better
+compare the predicted probabilities to the actual data, table
+\ref{tab:confusion} presents a confusion matrix with a 0.2 cutoff,
+which corresponds to the third quartile in the summary table.
+`r signif((93 / (259 + 93)) * 100, 3)` percent of onsets and
+`r signif((7902 / (7902 + 172)) * 100, 3)` percent of non-onsets are
+correctly predicted. The false positive rate is
+`r signif((172 / (7902 + 172)) * 100, 3)` percent and the false
+negative rate is `r signif((259 / (93 + 259)) * 100, 3)` percent.
+
+```{r fig.height = 2, fig.cap = "\\label{fig:phat} Density plot for predicted probabilities"}
+p_hat <- extract(reg, "p_hat")
+
+plot_dens(p_hat, groups = final.df$lepisode_onset) + xlim(c(0, 0.4))
+```
+
+```{r psum, warning = F}
+p_mean <- colMeans(p_hat)
+
+df <- bind_rows(summary(p_mean[final.df$lepisode_onset == 1]),
+                summary(p_mean[final.df$lepisode_onset == 0])) %>%
+    mutate_all(~signif(., 3))
+
+rownames(df) <- c("Conflict Onset", "No Onset")
+
+kable(df, "latex", align = "c", booktabs = T, linesep = "", escape = F,
+      caption = "Summarized posterior predicted probabilities")
+```
+
+```{r confusion}
+m <- table(p_mean > .2, final.df$lepisode_onset)
+
+rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
+colnames(m) <- c("Actual: No Onset", "Actual: Onset")
+
+kable(m, "latex", align = "c", booktabs = T, linesep = "", escape = F,
+      caption = "Confusion matrix with a 0.2 cutoff")
+```
+
+```{r results = "hide"}
+x <- evalmod(scores = p_mean, labels = final.df$lepisode_onset)
+
+auc <- attr(x$roc[[1]], "auc") %>% signif(3)
+auc_pr <- attr(x$prcs[[1]], "auc") %>% signif(3)
+```
+
+Figure \ref{fig:roc} plots a receiver operating characteristic (ROC)
+curve depicting the tradeoff between the true positive rate
+(*sensitivity*) and the false positive rate --- $1 -
+\mbox{specificity}$ where *specificity* refers to the true negative
+rate --- for varying probability thresholds according to the posterior
+means of the predicted probabilities. For example, for a threshold of
+approximately 0.1, the model will correctly predict 50 percent of
+onsets while misclassifying only 8 percent of non-onsets. The total
+performance of the model can be summarized by the area under the curve
+(AUC), which in this case is `r auc`.[^auc] However, for the present
+analysis the results from the ROC curve are misleading since there is
+a significant class imbalance in the dataset. Onsets comprise only 4
+percent of all observations meaning that the false positive rate is
+less sensitive to changing thresholds. Figure \ref{fig:roc} therefore
+also plots a precision-recall (PRC) curve, which depicts the tradeoff
+between *precision* --- the number of true positives divided by the
+sum of the true positives and false positives --- and recall --- the
+number of true positives divided by the sum of the true positives and
+false negatives --- for varying thresholds. The advantage is that
+*precision* is invariant to the number of non-onset years in the
+dataset. From this plot it is clear that the precision of the model,
+the ability to discriminate between onset country-years and false
+predicted non-set country-years, drops drastically as the *recall* of
+the model increases. The *AUC-PR* for this plot is only `r auc_pr`.
+[^auc_pr]
+
+[^auc]: An AUC of 1 for a ROC curve indicates that the model is able
+    to perfectly predict the outcome while an AUC of 0.5 denotes the
+    performance of a purely random classifier.
+
+[^auc_pr]: Similar to the ROC curve, an ACU of 1 for a PRC curve
+    denotes a perfect model. The AUC for the null model, or random
+    classifier, is the proportion of actual positive outcomes to negative
+    outcomes in the dataset. In this case this means approximately
+    .042.
+
+```{r fig.width = 7, fig.cap = "\\label{fig:roc} hello"}
+
+autoplot(x, curvetype = c("ROC", "PRC"))
+```
+
+Taken together, the evidence suggests a fairly poor fit to the
+underlying data. The confusion matrix demonstrates that even with a
+relatively low cutoff, the model is still only able to accurately
+predict 25 percent of conflict onsets. According to the ROC curve,
+decreasing this threshold improves the true positive rate with a
+relatively decent tradeoff in terms of the false positive rate, but as
+stated, this is largely due to the imbalance in the outcome
+variable. In terms of *precision*, the performance of the model
+quickly degrades as the threshold decreases.
+
+### Discussion
+
+- Link back to original theory, confirmatory
+
+- Small effect size
+- Poor fit
+- Noisy Data
+- Robustness?
+- Compared to democracy?
+
+Further analyses?
 

### 2020-02-02 R/thesis.utils/R/plot.R
Update R pkg, add ggplot themes

@@ -10,26 +10,27 @@
 plot_pars <- function(x, ...) UseMethod("plot_pars", x)
 
 #' @export
-plot_pars.matrix <- function(x) {
-    df <- post_summarise(x, probs = c(0.05, 0.16, 0.5, 0.84, 0.95)) %>%
-        rename(codelow95 = `5%`,
-               codelow68 = `16%`,
+plot_pars.matrix <- function(x, ylab = "Parameter Estimates", hline = F) {
+    df <- post_summarise(x, probs = c(0.10, 0.5, 0.9)) %>%
+        rename(codelow = `10%`,
                median = `50%`,
-               codehigh68 = `84%`,
-               codehigh95 = `95%`)
+               codehigh = `90%`)
 
-    ggplot(df, aes_(x = ~parameter, y = ~median)) +
-        #geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
-        geom_errorbar(aes_(x = ~parameter, ymin = ~codelow95, ymax = ~codehigh95),
-                      color = "#0072B2", linetype = "dotted", width = 0) +
-        geom_errorbar(aes_(x = ~parameter, ymin = ~codelow68, ymax = ~codehigh68),
+    p <- ggplot(df, aes_(x = ~parameter, y = ~median)) +
+        geom_errorbar(aes_(x = ~parameter, ymin = ~codelow, ymax = ~codehigh),
                       color = "#0072B2", width = 0, size = 1) +
-        geom_point(fill = "#56B4E9", color = "#0072B2", size = 2.5, shape = 21) +
+        geom_point(fill = "#56B4E9", color = "#0072B2", size = 1.5, shape = 21) +
+        theme_tufte() +
         theme(panel.background = element_blank(),
               axis.line = element_line(color = "black"),
               axis.text.x = element_text(angle = 45, hjust = 1),
               axis.title.x = element_blank()) +
-        ylab("Parameter Estimates")
+        ylab(ylab)
+
+    if (isTRUE(hline))
+        p + geom_hline(yintercept = 0, alpha = 0.5, color = "grey")
+
+    p
 }
 
 #' @export
@@ -55,6 +56,8 @@ plot_dens <- function(m, n = 50, groups = NULL) {
     ggplot(df, aes(value, group = interaction(iteration, id), colour = id)) +
         stat_density(position = "identity", geom = "line",
                      n = 1024, alpha = 0.2) +
+        theme_tufte() +
+        scale_color_colorblind() +
[…]

### 2020-02-02 R/thesis.utils/R/post.R

@@ -2,12 +2,12 @@
 #'
 #' Calculates quantile intervals from a posterior distribution
 #' represented either by a NumericMatrix where rows are posterior
-#' draws or extracted directly from a `stanfit` object.
+#' draws or extracted directly from a `posterior` object.
 #'
 #' @param x NumericMatrix or `stanfit` object
 #' @param ... Arguments passed to methods
 #'
-#' @details By default, `post_summarise` will calculate 68% and 95%
+#' @details By default, `post_summarise` will calculate 2.5% and 97.5%
 #'     quantile intervals in addition to the posterior median.
 #'
 #' @export
@@ -33,6 +33,14 @@ post_summarise.posterior <- function(x, pars = NULL, ...) {
     post_summarise(m, ...)
 }
 
+#' Load posterior object
+#'
+#' Reads a posterior file and creates a `posterior` object.
+#'
+#' @param file Target file. Can be a compress csv (`csv.gz`).
+#'
+#' @return DataFrame with the additional class, `posterior`.
+#'
 #' @export
 read_post <- function(file) {
     fit <- data.table::fread(file, data.table = F)
@@ -47,8 +55,18 @@ read_post <- function(file) {
     structure(fit, class = c(class(fit), "posterior"))
 }
 
+#' Extract parameters
+#'
+#' Extracts posteriors from a `posterior` object created using
+#' `read_post`.
+#'
+#' @param x DataFrame with the class `posterior`.
+#' @param pars CharacterVector of target parameters. If the parameter
+#'     is a vector or matrix, the function will extract all matching
+#'     values.
+#'
 #' @export
-extract <- function(x, ...) UseMethod("extract", x)
+extract <- function(x, pars = NULL) UseMethod("extract", x)
 
 #' @export
 extract.posterior <- function(x, pars = NULL) {

### 2020-02-02 R/thesis.utils/R/utils.R

@@ -130,6 +130,37 @@ explode <- function(df, from = NULL, to = NULL) {
     dplyr::bind_rows(ll)
 }
 
+#' Ordinal Values
+#'
+#' Checks a numeric vector if all the values can be considered
+#' ordinal. This is equivalent to saying that the function checks if a
+#' given numeric vector can be coerced to an integer vector without
+#' loss.
+#'
+#' @param x NumericVector
+#'
+#' @details This is only intended for NumericVectors. For all other
+#'     object types, the function will return `FALSE`.
+#'
+#' @examples
+#' x <- 1:3
+#' is.ordinal(x)
+#'
+#' y <- c(1:3, 0.5)
+#' is.ordinal(y)
+#'
+#' @export
+is.ordinal <- function(x) UseMethod("is.ordinal", x)
+
+#' @export
+is.ordinal.integer <- function(x) TRUE
+
+#' @export
+is.ordinal.double <- function(x) all(x %% 1 == 0)
+
+#' @export
+is.ordinal.default <- function(x) FALSE
+
 #' Normalize vector
 #'
 #' Simple wrapper around the default arguments to [scale()] which
@@ -162,6 +193,50 @@ partial <- function(fn, ...) {
     function(...) do.call(fn, c(dots, list(...)))
 }
 
+#' Calculate Summary Statistics
+#'
+#' Calculates summary statistics for all numeric columns in a data
+#' frame.
+#'
+#' @param x DataFrame
+#' @param vars CharacterVector. Calculate statistics for only a subset
+#'     of columns.
+#'
+#' @return A data frame with the following descriptive statistics: N,
+#'     mean, standard deviation, min, and max.
+#'
+#' @examples
+#' df <- data.frame(x = letters[1:3], y = 1:3, z = rnorm(3))
+#' summary_stats(df)
+#'
+#' @export
+summary_stats <- function(x, vars = colnames(x)) UseMethod("summary_stats", x)
+
+#' @export
+summary_stats.data.frame <- function(x, vars = colnames(x)) {
+    if (is.null(vars))
+        stop("Missing target columns")
+
+    df <- x[, vars, drop = F]
+    # N, mean, std.dev, min, max
+    ll <- lapply(colnames(df), function(s) {
+        col <- df[, s, drop = T]
+
+        if (!is.numeric(col))
+            return(NULL)
+
+        data.frame(Variable = s,
+                   N = sum(!is.na(col)),
[…]

### 2020-02-02 R/thesis.utils/tests/testthat/test_utils.R

@@ -37,31 +37,6 @@ test_that("calc_peace_yrs", {
                  "Unsorted years vector")
 })
 
-test_that("normalize", {
-    fn <- function(x) as.vector(scale(x))
-    x <- 1:10
-
-    expect_identical(normalize(x), fn(x))
-    expect_equal(mean(normalize(x)), 0)
-    expect_equal(stats::sd(normalize(x)), 1)
-    expect_equal(normalize(rep(5, 5)), rep(NaN, 5))
-    expect_error(suppressWarnings(normalize(letters[1:3])))
-
-    x <- c(NA, 1:10, NA)
-    expect_identical(normalize(x), fn(x))
-})
-
-test_that("to_idx", {
-    years <- rep(1900:1902, 3)
-    expect_identical(to_idx(years), c(1, 2, 3, 1, 2, 3, 1, 2, 3))
-
-    years <- c(1901, 1900, 1901, 1901, 1905)
-    expect_identical(to_idx(years), c(2, 1, 2, 2, 3))
-    expect_identical(to_idx(years, unique(years)), c(1, 2, 1, 1, 3))
-
-    countries <- c("SWE", "AFG", "FRA")
-    expect_identical(to_idx(countries), c(3, 1, 2))
-})
 
 test_that("consecutive", {
     years <- c(1900:1903, 1905:1906)
@@ -107,8 +82,73 @@ test_that("explode", {
     expect_equal(explode(x, x$from, x$to), out)
 })
 
+test_that("is.ordinal", {
+    # Force vector to be numeric
+    x <- c(0, 1, 2, 3)
+    y <- c(1L, 2L, 3L)
+
+    expect_true(is.ordinal(x))
+    expect_true(is.ordinal(y))
+
+    x <- c(0.5, 2)
+    y <- letters[1:3]
+    z <- c(T, F, T, F, F)
+
+    expect_false(is.ordinal(x))
+    expect_false(is.ordinal(y))
+    expect_false(is.ordinal(z))
+
+    expect_false(is.ordinal(data.frame()))
+    expect_false(is.ordinal(matrix()))
+    expect_false(is.ordinal(array()))
+})
+
+
+test_that("normalize", {
+    fn <- function(x) as.vector(scale(x))
+    x <- 1:10
+
+    expect_identical(normalize(x), fn(x))
+    expect_equal(mean(normalize(x)), 0)
+    expect_equal(stats::sd(normalize(x)), 1)
+    expect_equal(normalize(rep(5, 5)), rep(NaN, 5))
[…]

### 2020-02-02 R/merge.R
Add discriminated population variable grom GROWup

@@ -236,7 +236,7 @@ stopifnot(!is.na(merged.df$country_name))
 growup <- fread("./data/raw/growup/data.csv", data.table = F) %>%
     select(gwid = countries_gwid, country_name = countryname, year,
            onset_ko_flag, area_sqkm, meanelev, rlvt_groups_count,
-           incidence_flag, peaceyears)
+           discrimpop, incidence_flag, peaceyears)
 
 filter(growup, !gwid %in% merged.df$gwid) %$%
     unique(country_name) %>%

### 2020-02-02 R/model_data.R
Normalize final model dataset after dropping NAs

@@ -26,6 +26,7 @@ dbg_info(final.df)
 # Control variables
 X <- final.df %>%
     select(-country_name, -year, -matches("onset"), -reduced_idx) %>%
+    mutate_if(Negate(is.ordinal), normalize) %>%
     data.matrix
 
 if (schema$interaction != "") {

### 2020-02-02 R/transform.R
Add CINC as state capacity indicator

@@ -35,17 +35,16 @@ for (v2 in lg_vars) {
 # Final dataset for model
 final.df <- merged.df %>%
     select(country_name, year, lonset, lmajor_onset, lepisode_onset,
-           lepisode_major_onset, peace_yrs, one_of(constraint_vars),
-           one_of(paste0(constraint_vars, "_sd")), rgdpepc,
-           rgdpepc_gro, pop_density, meanelev, ongoing, cinc, independence,
-           rlvt_groups_count, neighbour_conflict, v2lgbicam) %>%
+           lepisode_major_onset, peace_yrs, v2lgbicam,
+           one_of(constraint_vars), one_of(paste0(constraint_vars, "_sd")),
+           rgdpepc, rgdpepc_gro, cinc, pop_density, meanelev, ongoing,
+           independence, rlvt_groups_count, discrimpop,
+           neighbour_conflict) %>%
     filter_at(constraint_vars, all_vars(!is.na(.))) %>%
-    mutate(rgdpepc = log(rgdpepc) %>% normalize,
-           rgdpepc_gro = normalize(rgdpepc_gro),
-           pop_density = log(pop_density) %>% normalize,
-           meanelev = log(meanelev) %>% normalize,
-           peace_yrs = log(peace_yrs + 1) %>% normalize,
-           cinc = normalize(cinc),
+    mutate(rgdpepc = log(rgdpepc),
+           pop_density = log(pop_density),
+           meanelev = log(meanelev),
+           peace_yrs = log(peace_yrs + 1),
            reduced_idx =
                do.call(paste, lapply(c("country_name", constraint_vars), as.symbol)) %>%
                    collapse_changes)

### 2020-02-02 paper.Rmd
First draft paper

@@
 library(ggplot2)
+library(ggthemes)
 library(gridExtra)
@@
                fig.height = 3)
-
 ```
@@
 # Introduction {-}
+
+Since the end of World War II, civil conflict has been the deadliest
+and most prevalent form of conflict. In that time period, upwards to
+two-thirds of all countries have experienced at least one minor
+insurgency and approximately one-third of all countries have
+experienced a major civil war with at least 1,000 battle deaths
+[@Pettersson2019]. Given the impact of civil conflict and the
+increasing availability of conflict related data, the last several
+decades has witnessed a veritable explosion in the number of
+quantitative studies investigating the causes of civil conflict
+[@Blattman2017]. Scholars have proposed numerous theories ranging
+from motivation-based arguments --- such as, for example, the role of
+horizontal inequalities [@Cederman2011] --- to opportunity-based
+explanations centered on economic factors [@Collier2012; @Collier2009]
+as well as the political and structural conditions underlying state
+weakness [@Hegre2001; @Fearon2003]. This latter strain of research in
+particular has spurred extensive work into examining the
+distinguishing features of weak states and whether political
+institutions play a significant role in the risk of civil conflict.
+Several studies in this vein have argued in favor of institutions
+utilizing disaggregated measures to also investigate the specific
+types of institutions relevant to the study of civil conflict
+[@Taydas2010; @Hegre2015a; @Bartusevic2018].
+
+This paper extends that work by taking a disaggregated approach to
+studying the effect of executive constraints on the risk of civil
+conflict onset. The motivation to specifically focus on executive
+constraints is two-fold: first, the executive is often times one of
+the key decision-makers in a state's regime. Institutions that provide
+a check against the types of actions which may exacerbate the risk of
+rebellion should be relevant to the study of civil conflict. In doing
+we can further our understanding of how political institutions shape
+the incentives for violence, especially through the influence of
+institutions on state behaviour.  Second, the link between executive
+constraints and civil conflict is relatively understudied. The
+previous literature that exists is fairly limited and contradictory in
+terms of results [@Getmansky2013; @Fjelde2009].
+
+The principle finding is that there is tentative evidence that
+executive constraints do matter and that there is a negative
+relationship between the latent level of constraints on the executive
+and the risk of civil conflict onset. However, this relationship
+largely disappears for low-income country-years when an interaction
+term with GDP per capita is included. If the primary theoretical
+mechanism involves the impact of executive constraints on rebel
+mobilization --- an unchecked executive may be more likely to engage
+in indiscriminate violence which could lead to a loss of legitimacy as
+well as the ability of rebel groups to offer local security guarantees
+to attract recruits --- then the incentives for violence in low-income
+country-years may be so great that it does not matter how constrained
+the executive may or may not be.
+
+We should not be overconfident with these results, however. Similar to
+most other studies examining conflict onset, the model is a poor fit
+to the underlying data. Modeling civil conflict is a difficult task,
+especially when trying to capture the dynamics of conflict with a
+standard time series cross-country regression. The results will
+require verification by additional studies using alternative
+methodological approaches.
+
+In what follows, a deeper justification for why political institutions
+are relevant to the study of civil conflict together with an overview
+of related empirical literature will be presented. The theoretical
+mechanisms linking executive constraints to conflict onset will then
+be discussed followed by an empirical analysis for which a latent
+estimate for executive constraints will be estimated and used in a
+cross-country time series logistic regression predicting conflict
+onset.
 
@@
 There are three general pathways which can theoretically link the
-concept of executive constraints to the onset of civil conflict:
+level of executive constraints to the onset of civil conflict:
 executive looting and state weakness, credible commitments to
@@
 benefits to offset the associated costs of fighting
-[@Mason1996]. [^rational] These benefits can include not only economic
-payoffs, but also protection from government violence. When the
-government engages in violent repression, especially in the form of
+[@Mason1996]. These benefits can include not only economic payoffs,
+but also protection from government violence. When the government
+engages in violent repression, especially in the form of
 indiscriminate violence, the cost of non-participation is
@@
 free to engage in reprisals.
-
-[^rational]: TODO: Assuming rational actors...
 
@@
 constraints should be associated with a lowered risk of civil conflict
-onset.  As discussed, this may be conditional though on the relative
-capacity of the state --- if executive constraints affect the
-mobilization potential for rebel groups, then in situations where the
-government is willing to engage in repression and has the strength to
-fully suppress the opposition, the impact of executive constraints on
-civil conflict onset should be diminished. Furthermore, even if the
-mobilization mechanism is not valid, the relationship could still be
-confounded by the overall opportunity for rebellion --- the degree of
-executive constraints may not be particularly relevant in relatively
-stable countries with, for example, high levels of economic
-development, little to no ethnic fractionalisation, and no easily
-lootable resources to finance a rebellion. Regardless of
-conditionalities, contrary results would indicate either that
-executive constraints are indeterminate to the risk of civil conflict
-or possibly that there is even a positive relationship since
-constrained executives may be less able to effectively deal with
-violent challenges due to risk aversion to violence and/or from the
-slowdown of having to seek approval from a plurality of stakeholders
-before committing to decisive action.
+onset. This proposition can be further extended if state weakness is a
+primary driver of civil conflict and the relative capacity of the
+state is effected by the levels of executive constraints, as depicted
+in the first theoretical pathway regarding executive looting, then
+state capacity could have a mediating affect on the relationship
+between executive constraints and conflict onset. Another extension
+would be if executive constraints affect the mobilization potential
+for rebel groups, then there could also be a confounding effect from
+the relative threshold for mobilization. Finally, the relationship
+could still be confounded by the overall opportunity for rebellion ---
+the degree of executive constraints may not be particularly relevant
+in relatively stable countries lacking the conditions conducive to
+carrying out a rebellion. Regardless of conditionalities, contrary
+results would indicate either that executive constraints are
+indeterminate to the risk of civil conflict or possibly that there is
+even a positive relationship since constrained executives may be less
+able to effectively deal with violent challenges due to risk aversion
+to violence and/or from the slowdown of having to seek approval from a
+plurality of stakeholders before committing to decisive action.
 
@@
     geom_line() +
+    theme_tufte() +
+    scale_color_colorblind() +
     theme(legend.title = element_blank()) +
@@
 outbreak of conflict. One possible mediator is the relative capacity
-of the state. For the present analysis I follow @Fearon2003 in using
-GDP per capita as a proxy for the state's administrative and military
-capacity by roughly capturing the revenue potential available to the
-state. This is far from being an ideal proxy for state capacity; GDP
-per capita could also capture economic grievances as well as the
-relative ease of financing a rebellion given low wages
-[@Hendrix2010]. Data on GDP per capita, however, is significantly more
-complete than other common competing measures such as bureaucratic
-quality or tax revenue as a percent of GDP.[^wbgi] Furthermore, it is
-more correlated with the administrative aspects of state capacity than
-measures focused solely on military strength, such as military
-expenditures [@Hendrix2010]. Thus, I use the natural log of the
-*RGDP^o* variable, constant price real GDP, from Penn World Tables
+of the state. For the present analysis I follow @Bartusevic2018 in
+using the Composite Indicator of National Capability (CINC) from the
+National Material Capabilities v5.0 dataset (1816-2012) [@Singer1972;
+@Singer1987]. The CINC index is aggregated from several demographic,
+industrial, and military indicators in order to capture the available
+resources a country can bring to bear in the event of a conflict. This
+is an imperfect measure that largely misses the state's bureaucratic
+and administrative capacity [@Hendrix2010]. [^wbgi] However, given
+that the dimension of interest for this analysis is the repressive
+capacity of the state, material capabilities, especially military
+resources, should still provide an adequate proxy.
+
+[^wbgi]: Furthermore, a common issue with many measures of
+    bureaucratic quality is missingness.  For example, the World Bank
+    governance indicator on government effectiveness extends back only
+    until 1996 (TODO: cite). Given that measures of institutional
+    quality often change fairly slowly, punctuated only by abrupt
+    events such as coups or revolutions, an extended time series is
+    desirable. Furthermore, in the specific case of bureaucratic
+    quality, there is the added issue of endogeneity. An unconstrained
+    executive may seek to promote government corruption and undermine
+    both meritocratic recruitment into the bureaucracy as well as its
+    impartiality.
+
+To capture the relative ease of mobilizing fighters, especially in
+terms of being able to offer economic selective benefits, I include
+GDP per capita. Specifically, I use the natural log of the $RGDP^o$
+variable, constant price real GDP per capita, from Penn World Tables
 (PWT) v9.1 (1950-2017) to allow for both cross-country and
-cross-temporal comparisons of GDP per capita [@Feenstra2015] and
-include an interaction term between GDP per capita and the latent
-measure for executive constraints to capture the conditional
+cross-temporal comparisons of GDP per capita [@Feenstra2015]. The
+justification for taking the natural log is that the marginal effect
+of the variable should decrease with higher values. In other words,
+the difference between the effects of relatively high levels of GDP
+per capita should be significantly smaller than the difference between
+the effects of low levels of GDP per capita. Finally, I also include
+an interaction term between GDP per capita and the latent measure for
+executive constraints to capture the potential conditional
 relationship.
 
-[^wbgi]: For example, the World Bank governance indicator on
-    government effectiveness extends back only until 1996 (TODO:
-    cite). Given that measures of institutional quality often change
-    fairly slowly, punctuated only by abrupt events such as coups or
-    revolutions, an extended time series is desirable. Furthermore, in
-    the specific case of bureaucratic quality, there is the added
-    issue of endogeneity. An unconstrained executive may seek to
-    promote government corruption and undermine both meritocratic
-    recruitment into the bureaucracy as well as its impartiality.
-
-In addition to state capacity, I also include a number of standard
-variables drawn largely from previous literature. The list of
-variables are not meant to provide an all encompassing theory of civil
-conflict, but rather roughly capture the opportunity structures for
-armed conflict.  This includes a single year lagged measure of GDP per
+In addition to the previous indicators, I also include a number of
+standard covariates drawn largely from previous literature. The list
+of indicators are not meant to provide an all encompassing theory of
+civil conflict, but rather roughly capture the opportunity structures
+for rebellion.  This includes a single year lagged measure of GDP per
 capita growth from PWT in order to capture short term economic shocks,
@@
 higher population density may require a more extensive and competent
-central administration and policing force in order to suppress violent
+central administration and police force in order to suppress violent
 dissent [@Fearon2003]. There's also the possibility that a high
@@
 state is identified as the source of a particular scarcity, the risk
-of civil conflict may increase. Similarly to GDP per capita, I also
-transform mean elevation and population by the natural log. Finally, I
-include measures of ethnic fractionalization in the form of the number
-of political relevant ethnic groups and the percent of the population
-facing discrimination. Where there are significant ethnic cleavages
-and state sponsored discrimination, ethnic ties can facilitate
-collective action thereby lowering the mobilization costs of rebel
-groups [@Cederman2011].
+of civil conflict may increase. Similar to GDP per capita, I also
+transform mean elevation and population by the natural log with the
+expectation that the scale increase should diminish as values get
+larger. Finally, I include measures of ethnic fractionalization in the
+form of the number of political relevant ethnic groups and the percent
+of the population facing discrimination. Where there are significant
+ethnic cleavages and state sponsored discrimination, ethnic ties can
+facilitate collective action thereby lowering the mobilization costs
+of rebel groups [@Cederman2011].
 
@@
 parameter estimates from the executive constraints measurement model
-will be presented followed by the parameters pertaining to the
-logistic regression and an assessment of model fit. The final section
-will be a discussion linking the results back to the outlined theories
-as well as an overview of the limitations of the presented model.
+will be presented. This will then be followed by the parameters
+pertaining to the logistic regression before concluding with an
+assessment of model fit.
 
@@
 
-```{r  fig.cap = "\\label{fig:fpars} hello", warning = F}
+```{r  fig.cap = "\\label{fig:fpars} Measurement Model Parameter Estimates"}
 load("./posteriors/full_model/input.RData")
@@
 
-p1 <- plot_pars(lambda)
-p2 <- plot_pars(psi)
-p3 <- plot_pars(gamma)
-p4 <- plot_pars(delta_eta)
+p1 <- plot_pars(lambda, "Lambda")
+p2 <- plot_pars(psi, "Psi")
+p3 <- plot_pars(gamma, "Gamma")
+p4 <- plot_pars(delta_eta, "")
 
@@
 
-```{r fa, fig.height = 6, fig.width = 6, warning = F}
+```{r fa, fig.height = 6, fig.width = 6}
 fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
@@
 
-```{r fig.width = 3, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for $\\theta$"}
+```{r fig.width = 6, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for executive constraints"}
 theta <- extract(fa, "theta")
-plot_dens(theta, groups = lgbicam)
+
+groups <- ifelse(lgbicam == 1, "Legislature Present", "No Legislature")
+plot_dens(theta, groups = groups)
 
@@
 
-```{r fig.cap = "\\label{fig:fa-line} Error"}
+```{r fig.cap = "\\label{fig:fa-line} Posterior shrinkage for \\textit{v2juncind}"}
 err <- read_post("posteriors/full_model/err_posteriors.csv.gz")
 
+target <- colnames(nonlg_mm$obs)[1]
 nonlg_est <- extract(err, "nonlg_est") %>%
@@
                 group = rep(1:length(idx), 2),
-                factor = c(rep(1, length(idx)), rep(2, length(idx))) %>% as.factor,
+                factor = c(rep("Observed Point Estimate", length(idx)),
+                           rep("Estimated Posterior Median", length(idx))) %>% as.factor,
                 theta = rep(theta_medians, 2))
@@
     geom_point(alpha = 0.8) +
-    geom_line(data = y, aes(theta, yhat, group = NA, col = NA), alpha = 1)
-
-```
-
-```{r fig.cap = "\\label{fig:err} Errors"}
+    geom_line(data = y, aes(theta, yhat, group = NA), colour = "black", alpha = 1) +
+    theme_tufte() +
+    scale_color_colorblind() +
+    theme(legend.title = element_blank()) +
+    ylab(target) +
+    xlab("Theta")
+
+```
+
+```{r fig.cap = "\\label{fig:err} Error plots"}
 p1 <- ggplot(nonlg_est, aes(se, interval)) +
-    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
+    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1) +
+    theme_tufte() +
+    ylab("Posterior CI Range") +
+    xlab("Observed Error")
 
 p2 <- ggplot(nonlg_est, aes(se, diff)) +
-    geom_point(col = "#56B4E9", alpha = 0.2)
+    geom_point(col = "#56B4E9", alpha = 0.2) +
+    theme_tufte() +
+    ylab("Pointwise difference") +
+    xlab("Observed Error")
 
@@
 
-```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Theta"}
+```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Executive Constraints Over Time"}
 theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
@@
     ggplot(aes(year, median, fill = country_name)) +
+    theme_tufte() +
+    scale_fill_colorblind() +
     geom_line() +
-    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2)
-
-```
-
-```{r fig.cap = "\\label{fig:tie} Tie Fighter"}
+    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
+    theme(legend.title = element_blank())
+
+```
+
+```{r fig.cap = "\\label{fig:tie} Executive Constraints 2000 vs 2010"}
 targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
@@
 
-df <- filter(theta_summarised, year %in% c(2017, 2007),
+df <- filter(theta_summarised, year %in% c(2010, 2000),
              country_name %in% targets) %>%
@@
 
-ggplot(df, aes(median, country_name, colour = year)) +
-    geom_point() +
-    geom_errorbarh(aes(xmax = codehigh, xmin = codelow))
+ggplot(df, aes(country_name, median, colour = year)) +
+    geom_errorbar(aes(ymax = codehigh, ymin = codelow),
+                  position = position_dodge(width = 0.5)) +
+    theme_tufte() +
+    scale_color_colorblind() +
+    theme(axis.title.y = element_blank(),
+          legend.title = element_blank()) +
+    ylab("Theta") +
+    coord_flip()
 
@@
 ```{r results = "hide"}
-romania_idx <- filter(final.df, country_name == "Romania", year %in% c(2007, 2017)) %$%
+chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2000, 2010)) %$%
     reduced_idx
 
-romania_p <- signif(mean(theta[, romania_idx[1]] > theta[, romania_idx[2]]) * 100, 2)
-
-chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2007, 2017)) %$%
+chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
+
+tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
     reduced_idx
 
-chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
+tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
 ```
@@
 To provide a further snapshot of a greater number of countries, figure
-\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
-the dataset, compared to 2007 by country. A full list of all countries
+\ref{fig:tie} plots the $\theta$ estimates for 2010, the last year in
+the dataset, compared to 2000 by country. A full list of all countries
 is available in [Appendix C](#appendix-c). Note, the statistical
@@
 strictly be judged by overlapping credible intervals
-[@Schenker2001]. For example, Romania, despite having overlapping
+[@Schenker2001]. For example, Chile, despite having overlapping
 credible intervals, still has a posterior probability of approximately
-$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
-the case of Chile with even more visual overlap, there is a
-$`r chile_p`\%$ posterior probability that the estimate for 2017 is
-higher than 2007.
+$`r chile_p`\%$ that the estimate for 2010 is lower than 2000. In the
+case of Tunisia with even more visual overlap, there is a
+$`r tunisia_p`\%$ posterior probability that the estimate for 2010 is
+higher than 2000.
 
@@
 
-```{r reg, fig.height = 3, warning = F}
+```{r reg, fig.height = 3}
 reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")
@@
 
-```{r fig.cap = "\\label{fig:reg} hello world"}
-plot_pars(beta)
+```{r fig.cap = "\\label{fig:reg} Regression parameters"}
+plot_pars(beta, hline = T)
 ```
@@
 ```{r results = "hide"}
-merged.df <- readRDS("data/merged_data.rds")
-
 gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
@@
 ```{r results = "hide"}
-gdp_eth <- filter(merged.df, country_name == "Ethiopia", year == 2012) %$%
+gdp_rwa <- filter(merged.df, country_name == "Rwanda", year == 2007) %$%
     rgdpepc %>%
@@
 
-gdp_india <- filter(merged.df, country_name == "India", year == 2014) %$%
+gdp_phl <- filter(merged.df, country_name == "Philippines", year == 2008) %$%
     rgdpepc %>%
@@
 
-gdp_nz <- filter(merged.df, country_name == "New Zealand", year == 2002) %$%
+gdp_po <- filter(merged.df, country_name == "Portugal", year == 2009) %$%
     rgdpepc %>%
@@
 values. These levels roughly correspond to the GDP per capita of
-Ethiopia in 2012 at $`r gdp_eth`$, India in 2014 at $`r gdp_india`$,
-and New Zealand in 2002 at $`r gdp_nz`$. In the dataset there are $75$
-conflict episode onsets below the first level, $198$ between the first
-and second, $73$ between the second and third, and $6$ above the
+Rwanda in 2007 at $`r gdp_rwa`$, the Philippines in 2008 at $`r gdp_phl`$,
+and Portugal in 2009 at $`r gdp_po`$. In the dataset there are 60
+conflict episode onsets below the first level, $152$ between the first
+and second, $55$ between the second and third, and $4$ above the
 third.
@@
 
-state <- quantile(final.df$rgdpepc, probs = c(0.1, .5, 0.9))
+state <- quantile(X[, "rgdpepc"], probs = c(0.1, .5, 0.9))
 
 f <- function(x, y) {
-    with(final.df, alpha +
+    with(as.data.frame(X),
+         alpha +
          beta[, 1] * x +
@@
          beta[, 3] * mean(rgdpepc_gro) +
-                   beta[, 4] * mean(pop_density) +
-                   beta[, 5] * mean(meanelev) +
-                   beta[, 6] * mean(rlvt_groups_count) +
-                   beta[, 7] * 0 +
-                   beta[, 8] * 0 +
-                   beta[, 9] * mean(peace_yrs) +
+         beta[, 4] * mean(cinc) +
+         beta[, 5] * mean(discrimpop) +
+         beta[, 6] * mean(pop_density) +
+         beta[, 7] * mean(meanelev) +
+         beta[, 8] * mean(rlvt_groups_count) +
+         beta[, 9] * 0 +
          beta[, 10] * 0 +
-                   beta[, 11] * x * y)
+         beta[, 11] * mean(peace_yrs) +
+         beta[, 12] * 0 +
+         beta[, 13] * x * y)
 }
@@
 df <- lapply(state, function(y) {
-    mclapply(theta_seq, function(theta) {
-        p_hat <- boot::inv.logit(f(theta, y)) * 100
+    mclapply(theta_seq, function(point) {
+        p_hat <- boot::inv.logit(f(point, y)) * 100
         post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
-            mutate(theta = theta,
+            mutate(theta = point,
                    y = y)
@@
     facet_grid(. ~ y) +
+    theme_tufte() +
     xlab("Executive Constraints") +
     ylab("Predicted Probability")
+
 ```
@@
 
-plot_dens(p_hat, groups = final.df$lepisode_onset) + xlim(c(0, 0.4))
-```
-
-```{r psum, warning = F}
+groups <- ifelse(final.df$lepisode_onset == 1,
+                 "Actual Onset",
+                 "No Onset")
+plot_dens(p_hat, groups = groups) +
+    xlim(c(0, 0.4))
+```
+
+```{r psum}
 p_mean <- colMeans(p_hat)
@@
 
-```{r fig.width = 7, fig.cap = "\\label{fig:roc} hello"}
+```{r fig.width = 7, fig.cap = "\\label{fig:roc} ROC & Precision-Recall Plots"}
 
@@
 
-### Discussion
-
-- Link back to original theory, confirmatory
-
-- Small effect size
-- Poor fit
-- Noisy Data
-- Robustness?
-- Compared to democracy?
-
-Further analyses?
-
-# Conclusion {-}
+# Discussion {-}
+
+How do the results from the presented model relate back to the
+theories potentially linking executive constraints to civil conflict?
+According to the model there is support for the general proposition
+that the risk of civil conflict onset decreases the more constrained
+the executive is by other government branches when controlling for the
+selected covariates. As stated previously, however, which theoretical
+mechanism(s) might be responsible for this relationship lies outside
+the scope of this paper in large part because the mechanisms that were
+discussed are not mutually exclusive. Furthermore, quantitative
+country-year regressions such as this one can only detect general
+patterns and provide evidence towards the associations of particular
+indicators. What they cannot do is examine the microfoundations
+underpinning a particular theory.
+
+That being said, on the basis of the mobilization theory I did argue
+that there may be a potential confounder in the relative threshold for
+mobilization. I operationalized this in the model as GDP per capita
+and included an interaction term with the latent construct for
+executive constraints. Accordingly, the negative relationship between
+executive constraints and conflict onset does not hold for low-income
+country-years. In these cases, the effect is practically
+non-existent. According to theory this would mean that where labor
+costs are low and the provisioning of economic selective benefits
+presents less of an obstacle for rebel leaders to mobilize fighters,
+the effect of executive constraints on the risk of civil conflict is
+largely indeterminate. This may be because mobilization in these
+situations is independent of the government's decision to engage in
+repression, assuming that the state does not have the capacity to
+completely eradicate the opposition immediately. More specifically,
+low-income countries may be vulnerable to the types of predatory
+insurgent groups which require significantly less fighters than rebel
+groups seeking regional independence or to topple the central
+regime. This interpretation could also hold even if GDP per capita was
+instead understood as proxying the state's administrative and military
+capacity by roughly capturing the revenue potential available to the
+state [@Fearon2003]. Where the state is particularly weak it simply
+might not matter how constrained the executive is.
+
+Despite these results, there is one significant caveat to the
+presented model. Namely, the model and the estimated predicted
+probabilities were a poor fit to the underlying data. The tradeoff
+between precision and recall was far from satisfactory; the model
+could easily predict non-onset country-years in large part because of
+the class imbalance in the dataset between onsets and non-onsets, but
+had difficultly predicting actual onsets with accuracy. This is
+unsurprising given that civil conflict onsets are relatively rare
+events and studies attempting to examine the impact of structural
+factors, such as political institutions, on onset risk are dealing
+with small effect sizes, noisy data, and relatively few variables of
+actual exogenous change. This is an issue that has plagued much of the
+conflict literature [@Ward2010] so the model presented here is far
+from being an outlier. But it does mean that any results should be
+considered tentative at best, at least until confirmed by additional
+studies.
+
+Finally, in terms of future research, if we want to better understand
+the impact of political institutions on the risk of civil conflict we
+need to directly model the behaviour of rebel groups and state
+actors. For executive constraints, and specifically the theoretical
+mechanism focused on rebel mobilization, one possible avenue that was
+unexplored in this paper would be to model the effects of
+repression. A future study could examine the temporal sequence of
+executive constraints on the likelihood of government repression and
+its effect in turn on the risk of civil conflict onset. Such an
+analysis would move beyond the standard country-year logistic
+regression, which is a poor model for the dynamics of civil conflict,
+and attempt to model country-date data, possibly even at the
+subnational level depending on data availability.
 
@@
 
+\newpage
+
+## D - Additional Descriptive Statistics {#appendix-d}
+
+## E - Robustness Checks {#appendix-e}
+
 # References {-}

### 2020-02-11 paper.Rmd
Draft edits

@@
 quantitative studies investigating the causes of civil conflict
-[@Blattman2017]. Scholars have proposed numerous theories ranging
-from motivation-based arguments --- such as, for example, the role of
+[@Blattman2017]. Scholars have proposed numerous theories ranging from
+motivation-based arguments --- such as, for example, the role of
 horizontal inequalities [@Cederman2011] --- to opportunity-based
@@
 institutions play a significant role in the risk of civil conflict.
-Several studies in this vein have argued in favor of institutions
+Several studies in this vein have argued in favor of institutions by
 utilizing disaggregated measures to also investigate the specific
@@
 using the Composite Indicator of National Capability (CINC) from the
-National Material Capabilities v5.0 dataset (1816-2012) [@Singer1972;
-@Singer1987]. The CINC index is aggregated from several demographic,
+National Material Capabilities v5.0 dataset (1816-2012)
+[@Singer1987]. The CINC index is aggregated from several demographic,
 industrial, and military indicators in order to capture the available
@@
 df <- merged.df %>%
-    select(rgdpepc, rgdpepc_gro, pop_density, meanelev,
+    select(rgdpepc, rgdpepc_gro, pop_density, meanelev, discrimpop,
            rlvt_groups_count, neighbour_conflict, ongoing, peace_yrs,
@@
 
-```{r fig.cap = "\\label{fig:err} Error plots"}
+```{r fig.cap = "\\label{fig:err} Posterior CI and point estimates vs. observed error"}
 p1 <- ggplot(nonlg_est, aes(se, interval)) +
@@
     select(pop, area_sqkm) %>%
-    mutate(area_sqlm = signif(area_sqkm, 3))
+    mutate(area_sqkm = signif(area_sqkm, 3),
+           pop = signif(pop, 3))
 
@@
     select(pop, area_sqkm, meanelev) %>%
-    mutate(area_sqkm = signif(area_sqkm, 3),
+    mutate(pop = signif(pop, 3),
+           area_sqkm = signif(area_sqkm, 3),
            meanelev = signif(area_sqkm, 3))
@@
 predicted probabilities. For onset country-years the interquartile
-interval is only 0.0485 to 0.209 with a max of 0.556. To better
+interval is only 0.0475 to 0.2 with a max of 0.516. To better
 compare the predicted probabilities to the actual data, table
@@
 
-```{r fig.width = 7, fig.cap = "\\label{fig:roc} ROC & Precision-Recall Plots"}
+```{r fig.width = 8, fig.cap = "\\label{fig:roc} ROC & Precision-Recall Plots"}
 
@@
 
+\newpage
+
 ## E - Robustness Checks {#appendix-e}
 
+```{r}
+models <- setdiff(list.files("./posteriors/"), c("full_model_new", "sim"))
+
+# TODO: this is killing our compile times
+all_betas <- mclapply(models, function(f) {
+    load(file.path("posteriors", f, "input.RData"))
+    beta <- file.path("posteriors", f, "reg_posteriors.csv.gz") %>%
+        read_post %>%
+        extract("beta")
+
+    colnames(beta)[1:(1 + ncol(X))] <- c("exec_constraints", colnames(X))
+    if (interaction_idx > 0)
+        colnames(beta)[ncol(beta)] <- paste("exec *", colnames(X)[interaction_idx])
+
+    post_summarise(beta, probs = c(0.1, 0.5, 0.9)) %>%
+        mutate(model = f)
+}, mc.cores = 4) %>% bind_rows
+
+all_betas %<>%
+    rename(codelow = `10%`,
+           median = `50%`,
+           codehigh = `90%`) %>%
+    mutate(model = to_idx(model) %>% as.factor)
+
+desc <- case_when(models == "full_model" ~ "Full Model",
+                  models == "full_no_cinc" ~ "Full Model Excl. CINC",
+                  models == "full_no_interaction" ~ "Full Model Without Interaction",
+                  models == "major_episodes" ~ "Major Conflict Episodes Only",
+                  models == "major_no_interaction" ~ "Major Conflict Episodes Without Interaction",
+                  models == "unique_no_interaction" ~ "Unique Onsets Without Interaction",
+                  models == "unique_onset" ~ "Unique Onsets Only")
+
+df <- data.frame(Model = 1:length(models),
+                 Description = desc,
+                 stringsAsFactors = F)
+
+kable(df, "latex", align = c("l", "l"), booktabs = T, linesep = "",
+      caption = "Model Descriptions") %>%
+    kable_styling(latex_options = c("hold_position", "striped"))
+
+```
+
+```{r fig.width = 7, fig.height = 6}
+ggplot(all_betas, aes(x = parameter, y = median, color = model)) +
+    geom_point(position = position_dodge(width = 0.8)) +
+    geom_errorbar(aes(x = parameter, ymin = codelow, ymax = codehigh),
+                  width = 0, size = 1, position = position_dodge(width = 0.8)) +
+    geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
+    theme_tufte() +
+    scale_color_colorblind() +
+    theme(panel.background = element_blank(),
+          legend.title = element_blank(),
+          axis.line = element_line(color = "black"),
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          axis.title.x = element_blank()) +
+    ylab("Parameter Estimate")
+
+```
+
+\newpage
+
 # References {-}

### 2020-02-12 utils/extract.c
Rename select to extract and speed up read with posix_fadvise

@@ -1,19 +1,21 @@
-// select.c
+// extract.c
 //
-// ./select [-h] [-n <nlines>] -s <parameter>... <FILE>...
+// ./extract [-h] [-n <nlines>] -s <regex> <FILE>...
 //
 // Simple program that extracts specified parameters from a Stan posterior
-// output file(s) and prints the result to stdout. 
+// output file(s) and concatenates the result to stdout.
 //
-// example: ./select -s '^alpha|^beta' -n 10 samples-chain_*.csv
+// example: ./extract -s '^alpha|^beta' -n 10 samples-chain_*.csv
 //
 // Fair warning: there's a lot of shortcuts in this program. Memory is not
 // freed and file descriptors are not closed with the assumption that the OS
-// will take of it when we exit. 
+// will take of it when we exit.
 
 #define _GNU_SOURCE
 
+#include <err.h>
 #include <errno.h>
+#include <fcntl.h>
 #include <regex.h>
 #include <stdbool.h>
 #include <stdlib.h>
@@ -22,45 +24,40 @@
 #include <string.h>
 #include <unistd.h>
 
-#define BUF_SIZE 1024 * 1024
-#define die(msg)                       \
-    do {                               \
-        fprintf(stderr, "%s\n", msg);  \
-        exit(EXIT_FAILURE);            \
-    } while (0);
+#define BUF_SIZE 1024 * 1024 * 8
 
 typedef struct bitarray {
-    size_t capacity;
-    uint64_t *data;
+size_t capacity;
+uint64_t *data;
 } bitarray;
 
 static char buf[BUF_SIZE];
-static size_t bufsize = BUF_SIZE;
-static int offset = 0;
+static size_t bufsize = BUF_SIZE,
+              offset = 0;
 
 void usage(void) {
-    printf("Usage: %s [-h] [-n nlines] -s <parameter>... <file>...\n",
+    printf("Usage: %s [-h] [-n nlines] -s <regex> <file>...\n",
             program_invocation_short_name);
 }
 
 void help(void) {
     usage();
     printf("\n");
-    printf("Extracts given parameters from stan posterior csv file(s) based on strict matching.\n");
+    printf("Extracts given parameters from stan posterior csv file(s) based on regex expression.\n");
     printf("\n");
     printf("Options:\n");
     printf("\t-h Useless help message\n");
-    printf("\t-n Optional number of lines to read per file [Default: 1000]\n");
-    printf("\t-s Target parameters. Can be specified multiple times.\n");
[…]

### 2020-02-12 paper.Rmd
Switch post file paths to use new csv extracts

@@
 load("./posteriors/full_model/input.RData")
-fa <- read_post("./posteriors/full_model/fa_posteriors.csv.gz")
+fa <- read_post("./posteriors/full_model/fa_posteriors.csv")
 
@@
 ```{r fig.cap = "\\label{fig:fa-line} Posterior shrinkage for \\textit{v2juncind}"}
-err <- read_post("posteriors/full_model/err_posteriors.csv.gz")
+err <- read_post("posteriors/full_model/err_posteriors.csv")
 
@@
 ```{r reg, fig.height = 3}
-reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")
+reg <- read_post("posteriors/full_model/reg_posteriors.csv")
 
@@
 ```{r fig.height = 2, fig.cap = "\\label{fig:phat} Density plot for predicted probabilities"}
-p_hat <- extract(reg, "p_hat")
+extra <- read_post("posteriors/full_model/extra_posteriors.csv")
+p_hat <- extract(extra, "p_hat")
 
@@
     load(file.path("posteriors", f, "input.RData"))
-    beta <- file.path("posteriors", f, "reg_posteriors.csv.gz") %>%
+    beta <- file.path("posteriors", f, "reg_posteriors.csv") %>%
         read_post %>%

### 2020-02-12 R/thesis.utils/R/post.R
Check that all columns are numeric for posterior file

@@ -37,7 +37,7 @@ post_summarise.posterior <- function(x, pars = NULL, ...) {
 #'
 #' Reads a posterior file and creates a `posterior` object.
 #'
-#' @param file Target file. Can be a compress csv (`csv.gz`).
+#' @param file Target file. Can be a compressed csv (`csv.gz`).
 #'
 #' @return DataFrame with the additional class, `posterior`.
 #'
@@ -51,6 +51,9 @@ read_post <- function(file) {
             sub("[.]", ",", .) %>%
             paste0("]")
     }
+
+    if (any(sapply(fit, Negate(is.numeric))))
+        warning("Some columns were not numeric")
 
     structure(fit, class = c(class(fit), "posterior"))
 }

### 2020-02-12 scripts/run.sh
Update extract binary name in run script

@@ -49,5 +49,5 @@ if [ "${CLEANUP:-0}" = 1 ]; then
 fi
 
 $CMD run -it $mount_opts $rm_opts \
-     -e cmdstan=/cmdstan -e select=/root/bin/select \
+     -e cmdstan=/cmdstan -e extract=/root/bin/extract \
      "jsks/conflict_onset:$TAG" make -j4 $@

### 2020-02-24 scripts/run.sh
BUGFIX: pass arguments directly to container

@@ -3,37 +3,15 @@
 # Convenience script to launch project image. Note: CLI args besides
 # the listed options are passed directly to Make, so no usage
 # function.
+#
+# By default, the latest tag will be used unless the IMG_TAG env variable is
+# set.
+#
+# If the env variable CLEANUP is defined and set to 1, then the container will
+# be automatically cleaned up after exit.
 ###
 
 set -e
-
-help() {
-    cat <<EOF
-$(usage)
-
-Launches an attached instance of 'jsks/conflict_onset'. Any
-commandline arguments beside the listed options are passed directly to
-'make -j4' in the container.
-
-If the env variable CLEANUP is defined and set to 1, then the
-container will be automatically removed following completion.
-
-Options:
-        -t Specify container tag [Default: latest;]
-EOF
-}
-
-while getopts 'ht:' opt; do
-    case $opt in
-        h)
-            help;;
-        t)
-            TAG="$OPTARG";;
-    esac
-done
-
-shift $(( $OPTIND - 1 ))
-: ${TAG:="latest"}
 
 which podman 2>&1 >/dev/null && CMD=podman || CMD=docker
 
@@ -50,4 +28,4 @@ fi
 
 $CMD run -it $mount_opts $rm_opts \
      -e cmdstan=/cmdstan -e extract=/root/bin/extract \
-     "jsks/conflict_onset:$TAG" make -j4 $@
+     "jsks/conflict_onset:${IMG_TAG:-latest}" make -j4 $@

### 2020-02-24 paper.Rmd
Add GP explainer + further edits to research design

@@
 horizontal inequalities [@Cederman2011] --- to opportunity-based
-explanations centered on economic factors [@Collier2012; @Collier2009]
+explanations centered on economic factors [@Collier2004; @Collier2009]
 as well as the political and structural conditions underlying state
@@
 Africa and argues that executive elections, even when they are of poor
-quality, reduce the risk of large-scale violent dissent. @Fjelde2010a
+quality, reduce the risk of large-scale violent dissent. @Fjelde2010
 breaks down authoritarian regime types in Africa attributing the
@@
 [^model]: The model code is written in Stan [@Carpenter2017] and was
-    run with the `CmdStan` interface [@Team2019a] using four parallel
-    chains at 2000 iterations each with a warmup of 1000. Convergence
-    was assessed according to the potential scale reduction factor
-    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well
-    as standard Stan diagnostics. The full code is available in
-    [Appendix A](#appendix-a) and a fully reproducible pipeline
-    compromised of replication scripts is available at
-    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis). A
-    simulation analysis assessing the model and it's ability to
-    recover simulated true values is available in [Appendix B](#appendix-b).
+    run using the `CmdStan` interface version 2.22.1 [@Stan2020] with
+    four parallel chains at 2000 iterations each with a warmup
+    of 1000. Convergence was assessed according to the potential scale
+    reduction factor $\widehat{R}$ and effective sample size (ESS)
+    [@Vehtari2019] in addition to standard Stan diagnostics. All
+    pre/post-processing was done using R 3.6.2 [@Core2019]. The Stan
+    model code is available in [Appendix A](#appendix-a) and the full
+    project in the form of a reproducible pipeline, including the
+    Rmarkdown source for this manuscript, is available at
+    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
+    Finally, a simulation analysis assessing the model and it's
+    ability to recover simulated true values is available in [Appendix
+    B](#appendix-b).
 
@@
 a function of a set of predictors, including executive constraints and
-an array of control variables, together with several grouping factors
-included as partially pooled varying intercepts. The units of analysis
-are all independent country-years from 1951 until 2017, as determined
-by data availability, with *I* countries indexed by $i = 1 \dotsc I$
-and *T* years indexed by $t = 1 \dotsc T$. The dependent variable,
-conflict onset, is a Bernoulli distributed binary variable where 1
-denotes the onset of at least one new civil conflict in a given
-country-year. The chosen link function transforming the linear
-regression output into the probability of conflict onset, $\nu_{it}$,
-is the inverse logistic such that $\nu_{it} \in [0, 1]$. [^invlogit]
-
-[^invlogit]: Technically, the inverse logistic transforms values to
+an array of control variables, together with several random
+effects. The units of analysis are all independent country-years from
+1951 until 2017, as determined by data availability, with *I*
+countries indexed by $i = 1 \dotsc I$ and *T* years indexed by $t = 1
+\dotsc T$. The dependent variable, conflict onset, is a Bernoulli
+distributed binary variable where 1 denotes the onset of at least one
+new civil conflict in a given country-year. The chosen link function
+transforming the linear regression output into the probability of
+conflict onset, $\nu_{it}$, is the inverse logit such that $\nu_{it}
+\in [0, 1]$. [^invlogit]
+
+[^invlogit]: Technically, the inverse logit transforms values to
     the open interval, $(0, 1)$, but due to overflow/underflow in
@@
 
-where **X** is a vector of predictors, $\beta$ the vector of
-regression slopes, and $\zeta^k$ the set of varying intercepts for *K*
-grouping factors indexed by $j_k$. For the present analysis there are
-only two grouping factors, year and country, to account for
+where **X** is the predictors, $\beta$ the vector of regression
+slopes, and $\zeta^k$ the *K* sets of varying intercepts indexed by
+$j_k$. All predictors are lagged by one year to control for time
+endogeneity and standardized by subtracting the mean and dividing by
+the standard deviation for each variable to aid in computation as well
+as comparison of the posterior $\beta$ estimates.
+
+For the present analysis there are three sets of varying intercepts,
+or random effects, included in the model. This includes two partially
+pooled grouping factors, year and country, to account for
 cross-sectional unmodeled heterogeneity between countries and over
-time [@Shor2007]. All predictors are also lagged by one year to
-control for time endogeneity and standardized by substracting the mean
-and dividing by the standard deviation for each variable to aid in
-computation as well as comparison of the final $\beta$ estimates.
-
-I place weakly informative priors on all parameters to allow the
-posterior estimates to be largely determined by the observed data
-while avoiding the more pathological values and hard constraints that
-can arise from uninformative priors. For the slopes and population
-intercept, given that the input data has been standardized, this
+time [@Shor2007], and the number of peace years since the last
+civil conflict incidence per country to control for the lingering
+endogenous effects of violence on the likelihood that additional
+conflict breaks
+out.
+
+I place weakly informative priors on the population intercept and
+slope parameters to allow the posterior estimates to be largely
+determined by the observed data while avoiding the more pathological
+values and hard constraints that can arise from uninformative
+priors. Given that the input data has been standardized, this
 translates to:
@@
 
-For the varying intercepts modeling group-specific variability, I use
-hierarchical priors to provide a regularizing effect against
-overfitting [@Gelman2007].
+For the varying intercepts modeling year and country specific
+variability, I use hierarchical priors to provide a regularizing
+effect against overfitting [@Gelman2007].
 \begin{align*}
-    \zeta^k_j & \sim \mathsf{Normal}(0, \sigma^k) \\
-    \sigma^k & \sim \mathsf{HalfCauchy}(0, 1)
+    \zeta^{year}_t & \sim \mathsf{Normal}(0, \sigma^{year}) \\
+    \zeta^{country}_i & \sim \mathsf{Normal}(0, \sigma^{country}) \\
+    \sigma^{year}, \sigma^{country} & \sim \mathsf{Half-Cauchy}(0, 1)
+\end{align*}
+
+Finally, for peace years I take a non-parametric approach and use
+gaussian process priors with an exponentiated quadratic kernel. This
+treats the number of peace years as an infinitely continuous dimension
+along which the covariance between any two given points is a function
+of their euclidean distance. Where that distance is smaller there is a
+higher correlation. Compared to other approaches to modeling time
+dependence this avoids the parametric assumption of a proportional
+rate in exponential decay functions, the difficulty in selecting
+interaction terms to account for non-proportionality for cubic
+polynomial approximation, and the somewhat arbitrary selection of
+knots for splines [*c.f.* @Raknerud1997; @Carter2010;
+@Beck1998]. Given the finite set of observed peace years in the
+dataset, the specification is as follows:
+\begin{align*}
+    \zeta^{peace\_years} & \sim \mathsf{MultiNormal}(0, K) \\
+    K_{ij} & = \eta^2 \mathsf{exp}(\minus \frac{1}{2 \rho^2} D_{ij}^2)
+\end{align*}
+
+where for $P$ unique peace years, $K \in \mathbb{R}^{P \times P}$ is
+the covariance matrix, $D \in \mathbb{R}^{P \times P}$ the distance
+matrix, and $i$ and $j$ the column and row indexes.[^positive] The
+hyperparameter, $\rho$, is the *length-scale* and corresponds roughly
+to the number of years before two peace years become uncorrelated. The
+hyperparameter, $\eta$, is the *marginal standard deviation* which
+determines the overall variation of the output intercepts
+[@Rasmussen2005].
+
+[^positive]: To ensure that the covariance matrix, $K$, is positive
+    definite a small positive term is added to the diagonal.
+
+I select a weakly informative prior for $\eta$ that is skeptical of
+large values given that $\nu$ is on the log-odds scale. For $\rho$ I
+use an informative prior that constrains an overwhelming majority of
+the prior probability mass to a *length-scale* between 2 and 12, which
+provides a strong smoothing effect while aiding in computation by
+avoiding unreasonable values [@Betancourt2017].
+\begin{align*}
+    \eta & \sim \mathsf{Half-Normal}(0, 1) \\
+    \rho & \sim \mathsf{InvGamma}(7.30124, 30.0086)
 \end{align*}
@@
 a linear function of the latent factor, $\theta$, transformed by the
-inverse logistic such that $\pi_i \in [0, 1]$. The practical effect is
+inverse logit such that $\pi_i \in [0, 1]$. The practical effect is
 that where there is data for the legislative manifest variables, the
@@
     governance indicator on government effectiveness extends back only
-    until 1996 (TODO: cite). Given that measures of institutional
+    until 1996 [@Kaufmann2010]. Given that measures of institutional
     quality often change fairly slowly, punctuated only by abrupt
@@
 terms of being able to offer economic selective benefits, I include
-GDP per capita. Specifically, I use the natural log of the $RGDP^o$
-variable, constant price real GDP per capita, from Penn World Tables
-(PWT) v9.1 (1950-2017) to allow for both cross-country and
-cross-temporal comparisons of GDP per capita [@Feenstra2015]. The
-justification for taking the natural log is that the marginal effect
-of the variable should decrease with higher values. In other words,
-the difference between the effects of relatively high levels of GDP
-per capita should be significantly smaller than the difference between
-the effects of low levels of GDP per capita. Finally, I also include
-an interaction term between GDP per capita and the latent measure for
-executive constraints to capture the potential conditional
-relationship.
+GDP per capita. Specifically, I use the natural log of the $RGDP^e$
+variable, constant price real GDP, from Penn World Tables (PWT) v9.1
+(1950-2017) to allow for both cross-country and cross-temporal
+comparisons of GDP per capita [@Feenstra2015].[^gdp] The justification
+for taking the natural log is that the marginal effect of the variable
+should decrease with higher values. In other words, the difference
+between the effects of relatively high levels of GDP per capita should
+be significantly smaller than the difference between the effects of
+low levels of GDP per capita. Finally, I also include an interaction
+term between GDP per capita and the latent measure for executive
+constraints to capture the potential conditional relationship.
+
+[^gdp]: Prior to taking the natural log, the GDP measure is of course
+    transformed to GDP per capita by dividing by the total population,
+    which is also available in PWT.  Note, PWT has two real GDP
+    measures available: $RGDP^o$, or output side real GDP, and
+    $RGDP^e$, or expenditure side real GDP. The former roughly
+    captures a country's productive capacity, while the latter is
+    better suited to measuring living standards [@Feenstra2015]. I use
+    $RGDP^e$ since I would argue that it serves as a better proxy for
+    the potential labour costs for rebel leaders.
 
@@
 [@Bazzi2014]. I also include a measure of mean elevation from the
-GROWup dataset [@Girardin2015] to capture terrain advantageous for
-conducting an insurgency [@Fearon2003] as well as a measure of
-population density derived from population data from PWT and the area
-in square kilometers for each country from GROWup. Countries with a
-higher population density may require a more extensive and competent
-central administration and police force in order to suppress violent
-dissent [@Fearon2003]. There's also the possibility that a high
-population density increases the competition for resources. If the
-state is identified as the source of a particular scarcity, the risk
-of civil conflict may increase. Similar to GDP per capita, I also
-transform mean elevation and population by the natural log with the
-expectation that the scale increase should diminish as values get
-larger. Finally, I include measures of ethnic fractionalization in the
-form of the number of political relevant ethnic groups and the percent
-of the population facing discrimination. Where there are significant
-ethnic cleavages and state sponsored discrimination, ethnic ties can
-facilitate collective action thereby lowering the mobilization costs
-of rebel groups [@Cederman2011].
-
-To control for the time endogeneity of conflict --- past and current
-civil conflicts can impact the present risk of conflict onset as well
-as the current level of executive constraints --- I include a dummy
-variable for ongoing conflict [^time] and the natural log of peace
-years [^peace_yrs] [@Hegre2015a]. To account for spillover effects, I
-include a dummy variable for whether a neighbouring country is
-currently experiencing a civil conflict. One notable lack is a general
-measure of political instability, which I omit since it could be
-highly correlated with executive constraints and would potentially
-bias results, at least for years of crisis where the risk of civil
-conflict is highest. I do, however, include a dummy variable for the
-independence year since newly sovereign states may be particularly
-weak.
+CShapes dataset [@Weidmann2010] as made available through the GROWup
+data platform [@Girardin2015] to capture terrain advantageous for
+conducting an insurgency [@Fearon2003]. In addition, I employ a
+measure of population density derived from population data from PWT
+and the area in square kilometers for each country from
+CShapes. Countries with a higher population density may require a more
+extensive and competent central administration and police force in
+order to suppress violent dissent [@Fearon2003]. There's also the
+possibility that a high population density increases the competition
+for resources. If the state is identified as the source of a
+particular scarcity, the risk of civil conflict may increase. Similar
+to GDP per capita, I also transform mean elevation and population
+density by the natural log with the expectation that the scale
+increase should diminish as values get larger. Finally, I include
+measures of ethnic fractionalization in the form of the number of
+political relevant ethnic groups and the percent of the population
+facing discrimination as taken from the Ethnic Power Relations dataset
+[@Vogt2015]. Where there are significant ethnic cleavages and state
+sponsored discrimination, ethnic ties can facilitate collective action
+thereby lowering the mobilization costs of rebel groups
+[@Cederman2011].
+
+As mentioned previously, to control for time dependence --- past and
+current civil conflicts can impact the present risk of conflict onset
+as well as the current level of executive constraints -- I include the
+number of peace years since the last conflict incident.[^time] To
+account for spillover effects, I include a dummy variable for whether
+a neighbouring country is currently experiencing a civil
+conflict.[^geo] Finally, one notable lack is a general measure of
+political instability, which I omit since it could be highly
+correlated with executive constraints and would potentially bias
+results, at least for years of crisis where the risk of civil conflict
+is highest. I do, however, include a dummy variable for the
+independence year since newly sovereign states may be especially weak.
+
+[^geo]: Country neighbours for each year were determined according to
+    CShape. Pre-independence conflicts were considered as neighbouring
+    to the countries bordering the territory location of the conflict,
+    not the government location.
 
@@
     potential confounder since it can influence both the probability
-    of the outcome as well as the executive constraints
-    measure.
-
-[^peace_yrs]: The peace years variable is calculated as the number of
-    years since the last conflict-year, including conflict-years from
-    extrasystemic conflicts. For countries left censored due to the
-    dataset starting date, I set the start of the peace year
+    of the outcome as well as the executive constraints measure. For
+    this purpose, the peace years variable is calculated as the number
+    of years since the last conflict-year, including conflict-years
+    from extrasystemic conflicts. For countries left censored due to
+    the dataset starting date, I set the start of the peace year
     calculation to 1946, the end of World War II. The rationale is
@@
 
-The $\delta$ and $\eta$ parameters correspond to the slope and
+The $\delta$ and $\kappa$ parameters correspond to the slope and
 intercept in the linear predictor for the presence of a legislature,

### 2020-02-27 paper.Rmd
Move sim summary to rmarkdown and remove old R scripts

@@
 ```{r eval = F, warning = F}
-load("posteriors/sim/simulated_summary.RData")
+true_values <- readRDS("posteriors/sim/simulated_parameters.rds")
+
+fa <- read_post("posteriors/sim/fa_posteriors.csv")
+reg <- read_post("posteriors/sim/reg_posteriors.csv")
+
+fit <- bind_cols(fa, reg)
+
+post <- post_summarise(fit, c("lambda", "psi", "gamma", "delta",
+                              "kappa", "rho", "eta", "alpha", "beta",
+                              "sigma")) %>%
+    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
+    mutate(type = "Posterior Estimate")
+
+pars.df <- filter(true_values, !grepl("theta", parameter)) %>%
+    bind_rows(post) %>%
+    mutate(type = as.factor(type))
 
@@
 ```{r eval = F, warning = F}
-latent.df <- arrange(latent.df, true_value) %>%
-    mutate(idx = 1:n())
-
-ggplot(latent.df, aes(idx, true_value)) +
-    geom_point(col = "darkgrey") +
-    geom_errorbar(aes(ymin = codelow, ymax = codehigh), alpha = 0.2, col = "darkblue") +
-    ylab("True Value") +
-    theme(axis.text.x = element_blank(),
-          axis.title.x = element_blank(),
-          axis.ticks.x = element_blank())
-
-```
-
-```{r eval = F, results = 'hide'}
-x <- with(latent.df, mean(true_value <= codehigh & true_value >= codelow))
-#x * 100
-
-#plot(df$lg_se[, 1], df$lg[, 1] - colMeans(lg_est))
-#plot(df$lg_se[, 1], lapply(lg_est, IQR) %>% unlist)
-
-#which(df$lg[, 1] + 2 * df$lg_se[, 1] < colMeans(lg_est) |
-#      df$lg[, 1] - 2 * df$lg_se[, 1] > colMeans(lg_est))
-
-#v <- lapply(theta_old, IQR) %>% unlist
-#plot(df$lg_se[, 1], v[df$obs_idx])
-
-#plot(density(p), xlim = c(0, 1), ylim = c(0, 2))
-#for (i in 1:100) lines(density(p_hat[i, ]), col = alpha("lightblue", 0.2))
-
-#table(colMeans(p_hat) > .5, y)
+latent.df <- post_summarise(fit, pars = "theta", probs = c(0.025, 0.975)) %>%
+    rename(codelow = `2.5%`, codehigh = `97.5`)
+
+x <- mean(theta <= latent.df$codehigh & theta >= latent.df$codelow)
 

### 2020-02-27 R/sim_data.R

@@ -20,7 +20,7 @@ lambda <- rlnorm(D, 0, 0.5)
 psi <- rweibull(D, 5, 1)
 
 # Legislative variables
-kappa <- rnorm(1)
+kappa <- rnorm(1, 0, 2.5)
 delta <- rlnorm(1, 0, 0.5)
 
 xi <- kappa + delta * theta
@@ -51,34 +51,55 @@ for (i in 1:ncol(nonlg))
     nonlg_obs[, i] <- rnorm(N, nonlg[, i], nonlg_err[, i])
 
 # Conflict regression
+rho <- rinvgamma(1, 7.30124, 30.0086)
+eta <- rhnorm(1, 0.5)
+
+D <- dist(1:N, diag = T, upper = T) %>% as.matrix
+
+K <- eta^2 * exp((-0.5 / (2 * rho^2)) * D^2)
+f <- mvrnorm(1, rep(0, N), K)
+
 alpha <- rnorm(1, 0, 5)
-beta <- rnorm(4, 0, 2.5)
+beta <- rnorm(3, 0, 2.5)
 
-X <- mvrnorm(N, c(0, 2), matrix(c(2, .5, .5, 3), 2, 2))
-X <- data.frame(a = rnorm(N, 0, 2), b = rnorm(N, 0, 3))
-interaction_idx  <- 2L
+X <- rnorm(N, 2, 2)
 
-n_years <- N %/% 10L
-n_countries <- 10L
+n_years <- N %/% 5L
+n_countries <- 5L
 
 sigma <- rhcauchy(2, 1)
 countries <- rnorm(n_countries, 0, sigma[1])
 years <- rnorm(n_years, 0, sigma[2])
 
-cy <- expand.grid(1:n_countries, 1:n_years)
-cy <- cy[sample(nrow(cy)), ]
-
+cy <- expand.grid(1:n_years, 1:n_countries)
 stopifnot(nrow(cy) == N)
 
-country_idx <- cy[, 1]
-year_idx <- cy[, 2]
+dataset <- setNames(cy, c("year", "country")) %>%
+    mutate(x = X,
+           exec_idx = 1:N,
+           peace_yrs = 1,
+           y = 0)
 
-p <- inv.logit(alpha + beta[1] * theta + beta[2] * X[, 1] +
-               beta[3] * X[, 2] + beta[4] * theta * X[, interaction_idx] +
-               years[year_idx] + countries[country_idx])
+country <- 0
+for (i in 1:nrow(dataset)) {
[…]

### 2020-02-27 R/summarise.R

@@ -1,22 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(ggplot2)
-library(rstan)
-
-load("data/prepped_data.RData")
-fit <- readRDS("posteriors/fit.rds")
-
-check_hmc_diagnostics(fit)
-
-theta <- as.matrix(fit, pars = "theta") %>%
-    apply(2, quantile, probs = c(0.25, 0.5, 0.75)) %>%
-    t %>%
-    as.data.frame %>%
-    rename(theta_codelow = `25%`,
-           theta = `50%`,
-           theta_codehigh = `75%`) %>%
-    mutate(country_name = final.df$country_name,
-           year = final.df$year)
-
-saveRDS(theta, "data/summarised_post.rds")

### 2020-02-27 R/summarise_sim.R

@@ -1,35 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(data.table)
-library(magrittr)
-library(thesis.utils)
-
-fit <- read_post("posteriors/sim/combined_posteriors.csv.gz")
-true_values <- readRDS("posteriors/sim/simulated_parameters.rds")
-
-###
-# Posterior predicted probabilities
-p_hat <- extract(fit, pars = "p_hat")
-
-###
-# Model parameters
-print("Summarizing model parameters")
-post <- post_summarise(fit, c("lambda", "psi", "gamma", "delta",
-                              "eta", "alpha", "beta", "sigma")) %>%
-    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
-    mutate(type = "Posterior Estimate")
-
-pars.df <- filter(true_values, !grepl("theta", parameter)) %>%
-    bind_rows(post) %>%
-    mutate(type = as.factor(type))
-
-###
-# Latent factor
-print("Summarizing theta")
-latent.df <- post_summarise(fit, pars = "theta") %>%
-    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`)
-
-latent.df$true_value <- filter(true_values, grepl("theta", parameter)) %$% point
-
-save(p_hat, pars.df, latent.df, file = "posteriors/sim/simulated_summary.RData")

### 2020-04-05 paper.Rmd
Update wordcount rule to include expanded refs and abstract

@@
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
+suppress-bibliography: false
 bibliography: library.bib

### 2020-04-05 R/thesis.utils/R/plot.R
BUGFIX: display horizontal line at zero for plot_pars

@@ -28,7 +28,8 @@ plot_pars.matrix <- function(x, ylab = "Parameter Estimates", hline = F) {
         ylab(ylab)
 
     if (isTRUE(hline))
-        p + geom_hline(yintercept = 0, alpha = 0.5, color = "grey")
+        p <- p + geom_hline(yintercept = 0, alpha = 0.5,
+                            color = "grey", linetype = "dotted")
 
     p
 }

### 2020-04-05 R/thesis.utils/R/plot.R
Display format for plot_pars and tables
- Preserve input ordering for plotted parameters.
- Pretty format numbers in summary tables

@@ -14,7 +14,8 @@ plot_pars.matrix <- function(x, ylab = "Parameter Estimates", hline = F) {
     df <- post_summarise(x, probs = c(0.10, 0.5, 0.9)) %>%
         rename(codelow = `10%`,
                median = `50%`,
-               codehigh = `90%`)
+               codehigh = `90%`) %>%
+        mutate(parameter = factor(parameter, levels = parameter))
 
     p <- ggplot(df, aes_(x = ~parameter, y = ~median)) +
         geom_errorbar(aes_(x = ~parameter, ymin = ~codelow, ymax = ~codehigh),

### 2020-04-05 R/thesis.utils/R/utils.R

@@ -31,6 +31,8 @@ calc_peace_yrs <- function(years, incidence) {
         stop("Unsorted years vector", call. = F)
     else if (any(consecutive(years) != 1))
         stop("Breaks found in years vector", call. = F)
+    else if (dplyr::n_distinct(incidence) > 2)
+        stop("Expected a binary variable for incidence", call. = F)
 
     locf_idx <- function(x) {
         x[x == 1] <- which(x == 1)
@@ -217,6 +219,8 @@ summary_stats.data.frame <- function(x, vars = colnames(x)) {
     if (is.null(vars))
         stop("Missing target columns")
 
+    fmt <- . %>% signif(3) %>% prettyNum(big.mark = ",")
+
     df <- x[, vars, drop = F]
     # N, mean, std.dev, min, max
     ll <- lapply(colnames(df), function(s) {
@@ -226,11 +230,11 @@ summary_stats.data.frame <- function(x, vars = colnames(x)) {
             return(NULL)
 
         data.frame(Variable = s,
-                   N = sum(!is.na(col)),
-                   Mean = mean(col, na.rm = T) %>% signif(3),
-                   `Std. dev.` = sd(col, na.rm = T) %>% signif(3),
-                   `Min.` = min(col, na.rm = T) %>% signif(3),
-                   `Max.` = max(col, na.rm = T) %>% signif(3),
+                   N = sum(!is.na(col)) %>% prettyNum(big.mark = ","),
+                   Mean = mean(col, na.rm = T) %>% fmt,
+                   `Std.dev.` = sd(col, na.rm = T) %>% fmt,
+                   `Min.` = min(col, na.rm = T) %>% fmt,
+                   `Max.` = max(col, na.rm = T) %>% fmt,
                    stringsAsFactors = F)
     })
 

### 2020-04-05 R/thesis.utils/tests/testthat/test_utils.R
Add additional unit tests

@@ -24,6 +24,12 @@ test_that("calc_peace_yrs", {
     expect_identical(calc_peace_yrs(years, ongoing),
                      c(0, 1:4, 0, 1:4))
 
+    years <- 1900:1906
+    ongoing <- c(0, 1, 1, 1, 0, 0, 1)
+
+    expect_identical(calc_peace_yrs(years, ongoing),
+                     c(0, 1, 0, 0, 0, 1, 2))
+
     years <- c(1900, 1902:1904)
     ongoing <- rep(0, 4)
 
@@ -35,6 +41,12 @@ test_that("calc_peace_yrs", {
 
     expect_error(calc_peace_yrs(years, ongoing),
                  "Unsorted years vector")
+
+    years <- 1900:1904
+    ongoing <- c(0, 1, 2, 0, 1)
+
+    expect_error(calc_peace_yrs(years, ongoing),
+                 "Expected a binary variable for incidence")
 })
 
 
@@ -127,11 +139,11 @@ test_that("partial", {
 test_that("summary_stats", {
     df <- data.frame(x = letters[1:10], y = c(1:9, NA))
     out <- data.frame(Variable = "y",
-                      N = 9,
-                      Mean = 5,
-                      `Std. dev.` = sd(1:9) %>% signif(3),
-                      `Min.` = 1,
-                      `Max.` = 9,
+                      N = "9",
+                      Mean = "5",
+                      `Std.dev.` = sd(1:9) %>% signif(3) %>% as.character,
+                      `Min.` = "1",
+                      `Max.` = "9",
                       stringsAsFactors = F)
 
     expect_equal(summary_stats(df), out)

### 2020-04-06 paper.Rmd
Draft edits

@@
 regression relating the latent construct to the observed outcome,
-civil conflict. The use of bayesian methods is particularly
-advantageous in this setting since the constructed dataset is
-comprised of the entire finite population of countries within a
-non-random subset of years given only non-stochastic
-missingness. Traditional frequentist estimators are inappropriate due
-to the assumption that the observed data is drawn from a random,
-repeatable generative mechanism. This is readily apparent when
-attempting to interpret a frequentist confidence interval in the present
-context. A standard confidence interval for a selected level of
-significance, $\alpha$, reports only that $(1 - \alpha) 100 \%$ of the
-calculated confidence intervals from additional inferences done on
-repeated samples are expected to contain the true population value
-[@Western1994]. However, for the current analysis the desired target
-is really to be able to calculate directly the probability that the
-current uncertainty interval contains the true value given the data
-and the model specification. Bayesianism offers a solution by casting
-probability as degrees of belief in some event wherein inference is
-then the process of updating those prior beliefs in light of new
-data. When using bayesian methods, parameters such as regression
-coefficients are considered to be random, meaning that prior beliefs
-in the form of probability distributions can be assigned and posterior
-probability distributions estimated from observed data [@Western1999].
+civil conflict.
+
+The use of bayesian methods is particularly advantageous in this
+setting since the constructed dataset is comprised of the entire
+finite population of countries within a non-random subset of years
+given only non-stochastic missingness. Traditional frequentist
+estimators are inappropriate due to the assumption that the observed
+data is drawn from a random, repeatable generative mechanism. This is
+readily apparent when attempting to interpret a frequentist confidence
+interval in the present context. A standard confidence interval for a
+selected level of significance, $\alpha$, reports only that $(1 -
+\alpha) 100 \%$ of the calculated confidence intervals from additional
+inferences done on repeated samples are expected to contain the true
+population value [@Western1994]. However, for the current analysis the
+desired target is really to be able to calculate directly the
+probability that the current uncertainty interval contains the true
+value given the data and the model specification. Bayesianism offers a
+solution by casting probability as degrees of belief in some event
+wherein inference is then the process of updating those prior beliefs
+in light of new data. When using bayesian methods, parameters such as
+regression coefficients are considered to be random, meaning that
+prior beliefs in the form of probability distributions can be assigned
+and posterior probability distributions estimated from observed data
+[@Western1999].
 
@@
 effects. The units of analysis are all independent country-years from
-1951 until 2017, as determined by data availability, with *I*
+1952 until 2013, as determined by data availability, with *I*
 countries indexed by $i = 1 \dotsc I$ and *T* years indexed by $t = 1
@@
 where **X** is the predictors, $\beta$ the vector of regression
-slopes, and $\zeta^k$ the *K* sets of varying intercepts indexed by
+slopes, and $\Zeta$ the *K* sets of varying intercepts indexed by
 $j_k$. All predictors are lagged by one year to control for time
 endogeneity and standardized by subtracting the mean and dividing by
-the standard deviation for each variable to aid in computation as well
-as comparison of the posterior $\beta$ estimates.
+the standard deviation to aid in computation as well as comparison of
+the posterior $\beta$ estimates.
 
@@
 knots for splines [*c.f.* @Raknerud1997; @Carter2010;
-@Beck1998]. Given the finite set of observed peace years in the
-dataset, the specification is as follows:
+@Beck1998].
+
+Given the finite set of observed peace years in the dataset, the
+specification is as follows:
 \begin{align*}
@@
 the covariance matrix, $D \in \mathbb{R}^{P \times P}$ the distance
-matrix, and $i$ and $j$ the column and row indexes.[^positive] The
+matrix, and $i$ and $j$ the column and row indexes. The
 hyperparameter, $\rho$, is the *length-scale* and corresponds roughly
-to the number of years before two peace years become uncorrelated. The
-hyperparameter, $\eta$, is the *marginal standard deviation* which
-determines the overall variation of the output intercepts
-[@Rasmussen2005].
-
-[^positive]: To ensure that the covariance matrix, $K$, is positive
-    definite a small positive term is added to the diagonal.
+to the number of years before two observations become
+uncorrelated. The hyperparameter, $\eta$, is the *marginal standard
+deviation* which determines the overall variation of the output
+intercepts [@Rasmussen2005].
 
@@
 use an informative prior that constrains an overwhelming majority of
-the prior probability mass to a *length-scale* between 2 and 12, which
-provides a strong smoothing effect while aiding in computation by
+the prior probability mass to a *length-scale* between 4 and 10, which
+provides a strong smoothing effect and aids in computation by
 avoiding unreasonable values [@Betancourt2017].
 \begin{align*}
-    \eta & \sim \mathsf{Half-Normal}(0, 1) \\
-    \rho & \sim \mathsf{InvGamma}(7.30124, 30.0086)
+    \eta & \sim \mathsf{HalfNormal}(0, 1) \\
+    \rho & \sim \mathsf{InvGamma}(15.5031, 78.2992)
 \end{align*}
@@
 UCDP/PRIO Armed Conflict Dataset v19.1 [@Gleditsch2002;
-@Pettersson2019] where intra-state conflict is defined as a violent
-contest over a government or territory between a state and non-state
-actor with a minimum threshold of 25 battle deaths within a single
-country-year with a coverage from 1946 to 2018.  The data was merged
-based on the government side for conflicts labeled as internal or
-internationalized internal --- *i.e.* where *type_of_conflict* was
-either 3 or 4. Multiple conflicts within the same years were
-aggregated to form a binary variable coded as 1 to denote the onset of
-at least one new civil conflict and 0 for all other years, including
-years of ongoing conflict.[^conflict_type]
+@Pettersson2019]. According to the dataset, intra-state conflict is
+defined as a violent contest over a government or territory between a
+state and non-state actor with a minimum threshold of 25 battle deaths
+within a single country-year with a coverage from 1946 to 2018.  The
+data was merged based on the government side for conflicts labeled as
+internal or internationalized internal --- *i.e.* where
+*type_of_conflict* was either 3 or 4. Multiple conflicts within the
+same years were aggregated to form a binary variable coded as 1 to
+denote the onset of at least one new civil conflict and 0 for all
+other years, including years of ongoing conflict.[^conflict_type]
 
 [^conflict_type]: Extrasystemic conflicts, *i.e.* imperial or colonial
-    conflicts, were excluded largely due to lack of data. Ideally,
-    such conflicts would be merged based on the territory location of
-    the conflict, however, for colonies, for example, there is very
-    little easily available economic data and it would be undesirable
-    to use the data for the colonizing power due to the often wide
-    discrepancies in development within colonial empires. TODO: maybe
-    a blurb on theory?
+    conflicts, were excluded largely due to lack of data. For example,
+    there is very little easily available economic data for colonial
+    territories and it would be undesirable to use the data for the
+    colonizing power due to the often wide discrepancies in
+    development within colonial empires.
 
@@
 
-```{r fig.cap = "\\label{fig:onsets} Proportion of countries with ongoing civil conflicts and new onsets"}
+```{r fig.cap = "\\label{fig:onsets} Proportion of countries experiencing ongoing civil conflicts and new onsets"}
 
@@
     theme(legend.title = element_blank()) +
-    ylab("Percentage of Countries") +
+    ylab("Proportion of Countries") +
     xlab("Year")
@@
 executive and the highest category, 4, denotes that it is certain or
-nearly certain. As such, on the aggregated, unbounded scale, higher
-values in the point estimates denote country-years where the
-legislature is more likely to investigate. Finally, figure
-\ref{fig:cor} also provides a correlation matrix demonstrating a
-fairly high level of correlation across all variables indicating some
-commonality in the covariation.
+nearly certain. On the aggregated, unbounded scale, after running the
+IRT model, this translates to higher values for the point estimates
+denoting country-years where the legislature is more likely to
+investigate.  Finally, figure \ref{fig:cor} also provides a
+correlation matrix demonstrating a fairly high level of correlation
+across all variables indicating some commonality in the covariation.
 
@@
 \mathbb{R}^{m \times n}$ be the matrix of manifest variables measured
-with an observed degree of error, $Z^\ast \in \mathbb{R}^{m \times n}$
-the matrix of unobserved true values, and $\theta$ the unidimensional
-latent factor of length $m$ measuring executive constraints.
+with an observed degree of error, $U \in \mathbb{R}^{m \times n}$, and
+$Z^\ast \in \mathbb{R}^{m \times n}$ the matrix of unobserved true
+values. Let $\theta$ then be the unidimensional latent factor of
+length $m$ measuring executive constraints.
 
@@
 where the location parameters are the unobserved true values and the
-scale, $u$, the observed posterior standard deviations derived from
-the IRT posterior distributions. [^err]
-\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, u_{ij}) \]
-
-[^err]: An ideal approach would be to directly embed the individual
+scale the observed posterior standard deviations derived from the IRT
+posterior distributions. [^err]
+\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, U_{ij}) \]
+
+[^err]: The ideal approach would be to directly embed the individual
     IRT models aggregating the manifest variables into the larger
@@
 the manifest variables exhibit the same scale direction and should be
-positively correlated with the latent factor. Finally, similar to the
-regression coefficients, I select the following weakly informative
-priors:
+positively correlated with the latent factor. This means that larger
+values in the point estimates for $\theta$ represent a more
+constrained executive.
+
+Similar to the regression coefficients, I select the following weakly
+informative priors:
 \begin{align*}
@@
 
-Finally, in the case of the legislative variables there is the added
+In the case of the legislative variables there is the added
 issue of missing observations for country-years lacking a
@@
 [^missing]: Missingness due to the lack of of a particular
-    institution, for example the suspension of high court, is an issue
-    for the other manifest variables as well. However, only the
+    institution, for example the suspension of a high court, is an
+    issue for the other manifest variables as well. However, only the
     existence of the legislature is coded by V-Dem as a separate
@@
 term between GDP per capita and the latent measure for executive
-constraints to capture the potential conditional relationship.
+constraints to capture the potential conditional relationship (TODO:
+expand this sentence).
 
@@
     $RGDP^e$ since I would argue that it serves as a better proxy for
-    the potential labour costs for rebel leaders.
-
-In addition to the previous indicators, I also include a number of
-standard covariates drawn largely from previous literature. The list
-of indicators are not meant to provide an all encompassing theory of
-civil conflict, but rather roughly capture the opportunity structures
-for rebellion.  This includes a single year lagged measure of GDP per
-capita growth from PWT in order to capture short term economic shocks,
-which may affect both the strength and legitimacy of the state
-[@Bazzi2014]. I also include a measure of mean elevation from the
-CShapes dataset [@Weidmann2010] as made available through the GROWup
-data platform [@Girardin2015] to capture terrain advantageous for
-conducting an insurgency [@Fearon2003]. In addition, I employ a
-measure of population density derived from population data from PWT
-and the area in square kilometers for each country from
-CShapes. Countries with a higher population density may require a more
-extensive and competent central administration and police force in
-order to suppress violent dissent [@Fearon2003]. There's also the
-possibility that a high population density increases the competition
-for resources. If the state is identified as the source of a
-particular scarcity, the risk of civil conflict may increase. Similar
-to GDP per capita, I also transform mean elevation and population
-density by the natural log with the expectation that the scale
-increase should diminish as values get larger. Finally, I include
-measures of ethnic fractionalization in the form of the number of
-political relevant ethnic groups and the percent of the population
-facing discrimination as taken from the Ethnic Power Relations dataset
-[@Vogt2015]. Where there are significant ethnic cleavages and state
-sponsored discrimination, ethnic ties can facilitate collective action
-thereby lowering the mobilization costs of rebel groups
-[@Cederman2011].
+    the potential recruitment costs for rebel leaders.
+
+In addition, I also include a number of standard covariates drawn
+largely from previous literature. The list of indicators are not meant
+to provide an all encompassing theory of civil conflict, but rather
+roughly capture the opportunity structures for rebellion.  This
+includes a single year lagged measure of GDP per capita growth from
+PWT in order to capture short term economic shocks, which may affect
+both the strength and legitimacy of the state [@Bazzi2014]. I also
+include a measure of mean elevation from the CShapes dataset
+[@Weidmann2010] as made available through the GROWup data platform
+[@Girardin2015] to capture terrain advantageous for conducting an
+insurgency [@Fearon2003]. In addition, I employ a measure of
+population density derived from population data from PWT and the area
+in square kilometers for each country from CShapes. Countries with a
+higher population density may require a more extensive and competent
+central administration and police force in order to suppress violent
+dissent [@Fearon2003]. There's also the possibility that a high
+population density increases the competition for resources. If the
+state is identified as the source of a particular scarcity, the risk
+of civil conflict may increase. Similar to GDP per capita, I also
+transform mean elevation and population density by the natural log
+with the expectation that the scale increase should diminish as values
+get larger. Finally, I include measures of ethnic fractionalization in
+the form of the number of politically relevant ethnic groups and the
+percent of the population facing discrimination as taken from the
+Ethnic Power Relations dataset [@Vogt2015].[^epr] Where there are
+significant ethnic cleavages and state sponsored discrimination,
+ethnic ties can facilitate collective action thereby lowering the
+mobilization costs of rebel groups [@Cederman2011].
+
+[^epr]: In this context, an ethnic group is considered politically
+    relevant if "either at least one significant political actor
+    claims to represent the interests of that group in the national
+    political arena or if group members are systematically and
+    intentionally discriminated against in the domain of public
+    politics" while discrimination denotes the formal political
+    exclusion of a specific ethnic group with citizen rights
+    [@EPR2019].
 
@@
 as well as the current level of executive constraints -- I include the
-number of peace years since the last conflict incident.[^time] To
-account for spillover effects, I include a dummy variable for whether
-a neighbouring country is currently experiencing a civil
-conflict.[^geo] Finally, one notable lack is a general measure of
-political instability, which I omit since it could be highly
-correlated with executive constraints and would potentially bias
-results, at least for years of crisis where the risk of civil conflict
-is highest. I do, however, include a dummy variable for the
-independence year since newly sovereign states may be especially weak.
-
-[^geo]: Country neighbours for each year were determined according to
-    CShape. Pre-independence conflicts were considered as neighbouring
-    to the countries bordering the territory location of the conflict,
-    not the government location.
+number of peace years since the last conflict incident.[^time] I also
+include a dummy variably for ongoing high intensity conflict since
+devastating civil ways may have a particularly destabilizing
+effect.[^intensity] To account for spillover effects, I include a
+dummy variable for whether a neighbouring country is currently
+experiencing a civil conflict.[^geo] Finally, one notable lack is a
+general measure of political instability, which I omit since it could
+be highly correlated with executive constraints and would potentially
+bias results, at least for years of crisis where the risk of civil
+conflict is highest. I do, however, include a dummy variable for the
+independence year since newly sovereign states may be especially
+vulnerable.
 
@@
 
+[^intensity]: High intensity conflict meaning conflict-years with at
+    least 1000 battle deaths.
+
+[^geo]: Country neighbours for each year were determined according to
+    CShape. Pre-independence conflicts were considered as neighbouring
+    to the countries bordering the territory location of the conflict,
+    not the government location.
+
 Table \ref{tab:reg-stats} lists the full descriptive statistics for
@@
 df <- merged.df %>%
-    select(rgdpepc, rgdpepc_gro, pop_density, meanelev, discrimpop,
-           rlvt_groups_count, neighbour_conflict, ongoing, peace_yrs,
-           independence)
+    select(rgdpepc, rgdpepc_gro, cinc, discrimpop, pop_density,
+           meanelev, rlvt_groups_count, neighbour_conflict,
+           major_ongoing, independence)
 
@@
 
-kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = T,
+kable(q, "latex", align = c("l", "r", "r", "r", "r", "r"),
+      booktabs = T, linesep = "", escape = T,
       caption = "Descriptive statistics for regression covariates") %>%
@@
 
-## Analysis
+## Results
 
@@
 pertaining to the logistic regression before concluding with an
-assessment of model fit.
+assessment of model fit and the presentation of several alternative
+specifications for the dependent variable, conflict onset.
 
@@
 
-```{r  fig.cap = "\\label{fig:fpars} Measurement Model Parameter Estimates"}
+```{r fig.height = 4, fig.cap = "\\label{fig:fpars} Measurement Model Parameter Estimates"}
 load("./posteriors/full_model/input.RData")
@@
 
-delta_eta <- extract(fa, pars = c("delta", "eta"))
-
-p1 <- plot_pars(lambda, "Lambda")
-p2 <- plot_pars(psi, "Psi")
-p3 <- plot_pars(gamma, "Gamma")
-p4 <- plot_pars(delta_eta, "")
+delta_eta <- extract(fa, pars = c("delta", "kappa"))
+
+p1 <- plot_pars(lambda, "") + ggtitle(expression(lambda))
+p2 <- plot_pars(psi, "") + ggtitle(expression(psi))
+p3 <- plot_pars(gamma, "") + ggtitle(expression(gamma))
+p4 <- plot_pars(delta_eta, "") + ggtitle(expression(delta * "-" * kappa)) +
+    theme(axis.text.x = element_text(angle = 0)) +
+    scale_x_discrete(labels = c(expression(delta), expression(kappa)))
 
@@
 Table \ref{tab:fa} and figure \ref{fig:fpars} summarise the posterior
-draws for the measurement model parameters according to $80$ percent
-quantile intervals. According to the results, the factor loadings,
-$\lambda$, for each manifest variable are fairly comparable, with the
-exception of the lower value for *v2lgfunds* ("legislature controls
-its own resources") and to a lesser extent *v2lgqstexp* ("legislature
+draws for the measurement model parameters with 80 percent quantile
+intervals. According to the results, the factor loadings, $\lambda$,
+for each manifest variable are fairly comparable, with the exception
+of the lower value for *v2lgfunds* ("legislature controls its own
+resources") and to a lesser extent *v2lgqstexp* ("legislature
 questions executive branch officials in practice"). Relative to the
@@
 ```{r fa, fig.height = 6, fig.width = 6}
-fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
+fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "kappa")) %>%
     post_summarise(probs = c(0.10, 0.9)) %>%
@@
 posterior standard deviations as reported by V-Dem. The practical
-effect is that where the reported error is larger, the uncertainty in
-$Z^{\ast}$ estimates increases and the mean is drawn towards the
-regression line in the factor analysis. In other words, the model is
-more skeptical of outliers in the manifest variables. This posterior
-shrinkage can be seen in figure \ref{fig:fa-line}. Figure
-\ref{fig:err} meanwhile plots the range of the upper and low bounds of
-the $80$ percent posterior quantile interval for an example manifest
-variable in $Z^{\ast}$ against the V-Dem standard deviations as well
-as the difference between the estimated posterior mean and the
-observed mean.
+effect is that for a given observation where the reported error is
+larger, the uncertainty in the $Z^{\ast}_{ij}$ estimate increases and
+the mean is drawn towards the regression line in the factor
+analysis. In other words, the model is more skeptical of outliers in
+the manifest variables. This posterior shrinkage can be seen in figure
+\ref{fig:fa-line}. Figure \ref{fig:err} meanwhile plots the range of
+the upper and low bounds of the $80$ percent posterior quantile
+intervals for an example manifest variable in $Z^{\ast}$ against the
+observed V-Dem standard deviations as well as the difference between
+the estimated posterior mean and the observed mean.
 
@@
 $\theta$, comprised of `r n_distinct(final.df$country_name)` unique
-countries and scaled from approximately $-2$ to $2$. Since not every
-single country can be presented, figure \ref{fig:theta} provides a
-time series plot for China, Brazil, Nigeria, Norway, and Russia so as
-to provide an example of the actual $\theta$ estimates. The countries
-were chosen to represent a diversity of geographic regions as well as
-political contexts. From the plot, China, being a single party
-authoritarian state, is consistently ranked the worst in executive
-constraints. The decline for Russia largely follows the ascent of
-Vladimir Putin at the end of 1999 and his gradual consolidation of
-power in the decades that followed. For both Nigeria and Brazil, the
-measure captures periods of military dictatorship such as the period
-following Johnson Aguiyi-Ironsi's seizure of power in Nigeria the 1966
-coup d'etat. The last country, Norway, represents a consolidated
-polyarchy for the time period under consideration with consistently
-strong constraints against the executive office.
+countries and scaled from approximately $-2$ to $2$. Since the
+complete time series for every single country can not be easily
+visualised, an exploratory overview on the performance of the
+executive constraints measure is provided in the form of several
+different plots.  The first, figure \ref{fig:theta}, is a restricted
+time series plot consisting of only China, Brazil, Nigeria, Norway,
+and Russia. The countries were chosen to represent a diverse set of
+geographic regions as well as political contexts. From the plot,
+China, being a single party authoritarian state, is consistently
+ranked the worst in executive constraints. The decline for Russia
+largely follows the ascent of Vladimir Putin to the presidency at the
+end of 1999 and his gradual consolidation of power in the decades that
+followed.[^russia] For both Nigeria and Brazil, the measure captures
+periods of military dictatorship such as the era of military rule in
+Brazil following the 1964 coup d'etat and the period following Johnson
+Aguiyi-Ironsi's seizure of power in the 1966 coup d'etat in Nigeria.
+The last country, Norway, is an example of a consolidated polyarchy
+for the time period under consideration with consistently strong
+constraints against the executive office.
+
+[^russia]: As can be seen in the plot, the observations for Russia
+    begin in 1991 with the fall of the Soviet Union. There is data
+    available in the V-Dem dataset to extend back the executive
+    constraints measure to cover the years of the Russian Soviet
+    Federative Socialist Republic. However, those observations are
+    lost when merging all of the covariates into the final dataset due
+    to the lack of data in PWT, most likely stemming from the
+    difficulty of converting the Soviet measure of economic
+    performance, Net Material Product, to GDP.
 
@@
     mutate(country_name = final.df$country_name,
+           gwid = final.df$gwid,
            year = final.df$year)
@@
 
+For a full cross-country comparison, figure \ref{fig:map} provides a
+snapshot of the posterior medians for $\theta$ for all countries in
+the year 2012. Lighter colors represent countries with higher levels
+of executive constraints, darker colors vice versa, and countries
+which are grey are missing from the dataset. Given the plot, regional
+differences are readily apparent; for example, many of the former
+Soviet Socialist Republics, with the notable exception of the Baltic
+states, are markedly lower, *i.e.* lacking in effective constraints
+against the executive, than the former Eastern bloc countries. This
+plot does not depict, however, the uncertainty in the estimates.
+
+```{r results = "hide", fig.width = 8, fig.cap = "\\label{fig:map} Executive Constraints - 2012. Grey indicates countries missing from the dataset."}
+map_data <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
+    filter(GWEYEAR == max(GWEYEAR)) %>%
+    select(country_name = CNTRY_NAME, gwid = GWCODE)
+
+df <- filter(theta_summarised, year == max(year)) %>%
+    select(-country_name) %>%
+    right_join(map_data) %>%
+    st_as_sf
+
+ggplot(df) +
+    geom_sf(aes(fill = median)) +
+    theme_tufte() +
+    theme(legend.title = element_blank())
+
+```
+
+```{r results = "hide"}
+chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2000, 2010)) %$%
+    reduced_idx
+
+chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
+
+tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
+    reduced_idx
+
+tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
+```
+
+Finally, to expand upon the country selection from \ref{fig:theta} and
+provide a temporal comparison, figure \ref{fig:tie} plots the $\theta$
+estimates for 2010 versus 2000 by country. A full list of all
+countries is available in [Appendix C](#appendix-c). Note, the
+statistical significance of differences between years for a single
+country cannot strictly be judged by overlapping credible intervals
+[@Schenker2001]. For example, Chile, despite having overlapping
+credible intervals, still has a posterior probability of approximately
+$`r chile_p`$ percent that the estimate for 2000 is lower
+than 2010. In the case of Tunisia with even more visual overlap, there
+is a $`r tunisia_p`$ percent posterior probability that the estimate
+for 2010 is higher than 2000.
+
 ```{r fig.cap = "\\label{fig:tie} Executive Constraints 2000 vs 2010"}
@@
            country_name = ifelse(country_name == "United States of America",
-                                 "USA", country_name) %>%
-               factor(levels = sort(c(targets, "USA"), T)))
+                                 "USA", country_name))
+
+levels <- filter(df, year == 2010) %>%
+    arrange(median) %$%
+    country_name
+
+df$country_name <- factor(df$country_name, levels = levels)
 
@@
 
-```{r results = "hide"}
-chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2000, 2010)) %$%
-    reduced_idx
-
-chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
-
-tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
-    reduced_idx
-
-tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
-```
-
-To provide a further snapshot of a greater number of countries, figure
-\ref{fig:tie} plots the $\theta$ estimates for 2010, the last year in
-the dataset, compared to 2000 by country. A full list of all countries
-is available in [Appendix C](#appendix-c). Note, the statistical
-significance of differences between years for a single country cannot
-strictly be judged by overlapping credible intervals
-[@Schenker2001]. For example, Chile, despite having overlapping
-credible intervals, still has a posterior probability of approximately
-$`r chile_p`\%$ that the estimate for 2010 is lower than 2000. In the
-case of Tunisia with even more visual overlap, there is a
-$`r tunisia_p`\%$ posterior probability that the estimate for 2010 is
-higher than 2000.
-
 ### Onset Regression
@@
 $\beta$ estimate is greater than zero and less than zero. Recall that
-the dependent variables are standardized to have mean $0$ and standard
-deviation $1$, with the exception of the ordinal/binary variables
-Relevant Group Count, Neighbouring Conflict, Ongoing Conflict, and
+the variables are standardized to have mean 0 and standard deviation
+1, with the exception of the ordinal/binary variables Relevant Group
+Count, Neighbouring Conflict, Ongoing Major Conflict, and
 Independence.
@@
 
-For the $\beta$ parameter for executive constraints, this means that
-there is an $80\%$ posterior probability that an increase of one
-standard deviation decreases the log-odds of conflict onset between
-$`r beta_s[1, 1]`$. A one standard deviation increase in this context
-corresponds roughly to the difference between the level of executive
-constraints for Zambia in 1967 under the administration of Kenneth
-Kaunda right before the banning of opposition parties following the
-1968 election with a score of approximately $0$ and Greece in 1976
-shortly after the end of military rule and establishment of a
-democratic republic and with a score of approximately $1$. The
-probability that the $\beta$ estimate for executive constraints is
-below $0$ is $0.984$.
+This means that for the $\beta$ parameter for executive constraints
+there is a `r as.numeric(beta_s[1, 2]) * 100` percent posterior
+probability that the parameter is less than zero and a 80 percent
+posterior probability that an increase of one standard deviation
+decreases the log-odds of conflict onset between $`r beta_s[1, 1]`$. A
+one standard deviation increase in this context corresponds roughly to
+the difference between the level of executive constraints for Zambia
+in 1967 under the administration of Kenneth Kaunda right before the
+banning of opposition parties following the 1968 election, with score
+of approximately 0, and Greece in 1976 shortly after the end of
+military rule and establishment of a democratic republic, with a score
+of approximately 1.
 
 ```{r results = "hide"}
-gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
-    rgdpepc %>%
-    prettyNum(big.mark = ",")
-
-gdp_czech <- filter(merged.df, country_name == "Czech Republic", year == 1992) %$%
-    rgdpepc %>%
-    prettyNum(big.mark = ",")
-
-gdpgro_korea <- filter(merged.df, country_name == "South Korea", year == 1990) %$%
-    rgdpepc_gro %>%
-    multiply_by(100) %>%
-    signif(3)
-
-gdpgro_usa <- filter(merged.df, country_name == "United States of America", year == 1992) %$%
-    rgdpepc_gro %>%
-    multiply_by(100) %>%
-    signif(3)
-
-lth <- filter(merged.df, country_name == "Lithuania", year == 2012) %>%
+argentina <- filter(merged.df, country_name == "Argentina", year == 1989) %$%
+    prettyNum(rgdpepc, big.mark = ",")
+
+slovakia <- filter(merged.df, country_name == "Slovakia", year == 2003) %$%
+    prettyNum(rgdpepc, big.mark = ",")
+
+cmb <- filter(merged.df, country_name == "Cambodia", year == 1987) %>%
     select(pop, area_sqkm) %>%
@@
 
-india <- filter(merged.df, country_name == "India", year == 1978) %>%
+india <- filter(merged.df, country_name == "India", year == 1976) %>%
     select(pop, area_sqkm, meanelev) %>%
@@
 
-afghan <- filter(merged.df, country_name == "Afghanistan", year == 2004) %>%
-    select(meanelev) %>%
-    mutate(meanelev = signif(meanelev, 3))
-
-```
-
-The slopes for the remaining covariates generally following the
-intuitive expectation of their relationship to civil conflict onset
-with high probabilities of being either greater than or less than
-zero. For example, GDP per capita is negative to which a one unit
-increase corresponds approximately to Nigeria in 2014 with a value of
-$`r gdp_nigeria`$ and Czech Republic in 1992 after the fall of
-communism with a value of $`r gdp_czech`$. In other words, an increase
-in GDP per capita on the logarithmic scale decreases the log-odds of
-conflict onset. Similarly, GDP per capita growth is negatively
-associated with conflict onset with an increase from 0 to 1 on the
-standardized scale matching the level of United States in 1992,
-$`r gdpgro_usa`$ percent, and South Korea in 1990, $`r gdpgro_korea`$ percent.
-Population density, mean elevation, number of politically relevant
-ethnic groups, presence of neighbouring conflict, and the dummy
+georgia <- filter(merged.df, country_name == "Georgia", year == 2012) %$%
+    signif(meanelev, 3)
+
+russia <- filter(merged.df, country_name == "Russia", year == 2012) %$%
+    signif(meanelev, 3)
+
+thai <- filter(merged.df, country_name == "Thailand", year == 2002) %$%
+    { discrimpop * 100 }
+
+bhutan <- filter(merged.df, country_name == "Bhutan", year == 1989) %$%
+    { discrimpop * 100 }
+
+```
+
+The slopes for the remaining covariates generally follow the intuitive
+expectations of their relationship to civil conflict onset with
+varying degrees of certainty. For example, GDP per capita is negative
+to which a one unit increase from 0 to 1 is roughly similar to the
+difference between Argentina in 1989 at the height of a hyperinflation
+crisis with a value of $`r argentina`$ 2011 US dollars and Slovakia in
+2003 with a value of $`r slovakia`$ 2011 US dollars at the beginning
+of a period of high GDP growth. In other words, an increase in GDP per
+capita on the logarithmic scale decreases the log-odds of conflict
+onset. GDP per capita growth and CINC are also negative, but the
+estimates are significantly less reliable. The posterior probability
+of being less than zero is $`r beta_s[3, 2]`$ for the former and only
+$`r beta_s[4, 2]`$ for the latter.
+
+Population density, mean elevation, number of political relevant
+ethnic groups, proportion of the population facing state
+discrimination, presence of neighbouring conflict, and the dummy
 variable for independence year all increase the risk of civil conflict
 onset. A one unit increase in population density is similar to going
-from Lithuania in 2012 with a population of $`r lth$pop`$ million and
-an area of $`r lth$area_sqkm`$ square kilometers to India in 1978 with
+from Cambodia in 1987 with a population of $`r cmb$pop`$ million and
+an area of $`r cmb$area_sqkm`$ square kilometers to India in 1976 with
 a population of $`r india$pop`$ million and an area of
 $`r india$area_sqkm`$ square kilometers. A one unit increase for mean
-elevation roughly matches the difference between India,
-$`r india$meanelev`$ meters, and Afghanistan, $`r afghan$meanelev`$
-meters. Finally, the number of peace years since the last civil
-conflict and the dummy variable for ongoing civil conflict decrease
-the risk of conflict onset. This latter result, though surprising, is
-a result of conditioning on the other covariates.
+elevation roughly matches the difference between Russia, $`r russia`$
+meters, and Georgia, $`r georgia`$ meters. And, a one unit increase
+for proportion discrimination is roughly equivalent to Thailand from
+2002 until present with $`r thai`$ percent and Bhutan from 1989 with
+$`r bhutan`$ percent.
+
+Finally, the GDP per capita and executive constraints interaction term
+and the dummy variable for ongoing high intensity conflict decrease
+the risk of conflict onset. This latter result, although surprising,
+may be due to the subsuming nature of high intensity civil wars. In
+the face of an all-encompassing civil war, there may be a plethora of
+factions with divergent interests active; however, their political
+goals and interests are structured according to the current
+incompatibility between the dominant political actors. Thus, there may
+be new armed groups emerging who are willing to challenge to the
+state, but that emergent may not necessarily be coded as a new civil
+conflict. That said, it should be mentioned that the there is a much
+higher degree of uncertainty in the estimate with the 80 percent
+posterior credible interval ranging from -0.558 to 0.064.
 
@@
          beta[, 10] * 0 +
-         beta[, 11] * mean(peace_yrs) +
-         beta[, 12] * 0 +
-         beta[, 13] * x * y)
+         beta[, 11] * 0 +
+         beta[, 12] * x * y)
 }
@@
 predicted probabilities. For onset country-years the interquartile
-interval is only 0.0475 to 0.2 with a max of 0.516. To better
-compare the predicted probabilities to the actual data, table
+interval is only 0.0475 to 0.2 with a max of 0.516. To better compare
+the predicted probabilities to the actual data, table
 \ref{tab:confusion} presents a confusion matrix with a 0.2 cutoff,
-which corresponds to the third quartile in the summary table.
-`r signif((93 / (259 + 93)) * 100, 3)` percent of onsets and
+which approximately corresponds to the third quartile in the summary
+table.  `r signif((93 / (259 + 93)) * 100, 3)` percent of onsets and
 `r signif((7902 / (7902 + 172)) * 100, 3)` percent of non-onsets are
@@
 between *precision* --- the number of true positives divided by the
-sum of the true positives and false positives --- and recall --- the
+sum of the true positives and false positives --- and *recall* --- the
 number of true positives divided by the sum of the true positives and

### 2020-04-06 R/thesis.utils/R/debug.R
Update debugging function

@@ -26,16 +26,20 @@ assert_cy <- function(df) {
 #' Merging stats
 #'
 #' Extremely simply debugging function that does nothing more than
-#' simply print out the number of conflict episodes, countries, and
+#' simply print out the number of conflicts, countries, and
 #' country-years in a given data frame. Completely useless for any
 #' generic task.
 #'
 #' @param df DataFrame
 #'
 #' @export
-dbg_info <- function(df) {
-    msg <- sprintf("%d conflict episodes, %d countries, %s country-years",
-                   sum(df$lepisode_onset, na.rm = T),
+dbg_info <- function(df, onset_col = "episode_onset") {
+    if (!onset_col %in% colnames(df))
+        sprintf("Can't find dependent variable: %s", onset_col) %>%
+            stop(call. = F)
+
+    msg <- sprintf("%d target onsets, %d countries, %s country-years",
+                   sum(df[, onset_col], na.rm = T),
                    dplyr::n_distinct(df$country_name),
                    prettyNum(nrow(df), big.mark = ","))
 

### 2020-04-29 paper.Rmd
Add missing sf import

@@
 library(precrec)
+library(sf)
 library(tidyr)

### 2020-04-29 paper.Rmd
Add section interpreting random intercepts

@@
 
+TODO: grab this programmatically
+
+#### Random Effects
+
+There are three sets of random effects included in the regression
+model. The first to be presented in figure \ref{fig:year} is the year
+intercepts with 80 percent posterior credible intervals. There is no
+discernible time trend in the year specific variability. The estimates
+are strongly pulled towards zero as a result of the partial pooling
+with upticks, such as the in the initial years of the 1990s during the
+dissolution of the Soviet Union and Yugoslavia, roughly mirroring the
+global incidence rate of onsets.
+
+```{r fig.cap = "\\label{fig:year} Log-odds year intercepts - $80\\%$ CIs"}
+Z_year <- extract(reg, "Z_year")
+colnames(Z_year) <- { unique(final.df$year) + 1 } %>% sort
+
+breaks <- seq(min(final.df$year), max(final.df$year), by = 4)
+plot_pars(Z_year, hline = T) +
+    scale_x_discrete(breaks = breaks)
+
+```
+
+```{r results = "hide"}
+alpha <- extract(reg, "alpha")
+
+alpha_s <- quantile(alpha, probs = c(0.1, 0.9)) %>% signif(3)
+alpha_interval <- sprintf("[%g,%g]", alpha_s[1], alpha_s[2])
+```
+
+Figure \ref{fig:country} presents the country intercepts. Since all of
+the countries cannot be easily visualized at once only the top ten and
+bottom ten depicted. From the results it is clear that there is
+considerably less uncertainty in the estimates for the top
+countries. The intercepts for these countries are consistently above
+zero meaning that the risk of civil conflict onset is higher than that
+predicted solely by the other predictors. Unsurprisingly, these
+countries tend to experience the highest number of civil conflict
+onsets in the dataset which, conditional on the data, accounts for the
+higher baseline risk. The countries at the bottom have experienced
+relatively few, if any, conflict onsets relative to their values for
+the other predictors. The majority of intercepts for these countries
+overlap zero possibly because the global intercept is already fairly
+low, `r alpha_interval` (80 percent log-odds CI).
+
+```{r, fig.cap = "\\label{fig:country} Log-odds country intercepts - $80\\%$ CIs"}
+
+Z_country <- extract(reg, "Z_country")
+colnames(Z_country) <- unique(final.df$country_name) %>% sort
+
+x <- apply(Z_country, 2, median) %>% order(decreasing = F)
+top <- head(x, 10)
+bottom <- tail(x, 10)
+
+fn <- function(x, title) {
+    plot_pars(x, hline = T) +
+        coord_flip() +
+        theme(axis.title.y = element_blank()) +
+        ggtitle(title)
+}
+
+p1 <- fn(Z_country[, top], "Bottom 10")
+p2 <- fn(Z_country[, bottom], "Top 10")
+
+grid.arrange(p1, p2, ncol = 2)
+
+```
+
+```{r results = "hide"}
+gp <- extract(reg, c("rho", "eta")) %>% post_summarise(probs = c(0.1, 0.9)) %>%
+    mutate(interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)))
+```
+
+Finally, figure \ref{fig:peace} plots the peace year intercepts.[^gp]
+The strongest effect is found in the years immediately following an
+occurrence of intrastate political violence resulting in an increased
+risk for an additional onset.  Approximately a decade after the last
+incidence of violence, the effect is drawn towards zero. The
+additional spike around approximately 30 years follows an uptick in
+conflict onsets in the dataset mostly among authoritarian regimes such
+as, for example, the 1989 coup d'etat overthrowing the Stroessner
+regime in Paraguay, the Saur Revolution in 1978 in Afghanistan, and
+the start of the Nepalese Civil War in 1996. After 30 years, the
+effect of each additional year of peace tends to be negative; however,
+there is a very high degree of uncertainty in the estimates that
+undermines any general conclusion regarding a pacifying effect.
+
+[^gp]: The 80 percent posterior credible intervals for the
+    hyperparameters from the gaussian process modeling the effect of
+    peace years were
+    `r filter(gp, parameter == "rho") %$% interval` and
+    `r filter(gp, parameter == "eta") %$% interval`
+    for the length-scale ($\rho$) and the marginal standard deviation
+    ($\eta$) respectively.
+
+```{r, fig.cap = "\\label{fig:peace} Log-odds peace year intercepts - $80\\%$ CIs"}
+f <- extract(reg, "f")
+colnames(f) <- unique(final.df$lpeace_yrs) %>% sort
+
+plot_pars(f, hline = T) +
+    scale_x_discrete(breaks = seq(0, ncol(f), by = 4))
+```
+
+#### Interaction Effect
+
 ```{r results = "hide"}

### 2020-05-01 R/merge.R
Focus on all episodes, high intensity, and cumulative intensity
This is far from comprehensive for all the different coding standards
that can be applied to civil conflict onset, but focuses on the "big
three" which has been used by much of the previous
literature (ex. Bazzi & Blattman (2014), although they also
incorporate a number of datasets besides just UCDP/Prio).

@@ -176,37 +176,27 @@ civil %<>% filter(type_of_conflict %in% 3:4)
 episode_ucdp <- find_onset(civil, c("conflict_id", "start_date2")) %>%
     rename(episode_onset = onset)
 
-# Episode onsets - major conflicts only
+# Episode onsets - high intensity
 episode_major_ucdp <- filter(civil, intensity_level == 2) %>%
     find_onset(c("conflict_id", "start_date2")) %>%
-    rename(episode_major_onset = onset)
-
-# Unique conflict onsets - all conflicts
-conflict_ucdp <- find_onset(civil, "conflict_id")
-
-# Unique conflict onsets - major conflicts only
-conflict_major_ucdp <- filter(civil, intensity_level == 2) %>%
-    find_onset("conflict_id") %>%
     rename(major_onset = onset)
 
-ucdp <- list(episode_ucdp, episode_major_ucdp, conflict_ucdp, conflict_major_ucdp) %>%
+# Episode onsets - cumulative intensity
+episode_cum_ucdp <- filter(civil, cumulative_intensity == 1) %>%
+    find_onset(c("conflict_id", "start_date2")) %>%
+    rename(cum_onset = onset)
+
+ucdp <- list(episode_ucdp, episode_major_ucdp, episode_cum_ucdp) %>%
     Reduce(partial(full_join, by = c("gwno_a", "year")), .)
 
 merged.df <- left_join(vdem, ucdp, by = c("gwid" = "gwno_a", "year")) %>%
     left_join(full_counts.df, by = c("gwid" = "gwno_a", "year")) %>%
-    mutate(ongoing = ifelse(is.na(intensity), 0, intensity),
+    mutate(ongoing = ifelse(is.na(intensity), 0, 1),
+           major_ongoing = ifelse(is.na(intensity) | intensity == 1, 0, 1),
            neighbour_conflict = ifelse(is.na(n_neighbour_conflicts), 0, 1),
-           onset = ifelse(is.na(onset), 0, 1),
-           major_onset = ifelse(is.na(major_onset), 0, 1),
            episode_onset = ifelse(is.na(episode_onset), 0, 1),
-           episode_major_onset = ifelse(is.na(episode_major_onset), 0, 1)) %>%
-    group_by(country_name, idx = consecutive(year)) %>%
-    mutate(lonset = lead(onset),
-           lmajor_onset = lead(major_onset),
-           lepisode_onset = lead(episode_onset),
[…]

### 2020-05-01 R/model_data.R

@@ -21,16 +21,20 @@ output_dir <- basename(args$schema) %>%
     file_path_sans_ext %>%
     file.path("posteriors", .)
 
+# Keep gwid for world plots in manuscript
 final.df %<>%
-    select(schema$outcome, country_name, year, reduced_idx,
-           one_of(schema$predictors)) %>%
+    select(country_name, gwid, year, reduced_idx, lpeace_yrs,
+           schema$outcome, one_of(schema$predictors)) %>%
     na.omit
-dbg_info(final.df)
+dbg_info(final.df, schema$outcome)
+
+stopifnot(schema$predictors %in% colnames(final.df))
 
 ###
 # Control variables
 X <- final.df %>%
-    select(-country_name, -year, -matches("onset"), -reduced_idx, -peace_yrs) %>%
+    select(-country_name, -year, -gwid, -matches("onset"),
+           -reduced_idx, -lpeace_yrs) %>%
     mutate_if(Negate(is.ordinal), normalize) %>%
     data.matrix
 
@@ -71,15 +75,15 @@ data <- list(
     interaction_idx = interaction_idx,
     exec_idx = final.df$reduced_idx,
 
-    n_peace_yrs = n_distinct(final.df$peace_yrs),
+    n_peace_yrs = n_distinct(final.df$lpeace_yrs),
     n_countries = n_distinct(final.df$country_name),
     n_years = n_distinct(final.df$year),
 
     country_id = to_idx(final.df$country_name),
     year_id = to_idx(final.df$year),
-    peace_yrs_id = to_idx(final.df$peace_yrs),
+    peace_yrs_id = to_idx(final.df$lpeace_yrs),
 
-    peace_yrs = unique(final.df$peace_yrs) %>% sort,
+    peace_yrs = unique(final.df$lpeace_yrs) %>% sort,
 
     y = final.df[[schema$outcome]]
 )

### 2020-05-01 R/transform.R

@@ -32,21 +32,29 @@ for (v2 in lg_vars) {
 }
 
 ###
-# Final dataset for model
+# Final dataset for model. Start by dropping country-years where we
+# have missingness in our manifest variables NOT caused by lack of
+# legislature (v2lgbicam) and log transforming key predictors.
 final.df <- merged.df %>%
-    select(country_name, year, lonset, lmajor_onset, lepisode_onset,
-           lepisode_major_onset, peace_yrs, v2lgbicam,
-           one_of(constraint_vars), one_of(paste0(constraint_vars, "_sd")),
-           rgdpepc, rgdpepc_gro, cinc, pop_density, meanelev, ongoing,
-           independence, rlvt_groups_count, discrimpop,
-           neighbour_conflict) %>%
     filter_at(constraint_vars, all_vars(!is.na(.))) %>%
+    arrange(country_name, year) %>%
     mutate(rgdpepc = log(rgdpepc),
            pop_density = log(pop_density),
            meanelev = log(meanelev),
            reduced_idx =
                do.call(paste, lapply(c("country_name", constraint_vars), as.symbol)) %>%
                    collapse_changes)
+
+# Rather than lag each predictor, take the lead of the outcome and
+# peace years.
+final.df %<>%
+    group_by(country_name, idx = consecutive(year)) %>%
+    mutate(lepisode_onset = lead(episode_onset),
+           lmajor_onset = lead(major_onset),
+           lcum_onset = lead(cum_onset),
+           lpeace_yrs = lead(peace_yrs)) %>%
+    ungroup %>%
+    select(-idx)
 
 dbg_info(final.df)
 

### 2020-05-01 R/sim_data.R
Update GP priors to match manuscript

@@ -51,12 +51,12 @@ for (i in 1:ncol(nonlg))
     nonlg_obs[, i] <- rnorm(N, nonlg[, i], nonlg_err[, i])
 
 # Conflict regression
-rho <- rinvgamma(1, 7.30124, 30.0086)
-eta <- rhnorm(1, 0.5)
+rho <- rinvgamma(1, 15.5031, 78.2992);
+eta <- rhnorm(1, 1)
 
 D <- dist(1:N, diag = T, upper = T) %>% as.matrix
 
-K <- eta^2 * exp((-0.5 / (2 * rho^2)) * D^2)
+K <- eta^2 * exp(-0.5 * (D / rho)^2)
 f <- mvrnorm(1, rep(0, N), K)
 
 alpha <- rnorm(1, 0, 5)
@@ -64,8 +64,8 @@ beta <- rnorm(3, 0, 2.5)
 
 X <- rnorm(N, 2, 2)
 
-n_years <- N %/% 5L
-n_countries <- 5L
+n_years <- N %/% 4L
+n_countries <- 4L
 
 sigma <- rhcauchy(2, 1)
 countries <- rnorm(n_countries, 0, sigma[1])
@@ -154,9 +154,10 @@ true_values <- data.frame(parameter = c(sprintf("lambda[%d]", seq_along(lambda))
                                         "alpha",
                                         sprintf("beta[%d]", seq_along(beta)),
                                         sprintf("sigma[%d]", seq_along(sigma)),
-                                        sprintf("theta[%d]", seq_along(theta))),
+                                        sprintf("theta[%d]", seq_along(theta)),
+                                        sprintf("y[%d]", seq_along(dataset$y))),
                           point = c(lambda, psi, gamma, delta, kappa, rho, eta,
-                                    alpha, beta, sigma, theta),
+                                    alpha, beta, sigma, theta, dataset$y),
                           type = "True Values",
                           stringsAsFactors = F)
 

### 2020-05-01 paper.Rmd
Manuscript edits

@@
 variables. Where the legislature is present, $\theta$ is distributed
-according to the $Normal(0, 1)$ prior.
+according to the $\mathsf{Normal}(0, 1)$ prior.
 
@@
 observed V-Dem standard deviations as well as the difference between
-the estimated posterior mean and the observed mean.
+the estimated posterior mean and the observed mean in order to further
+demonstrate this effect.
 
@@
 
-In total there are `r nrow(final.df)` country-year estimates of
-$\theta$, comprised of `r n_distinct(final.df$country_name)` unique
-countries and scaled from approximately $-2$ to $2$. Since the
-complete time series for every single country can not be easily
-visualised, an exploratory overview on the performance of the
-executive constraints measure is provided in the form of several
-different plots.  The first, figure \ref{fig:theta}, is a restricted
-time series plot consisting of only China, Brazil, Nigeria, Norway,
-and Russia. The countries were chosen to represent a diverse set of
-geographic regions as well as political contexts. From the plot,
-China, being a single party authoritarian state, is consistently
-ranked the worst in executive constraints. The decline for Russia
-largely follows the ascent of Vladimir Putin to the presidency at the
-end of 1999 and his gradual consolidation of power in the decades that
-followed.[^russia] For both Nigeria and Brazil, the measure captures
-periods of military dictatorship such as the era of military rule in
-Brazil following the 1964 coup d'etat and the period following Johnson
-Aguiyi-Ironsi's seizure of power in the 1966 coup d'etat in Nigeria.
-The last country, Norway, is an example of a consolidated polyarchy
-for the time period under consideration with consistently strong
-constraints against the executive office.
+In total there are `r nrow(final.df) %>% prettyNum(big.mark = ",")`
+country-year estimates of $\theta$, comprised of `r
+n_distinct(final.df$country_name)` unique countries and scaled from
+approximately $-2$ to $2$. Since the complete time series for every
+single country can not be easily visualised, an exploratory overview
+on the performance of the executive constraints measure is provided in
+the form of several different plots.  The first, figure
+\ref{fig:theta}, is a restricted time series plot consisting of only
+China, Brazil, Nigeria, Norway, and Russia. The countries were chosen
+to represent a diverse set of geographic regions as well as political
+contexts. From the plot, China, being a single party authoritarian
+state, is consistently ranked the worst in executive constraints. The
+decline for Russia largely follows the ascent of Vladimir Putin to the
+presidency at the end of 1999 and his gradual consolidation of power
+in the decades that followed.[^russia] For both Nigeria and Brazil,
+the measure is able to distinguish between democratic rule and periods
+of military dictatorship such as the era of military rule in Brazil
+following the 1964 coup d'etat and the period following Johnson
+Aguiyi-Ironsi's seizure of power in the 1966 coup d'etat in
+Nigeria. The last country, Norway is an example of a consolidated
+polyarchy for the time period under consideration with consistently
+strong constraints against the executive office which are accurately
+captured by the measure.
 
@@
 
-Figure \ref{fig:reg} plots the slope coefficients on the log-odds
-scale from the logistic regression linking the latent construct for
-executive constraints to the outcome, civil conflict onset. Credible
+Figure \ref{fig:reg} plots the slope coefficients from the logistic
+regression linking the latent construct for executive constraints to
+the outcome, civil conflict onset, on the log-odds scale. Credible
 intervals are summarised the same as previous plots, 80 percent
@@
 
-This means that for the $\beta$ parameter for executive constraints
-there is a `r as.numeric(beta_s[1, 2]) * 100` percent posterior
-probability that the parameter is less than zero and a 80 percent
-posterior probability that an increase of one standard deviation
-decreases the log-odds of conflict onset between $`r beta_s[1, 1]`$. A
-one standard deviation increase in this context corresponds roughly to
-the difference between the level of executive constraints for Zambia
-in 1967 under the administration of Kenneth Kaunda right before the
-banning of opposition parties following the 1968 election, with score
-of approximately 0, and Greece in 1976 shortly after the end of
-military rule and establishment of a democratic republic, with a score
-of approximately 1.
+For the $\beta$ parameter for executive constraints there is a
+`r as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
+parameter is less than zero and a 80 percent posterior probability
+that an increase of one standard deviation decreases the log-odds of
+conflict onset between $`r beta_s[1, 1]`$. A one standard deviation
+increase in this context corresponds roughly to the difference between
+the level of executive constraints for Zambia in 1967 under the
+administration of Kenneth Kaunda right before the banning of
+opposition parties following the 1968 election, with a score of
+approximately 0, and Greece in 1976 shortly after the end of military
+rule and the establishment of a democratic republic with a score of
+approximately 1.
 
@@
 $`r india$area_sqkm`$ square kilometers. A one unit increase for mean
-elevation roughly matches the difference between Russia, $`r russia`$
-meters, and Georgia, $`r georgia`$ meters. And, a one unit increase
-for proportion discrimination is roughly equivalent to Thailand from
-2002 until present with $`r thai`$ percent and Bhutan from 1989 with
-$`r bhutan`$ percent.
+elevation roughly matches the difference between Russia at $`r russia`$
+meters and Georgia at $`r georgia`$ meters. And, a one unit increase
+for proportion of the population facing discrimination is roughly
+equivalent to Thailand from 2002 until present with $`r thai`$ percent
+and Bhutan from 1989 with $`r bhutan`$ percent.
 
@@
 the face of an all-encompassing civil war, there may be a plethora of
-factions with divergent interests active; however, their political
+active factions with divergent interests; however, their political
 goals and interests are structured according to the current
@@
 be new armed groups emerging who are willing to challenge to the
-state, but that emergent may not necessarily be coded as a new civil
-conflict. That said, it should be mentioned that the there is a much
-higher degree of uncertainty in the estimate with the 80 percent
-posterior credible interval ranging from -0.558 to 0.064.
+state, but that emergence may not necessarily be coded as a new civil
+conflict. It should be noted, however, that there is a high degree of
+uncertainty in the estimate with the 80 percent posterior credible
+interval ranging from -0.558 to 0.064.
 
@@
 capita while holding constant the other covariates.  In this case, I
-took the mean of GDP per capita growth, population density, mean
-elevation, politically relevant ethnic group count, and peace years
-while setting the dummy indicators for independence, ongoing conflict,
-and neighbouring conflict and the random intercepts for year and
-country to 0. The levels of GDP per capita were selected based on the
-10\%, 50\%, and 90\% quantiles of the standardized log transformed
-values. These levels roughly correspond to the GDP per capita of
-Rwanda in 2007 at $`r gdp_rwa`$, the Philippines in 2008 at $`r gdp_phl`$,
-and Portugal in 2009 at $`r gdp_po`$. In the dataset there are 60
-conflict episode onsets below the first level, $152$ between the first
-and second, $55$ between the second and third, and $4$ above the
-third.
+take the mean of GDP per capita growth, CINC, the proportion of
+discriminated population, population density, mean elevation, and the
+number of politically relevant ethnic groups. I also set the dummy
+indicators for independence, major ongoing conflict, and neighbouring
+conflict and the random intercepts for year and country to zero. For
+peace years I take the intercept for the average number of peace years
+rounded to the nearest integer, which turns out to be 19. The levels
+of GDP per capita were selected based on the 10\%, 50\%, and 90\%
+quantiles of the standardized log transformed values. These levels
+roughly correspond to the GDP per capita of Rwanda in 2007 at
+$`r gdp_rwa`$ 2011 US dollars, the Philippines in 2008 at $`r gdp_phl`$
+2011 US dollars, and Portugal in 2009 at $`r gdp_po`$ 2011 US
+dollars. In the dataset there are 60 conflict episode onsets below the
+first level, 152 between the first and second, 55 between the second
+and third, and 4 above the third.
 
@@
 # Interaction plot
-alpha <- extract(reg, "alpha")
 theta_seq <- seq(-2.5, 2.5, .01)
-
 state <- quantile(X[, "rgdpepc"], probs = c(0.1, .5, 0.9))
 
-f <- function(x, y) {
-    with(as.data.frame(X),
+marginal_data.df <- as.data.frame(X) %>%
+    summarise_all(~mean(.)) %>%
+    mutate(rlvt_groups_count = round(rlvt_groups_count),
+           neighbour_conflict = 0,
+           major_ongoing = 0,
+           independence = 0)
+
+peace_yr <- f[, mean(final.df$lpeace_yrs) %>% round]
+
+fn <- function(x, y) {
+    with(marginal_data.df,
          alpha +
@@
          beta[, 2] * y +
-         beta[, 3] * mean(rgdpepc_gro) +
-         beta[, 4] * mean(cinc) +
-         beta[, 5] * mean(discrimpop) +
-         beta[, 6] * mean(pop_density) +
-         beta[, 7] * mean(meanelev) +
-         beta[, 8] * mean(rlvt_groups_count) +
-         beta[, 9] * 0 +
-         beta[, 10] * 0 +
-         beta[, 11] * 0 +
-         beta[, 12] * x * y)
+         beta[, 3] * rgdpepc_gro +
+         beta[, 4] * cinc +
+         beta[, 5] * discrimpop +
+         beta[, 6] * pop_density +
+         beta[, 7] * meanelev +
+         beta[, 8] * rlvt_groups_count +
+         beta[, 9] * neighbour_conflict +
+         beta[, 10] * major_ongoing +
+         beta[, 11] * independence +
+         beta[, 12] * x * y +
+         peace_yr)
 }
@@
     mclapply(theta_seq, function(point) {
-        p_hat <- boot::inv.logit(f(point, y)) * 100
+        p_hat <- boot::inv.logit(fn(point, y)) * 100
         post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%

### 2020-05-02 paper.Rmd
Remove old robustness checks
This paves the way to eventually rework the appendices and incorporate
additional model specifications directly in the manuscript.

@@
 
+```
+
 \newpage
 
-## C - Country Comparison {#appendix-c}
-
-\newpage
-
-## D - Additional Descriptive Statistics {#appendix-d}
-
-\newpage
-
-## E - Robustness Checks {#appendix-e}
-
-```{r}
-models <- setdiff(list.files("./posteriors/"), c("full_model_new", "sim"))
-
-# TODO: this is killing our compile times
-all_betas <- mclapply(models, function(f) {
-    load(file.path("posteriors", f, "input.RData"))
-    beta <- file.path("posteriors", f, "reg_posteriors.csv") %>%
-        read_post %>%
-        extract("beta")
-
-    colnames(beta)[1:(1 + ncol(X))] <- c("exec_constraints", colnames(X))
-    if (interaction_idx > 0)
-        colnames(beta)[ncol(beta)] <- paste("exec *", colnames(X)[interaction_idx])
-
-    post_summarise(beta, probs = c(0.1, 0.5, 0.9)) %>%
-        mutate(model = f)
-}, mc.cores = 4) %>% bind_rows
-
-all_betas %<>%
-    rename(codelow = `10%`,
-           median = `50%`,
-           codehigh = `90%`) %>%
-    mutate(model = to_idx(model) %>% as.factor)
-
-desc <- case_when(models == "full_model" ~ "Full Model",
-                  models == "full_no_cinc" ~ "Full Model Excl. CINC",
-                  models == "full_no_interaction" ~ "Full Model Without Interaction",
-                  models == "major_episodes" ~ "Major Conflict Episodes Only",
-                  models == "major_no_interaction" ~ "Major Conflict Episodes Without Interaction",
-                  models == "unique_no_interaction" ~ "Unique Onsets Without Interaction",
-                  models == "unique_onset" ~ "Unique Onsets Only")
-
-df <- data.frame(Model = 1:length(models),
-                 Description = desc,
-                 stringsAsFactors = F)
-
-kable(df, "latex", align = c("l", "l"), booktabs = T, linesep = "",
-      caption = "Model Descriptions") %>%
-    kable_styling(latex_options = c("hold_position", "striped"))
-
-```
-
-```{r fig.width = 7, fig.height = 6}
-ggplot(all_betas, aes(x = parameter, y = median, color = model)) +
-    geom_point(position = position_dodge(width = 0.8)) +
-    geom_errorbar(aes(x = parameter, ymin = codelow, ymax = codehigh),
-                  width = 0, size = 1, position = position_dodge(width = 0.8)) +
-    geom_hline(yintercept = 0, alpha = 0.5, color = "grey") +
-    theme_tufte() +
-    scale_color_colorblind() +
-    theme(panel.background = element_blank(),
-          legend.title = element_blank(),
-          axis.line = element_line(color = "black"),
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          axis.title.x = element_blank()) +
-    ylab("Parameter Estimate")
-
-```
-
-\newpage
-
 # References {-}

### 2020-05-21 paper.Rmd
Update stan file path in manuscript
Draft edits

@@
 there is a very high degree of uncertainty in the estimates that
-undermines any general conclusion regarding a pacifying effect.
+undermines any general conclusion regarding a pacifying effect in the
+model.
 
@@
 
-To interpret the results for the interaction term between the slopes
-for executive constraints and GDP per Capita I calculate the marginal
-effect of $\theta$, executive constraints, on the predicted
-probability of civil conflict onset for different levels of GDP per
-capita while holding constant the other covariates.  In this case, I
-take the mean of GDP per capita growth, CINC, the proportion of
-discriminated population, population density, mean elevation, and the
-number of politically relevant ethnic groups. I also set the dummy
-indicators for independence, major ongoing conflict, and neighbouring
-conflict and the random intercepts for year and country to zero. For
-peace years I take the intercept for the average number of peace years
-rounded to the nearest integer, which turns out to be 19. The levels
-of GDP per capita were selected based on the 10\%, 50\%, and 90\%
-quantiles of the standardized log transformed values. These levels
+To interpret the interaction term between executive constraints and
+GDP per Capita I calculate the marginal effect of $\theta$, executive
+constraints, on the predicted probability of civil conflict onset for
+different levels of GDP per capita while holding constant the other
+covariates. In this case, I take the mean of GDP per capita growth,
+CINC, the proportion of discriminated population, population density,
+mean elevation, and the number of politically relevant ethnic groups
+rounded to the nearest integer. I also set the dummy indicators for
+independence, major ongoing conflict, and neighbouring conflict to
+zero following the majority of observations in the dataset. I set the
+random intercepts for year and country to zero, which is the average
+effect. For peace years I take the intercept corresponding to the
+average number of peace years rounded to the nearest integer. The
+levels of GDP per capita were select based on the 10\%, 50\%, and 90
+\% quantiles of the standardized log transformed values. These levels
 roughly correspond to the GDP per capita of Rwanda in 2007 at
@@
 first level, 152 between the first and second, 55 between the second
-and third, and 4 above the third.
-
-```{r fig.cap = "\\label{fig:int} Interaction between executive constraints and GDP per capita with 80\\% CIs"}
+and third, and 4 above the third. The results are plotted in figure
+\ref{fig:int}.
+
+```{r fig.cap = "\\label{fig:int} Interaction between executive constraints and GDP per capita with 80 percent CIs"}
 # Interaction plot
-theta_seq <- seq(-2.5, 2.5, .01)
+medians <- apply(theta, 2, quantile, probs = 0.5)
+lower_bound <- min(medians) %>% round(digits = 1)
+upper_bound <- max(medians) %>% round(digits = 1)
+
+theta_seq <- seq(lower_bound, upper_bound, .01)
 state <- quantile(X[, "rgdpepc"], probs = c(0.1, .5, 0.9))
@@
 
-From the results in figure \ref{fig:int} it is clear that the overall
-effect of executive constraints when interacted with GDP per capita on
-the predicted probability of conflict onset is negative. For countries
-with a relatively high level of GDP per capita, going from the worst
-to best in executive constraints decreases the predicted probability
-of conflict onset from [1.5, 4.5] down to below 0.5. This effect
-diminishes, however, for lower values of GDP per capita to the point
-that executive constraints are practically indeterminate to the risk
-of conflict onset for low-income country-years.
+From the listed plots it is clear that for medium to high income
+country-years the overall effect of executive constraints when
+interacted with GDP per capita on the predicted probability of
+conflict onset is negative --- that is to say a more constrained
+executive is associated with a decreased risk of civil conflict onset.
+For example, for country-years with a relatively high level of GDP per
+capita, going from the worst to best in executive constraints
+decreases the predicted probability from [1.5, 4.5] down to below
+0.5. This effect diminishes, however, for lower values of GDP per
+capita to the point that executive constraints are practically
+indeterminate to the risk of conflict onset for country-years situated
+around the 10\% quantile for GDP per capita due to the high degree of
+posterior uncertainty. In the low GDP per capita plot, there is,
+nonetheless, a slight increase in the median predicted probability,
+which begs the question: for GDP per capita values lower than the 10\%
+quantile is there a distinguishable reversal in the relationship
+between executive constraints and civil conflict such that there is
+actually a *positive* increase in the predicted probability for higher
+levels of executive constraints?
+
+```{r fig.cap="\\label{fig:gdp} Difference in predicted probabilities between the lowest and highest level of executive constraints for different values of GDP per capita"}
+rgdpepc <- seq(min(X[, "rgdpepc"]), state[1], 0.01)
+df <- mclapply(rgdpepc, function(y) {
+    low <- boot::inv.logit(fn(lower_bound, y)) * 100
+    high <- boot::inv.logit(fn(upper_bound, y)) * 100
+
+    post_summarise(high - low, probs = c(0.1, 0.5, 0.9)) %>%
+        mutate(y = y,
+               prob = mean(high > low))
+}, mc.cores = 4) %>% bind_rows
+
+df %<>%
+    rename(codelow = `10%`, median = `50%`, codehigh = `90%`) %>%
+    mutate(interval = sprintf("[%.3f, %.3f]",
+                              signif(codelow, 3),
+                              signif(codehigh, 3)))
+
+df$original <- exp(df$y * sd(final.df$rgdpepc) + mean(final.df$rgdpepc))
+
+ggplot(df, aes(original, median)) +
+    geom_line() +
+    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
+    theme_tufte() +
+    xlab("GDP per Capita")
+```
+
+To investigate this possibility, \ref{fig:gdp} plots the pointwise
+difference between the predicted probability of civil conflict given
+the highest level of executive constraints and the predicted
+probability given the lowest level of executive constraints for
+differing values of GDP per capita below the 10\% quantile and down to
+the minimum observed value in the dataset while holding constant the
+remaining covariates in the same fashion as the interaction plots in
+figure \ref{fig:int}.[^theta_bounds] At the far end of the X-axis,
+given a GDP per capita of approximately $1,109$ 2011 US dollars, there
+is an 80 percent posterior probability that the difference lies within
+the interval `r last(df$interval)`. At the opposite end, given a GDP
+per capita of approximately $223$ 2011 US dollars, there is an 80
+percent posterior probability that the difference is between
+`r first(df$interval)`. The lower bounds of the credible intervals remain
+fairly consistent across changing values of GDP per capita while the
+upper bounds increases due to the distributions of predicted
+probabilities being progressively skewed as GDP per capita
+decreases. The high degree of uncertainty precludes the ability to
+conclusively assert that there is a positive relationship between
+civil conflict risk and executive constraints for low income
+country-years. This is further bolstered by the fact that there are
+relatively few observations falling within the extreme tail ends of
+either the executive constraints posterior distribution or the
+empirical distribution of GDP per capita --- in the case of the
+latter, only `r signif(mean(X[, "rgdpepc"] < mean(df$y)), 3) * 100`
+percent of the observations lay below the midpoint of the X-axis at
+600 2011 US dollars.
+
+[^theta_bounds]: Since the executive constraints parameter, $\theta$,
+    is on the unbounded scale, the highest and lowest levels used in
+    the calculations for \ref{fig:gdp} are drawn from the minimum and
+    maximum of the median posterior estimates.
 
 ### Assessing Model Fit
-
-Figure \ref{fig:phat} plots the densities of the first 50 posterior
-draws for the predicted probabilities of conflict onset for two
-groups, country-years that actually experienced an onset of a civil
-conflict and country-years that did not. On average, the predicted
-probabilities for the actual onset country-years are higher albeit
-still extremely low. This can be seen in the summary statistics
-reported in table \ref{tab:psum} for the posterior means of the
-predicted probabilities. For onset country-years the interquartile
-interval is only 0.0475 to 0.2 with a max of 0.516. To better compare
-the predicted probabilities to the actual data, table
-\ref{tab:confusion} presents a confusion matrix with a 0.2 cutoff,
-which approximately corresponds to the third quartile in the summary
-table.  `r signif((93 / (259 + 93)) * 100, 3)` percent of onsets and
-`r signif((7902 / (7902 + 172)) * 100, 3)` percent of non-onsets are
-correctly predicted. The false positive rate is
-`r signif((172 / (7902 + 172)) * 100, 3)` percent and the false
-negative rate is `r signif((259 / (93 + 259)) * 100, 3)` percent.
 
@@
 ```{r psum}
-p_mean <- colMeans(p_hat)
-
-df <- bind_rows(summary(p_mean[final.df$lepisode_onset == 1]),
-                summary(p_mean[final.df$lepisode_onset == 0])) %>%
+p_median <- apply(p_hat, 2, median)
+
+df <- bind_rows(summary(p_median[final.df$lepisode_onset == 1]),
+                summary(p_median[final.df$lepisode_onset == 0])) %>%
     mutate_all(~signif(., 3))
@@
 ```{r confusion}
-m <- table(p_mean > .2, final.df$lepisode_onset)
+m <- table(p_median > .168, final.df$lepisode_onset)
 
@@
 
+tpr <- signif(100 * m["Predicted: Onset", "Actual: Onset"] /
+              sum(m[, "Actual: Onset"]), 3)
+tnr <- signif(100 * m["Predicted: No Onset", "Actual: No Onset"] /
+              sum(m[, "Actual: No Onset"]), 3)
+
+fpr <- signif(100 * m["Predicted: Onset", "Actual: No Onset"] /
+              sum(m[, "Actual: No Onset"]), 3)
+fnr <- signif(100 * m["Predicted: No Onset", "Actual: Onset"] /
+              sum(m[, "Actual: Onset"]), 3)
+
 kable(m, "latex", align = "c", booktabs = T, linesep = "", escape = F,
@@
 
+To assess how well the presented model fits the observed data, I first
+consider the predicted probabilities of conflict onset by plotting the
+first 50 posterior draws for two groups, country-years that actually
+experienced an onset of civil conflict and country-years that did
+not. On average, the predicted probabilities for the actual onset
+country-years are higher albeit still extremely low. This can be seen
+in the summary statistics reported in table \ref{tab:psum} for the
+posterior medians of the predicted probabilities. For onset
+country-years the interquartile interval is only `r df[1, "1st Qu."]`
+to `r df[1, "3rd Qu."]` with a max of `r df[1, "Max."]`. To better
+compare the predicted probabilities to the actual data, table
+\ref{tab:confusion} presents a confusion matrix with a 0.17 cutoff,
+which approximately corresponds to the third quartile in the summary
+table. `r tpr` percent of onsets and `r tnr` percent of non-onsets are
+correctly predicted. The false positive rate is `r fpr` percent and
+the false negative rate is `r fnr` percent.
+
 ```{r results = "hide"}
-x <- evalmod(scores = p_mean, labels = final.df$lepisode_onset)
+x <- evalmod(scores = p_median, labels = final.df$lepisode_onset)
 
@@
 curve depicting the tradeoff between the true positive rate
-(*sensitivity*) and the false positive rate --- $1 -
-\mbox{specificity}$ where *specificity* refers to the true negative
-rate --- for varying probability thresholds according to the posterior
-means of the predicted probabilities. For example, for a threshold of
-approximately 0.1, the model will correctly predict 50 percent of
-onsets while misclassifying only 8 percent of non-onsets. The total
-performance of the model can be summarized by the area under the curve
-(AUC), which in this case is `r auc`.[^auc] However, for the present
-analysis the results from the ROC curve are misleading since there is
-a significant class imbalance in the dataset. Onsets comprise only 4
-percent of all observations meaning that the false positive rate is
-less sensitive to changing thresholds. Figure \ref{fig:roc} therefore
-also plots a precision-recall (PRC) curve, which depicts the tradeoff
-between *precision* --- the number of true positives divided by the
-sum of the true positives and false positives --- and *recall* --- the
-number of true positives divided by the sum of the true positives and
-false negatives --- for varying thresholds. The advantage is that
+(*sensitivity*) and the false positive rate --- denoted by the formula
+$1 - \mbox{specificity}$, where *specificity* refers to the true
+negative rate --- for varying probability thresholds according to the
+posterior medians of the predicted probabilities. For example, given a
+threshold of approximately 0.085, the model will correctly predict
+roughly 50 percent of onsets while misclassifying only 8 percent of
+non-onsets. The total performance of the model can be summarized by
+the area under the curve (AUC), which in this case is `r auc`.[^auc]
+However, for the present analysis the results from the ROC curve are
+misleading since there is a significant class imbalance in the
+dataset. Onsets comprise only
+`r signif(100 * mean(final.df$lepisode_onset == 1), 3)` percent of all
+observations meaning that the false positive rate is less sensitive to
+changing thresholds. Figure \ref{fig:roc} therefore also plots a
+precision-recall (PRC) curve, which depicts the tradeoff between
+*precision* --- the number of true positives divided by the sum of the
+true positives and false positives --- and *recall* --- the number of
+true positives divided by the sum of the true positives and false
+negatives --- for varying thresholds. The advantage is that
 *precision* is invariant to the number of non-onset years in the
 dataset. From this plot it is clear that the precision of the model,
-the ability to discriminate between onset country-years and false
-predicted non-set country-years, drops drastically as the *recall* of
-the model increases. The *AUC-PR* for this plot is only `r auc_pr`.
-[^auc_pr]
+the ability to discriminate between onset country-years and falsely
+predicted non-onset country-years, drops drastically as the *recall* of
+the model increases. The *AUC-PR* for this plot is only
+`r auc_pr`.[^auc_pr]
 
@@
 relatively low cutoff, the model is still only able to accurately
-predict 25 percent of conflict onsets. According to the ROC curve,
-decreasing this threshold improves the true positive rate with a
-relatively decent tradeoff in terms of the false positive rate, but as
-stated, this is largely due to the imbalance in the outcome
+predict `r tpr` percent of conflict onsets. According to the ROC
+curve, decreasing this threshold improves the true positive rate with
+a relatively decent tradeoff in terms of the false positive rate, but
+as stated, this is largely due to the imbalance in the outcome
 variable. In terms of *precision*, the performance of the model
@@
 
-```{stan, code = readLines("stan/model.stan"), eval = F, output.var = "x", echo = T}
+```{stan, code = readLines("stan/sem.stan"), eval = F, output.var = "x", echo = T}
 ```

### 2020-05-26 paper.Rmd
Fix latex compilation of manuscript
Rownames were not being added anymore causing the table to fail. I
suspect the latest minor dplyr update from the last image rebuild.

@@
 
-rownames(fa_pars) <- sprintf("\\textit{%s}", fa_pars$manifest)
-fa_pars$manifest <- NULL
-
-kable(fa_pars, "latex", align = "c", booktabs = T, linesep = "", escape = F,
-      caption = "Posterior credible intervals for measurement model parameters") %>%
+# I absolutely hate tibbles
+fa.mm <- select(fa_pars, -manifest) %>% as.matrix
+rownames(fa.mm) <-sprintf("\\textit{%s}", fa_pars$manifest)
+
+kable(fa.mm, "latex", align = "c", booktabs = T, linesep = "", escape = F,
     kable_styling(latex_options = c("striped", "scale_down"), full_width = F,

### 2020-05-26 paper.Rmd
Add text references to manuscript

@@
         pandoc_args: ["--syntax-definition", "assets/stan.xml",
-                      "--highlight-style", "kate"]
+                      "--highlight-style", "kate",
+                      "--lua-filter", "utils/text_refs/text_refs.lua"]
     html_document:
@@
 constraints, I construct a dataset of relevant variables and run a
-bayesian structural equation model with a measurement portion estimating
-executive constraints as a latent construct and a multilevel logistic
-regression relating the latent construct to the observed outcome,
-civil conflict.
+bayesian latent variable model with a measurement portion
+estimating executive constraints as a latent construct and a
+multilevel logistic regression relating the latent construct to the
+observed outcome, civil conflict.
 
@@
 [^model]: The model code is written in Stan [@Carpenter2017] and was
-    run using the `CmdStan` interface version 2.22.1 [@Stan2020] with
+    run using the `CmdStan` interface version 2.23.0 [@Stan2020] with
     four parallel chains at 2000 iterations each with a warmup
@@
 
-```{r fig.cap = "\\label{fig:onsets} Proportion of countries experiencing ongoing civil conflicts and new onsets"}
+(ref:onsets) Proportions of countries per year experiencing at least
+    one ongoing civil conflict and at least one new onset.
+
+```{r fig.cap = "\\label{fig:onsets} (ref:onsets)"}
 
@@
 
-For the purposes of this paper I estimate an index measuring
-executive constraints from a factor analysis of multiple manifest
-variables drawn from the V-Dem dataset. The aim is to capture the
-extent to which the executive is *de facto* constrained by other
-government institutions --- more specifically, the legislature,
-judiciary, and independent investigative bodies --- rather than the
-*de jure* powers of any one institution *vis-à-vis* the executive, and
-to do so using a latent variable aggregation that avoids the arbitrary
-weighting of linear additive indices.[^additive] The aggregation
-method is largely similar to the accountability index created by
-@Luhrmann2017 and the presidentialism index created by @Sigman2018. By
-creating a separate index, however, I am able to adjust the list of
-included manifest variables as well as embed the factor analysis
-within the larger regression model such that the uncertainty from the
-latent factor propagates as a result of the estimation process.
+For the purposes of this paper I estimate an index measuring executive
+constraints from a factor analysis of multiple manifest variables
+drawn from the V-Dem dataset. The aim is to capture the extent to
+which the executive is *de facto* constrained by other government
+institutions --- more specifically, the legislature, judiciary, and
+independent investigative bodies --- rather than the *de jure* powers
+of any one institution *vis-à-vis* the executive, and to do so using a
+latent variable aggregation that avoids the arbitrary weighting of
+linear additive indices.[^additive] The aggregation method is largely
+similar to the accountability index created by @Luhrmann2017 and the
+presidentialism index created by @Sigman2018 already available in the
+V-Dem dataset. By creating a separate index, however, I am able to
+adjust the list of included manifest variables as well as embed the
+factor analysis within the larger regression model such that the
+uncertainty from the latent factor propagates as a result of the
+estimation process.
 
@@
 
+(ref:fa-sum) Descriptive statistics for manifest variables
+
 ```{r fa-sum}
@@
 kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = F,
-      caption = "Descriptive statistics for manifest variables") %>%
+      caption = "(ref:fa-sum)") %>%
     kable_styling(latex_options = "striped")
@@
 
-```{r fig.height = 5, fig.width = 7, fig.cap = "\\label{fig:cor} Correlation Heatmap"}
+(ref:cor) Correlation heatmap of manifest variables
+
+```{r fig.height = 5, fig.width = 7, fig.cap = "\\label{fig:cor} (ref:cor)"}
 m <- cor(manifest, use = "pairwise.complete.obs") %>%
@@
 
+(ref:reg-stats) Descriptive statistics for regression covariates
+
 ```{r reg-stats}
@@
       booktabs = T, linesep = "", escape = T,
-      caption = "Descriptive statistics for regression covariates") %>%
+      caption = "(ref:reg-stats)") %>%
     kable_styling(latex_options = "striped")
@@
 
-```{r fig.height = 4, fig.cap = "\\label{fig:fpars} Measurement Model Parameter Estimates"}
+(ref:fpars) Executive constraints measurement model parameter
+    estimates with 80 percent posterior credible intervals.
+
+```{r fig.height = 4, fig.cap = "\\label{fig:fpars} (ref:fpars)"}
 load("./posteriors/full_model/input.RData")
@@
 
+(ref:fa)
+
+Executive constraints measurement model parameter
+    estimates with 80 percent posterior credible intervals
+
 ```{r fa, fig.height = 6, fig.width = 6}
@@
 kable(fa.mm, "latex", align = "c", booktabs = T, linesep = "", escape = F,
+      caption = "(ref:fa)") %>%
     kable_styling(latex_options = c("striped", "scale_down"), full_width = F,
@@
 
-```{r fig.width = 6, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for executive constraints"}
+(ref:bicam) Posterior density plots for the executive constraints
+    measure, $\theta$.
+
+```{r fig.width = 6, fig.height = 2, fig.cap = "\\label{fig:bicam} (ref:bicam)"}
 theta <- extract(fa, "theta")
@@
 
-```{r fig.cap = "\\label{fig:fa-line} Posterior shrinkage for \\textit{v2juncind}"}
+(ref:fa-line) Posterior shrinkage for *v2juncind*. The observed point
+    estimates are the data points reported by V-Dem.
+
+```{r fig.cap = "\\label{fig:fa-line} (ref:fa-line)"}
 err <- read_post("posteriors/full_model/err_posteriors.csv")
@@
 
-```{r fig.cap = "\\label{fig:err} Posterior CI and point estimates vs. observed error"}
+(ref:err) On the left, the posterior uncertainty for *v2juncind* as
+    estimated by the model plotted against the observed error reported
+    by V-Dem as *v2juncind_sd*. On the right, the difference between
+    the point estimates estimated by the model and the point estimates
+    reported by V-Dem versus the observed error for *v2juncind*.
+
+```{r fig.cap = "\\label{fig:err} (ref:err)"}
 p1 <- ggplot(nonlg_est, aes(se, interval)) +
@@
 
-```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Executive Constraints Over Time"}
+(ref:theta) Executive constraints over time for several selected
+    countries.
+
+```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} (ref:theta)"}
 theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
@@
 
-```{r results = "hide", fig.width = 8, fig.cap = "\\label{fig:map} Executive Constraints - 2012. Grey indicates countries missing from the dataset."}
+(ref:map) Executive constraints in 2012 for all countries present in
+    the dataset. Missingness indicated by the color grey.
+
+```{r results = "hide", fig.width = 8, fig.cap = "\\label{fig:map} (ref:map)"}
 map_data <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
@@
 
-```{r fig.cap = "\\label{fig:tie} Executive Constraints 2000 vs 2010"}
+(ref:cap) Executive constraints in 2000 versus 2010 for several
+    selected countries.
+
+```{r fig.cap = "\\label{fig:tie} (ref:cap)"}
 targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
@@
 
+(ref:reg) 80 percent posterior credible intervals for regression
+    slopes
+
 ```{r reg, fig.height = 3}
@@
       linesep = "", escape = F,
-      caption = "Posterior credible intervals for regression slopes") %>%
+      caption = "(ref:reg)") %>%
     kable_styling(latex_options = "striped", "scale_down")
@@
 
-```{r fig.cap = "\\label{fig:reg} Regression parameters"}
+(ref:fig-reg) Regression slopes with 80 percent posterior credible
+    intervals.
+
+```{r fig.cap = "\\label{fig:reg} (ref:fig-reg)"}
 plot_pars(beta, hline = T)
@@
 
-```{r fig.cap = "\\label{fig:year} Log-odds year intercepts - $80\\%$ CIs"}
+(ref:year) Log-odds year intercepts with 80 percent credible intervals
+
+```{r fig.cap = "\\label{fig:year} (ref:year)"}
 Z_year <- extract(reg, "Z_year")
@@
 
-```{r, fig.cap = "\\label{fig:country} Log-odds country intercepts - $80\\%$ CIs"}
+(ref:country) Log-odds country intercepts with 80 percent credible
+    intervals
+
+```{r, fig.cap = "\\label{fig:country} (ref:country)"}
 
@@
 
-```{r, fig.cap = "\\label{fig:peace} Log-odds peace year intercepts - $80\\%$ CIs"}
+(ref:peace) Log-odds peace year intercepts with 80 percent credible
+    intervals
+
+```{r, fig.cap = "\\label{fig:peace} (ref:peace)"}
 f <- extract(reg, "f")
@@
 
-```{r fig.cap = "\\label{fig:int} Interaction between executive constraints and GDP per capita with 80 percent CIs"}
+(ref:int) Interaction effect between executive constraints and GDP per
+    capita with 80 percent credible intervals.
+
+```{r fig.cap = "\\label{fig:int} (ref:int)"}
 # Interaction plot
@@
 
-```{r fig.cap="\\label{fig:gdp} Difference in predicted probabilities between the lowest and highest level of executive constraints for different values of GDP per capita"}
+(ref:gdp) Difference in predicted probabilities between the lowest and
+    highest level of executive constraints for different values of GDP
+    per capita.
+
+```{r fig.cap="\\label{fig:gdp} (ref:gdp)"}
 rgdpepc <- seq(min(X[, "rgdpepc"]), state[1], 0.01)
@@
 
-```{r fig.height = 2, fig.cap = "\\label{fig:phat} Density plot for predicted probabilities"}
+(ref:phat) Density plot of predicted probabilities for actual onset
+    country-years and no onset country-years.
+
+```{r fig.height = 2, fig.cap = "\\label{fig:phat} (ref:phat)"}
 extra <- read_post("posteriors/full_model/extra_posteriors.csv")
@@
 
+(ref:psum) Summarized posterior predicted probabilities
+
 ```{r psum}
@@
 kable(df, "latex", align = "c", booktabs = T, linesep = "", escape = F,
-      caption = "Summarized posterior predicted probabilities")
-```
+      caption = "(ref:psum)")
+```
+
+(ref:confusion) Confusion matrix with a 0.168 cutoff
 
@@
 kable(m, "latex", align = "c", booktabs = T, linesep = "", escape = F,
-      caption = "Confusion matrix with a 0.2 cutoff")
+      caption = "(ref:confusion)")
 ```
@@
 
-```{r fig.width = 8, fig.cap = "\\label{fig:roc} ROC & Precision-Recall Plots"}
-
+(ref:roc) ROC & Precision-Recall plots
+
+```{r fig.width = 8, fig.cap = "\\label{fig:roc} (ref:roc)"}
 autoplot(x, curvetype = c("ROC", "PRC"))

### 2020-07-03 paper.Rmd
Draft edits

@@
 Since the end of World War II, civil conflict has been the deadliest
-and most prevalent form of conflict. In that time period, upwards to
-two-thirds of all countries have experienced at least one minor
-insurgency and approximately one-third of all countries have
+and most prevalent form of state-based conflict. In that time period,
+upwards to two-thirds of all countries have experienced at least one
+minor insurgency and approximately one-third of all countries have
 experienced a major civil war with at least 1,000 battle deaths
@@
 motivation-based arguments --- such as, for example, the role of
-horizontal inequalities [@Cederman2011] --- to opportunity-based
-explanations centered on economic factors [@Collier2004; @Collier2009]
-as well as the political and structural conditions underlying state
-weakness [@Hegre2001; @Fearon2003]. This latter strain of research in
-particular has spurred extensive work into examining the
-distinguishing features of weak states and whether political
-institutions play a significant role in the risk of civil conflict.
-Several studies in this vein have argued in favor of institutions by
-utilizing disaggregated measures to also investigate the specific
-types of institutions relevant to the study of civil conflict
-[@Taydas2010; @Hegre2015a; @Bartusevic2018].
+horizontal inequalities and ethnic identities [@Cederman2011] --- to
+opportunity-based explanations centered on economic factors
+[@Collier2004; @Collier2009] as well as the political and structural
+conditions underlying state weakness [@Hegre2001; @Fearon2003]. This
+latter strain of research in particular has spurred extensive work
+into examining the distinguishing features of weak states and whether
+political institutions play a significant role in the risk of civil
+conflict. Several studies in this vein have argued in favor of
+institutions by investigating, for example, aspects of governance,
+institutional quality, and opportunities for political participation
+[@Fjelde2009; @Taydas2010; @Hegre2015a; @Bartusevic2018].
 
 This paper extends that work by taking a disaggregated approach to
-studying the effect of executive constraints on the risk of civil
-conflict onset. The motivation to specifically focus on executive
-constraints is two-fold: first, the executive is often times one of
-the key decision-makers in a state's regime. Institutions that provide
-a check against the types of actions which may exacerbate the risk of
-rebellion should be relevant to the study of civil conflict. In doing
-we can further our understanding of how political institutions shape
-the incentives for violence, especially through the influence of
-institutions on state behaviour.  Second, the link between executive
-constraints and civil conflict is relatively understudied. The
-previous literature that exists is fairly limited and contradictory in
-terms of results [@Getmansky2013; @Fjelde2009].
-
-The principle finding is that there is tentative evidence that
-executive constraints do matter and that there is a negative
-relationship between the latent level of constraints on the executive
-and the risk of civil conflict onset. However, this relationship
-largely disappears for low-income country-years when an interaction
-term with GDP per capita is included. If the primary theoretical
-mechanism involves the impact of executive constraints on rebel
-mobilization --- an unchecked executive may be more likely to engage
-in indiscriminate violence which could lead to a loss of legitimacy as
-well as the ability of rebel groups to offer local security guarantees
-to attract recruits --- then the incentives for violence in low-income
-country-years may be so great that it does not matter how constrained
-the executive may or may not be.
+studying political institutions by focusing specifically on the effect
+of executive constraints on the risk of civil conflict onset. The
+motivation to focus on executive constraints is two-fold: first, the
+executive is often times one of the key decision-makers in a state's
+regime. Institutions that provide a check against the types of actions
+which may exacerbate the risk of rebellion should be relevant to the
+study of civil conflict. Such an approach can further our
+understanding of how specific political institutions may shape the
+incentives for violence, especially through the influence on state
+behaviour. Second, the link between executive constraints is
+relatively understudied. The previous literature that exists is fairly
+limited and contradictory in terms of results.
+
+The principle finding of this paper is that there is tentative
+evidence that executive constraints do matter and that there is a
+negative relationship between the latent level of constraints on the
+executive and the risk of civil conflict onset. However, this
+relationship largely disappears for low-income country-years when an
+interaction term with GDP per capita is included. If the primary
+theoretical mechanism involves the impact of executive constraints on
+rebel mobilization --- an unchecked executive may be more likely to
+engage in indiscriminate violence which could lead to a loss of
+legitimacy as well as the ability of rebel groups to offer local
+security guarantees to attract recruits --- then the incentives for
+violence in low-income country-years may be so great that it does not
+matter how constrained the executive may or may not be.
 
 We should not be overconfident with these results, however. Similar to
-most other studies examining conflict onset, the model is a poor fit
-to the underlying data. Modeling civil conflict is a difficult task,
-especially when trying to capture the dynamics of conflict with a
-standard time series cross-country regression. The results will
-require verification by additional studies using alternative
-methodological approaches.
+most other studies examining conflict onset, the presented models are
+a poor fit to the underlying data. Modeling civil conflict is a
+difficult task, especially when trying to capture the dynamics of
+conflict with a standard time series cross-country regression. The
+results will require verification by additional studies using
+alternative methodological approaches.
 
@@
 mechanisms linking executive constraints to conflict onset will then
-be discussed followed by an empirical analysis for which a latent
-estimate for executive constraints will be estimated and used in a
-cross-country time series logistic regression predicting conflict
-onset.
+be discussed followed by an empirical analysis using a bayesian latent
+variable model to estimate executive constraints from the basis of
+several observed manifest variables. This latent measure will then be
+used to predict civil conflict onset using a cross-country time series
+logistic regression.
 
@@
 
-Unsurprisingly, insofar as the role of institutions has been examined
-by previous literature the focus has largely been on the broader
-notion of democracy, which at a minimum is a system regulating access
-to political power through popular representation determined by free
-and competitive elections [@Schumpeter1962]. Democracy serves as a
-proxy for a number of normatively desirable institutions and has been
-hypothesised to have a pacifying domestic effect according to two
-predominant mechanisms. The first is centered on the concept of
-relative deprivation, or inequalities in value expectations and value
-capabilities, as a primary source of discontent that can motivate
-collective violence [@Gurr1970]. [^gurr] Democracy is thought to
-decrease the amount of political discontent by providing avenues for
-free expression and adjudicating group conflicts through non-violent
-institutional means --- for example, through the regular occurrence of
-free and fair elections [@Hegre2014]. The second mechanism assumes
-that violent conflict is costly and that there exists a bargained
-solution preferable to fighting, assuming that actors are rational and
-able to internalize the cost of violence [@Fearon1995]. One of the
-reasons why a bargained solution may be impossible is the risk of
-future shifts in the distribution of power. More specifically, a
-future change in favor of the relative strength of one of the parties
-can provide an incentive to later renege. This influences the present
-degree to which concessions can be offered and parties are willing to
-credibly commit [@Powell2006]. Within the domestic sphere, democracy
-offers a possible solution by formalising the regular contest of power
-with full enfranchisement, thereby providing a *de jure* guarantee for
-the balance of power [@Origins2006], and should, likewise, decrease
-the likelihood of civil conflict.
+Unsurprisingly, insofar as the role of institutions on the likelihood
+of civil conflict has been examined by previous literature the focus
+has largely been on the broader notion of democracy, which at a
+minimum is a system regulating access to political power through
+popular representation determined by free and competitive elections
+[@Schumpeter1962]. Democracy serves as a proxy for a number of
+normatively desirable institutions and has been hypothesised to have a
+pacifying domestic effect according to two predominant mechanisms. The
+first is centered on the concept of relative deprivation, or
+inequalities in value expectations and value capabilities, as a
+primary source of discontent that can motivate collective violence
+[@Gurr1970]. [^gurr] Democracy is thought to decrease the amount of
+political discontent by providing avenues for free expression and
+adjudicating group conflicts through non-violent institutional means
+--- for example, through the regular occurrence of free and fair
+elections [@Hegre2014]. The second mechanism assumes that violent
+conflict is costly and that there exists a bargained solution
+preferable to fighting, assuming that actors are rational and able to
+internalize the cost of violence [@Fearon1995]. One of the reasons why
+a bargained solution may be impossible is the risk of future shifts in
+the distribution of power. More specifically, a future change in favor
+of the relative strength of one of the parties can provide an
+incentive to later renege. This influences the present degree to which
+concessions can be offered and parties are willing to credibly commit
+[@Powell2006]. Within the domestic sphere, democracy offers a possible
+solution by formalising the regular contest of power with full
+enfranchisement, thereby providing a *de jure* guarantee for the
+balance of power [@Origins2006], and should, likewise, decrease the
+likelihood of civil conflict.
 
@@
 are still ubiquitous enough to be a poor explanation for why conflict
-breaks out in only a relatively small subset of countries. Instead,
-the determining factors for conflict onset are the structural
-conditions conducive to rebellion.[^opportunities]
+breaks out in only a relatively small subset of countries at any given
+point in time. Instead, the determining factors for conflict onset are
+the structural conditions conducive to rebellion.[^opportunities]
 
@@
     the proposition, "where a rebellion is financially and militarily
-    feasible it will occur." Both approaches, however, reach the same
-    conclusion concerning the empirical evidence surrounding the role
-    of democracy.
+    feasible it will occur" [@Collier2009]. Both approaches, however,
+    reach the same conclusion concerning the empirical evidence
+    surrounding the role of democracy.
 
@@
 may have been hamstrung by aggregate measures which poorly reflect the
-within-group variation of political regimes. In a larger sample
-including democratic polities, @Goldstone2010 create a regime
-classification scheme based on measures for openness of executive
-recruitment and competitiveness of political participation, as well as
-factionalism in order to distinguish between partial democracies. They
-offer a theory of civil conflict occurring in countries where elite
-divisions undermine the capacity of the state to defeat an insurgency
-and claim that a relatively parsimonious model including their regime
-classification offers better predictive accuracy than previous
-analyses and provides evidence that political institutions are a
-dominant driving factor for both political instability and civil
-conflict. Finally, @Bartusevic2018 construct a regime typology focused
-specifically on electoral features, which are argued form the
-essential core of the concept of democracy. Where there are
-constraints on electoral contestation, the authors contend that such
-constraints restrict the opportunities for non-violent political
-participation thereby legitimising anti-government violence and
-self-selecting recruits into rebel groups willing to engage in armed
-conflict. Evidence is presented which the authors claim supports not
-just the significance of regime type, but a negative monotonic
+within-group variation of political regimes.
+
+In a larger sample including democratic polities, @Goldstone2010
+create a regime classification scheme based on measures for openness
+of executive recruitment and competitiveness of political
+participation, as well as factionalism in order to distinguish between
+partial democracies. They offer a theory of civil conflict occurring
+in countries where elite divisions undermine the capacity of the state
+to defeat an insurgency and claim that a relatively parsimonious model
+containing their regime classification offers better predictive
+accuracy than previous analyses and provides evidence that political
+institutions are a dominant driving factor for both political
+instability and civil conflict. Finally, @Bartusevic2018 construct a
+regime typology focused specifically on electoral features, which are
+argued form the essential core of the concept of democracy. Where
+there are constraints on electoral contestation, the authors contend
+that such constraints restrict the opportunities for non-violent
+political participation thereby legitimising anti-government violence
+and self-selecting recruits into rebel groups willing to engage in
+armed conflict. Evidence is presented which the authors claim supports
+not just the significance of regime type, but a negative monotonic
 relationship between democracy and the risk of civil conflict.
@@
 democratic institutions and find both important for reducing the
-incidence of civil conflict, or in this specific case the recurrence
-of conflict.
+incidence of recurring civil conflict.
 
@@
 
-[^ODonnell]: Adapted from the definition of horizontal accountability
-    offered by @ODonnell1998 who further argues that horizontal
-    accountability can potentially be a distinguishing feature between
-    consolidated democracies and newly democratic states, the latter
-    of which may have free and open elections, but are considerably
-    weaker in upholding liberal notions of individual freedom against
-    state power and republican notions of representation. In this way,
-    executive constraints can possibly aid in discriminating the types
-    of partially democratized regimes which may be at a heightened
-    risk for civil conflict.
+[^ODonnell]: This is adapted from the definition of horizontal
+    accountability offered by @ODonnell1998 who further argues that
+    horizontal accountability can potentially be a distinguishing
+    feature between consolidated democracies and newly democratic
+    states, the latter of which may have free and open elections, but
+    are considerably weaker in upholding liberal notions of individual
+    freedom against state power and republican notions of
+    representation. In this way, executive constraints can possibly
+    aid in discriminating the types of partially democratized regimes
+    which may be at a heightened risk for civil conflict.
 
@@
 constraints should be associated with a lowered risk of civil conflict
-onset. This proposition can be further extended if state weakness is a
+onset. This proposition can be further extended: if state weakness is a
 primary driver of civil conflict and the relative capacity of the
@@
 the prior probability mass to a *length-scale* between 4 and 10, which
-provides a strong smoothing effect and aids in computation by
-avoiding unreasonable values [@Betancourt2017].
+provides a strong smoothing effect and aids in computation by avoiding
+unreasonably large values far beyond the observed range
+[@Betancourt2017].
 \begin{align*}
@@
     for a given variable is mapped on to the ordinal answer categories
-    presented on the survey. Due to differential item functioning,
+    presented in the survey. Due to differential item functioning,
     experts may perceive the magnitude of difference between answer
@@
 
-In the case of the legislative variables there is the added
-issue of missing observations for country-years lacking a
+In the case of the legislative variables there is the added issue of
+missing observations for country-years lacking a
 legislature.[^missing] For those observations, given the centrality of
@@
 manifest variables. The resulting model specification, including
-priors, is as follows:
+priors, can be summarised as follows:
 \begin{align*}
@@
 include a dummy variably for ongoing high intensity conflict since
-devastating civil ways may have a particularly destabilizing
+devastating civil wars may have a particularly destabilizing
 effect.[^intensity] To account for spillover effects, I include a
@@
 Table \ref{tab:fa} and figure \ref{fig:fpars} summarise the posterior
-draws for the measurement model parameters with 80 percent quantile
-intervals. According to the results, the factor loadings, $\lambda$,
-for each manifest variable are fairly comparable, with the exception
-of the lower value for *v2lgfunds* ("legislature controls its own
-resources") and to a lesser extent *v2lgqstexp* ("legislature
+draws for the measurement model parameters with 80 percent posterior
+quantile intervals. According to the results, the factor loadings,
+$\lambda$, for each manifest variable are fairly comparable, with the
+exception of the lower value for *v2lgfunds* ("legislature controls
+its own resources") and to a lesser extent *v2lgqstexp* ("legislature
 questions executive branch officials in practice"). Relative to the
@@
 
-(ref:fa)
-
-Executive constraints measurement model parameter
-    estimates with 80 percent posterior credible intervals
+(ref:fa) Executive constraints measurement model parameter
+    estimates with 80 percent posterior credible intervals.
 
@@
 In total there are `r nrow(final.df) %>% prettyNum(big.mark = ",")`
-country-year estimates of $\theta$, comprised of `r
-n_distinct(final.df$country_name)` unique countries and scaled from
+country-year estimates of $\theta$, comprised of
+`r n_distinct(final.df$country_name)` unique countries and scaled from
 approximately $-2$ to $2$. Since the complete time series for every
@@
 
-```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} (ref:theta)"}
+```{r fig.width = 6, fig.height = 4, fig.cap = "\\label{fig:theta} (ref:theta)"}
 theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
@@
 
-(ref:cap) Executive constraints in 2000 versus 2010 for several
+(ref:tie) Executive constraints in 2000 versus 2010 for several
     selected countries.
 
-```{r fig.cap = "\\label{fig:tie} (ref:cap)"}
+```{r fig.cap = "\\label{fig:tie} (ref:tie)"}
 targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
@@
 
-levels <- filter(df, year == 2010) %>%
-    arrange(median) %$%
-    country_name
-
-df$country_name <- factor(df$country_name, levels = levels)
-
-ggplot(df, aes(country_name, median, colour = year)) +
-    geom_errorbar(aes(ymax = codehigh, ymin = codelow),
-                  position = position_dodge(width = 0.5)) +
-    theme_tufte() +
-    scale_color_colorblind() +
-    theme(axis.title.y = element_blank(),
-          legend.title = element_blank()) +
-    ylab("Theta") +
-    coord_flip()
-
+tfplot(df) + xlab("Theta")
 ```
@@
 
-(ref:reg) 80 percent posterior credible intervals for regression
-    slopes
-
-```{r reg, fig.height = 3}
+(ref:reg) Regression slopes with 80 percent posterior credible
+    intervals and the posterior probability of each parameter being
+    less than and greater than zero.
+
+```{r reg}
 reg <- read_post("posteriors/full_model/reg_posteriors.csv")
@@
 kable(beta_s, "latex", align = c("c", "r", "r"), booktabs = T,
-      linesep = "", escape = F,
-      caption = "(ref:reg)") %>%
+      linesep = "", escape = F, caption = "(ref:reg)") %>%
     kable_styling(latex_options = "striped", "scale_down")

### 2020-08-08 paper.Rmd
Switch to Chicago style citations and add proper title page

@@
 title: Executive Constraints and Civil Conflict Onset
-author: J.L.A. Krusell
-date: "`r format(Sys.time(), '%Y-%m-%d')`"
+author:
+    name: J.L.A. Krusell
+    affiliation: Department of Peace and Conflict Research, Uppsala University
+    supervisor: Kristine Eck
+    degree: Master's Thesis
+date: "Spring 2020"
+titlelogo: |
+    ```{=latex}
+    assets/UU_logo_CMYK-eps-converted-to.pdf
+    ```
+wordcount: 13052
 suppress-bibliography: false
 bibliography: library.bib
-csl: assets/american-political-science-association.csl
+csl: assets/chicago-author-date.csl
 link-citations: true
 documentclass: scrreprt
+classoption: paper=letter
+geometry:
+    - top=30mm
+    - heightrounded
 fontsize: 11pt
@@
         toc_depth: 2
-        pandoc_args: ["--syntax-definition", "assets/stan.xml",
+        pandoc_args: ["--template", "assets/thesis_template.latex",
+                      "--syntax-definition", "assets/stan.xml",
                       "--highlight-style", "kate",

### 2020-08-24 paper.Rmd
Manuscript edits

--- 2020-07-03 paper.Rmd
@@
 title: Executive Constraints and Civil Conflict Onset
-author: J.L.A. Krusell
-date: "`r format(Sys.time(), '%Y-%m-%d')`"
+author:
+    name: J.L.A. Krusell
+    affiliation: Department of Peace and Conflict Research
+    supervisor: Kristine Eck
+    degree: "Master's Thesis"
+date: "Spring 2020"
+titlelogo: |
+    ```{=latex}
+    assets/UU_logo_CMYK-eps-converted-to.pdf
+    ```
+wordcount: 13052
 suppress-bibliography: false
 bibliography: library.bib
-csl: assets/american-political-science-association.csl
+csl: assets/chicago-author-date.csl
 link-citations: true
 documentclass: scrreprt
-fontsize: 11pt
+classoption: paper=letter
+geometry:
+    - top=30mm
+    - heightrounded
+fontsize: 12pt
 linestretch: 1.5
@@
     - \counterwithout{figure}{chapter}
+    - \usepackage[font={small}]{caption}
 output:
@@
 options(knitr.kable.NA = ".")
-opts_chunk$set(echo = F, eval = T, warning = F, fig.width = 5,
-               fig.height = 3)
+opts_chunk$set(echo = F, message = F, warning = F,
+               fig.width = 5, fig.height = 3, cache = T)
+
+# TODO: this has been added to package
+fmt <- function(x, digits = 2) {
+    prettyNum(round(x, digits), big.mark = ",")
+}
+
+# Create initial descriptive stats for introduction
+ucdp <- readRDS("./data/raw/UcdpPrioConflict_v19_1.rds")
+
+total_conflicts <- n_distinct(ucdp$conflict_id)
+intrastate_conflicts <- filter(ucdp, type_of_conflict %in% 3:4) %$%
+    n_distinct(conflict_id)
+intrastate_percent <- 100 * intrastate_conflicts / total_conflicts
+
+n_states <- filter(ucdp, type_of_conflict %in% 3:4) %$%
+    n_distinct(gwno_a)
+
+battle_deaths <- readRDS("./data/raw/BattleDeaths_v19_1.rds")
+nonstate_deaths <- readRDS("./data/raw/NonState_v19_1.rds")
+
+total_deaths <- sum(battle_deaths$bd_best) + sum(nonstate_deaths$best_fatality_estimate)
+civil_deaths <- filter(battle_deaths, type_of_conflict %in% 3:4) %$%
+    sum(bd_best)
+
+civil_deaths_percent <- 100 * civil_deaths / total_deaths
 ```
@@
 
-Since the end of World War II, civil conflict has been the deadliest
-and most prevalent form of state-based conflict. In that time period,
-upwards to two-thirds of all countries have experienced at least one
-minor insurgency and approximately one-third of all countries have
-experienced a major civil war with at least 1,000 battle deaths
-[@Pettersson2019]. Given the impact of civil conflict and the
-increasing availability of conflict related data, the last several
-decades has witnessed a veritable explosion in the number of
-quantitative studies investigating the causes of civil conflict
-[@Blattman2017]. Scholars have proposed numerous theories ranging from
-motivation-based arguments --- such as, for example, the role of
-horizontal inequalities and ethnic identities [@Cederman2011] --- to
+Since the end of World War II, civil conflict has been the most
+prevalent form of state-based conflict. Of the
+`r fmt(total_conflicts)` conflicts recorded by the UCDP/PRIO Armed
+Conflict Database v19.1 [@Gleditsch2002; @Pettersson2019] for the
+years covering 1946 until 2018, `r fmt(intrastate_conflicts)` of
+those, or approximately `r fmt(intrastate_percent, 1)` percent, are
+classified as intrastate. These civil conflicts are distributed across
+`r fmt(n_states)` countries, more than half of which having
+experienced at least one major civil war with 1,000 battle deaths at a
+minimum.  Where data is available, civil conflicts are also
+empirically among the deadliest. Out of `r fmt(total_deaths)`
+cumulative battles deaths from organized violence recorded
+since 1989 [@Pettersson2019; @Eck2007; @Sundberg2012],
+`r fmt(civil_deaths)`, or `r fmt(civil_deaths_percent, 1)` percent, are
+the result of civil conflict.
+
+Given the impact of civil conflict and the increasing availability of
+conflict related data, the last several decades has witnessed a
+veritable explosion in the number of quantitative studies
+investigating the causes of civil conflict [@Blattman2017]. Scholars
+have proposed numerous theories ranging from motivation-based
+arguments --- such as, for example, the role of horizontal
+inequalities and ethnic identities [@Cederman2011] --- to
 opportunity-based explanations centered on economic factors
@@
 political institutions play a significant role in the risk of civil
-conflict. Several studies in this vein have argued in favor of
-institutions by investigating, for example, aspects of governance,
-institutional quality, and opportunities for political participation
-[@Fjelde2009; @Taydas2010; @Hegre2015a; @Bartusevic2018].
+conflict. Several studies have argued in favor of institutions by
+investigating, for example, aspects of governance, institutional
+quality, and opportunities for political participation [@Fjelde2009;
+@Taydas2010; @Hegre2015a; @Bartusevic2018].
 
@@
 understanding of how specific political institutions may shape the
-incentives for violence, especially through the influence on state
-behaviour. Second, the link between executive constraints is
-relatively understudied. The previous literature that exists is fairly
-limited and contradictory in terms of results.
+incentives for violence by affecting state behaviour. Second, the link
+between executive constraints is relatively understudied. The previous
+literature that exists is fairly limited and contradictory in terms of
+results.
 
@@
 co-opt potential opposition. Although the preceding examples are far
-from forming an exhaustive list, there are thus a number of different
-ways in which institutions can enter into theoretical discussions
-about civil conflict.
+from forming an exhaustive list, they illustrate the number of
+different ways in which institutions can enter into theoretical
+discussions about civil conflict.
 
@@
 presence of institutions conducive to conflict resolution such that
-the opportunity for rebellion may be present, given the presence of
-civil liberties conducive to organizing, but there is a notable lack
-in motivation for collective violence. For the latter, authoritarian
-regimes may produce a significant amount of political discontent;
-however, the opportunity is lacking due to the state's capacity and
-willingness to violently repress dissent. Anocratic regimes are by
-their very nature transitory, partly open and partly repressive,
-encouraging political discontent against the state without the
-capacity to either accommodate or suppress opposition.
+the opportunity for rebellion may be present, given the availability
+of civil liberties conducive to organizing, but there is a notable
+lack in motivation for collective violence. For the latter,
+authoritarian regimes may produce a significant amount of political
+discontent; however, the opportunity is lacking due to the state's
+capacity and willingness to violently repress dissent. Anocratic
+regimes are by their very nature transitory, partly open and partly
+repressive, encouraging political discontent against the state without
+the capacity to either accommodate or suppress opposition.
 
@@
 conflict onset both in terms of political discontent --- a regime's
-inability to govern effectively and respond to public needs leading to
+inability to govern effectively and respond to public needs leads to
 grievances and a loss in legitimacy --- and political opportunities
@@
     representation. In this way, executive constraints can possibly
-    aid in discriminating the types of partially democratized regimes
-    which may be at a heightened risk for civil conflict.
+    aid in discriminating between the types of partially democratized
+    regimes which may be at a heightened risk for civil conflict.
 
@@
 The dependent variable, *civil conflict onset*, is taken from the
-UCDP/PRIO Armed Conflict Dataset v19.1 [@Gleditsch2002;
-@Pettersson2019]. According to the dataset, intra-state conflict is
-defined as a violent contest over a government or territory between a
-state and non-state actor with a minimum threshold of 25 battle deaths
-within a single country-year with a coverage from 1946 to 2018.  The
-data was merged based on the government side for conflicts labeled as
-internal or internationalized internal --- *i.e.* where
-*type_of_conflict* was either 3 or 4. Multiple conflicts within the
-same years were aggregated to form a binary variable coded as 1 to
-denote the onset of at least one new civil conflict and 0 for all
-other years, including years of ongoing conflict.[^conflict_type]
+UCDP/PRIO Armed Conflict Dataset (ACD) v19.1.[^onset_dataset]
+According to the dataset, intra-state conflict is defined as a violent
+contest over a government or territory between a state and non-state
+actor with a minimum threshold of 25 battle deaths within a single
+country-year with a coverage from 1946 to 2018.  The data was merged
+based on the government side for conflicts labeled as internal or
+internationalized internal --- *i.e.* where *type_of_conflict* was
+either 3 or 4. Multiple conflicts within the same years were
+aggregated to form a binary variable coded as 1 to denote the onset of
+at least one new civil conflict and 0 for all other years, including
+years of ongoing conflict.[^conflict_type]
+
+[^onset_dataset]: UCDP also makes available a pre-aggregated onset
+    dataset for internal and internationalized internal armed
+    conflicts. However, there are a number of coded onsets which do
+    not correspond to the conflict years listed in the standard
+    ACD. For example, China is coded as experiencing episode onsets in
+    1969, 1974, 1978, 1983, 1986 with unique conflict starts in 1969
+    and 1974 according to the *onset1* and *newconf* variables in the
+    onset dataset.  In the ACD, however, China is only coded as being
+    engaged in an interstate conflict with Myanmar/Burma (1969),
+    Russia (1969), and Vietnam (1974 - 1988, with several gap
+    years). Likewise discrepancies exist for a number of other
+    countries affecting 80 country-years in total. Since the onset
+    dataset does not inlude identifying information on each conflict,
+    I use the larger ACD to programmatically calculate the conflict
+    onsets.
 
 [^conflict_type]: Extrasystemic conflicts, *i.e.* imperial or colonial
-    conflicts, were excluded largely due to lack of data. For example,
+    conflicts, were excluded largely due to lack of data. Notably,
     there is very little easily available economic data for colonial
@@
 In total there are `r sum(merged.df$episode_onset)` onsets in the
-dataset according to the episodic definition affecting
+assembled dataset according to the episodic definition affecting
 `r filter(merged.df, episode_onset == 1) %$% n_distinct(country_name)`
@@
 principal data is drawn from expert coded surveys involving a global
-network of over 3000 coders, most of whom are academics or
+network of over 3,000 coders, most of whom are academics or
 professionals working in policy related fields. Individual ratings are
@@
     that World War II was a significant international event that
-    reshaped the world order and even had domestic consequences for
+    reshaped the world order and had domestic consequences for even
     countries not directly involved. For newly independent countries,
@@
 although still conditional on the values for the remaining manifest
-variables. Where the legislature is present, $\theta$ is distributed
-according to the $\mathsf{Normal}(0, 1)$ prior.
+variables.
 
@@
 
-Finally, to expand upon the country selection from \ref{fig:theta} and
-provide a temporal comparison, figure \ref{fig:tie} plots the $\theta$
-estimates for 2010 versus 2000 by country. A full list of all
-countries is available in [Appendix C](#appendix-c). Note, the
-statistical significance of differences between years for a single
-country cannot strictly be judged by overlapping credible intervals
-[@Schenker2001]. For example, Chile, despite having overlapping
-credible intervals, still has a posterior probability of approximately
-$`r chile_p`$ percent that the estimate for 2000 is lower
-than 2010. In the case of Tunisia with even more visual overlap, there
-is a $`r tunisia_p`$ percent posterior probability that the estimate
-for 2010 is higher than 2000.
+Finally, to expand upon the country selection from figure
+\ref{fig:theta} and provide a temporal comparison, figure
+\ref{fig:tie} plots the $\theta$ estimates for 2010 versus 2000 by
+country. A full list of all countries is available in [Appendix
+C](#appendix-c). Note, the statistical significance of differences
+between years for a single country cannot strictly be judged by
+overlapping credible intervals [@Schenker2001]. For example, Chile,
+despite having overlapping credible intervals, still has a posterior
+probability of approximately $`r chile_p`$ percent that the estimate
+for 2000 is lower than 2010. In the case of Tunisia with even more
+visual overlap, there is a $`r tunisia_p`$ percent posterior
+probability that the estimate for 2010 is higher than 2000.
 
@@
 are strongly pulled towards zero as a result of the partial pooling
-with upticks, such as the in the initial years of the 1990s during the
+with upticks, such as the initial years of the 1990s during the
 dissolution of the Soviet Union and Yugoslavia, roughly mirroring the
@@
 different levels of GDP per capita while holding constant the other
-covariates. In this case, I take the mean of GDP per capita growth,
+covariates. This means that I take the mean of GDP per capita growth,
 CINC, the proportion of discriminated population, population density,
@@
     rename(codelow = `10%`, median = `50%`, codehigh = `90%`) %>%
-    mutate(interval = sprintf("[%.3f, %.3f]",
-                              signif(codelow, 3),
-                              signif(codehigh, 3)))
+    mutate(interval = sprintf("[%s, %s]",
+                              fmt(codelow, 3),
+                              fmt(codehigh, 3)))
 
@@
 empirical distribution of GDP per capita --- in the case of the
-latter, only `r signif(mean(X[, "rgdpepc"] < mean(df$y)), 3) * 100`
+latter, only `r fmt(mean(X[, "rgdpepc"] < mean(df$y)) * 100)`
 percent of the observations lay below the midpoint of the X-axis at
@@
     is on the unbounded scale, the highest and lowest levels used in
-    the calculations for \ref{fig:gdp} are drawn from the minimum and
-    maximum of the median posterior estimates.
+    the calculations for figure \ref{fig:gdp} are drawn from the
+    minimum and maximum of the median posterior estimates.
 
@@
 
-tpr <- signif(100 * m["Predicted: Onset", "Actual: Onset"] /
-              sum(m[, "Actual: Onset"]), 3)
-tnr <- signif(100 * m["Predicted: No Onset", "Actual: No Onset"] /
-              sum(m[, "Actual: No Onset"]), 3)
-
-fpr <- signif(100 * m["Predicted: Onset", "Actual: No Onset"] /
-              sum(m[, "Actual: No Onset"]), 3)
-fnr <- signif(100 * m["Predicted: No Onset", "Actual: Onset"] /
-              sum(m[, "Actual: Onset"]), 3)
+tpr <- fmt(100 * m["Predicted: Onset", "Actual: Onset"] /
+              sum(m[, "Actual: Onset"]))
+tnr <- fmt(100 * m["Predicted: No Onset", "Actual: No Onset"] /
+              sum(m[, "Actual: No Onset"]))
+
+fpr <- fmt(100 * m["Predicted: Onset", "Actual: No Onset"] /
+              sum(m[, "Actual: No Onset"]))
+fnr <- fmt(100 * m["Predicted: No Onset", "Actual: Onset"] /
+              sum(m[, "Actual: Onset"]))
 
@@
 compare the predicted probabilities to the actual data, table
-\ref{tab:confusion} presents a confusion matrix with a 0.17 cutoff,
-which approximately corresponds to the third quartile in the summary
-table. `r tpr` percent of onsets and `r tnr` percent of non-onsets are
-correctly predicted. The false positive rate is `r fpr` percent and
-the false negative rate is `r fnr` percent.
+\ref{tab:confusion} presents a confusion matrix with a 0.168 cutoff,
+which corresponds to the third quartile in the summary table. `r tpr`
+percent of onsets and `r tnr` percent of non-onsets are correctly
+predicted. The false positive rate is `r fpr` percent and the false
+negative rate is `r fnr` percent.
 
@@
 
-auc <- attr(x$roc[[1]], "auc") %>% signif(3)
-auc_pr <- attr(x$prcs[[1]], "auc") %>% signif(3)
+auc <- attr(x$roc[[1]], "auc") %>% fmt
+auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt
 ```
@@
 dataset. Onsets comprise only
-`r signif(100 * mean(final.df$lepisode_onset == 1), 3)` percent of all
+`r fmt(100 * mean(final.df$lepisode_onset == 1))` percent of all
 observations meaning that the false positive rate is less sensitive to
@@
 
-[^auc_pr]: Similar to the ROC curve, an ACU of 1 for a PRC curve
+[^auc_pr]: Similar to the ROC curve, an AUC of 1 for a PRC curve
     denotes a perfect model. The AUC for the null model, or random
-    classifier, is the proportion of actual positive outcomes to negative
-    outcomes in the dataset. In this case this means approximately
-    .042.
+    classifier, is the proportion of actual positive outcomes to
+    negative outcomes in the dataset. For the present model this is
+    approximately
+    `r fmt(sum(final.df$lepisode_onset == 1) / sum(final.df$lepisode_onset == 0))`.
 

### 2020-08-29 paper.Rmd
Manuscript edits

@@
     ```
-wordcount: 13052
+wordcount: 14000
 suppress-bibliography: false
@@
 executive constraints to the incentives for political violence should
-still hold.[^robust]
-
-[^robust]: Due to the potential sensitivity of results to changing
-    conflict definitions [@Sambanis2004b], all analyses were also run
-    using only unique onsets as well as major episodic onsets with a
-    threshold of 1000 battle deaths. The results are available in
-    [Appendix C](#appendix-c).
+still hold.
 
@@
 
-(ref:confusion) Confusion matrix with a 0.168 cutoff
+(ref:confusion) Confusion matrix with a 0.168 cutoff for the median
+    posterior predicted probabilities.
 
@@
 
-# Discussion {-}
+## Alternative Specifications
+
+An additional issue which may affect the validity of the model is the
+potential sensitivity of the results to changing operational
+definitions of civil conflict. Operationalizing civil conflict onset
+is a difficult endeavour [@Sambanis2004b]. For example, when exactly
+does a conflict start? Is it at the initial formation of a rebel
+group? The first battle death? Or, when a certain threshold in
+conflict intensity has been achieved? Similar questions can be raised
+about when a conflict ends and how to clearly distinguish conflict
+termination from brief lulls in fighting. Different operational
+definitions can affect estimates and hinder the ability to compare
+results across studies. While datasets such as UCDP/PRIO have done
+much to standardise the operational definition of civil conflict,
+there is still a significant degree of researcher freedom in how onset
+is included as the outcome variable in a model.
+
+For the results presented so far the outcome variable has followed the
+episodic definition of civil conflict onset from UCDP/PRIO --- 25
+battle death threshold with new episodes coded after one year of
+inactivity --- which captures both high and low intensity civil
+conflicts. How robust are these findings when only examining
+full-scale civil wars? There is no *a prior* theoretical reason to
+expect that the relationship between executive constraints and civil
+conflict onset should not be negative when only examining civil
+wars. In fact, at least one of the mentioned theoretical mechanisms,
+namely that executive constraints may affect the ability of rebel
+groups to mobilize potential fighters by offering local security
+guarantees against government repression, posited that the theory may
+*not* hold for low intensity conflicts featuring predatory rebel
+groups. Thus, if the relationship were not to be found when re-running
+the model with only high intensity conflict onsets that would be a
+significant challenge to the external validity of the initial results.
+
+In the UCDP/PRIO armed conflict dataset, high intensity is defined at
+the threshold of 1,000 battle deaths. High intensity conflicts, or
+civil wars, can be coded according to when the threshold for high
+intensity is crosses either by a single year or by the cumulative
+battle deaths.[^battle_deaths] I therefore re-run the same model,
+including the same priors and runtime parameters, twice for the two
+different operational definitions. In both cases, I retain an episodic
+delineation between separate conflicts --- in other words, conflicts
+over the same incompatibility can have multiple onsets when there are
+lulls in fighting.[^episode_gaps] For the former, the onset year is
+set as the first year which crosses the high intensity threshold for
+each period of sustained fighting of at least 25 battle deaths. For
+the latter, onset is the coded conflict year for the cumulative period
+passing the high intensity threshold.
+
+[^battle_deaths]: A more robust modeling strategy would be to directly
+    predict the number of battle deaths and then apply *a posteriori*
+    decision criteria to the predicted results to classify
+    country-years as low-intensity or high intensity conflict. This
+    would avoid the arbitrary nature of battle deaths thresholds for
+    determining the intensity types; however, battle death data tends
+    to be limited in time span and there would be the additional issue
+    of data quality.
+
+[^episode_gaps]: Another robustness check would be to re-run the model
+    for different episode definitions based on varying intervals of
+    fighting inactivity. For example, coding termination after one,
+    two, or three years of inactivity. This remains outside the scope
+    of the present paper, however, due to lack of computational
+    resources for such a high number of model runs.
+
+(ref:overlay) Comparison of regression slopes between the original
+    model run, "Low & High Intensity", and the model runs for the two
+    different operational definitions of high intensity
+    conflict. Credible intervals are 80 percent posterior quantile
+    intervals.
+
+```{r fig.width = 8, fig.height = 4, fig.cap = "\\label{fig:overlay} (ref:overlay)"}
+models <- c("full_model", "major_only", "cumulative_only")
+ll <- lapply(models, function(model) {
+    post <- file.path("posteriors", model, "reg_posteriors.csv") %>%
+        read_post
+})
+
+args <- list(pars = "beta",
+             legend_labels = c("Low & High Intensity",
+                               "High Intensity",
+                               "Cumulatively High Intensity"))
+do.call(overlay_pars, c(ll, args)) +
+    scale_x_discrete(labels = colnames(beta))
+```
+
+```{r}
+ix_probs <- lapply(seq_along(models), function(i) {
+    beta <- extract(ll[[i]], "beta")
+
+    # Really could just grab this from the existing `beta` variable,
+    # but just to be safe...
+    load(file.path("posteriors", models[i], "input.RData"),
+         envir = e <- new.env())
+    colnames(beta) <- c("exec_constraints", colnames(e$X),
+                        sprintf("exec * %s",
+                                colnames(e$X)[e$interaction_idx]))
+
+    mean(beta[, "exec * rgdpepc"] < 0)
+})
+
+ix_probs <- unlist(ix_probs)
+names(ix_probs) <- models
+```
+
+Figure \ref{fig:overlay} plots the original results compared to the
+results from running the model with the two different operational
+definitions of high intensity conflict.[^theta] The majority of
+depicted parameter estimates are fairly stable across runs.
+Notably, the relationship between executive constraints and conflict onsets remains negative with a high posterior probability of being less than zero.
+The estimates for the interaction effect between executive constraints and GDP per capita, however, become far to uncertain to be able to conclusively assert an association.
+The probability that the interaction term is less than zero for the high intensity model run, *i.e.* where at least a single year crosses the high intensity threshold, is only `r fmt(ix_probs["major_only"])` while the probability for the cumulative intensity model is only
+`r fmt(ix_probs["cumulative_only"])`.
+
+[^theta]: Since the measurement model for executive constraints is
+    embedded within the larger regression predicting civil conflict
+    onset, the latent factor $\theta$ is re-estimated for each model
+    run. The stochastic nature of posterior sampling ensures that the
+    estimates are never be exactly the same; however, the differences
+    are inconsequential. In fact the correlation between the main
+    model and both the cumulatively high intensity model and the major
+    conflicts only model is effectively one.
+
+(ref:confusion_major) Confusion matrix for conflicts with at least a
+    single year passing the high intensity threshold with a 0.168
+    cutoff for the median posterior predicted probabilities.
+
+```{r confusion_major}
+extra <- read_post("./posteriors/major_only/extra_posteriors.csv")
+
+p_hat <- extract(extra, "p_hat")
+p_median <- apply(p_hat, 2, median)
+
+load("./posteriors/major_only/input.RData")
+
+m <- table(p_median > .168, final.df$lmajor_onset)
+kable(m, "latex", align = "c", booktabs = T,
+      linesep = "", escape = F, caption = "hello")
+```
+
+(ref:confusion_cumulative) Confusion matrix for conflicts with a
+    cumulative high intensity with a 0.168 cutoff for the median
+    posterior predicted probabilities.
+
+```{r confusion_cumulative}
+extra <- read_post("./posteriors/cumulative_only/extra_posteriors.csv")
+
+p_hat <- extract(extra, "p_hat")
+p_median <- apply(p_hat, 2, median)
+
+load("./posteriors/cumulative_only/input.RData")
+
+m <- table(p_median > .168, final.df$lcum_onset)
+kable(m, "latex", align = "c", booktabs = T,
+      linesep = "", escape = F,
+      caption = "world")
+```
+
+How well does the model fit the data when only examining high intensity conflicts?
+
+# Conclusion {-}
 

### 2020-08-30 paper.Rmd
Finish alternative specifications & update R dependencies

@@
     ```
-wordcount: 14000
+wordcount: 14300
 suppress-bibliography: false
@@
                fig.width = 5, fig.height = 3, cache = T)
-
-# TODO: this has been added to package
-fmt <- function(x, digits = 2) {
-    prettyNum(round(x, digits), big.mark = ",")
-}
-
+```
+
+# Introduction {-}
+
+```{r intro-descriptive-stats}
 # Create initial descriptive stats for introduction
@@
 ```
-
-# Introduction {-}
 
@@
 
+```{r}
+load("./posteriors/cumulative_only/input.RData",
+     envir = cumulative <- new.env())
+
+load("./posteriors/major_only/input.RData",
+     envir = major <- new.env())
+```
+
 In the UCDP/PRIO armed conflict dataset, high intensity is defined at
@@
 civil wars, can be coded according to when the threshold for high
-intensity is crosses either by a single year or by the cumulative
-battle deaths.[^battle_deaths] I therefore re-run the same model,
+intensity is crossed either by a single year or by the cumulative
+battle death count.[^battle_deaths] I therefore re-run the same model,
 including the same priors and runtime parameters, twice for the two
@@
 each period of sustained fighting of at least 25 battle deaths. For
-the latter, onset is the coded conflict year for the cumulative period
-passing the high intensity threshold.
+the latter, onset is the first coded conflict year for the cumulative
+period passing the high intensity threshold. In total there
+`r sum(major$final.df$lmajor_onset)` onsets for the single year 1,000
+battle deaths definition, and `r sum(cumulative$final.df$lcum_onset)`
+onsets for the cumulative definition.
 
@@
                                "High Intensity",
-                               "Cumulatively High Intensity"))
+                               "Cumulative High Intensity"))
 do.call(overlay_pars, c(ll, args)) +
@@
 
-    # Really could just grab this from the existing `beta` variable,
-    # but just to be safe...
-    load(file.path("posteriors", models[i], "input.RData"),
-         envir = e <- new.env())
-    colnames(beta) <- c("exec_constraints", colnames(e$X),
+    if (models[i] == "major_only") {
+        colnames(beta) <- c("exec_constraints", colnames(major$X),
                             sprintf("exec * %s",
-                                colnames(e$X)[e$interaction_idx]))
+                                    colnames(major$X)[major$interaction_idx]))
+    } else {
+        colnames(beta) <- c("exec_constraints", colnames(cumulative$X),
+                            sprintf("exec * %s",
+                                    colnames(cumulative$X)[cumulative$interaction_idx]))
+    }
 
@@
 
-Figure \ref{fig:overlay} plots the original results compared to the
-results from running the model with the two different operational
-definitions of high intensity conflict.[^theta] The majority of
-depicted parameter estimates are fairly stable across runs.
-Notably, the relationship between executive constraints and conflict onsets remains negative with a high posterior probability of being less than zero.
-The estimates for the interaction effect between executive constraints and GDP per capita, however, become far to uncertain to be able to conclusively assert an association.
-The probability that the interaction term is less than zero for the high intensity model run, *i.e.* where at least a single year crosses the high intensity threshold, is only `r fmt(ix_probs["major_only"])` while the probability for the cumulative intensity model is only
+Figure \ref{fig:overlay} plots the original regression slopes compared
+to the results from running the model with the two different
+operational definitions of high intensity conflict.[^theta] The
+majority of depicted parameter estimates are fairly stable across
+runs. Notably, the relationship between executive constraints and
+conflict onsets remains negative with a high posterior probability of
+being less than zero. The estimates for the interaction effect between
+executive constraints and GDP per capita, however, become far too
+uncertain to be able to conclusively assert an association. The
+probability that the interaction term is less than zero for the high
+intensity model run, *i.e.* where at least a single year crosses the
+high intensity threshold, is only `r fmt(ix_probs["major_only"])`
+while the probability for the cumulative intensity model run is only
 `r fmt(ix_probs["cumulative_only"])`.
@@
 
-(ref:confusion_major) Confusion matrix for conflicts with at least a
-    single year passing the high intensity threshold with a 0.168
-    cutoff for the median posterior predicted probabilities.
-
-```{r confusion_major}
+(ref:confusion-major) Confusion matrix with a 0.168 cutoff for the
+    median posterior predicted probabilities from the single year high
+    intensity model run.
+
+```{r confusion-major}
 extra <- read_post("./posteriors/major_only/extra_posteriors.csv")
@@
 
-load("./posteriors/major_only/input.RData")
-
-m <- table(p_median > .168, final.df$lmajor_onset)
+m <- table(p_median > .168, major$final.df$lmajor_onset)
+rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
+colnames(m) <- c("Actual: No Onset", "Actual: Onset")
+
+# Calculate the AUCs as well
+x <- evalmod(scores = p_median, labels = major$final.df$lmajor_onset)
+major_auc <- attr(x$roc[[1]], "auc") %>% fmt
+major_auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt
+
 kable(m, "latex", align = "c", booktabs = T,
-      linesep = "", escape = F, caption = "hello")
-```
-
-(ref:confusion_cumulative) Confusion matrix for conflicts with a
-    cumulative high intensity with a 0.168 cutoff for the median
-    posterior predicted probabilities.
-
-```{r confusion_cumulative}
+      linesep = "", escape = F, caption = "(ref:confusion-major)")
+```
+
+Table \ref{tab:confusion-major} presents the confusion matrix for the
+single year high intensity model run using the median posterior
+predicted probabilities. A cutoff of 0.168 was reused for comparative
+purposes to the original results. In this case, the model is unable to
+predict even a single major conflict onset with the highest median
+posterior predicted probability for an actual onset being only
+`r fmt(max(p_median[major$final.df$lmajor_onset == 1]))`. Furthermore,
+although the ROC AUC is `r major_auc`, this is again due to the major
+class imbalance between positive versus negative outcomes in the
+dataset. A precision-recall AUC of `r major_auc_pr` demonstrates the
+poor performance of the model.
+
+(ref:confusion-cumulative) Confusion matrix with a 0.168 cutoff for
+    the median posterior predicted probabilities from the cumulative
+    high intensity model run.
+
+```{r confusion-cumulative}
 extra <- read_post("./posteriors/cumulative_only/extra_posteriors.csv")
@@
 
-load("./posteriors/cumulative_only/input.RData")
-
-m <- table(p_median > .168, final.df$lcum_onset)
+m <- table(p_median > .168, cumulative$final.df$lcum_onset)
+rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
+colnames(m) <- c("Actual: No Onset", "Actual: Onset")
+
+x <- evalmod(scores = p_median, labels = cumulative$final.df$lcum_onset)
+cum_auc <- attr(x$roc[[1]], "auc") %>% fmt
+cum_auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt
+
 kable(m, "latex", align = "c", booktabs = T,
       linesep = "", escape = F,
-      caption = "world")
-```
-
-How well does the model fit the data when only examining high intensity conflicts?
+      caption = "(ref:confusion-cumulative)")
+```
+
+Similarly, table \ref{tab:confusion-cumulative} shows the confusion
+matrix for the cumulative high intensity model run with the same
+median posterior predicted probability cutoff. Although the model in
+this case is able to correctly classify 17 percent of onsets with an
+ROC AUC of `r cum_auc`, the overall performance is fairly limited with
+a precision-recall AUC of only `r cum_auc_pr`.
+
+Taken together the model runs examining high intensity conflicts
+suggest that the negative relationship between executive constraints
+and civil conflict onset is fairly robust. However, the same issues of
+model fit continue to plague
 

### 2020-08-30 R/thesis.utils/R/utils.R

@@ -269,3 +269,9 @@ summary_stats.data.frame <- function(x, vars = colnames(x)) {
 #'
 #' @export
 to_idx <- function(x, ...) factor(x, ...) %>% as.numeric
+
+
+#' @export
+fmt <- function(x, digits = 2) {
+    prettyNum(round(x, digits), big.mark = ",")
+}

### 2020-08-30 paper.Rmd
Manuscript edits

@@
     ```
-wordcount: 14300
+wordcount: 15000
 suppress-bibliography: false
@@
         toc: true
-abstract: Lorem ipsum
+abstract: |
+    **Abstract**: Do institutional constraints on a regime's executive decrease the
+    likelihood of civil conflict onset? An unconstrained executive is free
+    to loot state resources, undermining the state's capacity to
+    effectively deal with nascent rebellions, and may be more likely to
+    engage in violent repression, especially in the form of indiscriminate
+    violence. This can encourage political grievances, lead to a loss of
+    legitimacy for the regime, and provide the opportunity for rebel
+    groups to attract new members by offering protection against
+    government violence. Furthermore, the lack of guaranteed checks to
+    executive power may incentivise actors to avoid bargained solutions if
+    they fear future extralegal retributions. Taken together, where there
+    is a lack of effective constraints against the executive there may be
+    an increased risk for the onset of civil conflict. To test this
+    proposition I employ a bayesian latent variable model which estimates
+    executive constraint as a latent factor derived from several manifest
+    variables and relates it to a binary measure of civil conflict onset
+    through a logistic regression.  The primary finding is that there is a
+    negative relationship such that the predicted probability for civil
+    conflict is lower where there exists higher levels of executive
+    constraints. However, this is conditional on the level of GDP per
+    capita; for low-income country-years the relationship between
+    executive constraints and civil conflict onset is indeterminate
+    possibly because it is easier to recruit and mobilize fighters
+    regardless of the overall level of executive constraints. The issue of
+    poor model fit means that the presented results should be considered
+    tentative at best.
 ---
@@
 empirically among the deadliest. Out of `r fmt(total_deaths)`
-cumulative battles deaths from organized violence recorded
-since 1989 [@Pettersson2019; @Eck2007; @Sundberg2012],
-`r fmt(civil_deaths)`, or `r fmt(civil_deaths_percent, 1)` percent, are
-the result of civil conflict.
+cumulative battles deaths from organized violence recorded since 1989
+[@Pettersson2019; @Eck2007; @Sundberg2012], `r
+fmt(civil_deaths_percent, 1)` percent are the result of civil
+conflict.
 
 Given the impact of civil conflict and the increasing availability of
-conflict related data, the last several decades has witnessed a
+conflict related data, the last several decades have witnessed a
 veritable explosion in the number of quantitative studies
@@
 political institutions play a significant role in the risk of civil
-conflict. Several studies have argued in favor of institutions by
-investigating, for example, aspects of governance, institutional
-quality, and opportunities for political participation [@Fjelde2009;
-@Taydas2010; @Hegre2015a; @Bartusevic2018].
+conflict. Several studies have argued in favor of institutions and
+have investigated, for example, aspects of governance, institutional
+quality, and opportunities for political participation as factors
+influencing the risk of civil conflict [@Fjelde2009; @Taydas2010;
+@Hegre2015a; @Bartusevic2018].
 
@@
 incentives for violence by affecting state behaviour. Second, the link
-between executive constraints is relatively understudied. The previous
-literature that exists is fairly limited and contradictory in terms of
-results.
+between executive constraints and civil conflict is relatively
+understudied. The previous literature that exists is fairly limited
+and contradictory in terms of results.
 
@@
 relationship largely disappears for low-income country-years when an
-interaction term with GDP per capita is included. If the primary
-theoretical mechanism involves the impact of executive constraints on
-rebel mobilization --- an unchecked executive may be more likely to
-engage in indiscriminate violence which could lead to a loss of
-legitimacy as well as the ability of rebel groups to offer local
-security guarantees to attract recruits --- then the incentives for
-violence in low-income country-years may be so great that it does not
-matter how constrained the executive may or may not be.
+interaction term between executive constraints and gross domestic
+product (GDP) per capita is included. If the primary theoretical
+mechanism involves the impact of executive constraints on rebel
+mobilization --- an unchecked executive may be more likely to engage
+in indiscriminate violence which could lead to a loss of legitimacy as
+well as the ability of rebel groups to offer local security guarantees
+to attract recruits --- then the incentives for violence in low-income
+country-years due to say economic factors may be so great that it does
+not matter how constrained the executive may or may not be.
 
@@
 
-# Theoretical Overview {-}
-
-## Institutions and Civil Conflict
+# Theoretical Background {-}
+
+## Political Institutions and Civil Conflict
 
@@
 social interactions" [@Hodgson2006]. More informally, we can consider
-them to be the "rules of the game", encompassing for example codified
+them to be the "rules of the game", encompassing, for example, codified
 laws, norms, and shared understandings, all of which can constrain or
@@
 is seen as accountable to and representative of the larger population,
-thereby influencing the legitimacy of anti-government violence.  And,
-their quality can shape the capacity of the state to suppress or
-co-opt potential opposition. Although the preceding examples are far
-from forming an exhaustive list, they illustrate the number of
-different ways in which institutions can enter into theoretical
-discussions about civil conflict.
-
-Unsurprisingly, insofar as the role of institutions on the likelihood
-of civil conflict has been examined by previous literature the focus
-has largely been on the broader notion of democracy, which at a
-minimum is a system regulating access to political power through
-popular representation determined by free and competitive elections
-[@Schumpeter1962]. Democracy serves as a proxy for a number of
-normatively desirable institutions and has been hypothesised to have a
-pacifying domestic effect according to two predominant mechanisms. The
-first is centered on the concept of relative deprivation, or
-inequalities in value expectations and value capabilities, as a
-primary source of discontent that can motivate collective violence
-[@Gurr1970]. [^gurr] Democracy is thought to decrease the amount of
-political discontent by providing avenues for free expression and
-adjudicating group conflicts through non-violent institutional means
---- for example, through the regular occurrence of free and fair
-elections [@Hegre2014]. The second mechanism assumes that violent
-conflict is costly and that there exists a bargained solution
-preferable to fighting, assuming that actors are rational and able to
-internalize the cost of violence [@Fearon1995]. One of the reasons why
-a bargained solution may be impossible is the risk of future shifts in
-the distribution of power. More specifically, a future change in favor
-of the relative strength of one of the parties can provide an
-incentive to later renege. This influences the present degree to which
-concessions can be offered and parties are willing to credibly commit
-[@Powell2006]. Within the domestic sphere, democracy offers a possible
-solution by formalising the regular contest of power with full
-enfranchisement, thereby providing a *de jure* guarantee for the
-balance of power [@Origins2006], and should, likewise, decrease the
-likelihood of civil conflict.
+thereby influencing the legitimacy of anti-government violence. In
+addition, institutional quality can shape the capacity of the state to
+suppress or co-opt potential opposition. Although the preceding
+examples are far from forming an exhaustive list, they illustrate the
+number of different ways in which institutions can enter into
+theoretical discussions about civil conflict.
+
+Unsurprisingly, insofar as the role of institutions in shaping the
+likelihood of civil conflict has been examined by previous
+literature, the focus has largely been on the broader notion of
+democracy, which at a minimum is a system regulating access to
+political power through popular representation determined by free and
+competitive elections [@Schumpeter1962]. Democracy serves as a proxy
+for a number of normatively desirable institutions and has been
+hypothesised to have a pacifying domestic effect according to two
+predominant mechanisms. The first is centered on the concept of
+relative deprivation, or inequalities in value expectations and value
+capabilities, as a primary source of discontent that can motivate
+collective violence [@Gurr1970]. [^gurr] Democracy is thought to
+decrease the amount of political discontent by providing avenues for
+free expression and adjudicating group conflicts through non-violent
+institutional means --- for example, through the regular occurrence of
+free and fair elections [@Hegre2014]. The second mechanism assumes
+that violent conflict is costly and that there exists a bargained
+solution preferable to fighting, assuming that actors are rational and
+able to internalize the cost of violence [@Fearon1995]. One of the
+reasons why a bargained solution may be impossible is the risk of
+future shifts in the distribution of power. More specifically, a
+future change in favor of the relative strength of one of the parties
+can provide an incentive to later renege. This influences the present
+degree to which concessions can be offered and parties are willing to
+credibly commit [@Powell2006]. Within the domestic sphere, democracy
+offers a possible solution by formalising the regular contest of power
+with full enfranchisement, thereby providing a *de jure* guarantee for
+the balance of power [@Origins2006], and should, likewise, decrease
+the likelihood of civil conflict.
 
@@
     stage, the propensity for political regimes to generate
-    discontent, given the difficultly of measuring the variables that
+    discontent, given the difficultly of measuring the factors that
     may affect politicisation and actualisation.
 
-Despite the theoretical basis linking democracy to reduced risks for
+Despite the theoretical basis linking democracy to reduced risks of
 civil conflict, the empirical results tend to be mixed. @Hegre2001,
@@
 states. Instead, anocratic regimes in between democracy and
-authoritarianism are considered to be the most prone. The explanatory
-logic is that both democratic and fully authoritarian states tend to
-be consolidated regimes. In the case of the former, this means the
-presence of institutions conducive to conflict resolution such that
-the opportunity for rebellion may be present, given the availability
-of civil liberties conducive to organizing, but there is a notable
-lack in motivation for collective violence. For the latter,
-authoritarian regimes may produce a significant amount of political
-discontent; however, the opportunity is lacking due to the state's
-capacity and willingness to violently repress dissent. Anocratic
-regimes are by their very nature transitory, partly open and partly
-repressive, encouraging political discontent against the state without
-the capacity to either accommodate or suppress opposition.
+authoritarianism are considered to be the most prone to civil conflict
+onset. The explanatory logic is that both democratic and fully
+authoritarian states tend to be consolidated regimes. In the case of
+the former, this means the presence of institutions conducive to
+conflict resolution such that the opportunity for rebellion may be
+present, given the availability of civil liberties conducive to
+organizing, but there is a notable lack in motivation for collective
+violence. For the latter, authoritarian regimes may produce a
+significant amount of political discontent; however, the opportunity
+is lacking due to the state's capacity and willingness to violently
+repress dissent. Anocratic regimes are by their very nature
+transitory, partly open and partly repressive, encouraging political
+discontent against the state without the capacity to either
+accommodate or suppress opposition.
 
@@
 conflict where explanatory primacy is given to either greed-based
-motivation or economic conditions conducive for rebellion, the state
+motivation or economic conditions conducive to rebellion, the state
 is a key actor. How a state responds to dissent structures the
@@
 
-There are of course a large variety of political institutions whose
-presence and quality can influence the behaviour of the state.  Given
-the importance of the executive as a central authority, however,
-especially in times of political crisis, institutions which affect the
-executive's decision making should be pertinent. Chief among these is
-the idea of executive constraints --- more formally, the institutional
-constraints by other government bodies on the executive's ability to
-abuse power and act unilaterally. Executive constraints can be
-considered a subset of horizontal accountability and imply the
-existence of state agents --- such as the legislature or the judiciary
---- with both the legal mandate and willingness to oversee or sanction
-the executive thus serving as deterrents to the forms of
-authoritarian excesses in executive power which may be linked to civil
-conflict. [^ODonnell]
+There are of course a large variety of political institutions for
+which the presence and quality can influence the behaviour of the
+state.  Given the importance of the executive as a central authority,
+however, especially in times of political crisis, institutions which
+affect the executive's decision making should be pertinent. Chief
+among these is the idea of executive constraints --- more formally,
+the institutional constraints by other government bodies on the
+executive's ability to abuse power and act unilaterally. Executive
+constraints can be considered a subset of horizontal accountability
+and imply the existence of state agents --- such as the legislature or
+the judiciary --- with both the legal mandate and willingness to
+oversee or sanction the executive thus serving as deterrents to the
+forms of authoritarian excesses in executive power which may be linked
+to civil conflict. [^ODonnell]
 
@@
 conflicted. @Getmansky2013, for example, offers a theory of executive
-constraints playing a role in risk aversion in democratic polities;
-however, they find their measure to be insignificant in predicting the
-outbreak of insurgencies. On the other hand, @Fjelde2009 with a theory
-tying executive constraints to public goods provisioning reports a
-significant interaction term with public spending. Both cases use the
-ordinal measure of executive constraints form Polity IV
-[@Marshall2019]. This leaves open further work on additional
-theoretical mechanisms and analyses using other datasets.
-
-## Executive Constraints
+constraints playing a role in risk aversion in democratic polities.
+When the executive is constrained by other government institutions a
+regime may be unwilling to resort to force against a growing
+insurgency and may instead seek a peaceful solution to avoid
+fighting.[^weakness] They use a dichotomized measure of XCONST,
+"Constraint on Chief Executive", from Polity IV [@Marshall2019] in
+order to test this theory, finding their measure to be insignificant
+in predicting the outbreak of insurgencies.
+
+[^weakness]: If the theory were to hold, it is unclear if risk aversion
+    really translates to a lower incidence rate of civil conflict
+    since such regimes may be subject to opportunistic rebel groups
+    ready to exploit precisely those weaknesses. In other words,
+    democratic regimes may actively attempt to avoid armed conflict;
+    however, they can still be forced into such situations by rebel
+    groups.
+
+In another study, @Fjelde2009 put forth a theory tying executive
+constraints to public provisioning. Increased government expenditures
+in the presence of strong institutional checks against executive
+corruption is thought to decrease the risk of civil conflict by
+reducing potential grievances against the state. In contrast to
+Getmansky, the authors find a significant negative effect for an
+interaction term between executive constraints and government
+expenditures when also using XCONST from Polity IV, although the
+regression slope for the executive constraints term is insignificant.
+
+In both studies, using an ordinal measure, and in the case of
+@Getmansky2013 a binary measure, for executive constraints risks
+losing a significant amount of information concerning the varying
+degrees of effectiveness in the constraints against executives among
+different countries. A possible improvement would be the use of a
+continuous measure to be more able to capture the quality of
+constraints and better adjudicate whether there exists a measurable
+effect on civil conflict onset.
+
+## Executive Constraints and Civil Conflict
 
@@
 conflicts are protracted events against an often times much stronger
-opponent the short-term utility from participation is relatively
+opponent, the short-term utility from participation is relatively
 low. Thus, rebel leaders must be able to offer proportionate selective
@@
 constraints on the executive, the opportunity cost for joining the
-rebel side should be higher. Rebel leaders will then have to rely more
-heavily on economic selective benefits [@Regan2005], creating a higher
-barrier for mobilization and decreasing the risk of civil conflict
-onset. It should be noted though, that where this relationship may not
-hold is in situations of low intensity insurgencies involving
-predatory rebel groups solely motivated by rent-seeking. The threshold
-for mobilization in such scenarios is significantly lower than in full
-scale civil wars, meaning that rebel groups can initiate a conflict
-regardless of the current level of state repression.
+rebel side should be higher since the lack of indiscriminate violence
+makes non-participation an attractive alternative. Rebel leaders will
+then have to rely more heavily on economic selective benefits
+[@Regan2005], creating a higher barrier for mobilization and
+decreasing the risk of civil conflict onset. It should be noted
+though, that where this relationship may not hold is in situations of
+low intensity insurgencies involving predatory rebel groups solely
+motivated by rent-seeking, or resource extraction, rather than
+secession or government overthrow. The threshold for mobilization in
+such scenarios is significantly lower than in full scale civil wars
+since such rebel groups are not actively challenging the state for
+control, meaning that rebel groups can initiate a conflict regardless
+of the current level of state repression.
 
@@
 constraints should be associated with a lowered risk of civil conflict
-onset. This proposition can be further extended: if state weakness is a
-primary driver of civil conflict and the relative capacity of the
+onset. This proposition can be further extended: if state weakness is
+a primary driver of civil conflict and the relative capacity of the
 state is effected by the levels of executive constraints, as depicted
@@
 state capacity could have a mediating affect on the relationship
-between executive constraints and conflict onset. Another extension
-would be if executive constraints affect the mobilization potential
-for rebel groups, then there could also be a confounding effect from
-the relative threshold for mobilization. Finally, the relationship
-could still be confounded by the overall opportunity for rebellion ---
-the degree of executive constraints may not be particularly relevant
-in relatively stable countries lacking the conditions conducive to
-carrying out a rebellion. Regardless of conditionalities, contrary
-results would indicate either that executive constraints are
-indeterminate to the risk of civil conflict or possibly that there is
-even a positive relationship since constrained executives may be less
-able to effectively deal with violent challenges due to risk aversion
-to violence and/or from the slowdown of having to seek approval from a
-plurality of stakeholders before committing to decisive action.
+between executive constraints and civil conflict onset. Another
+extension would be if executive constraints affect the mobilization
+potential for rebel groups, then there could also be a confounding
+effect from the relative threshold for mobilization. Finally, the
+relationship could still be confounded by the overall opportunity for
+rebellion --- the degree of executive constraints may not be
+particularly relevant in relatively stable countries lacking the
+conditions conducive to carrying out a rebellion. Regardless of
+conditionalities, contrary results would indicate either that
+executive constraints are indeterminate to the risk of civil conflict
+or possibly that there is even a positive relationship to conflict
+onset since constrained executives may be less able to effectively
+deal with violent challenges due to risk aversion to violence and/or
+from the slowdown of having to seek approval from a plurality of
+stakeholders before committing to decisive action.
 
@@
 the role of elections or the media; indeed, vertical accountability
-can plausibly play a role in all of the mentioned mechanisms. For
+can plausibly play a role in all of the aforementioned mechanisms. For
 example, studies on state repression have emphasised the importance of
@@
 effective deterrent against negative sanctioning by leaders
-[@BuenoDeMesquita2005; @Davenport2007]. But elections occur only
+[@BuenoDeMesquita2005; @Davenport2007]. But, elections occur only
 intermittently and it is not an automatic given that voters will
@@
 through government bodies [@ODonnell1998]. Thus, executive constraints
-are interesting in their own right.
+are interesting in their own right and vertical accountability should
+not confound any potential negative relationship to civil conflict
+onset.
 
@@
 a function of a set of predictors, including executive constraints and
-an array of control variables, together with several random
-effects. The units of analysis are all independent country-years from
-1952 until 2013, as determined by data availability, with *I*
+an array of control variables, together with several varying
+intercepts. The units of analysis are all independent country-years
+from 1952 until 2013, as determined by data availability, with *I*
 countries indexed by $i = 1 \dotsc I$ and *T* years indexed by $t = 1
@@
 conflict onset, $\nu_{it}$, is the inverse logit such that $\nu_{it}
-\in [0, 1]$. [^invlogit]
-
-[^invlogit]: Technically, the inverse logit transforms values to
+\in [0, 1]$. [^invlogit] The choice of a logit link function avoids
+the truncation necessary for a linear probability model. Being a
+sigmoid function it also ensures that the marginal contributions from
+the predictors are reduced at the tails for higher and lower
+probability values. That is to say the added value of a one unit
+increase in a predictor is higher when the predicted probability is
+already at 0.5 at the boundary between a positive and negative outcome
+versus when it is at 0.95.
+
+[^invlogit]: Technically, the probabilities in a logit function are on
     the open interval, $(0, 1)$, but due to overflow/underflow in
-    floating point calculations by computers the endpoints can be
+    floating point calculations done by computers the endpoints can be
     observed.
@@
 
-where **X** is the predictors, $\beta$ the vector of regression
-slopes, and $\Zeta$ the *K* sets of varying intercepts indexed by
-$j_k$. All predictors are lagged by one year to control for time
-endogeneity and standardized by subtracting the mean and dividing by
-the standard deviation to aid in computation as well as comparison of
-the posterior $\beta$ estimates.
+where **X** is the predictors including the latent measure for
+executive constraints, $\beta$ the vector of regression slopes, and
+$\Zeta$ the *K* sets of varying intercepts indexed by $j_k$. All
+predictors are lagged by one year to control for time endogeneity and
+standardized by subtracting the mean and dividing by the standard
+deviation to aid in computation as well as comparison of the posterior
+$\beta$ estimates.
 
@@
 cross-sectional unmodeled heterogeneity between countries and over
-time [@Shor2007], and the number of peace years since the last
-civil conflict incidence per country to control for the lingering
+time [@Shor2007]. Also included is the number of peace years since the
+last civil conflict incidence per country to control for the lingering
 endogenous effects of violence on the likelihood that additional
-conflict breaks
-out.
+conflict breaks out.
 
@@
     \zeta^{country}_i & \sim \mathsf{Normal}(0, \sigma^{country}) \\
-    \sigma^{year}, \sigma^{country} & \sim \mathsf{Half-Cauchy}(0, 1)
+    \sigma^{year}, \sigma^{country} & \sim \mathsf{Cauchy_{+}}(0, 1)
 \end{align*}
@@
 of their euclidean distance. Where that distance is smaller there is a
-higher correlation. Compared to other approaches to modeling time
-dependence this avoids the parametric assumption of a proportional
-rate in exponential decay functions, the difficulty in selecting
-interaction terms to account for non-proportionality for cubic
-polynomial approximation, and the somewhat arbitrary selection of
-knots for splines [*c.f.* @Raknerud1997; @Carter2010;
-@Beck1998].
+higher correlation. Compared to other methods for modeling time
+dependence this approach avoids the parametric assumption of a
+proportional rate in exponential decay functions, the difficulty in
+selecting interaction terms to account for non-proportionality for
+cubic polynomial approximation, and the somewhat arbitrary selection
+of knots for splines [*c.f.* @Raknerud1997; @Carter2010; @Beck1998].
 
@@
 
-where for $P$ unique peace years, $K \in \mathbb{R}^{P \times P}$ is
+where given $P$ unique peace years, $K \in \mathbb{R}^{P \times P}$ is
 the covariance matrix, $D \in \mathbb{R}^{P \times P}$ the distance
@@
 \begin{align*}
-    \eta & \sim \mathsf{HalfNormal}(0, 1) \\
+    \eta & \sim \mathsf{Normal_{+}}(0, 1) \\
     \rho & \sim \mathsf{InvGamma}(15.5031, 78.2992)
@@
 phases of fighting which can make it difficult to accurately code the
-end of a conflict. The episodic definition avoids potentially
-conflating briefs lulls involving the same rebel groups and leaders
-versus the re-emergence of previously long dormant
-conflicts. Furthermore, in both scenarios, the theory linking
-executive constraints to the incentives for political violence should
-still hold.
-
-In total there are `r sum(merged.df$episode_onset)` onsets in the
+end of a conflict. The episodic definition avoids missing onsets from
+the re-emergence of previously long dormant conflicts.  The risk
+though is the inclusion of added onsets from low intensity conflicts
+due simply to intermittent drops in the recorded number of battle
+deaths per year. However, the posited theories linking executive
+constraints to civil conflict should still hold even in such
+scenarios. The incentive structures surrounding political violence are
+still relevant to rebel groups even in intermittent years. Groups need
+to be able to continue to attract fighters in order to maintain a
+credible threat, in which case the level of executive constraints
+should still have its theorized effect.
+
+In total there are `r sum(merged.df$episode_onset)` civil conflict onsets in the
 assembled dataset according to the episodic definition affecting
@@
 institutions --- more specifically, the legislature, judiciary, and
-independent investigative bodies --- rather than the *de jure* powers
-of any one institution *vis-à-vis* the executive, and to do so using a
-latent variable aggregation that avoids the arbitrary weighting of
-linear additive indices.[^additive] The aggregation method is largely
-similar to the accountability index created by @Luhrmann2017 and the
-presidentialism index created by @Sigman2018 already available in the
-V-Dem dataset. By creating a separate index, however, I am able to
-adjust the list of included manifest variables as well as embed the
-factor analysis within the larger regression model such that the
-uncertainty from the latent factor propagates as a result of the
-estimation process.
+constitution --- rather than the *de jure* powers of any one
+institution *vis-à-vis* the executive, and to do so using a latent
+variable aggregation that avoids the arbitrary weighting of linear
+additive indices.[^additive] The aggregation method is largely similar
+to the horizontal accountability index, *v2x_horacc*, created by
+@Luhrmann2017 and the presidentialism index, *v2xnp_pres*, created by
+@Sigman2018 already available in the V-Dem dataset. By creating a
+separate index, however, I am able to adjust the list of included
+manifest variables as well as embed the factor analysis within the
+larger regression model such that the uncertainty from the latent
+factor propagates as a result of the estimation process.
 
@@
 the degree to which the government complies with lower court
-decisions. And for the final dimension, independent investigative
-bodies, I include the variable *v2lgotovst*, executive oversight by
-bodies other than the legislature.
+decisions. And for the final dimension, I include the variable
+*v2exrescon*, executive respects constitution.[^lgotovst]
+
+[^lgotovst]: Both *v2x_horacc* and *v2xnp_pres* also include the
+    variable *v2lgotovst*, executive oversight by bodies other than
+    the legislature such as comptroller general, general prosecutor,
+    or ombudsman. While this is certainly a relevant variable, there is
+    an error in V-Dem's v9.0 dataset whereby it has been cleaned
+    according to when the legislature is missing. The result is that
+    1,619 country-years, or approximately $13.5\%$ of observations,
+    are missing. Since this pattern of missingness is not random, and
+    I lack additional information predicting the values for this
+    variable, I cannot implement a general imputation
+    strategy. Instead, I avoid including it in the index. Due to the
+    already large number of manifest variables and *v2lgotovst* being
+    fairly colinear with *v2lginvstp*, legislature investigates in
+    practice, with a correlation of 0.84, the result of dropping
+    *v2lgotovst* is largely inconsequential.
 
@@
     select(v2lgoppart, v2lgfunds, v2lginvstp, v2lgqstexp, v2juhcind,
-           v2juncind, v2juhccomp, v2jucomp, v2lgotovst)
+           v2juncind, v2juhccomp, v2jucomp, v2exrescon)
 
@@
 V-Dem indicators --- namely, low to high, *i.e.*, 0 is the "worst" and
-the highest value the "best." For example, for $v2lginvstp$,
+the highest value the "best." For example, for *v2lginvstp*,
 legislature investigates in practice, the answer category 0 denotes
@@
 IRT model, this translates to higher values for the point estimates
-denoting country-years where the legislature is more likely to
+representing country-years where the legislature is more likely to
 investigate.  Finally, figure \ref{fig:cor} also provides a
@@
 
-(ref:cor) Correlation heatmap of manifest variables
+(ref:cor) Correlation matrix of manifest variables
 
@@
 In the preceding equation, $\lambda$ is the variable specific factor
-loading, $\gamma$ the population intercepts, and $\psi$ the scale of
+loadings, $\gamma$ the population intercepts, and $\psi$ the scale of
 the error term. I assume local item independence such that the error
-terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
-Novick). Furthermore, for the purposes of identifying the likelihood,
-the scale of the latent factor, $\theta$, is fixed to mean $0$ and
-standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
-\lambda_n)$, are constrained to be positive [@Jackman2009]. The
-positive constraints follow theory since, as stated previously, all of
-the manifest variables exhibit the same scale direction and should be
-positively correlated with the latent factor. This means that larger
-values in the point estimates for $\theta$ represent a more
-constrained executive.
+terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated. Furthermore, for
+the purposes of identifying the likelihood, the scale of the latent
+factor, $\theta$, is fixed to mean $0$ and standard deviation $1$
+while the factor loadings, $(\lambda_1, \dotsc, \lambda_n)$, are
+constrained to be positive [@Jackman2009]. The positive constraints
+follow theory since, as stated previously, all of the manifest
+variables exhibit the same scale direction and should be positively
+correlated with the latent factor. This means that larger values in
+the point estimates for $\theta$ represent a more constrained
+executive.
 
@@
 take a cue from *v2x_horacc* and include into the model a dichotomized
-version of *v2lgbicam*, how many chambers does the legislature
-contain, denoted by the Bernoulli distributed variable, $w$, where
+version of *v2lgbicam*, measuring how many chambers the legislature
+contains, denoted by the Bernoulli distributed variable, $w$, where
 country-years with a sitting legislature are given a $1$ and
 country-years where the legislature is missing or has been suspended,
-$0$. The probability of a legislature being present, $\pi_i$, is then
-a linear function of the latent factor, $\theta$, transformed by the
-inverse logit such that $\pi_i \in [0, 1]$. The practical effect is
-that where there is data for the legislative manifest variables, the
-latent factor is informed by the measured institutional quality of the
-legislature. Where data is missing due to the suspension or lack of a
-legislature, the latent factor is down weighted by the linear function
-predicting $w$, conditional on the observed values of the remaining
-manifest variables. The resulting model specification, including
-priors, can be summarised as follows:
-\begin{align*}
-    w_i & \sim \mathsf{Bernoulli}(\pi_i) \\
-    \pi_i & = \mathsf{logit}^{-1}(\eta + \delta \theta_i) \\
-    \eta & \sim \mathsf{Normal}(0, 5) \\
-    \delta & \sim \mathsf{LogNormal}(0, 1)
+are assigned $0$. The probability of a legislature being present,
+$\pi_i$, is then a linear function of the latent factor, $\theta$,
+transformed by the inverse logit such that $\pi_i \in [0, 1]$. The
+practical effect is that where there is data for the legislative
+manifest variables, the latent factor is informed by the measured
+institutional quality of the legislature. Where data is missing due to
+the suspension or lack of a legislature, the latent factor is down
+weighted by the linear function predicting $w$, conditional on the
+observed values of the remaining manifest variables. The resulting
+model specification, including priors, can be summarised as follows:
+\begin{align*} w_i & \sim \mathsf{Bernoulli}(\pi_i) \\ \pi_i & =
+\mathsf{logit}^{-1}(\eta + \delta \theta_i) \\ \eta & \sim
+\mathsf{Normal}(0, 5) \\ \delta & \sim \mathsf{LogNormal}(0, 1)
 \end{align*}
@@
     are missing data for at least one of the judicial variables
-    ($1\%$) and only 58 are missing for *v2exrescon* ($0.5\%$). The
-    variable *v2lgotovst*, executive oversight by bodies other than
-    the legislature, is an outlier with 1619 country-years missing
-    ($13.5\%$ missingness) since there is an error in V-Dem's v9.0
-    dataset whereby it has been cleaned according to when the
-    legislature is missing. Thus, I impute *v2lgotovst* within the
-    larger model where the legislature is missing by drawing missing
-    values from a normal distribution with a mean and standard
-    deviation estimated from the observed *v2lgotovst* data. No other
-    country-years or variables are imputed. The rationale is that,
-    disregarding data processing issues, missingness in V-Dem can
-    arise either from a lack of expert coders or lack of
-    information. The latter scenario will most likely be correlated
-    with country-years where there is a crisis or significant
-    political instability, such as coups or civil war, where extreme
-    values are more probable. Since it cannot be assumed that the
-    values are completely missing at random, the data generative
-    process for missingness cannot be modeled without potentially
-    introducing bias into the estimates.
+    ($1\%$) and only 58 are missing for *v2exrescon* ($0.5\%$). I do
+    not implement any imputation strategy for the manifest
+    variables. The rationale is that, disregarding data processing
+    issues, missingness in V-Dem can arise either from a lack of
+    expert coders or lack of information. The latter scenario will
+    most likely be correlated with country-years where there is a
+    crisis or significant political instability, such as coups or
+    civil war, where extreme values are more probable. Since it cannot
+    be assumed that the values are completely missing at random, the
+    data generative process for missingness cannot be modeled without
+    potentially introducing bias into the estimates.
 
@@
 As mentioned during the theoretical discussion of possible mechanisms,
-executive constraints alone are insufficient in determining the
-outbreak of conflict. One possible mediator is the relative capacity
-of the state. For the present analysis I follow @Bartusevic2018 in
-using the Composite Indicator of National Capability (CINC) from the
-National Material Capabilities v5.0 dataset (1816-2012)
-[@Singer1987]. The CINC index is aggregated from several demographic,
-industrial, and military indicators in order to capture the available
-resources a country can bring to bear in the event of a conflict. This
-is an imperfect measure that largely misses the state's bureaucratic
-and administrative capacity [@Hendrix2010]. [^wbgi] However, given
-that the dimension of interest for this analysis is the repressive
-capacity of the state, material capabilities, especially military
-resources, should still provide an adequate proxy.
+the presence of executive constraints alone are insufficient in
+determining the outbreak of conflict. One possible moderator is the
+relative capacity of the state. For the present analysis I follow
+@Bartusevic2018 in using the Composite Indicator of National
+Capability (CINC) from the National Material Capabilities v5.0 dataset
+(1816-2012) [@Singer1987]. The CINC index is aggregated from several
+demographic, industrial, and military indicators in order to capture
+the available resources a country can bring to bear in the event of a
+conflict. This is an imperfect measure that largely misses the state's
+bureaucratic and administrative capacity [@Hendrix2010]. [^wbgi]
+However, given that the dimension of interest for this analysis is the
+repressive capacity of the state, material capabilities, especially
+military resources, should still provide an adequate proxy.
 
@@
 term between GDP per capita and the latent measure for executive
-constraints to capture the potential conditional relationship (TODO:
-expand this sentence).
-
-[^gdp]: Prior to taking the natural log, the GDP measure is of course
+constraints to capture any potential conditional relationship to
+specifically test whether the effect from executive constraints
+diminishes with lower levels of GDP per capita.
+
+[^gdp]: Prior to taking the natural log, the GDP measure is
     transformed to GDP per capita by dividing by the total population,
@@
 central administration and police force in order to suppress violent
-dissent [@Fearon2003]. There's also the possibility that a high
+dissent [@Fearon2003]. There is also the possibility that a high
 population density increases the competition for resources. If the
@@
 number of peace years since the last conflict incident.[^time] I also
-include a dummy variably for ongoing high intensity conflict since
+include a dummy variable for ongoing high intensity conflict since
 devastating civil wars may have a particularly destabilizing
@@
 \ref{fig:fa-line}. Figure \ref{fig:err} meanwhile plots the range of
-the upper and low bounds of the $80$ percent posterior quantile
+the upper and lower bounds of the $80$ percent posterior quantile
 intervals for an example manifest variable in $Z^{\ast}$ against the
@@
 single country can not be easily visualised, an exploratory overview
-on the performance of the executive constraints measure is provided in
+of the performance of the executive constraints measure is provided in
 the form of several different plots.  The first, figure
@@
 the outcome, civil conflict onset, on the log-odds scale. Credible
-intervals are summarised the same as previous plots, 80 percent
-quantile intervals, and exact values are reported in table
+intervals are summarised in the same way as previous plots with 80
+percent quantile intervals, and exact values are reported in table
 \ref{tab:reg} together with the posterior probabilities that each
@@
 
-For the $\beta$ parameter for executive constraints there is a
-`r as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
+For the $\beta$ parameter for executive constraints there is a `r
+as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
 parameter is less than zero and a 80 percent posterior probability
 that an increase of one standard deviation decreases the log-odds of
-conflict onset between $`r beta_s[1, 1]`$. A one standard deviation
-increase in this context corresponds roughly to the difference between
-the level of executive constraints for Zambia in 1967 under the
-administration of Kenneth Kaunda right before the banning of
-opposition parties following the 1968 election, with a score of
-approximately 0, and Greece in 1976 shortly after the end of military
-rule and the establishment of a democratic republic with a score of
-approximately 1.
+civil conflict onset between $`r beta_s[1, 1]`$. A one standard
+deviation increase in this context corresponds roughly to the
+difference between the level of executive constraints for Zambia in
+1967 under the administration of Kenneth Kaunda right before the
+banning of opposition parties following the 1968 election, with a
+score of approximately 0, and Greece in 1976 shortly after the end of
+military rule and the establishment of a democratic republic with a
+score of approximately 1.
 
@@
 
-Population density, mean elevation, number of political relevant
+Population density, mean elevation, number of politically relevant
 ethnic groups, proportion of the population facing state
-discrimination, presence of neighbouring conflict, and the dummy
-variable for independence year all increase the risk of civil conflict
-onset. A one unit increase in population density is similar to going
-from Cambodia in 1987 with a population of $`r cmb$pop`$ million and
-an area of $`r cmb$area_sqkm`$ square kilometers to India in 1976 with
-a population of $`r india$pop`$ million and an area of
-$`r india$area_sqkm`$ square kilometers. A one unit increase for mean
-elevation roughly matches the difference between Russia at $`r russia`$
-meters and Georgia at $`r georgia`$ meters. And, a one unit increase
-for proportion of the population facing discrimination is roughly
-equivalent to Thailand from 2002 until present with $`r thai`$ percent
-and Bhutan from 1989 with $`r bhutan`$ percent.
+discrimination, presence of at least one neighbouring conflict, and
+the dummy variable for independence year all increase the risk of
+civil conflict onset. A one unit increase in population density is
+similar to going from Cambodia in 1987 with a population of $`r
+cmb$pop`$ million and an area of $`r cmb$area_sqkm`$ square kilometers
+to India in 1976 with a population of $`r india$pop`$ million and an
+area of $`r india$area_sqkm`$ square kilometers. A one unit increase
+for mean elevation roughly matches the difference between Russia at
+$`r russia`$ meters and Georgia at $`r georgia`$ meters. And, a one
+unit increase for proportion of the population facing discrimination
+is roughly equivalent to Thailand from 2002 until present with $`r
+thai`$ percent and Bhutan from 1989 until present with $`r bhutan`$
+percent.
 
@@
 
-TODO: grab this programmatically
-
 #### Random Effects
@@
 the countries cannot be easily visualized at once only the top ten and
-bottom ten depicted. From the results it is clear that there is
+bottom ten are depicted. From the results it is clear that there is
 considerably less uncertainty in the estimates for the top
@@
 To interpret the interaction term between executive constraints and
-GDP per Capita I calculate the marginal effect of $\theta$, executive
+GDP per capita I calculate the marginal effect of $\theta$, executive
 constraints, on the predicted probability of civil conflict onset for
@@
 around the 10\% quantile for GDP per capita due to the high degree of
-posterior uncertainty. In the low GDP per capita plot, there is,
-nonetheless, a slight increase in the median predicted probability,
-which begs the question: for GDP per capita values lower than the 10\%
-quantile is there a distinguishable reversal in the relationship
-between executive constraints and civil conflict such that there is
-actually a *positive* increase in the predicted probability for higher
-levels of executive constraints?
+posterior uncertainty. In the plot depicting low GDP per capita, there
+is, nonetheless, a slight increase in the median predicted
+probability, which begs the question: for GDP per capita values lower
+than the 10\% quantile is there a distinguishable reversal in the
+relationship between executive constraints and civil conflict such
+that there is actually a *positive* increase in the predicted
+probability for higher levels of executive constraints?
 
@@
 fairly consistent across changing values of GDP per capita while the
-upper bounds increases due to the distributions of predicted
+upper bounds increase due to the distributions of predicted
 probabilities being progressively skewed as GDP per capita
@@
 inactivity --- which captures both high and low intensity civil
-conflicts. How robust are these findings when only examining
-full-scale civil wars? There is no *a prior* theoretical reason to
-expect that the relationship between executive constraints and civil
-conflict onset should not be negative when only examining civil
-wars. In fact, at least one of the mentioned theoretical mechanisms,
-namely that executive constraints may affect the ability of rebel
-groups to mobilize potential fighters by offering local security
-guarantees against government repression, posited that the theory may
-*not* hold for low intensity conflicts featuring predatory rebel
-groups. Thus, if the relationship were not to be found when re-running
-the model with only high intensity conflict onsets that would be a
-significant challenge to the external validity of the initial results.
+conflicts. However, it is unclear if these findings will remain robust
+when only examing full-scale civil wars. There re no *a prior*
+theoretical reasons to expect that the relationship between executive
+constraints and civil conflict onset should not be negative when only
+examining civil wars. In fact, at least one of the mentioned
+theoretical mechanisms, namely that executive constraints may affect
+the ability of rebel groups to mobilize potential fighters by offering
+local security guarantees against government repression, posited that
+the theory may *not* hold for low intensity conflicts featuring
+predatory rebel groups. Thus, if the relationship were not to be found
+when re-running the model with only high intensity conflict onsets
+that would be a significant challenge to the external validity of the
+initial results.
 
@@
 the latter, onset is the first coded conflict year for the cumulative
-period passing the high intensity threshold. In total there
+period passing the high intensity threshold. In total there are
 `r sum(major$final.df$lmajor_onset)` onsets for the single year 1,000
@@
 majority of depicted parameter estimates are fairly stable across
-runs. Notably, the relationship between executive constraints and
-conflict onsets remains negative with a high posterior probability of
-being less than zero. The estimates for the interaction effect between
-executive constraints and GDP per capita, however, become far too
-uncertain to be able to conclusively assert an association. The
-probability that the interaction term is less than zero for the high
-intensity model run, *i.e.* where at least a single year crosses the
-high intensity threshold, is only `r fmt(ix_probs["major_only"])`
-while the probability for the cumulative intensity model run is only
-`r fmt(ix_probs["cumulative_only"])`.
+runs. Notably, the relationship between executive constraints and the
+risk of civil conflict onsets remains negative with a high posterior
+probability of being less than zero. The estimates for the interaction
+effect between executive constraints and GDP per capita, however,
+become far too uncertain to be able to conclusively assert an
+association. The probability that the interaction term is less than
+zero for the high intensity model run, *i.e.* where at least a single
+year crosses the high intensity threshold, is only `r
+fmt(ix_probs["major_only"])` while the probability for the cumulative
+intensity model run is only `r fmt(ix_probs["cumulative_only"])`.
 
@@
     run. The stochastic nature of posterior sampling ensures that the
-    estimates are never be exactly the same; however, the differences
-    are inconsequential. In fact the correlation between the main
-    model and both the cumulatively high intensity model and the major
+    estimates are never exactly the same; however, the differences are
+    inconsequential. In fact the correlation between the main model
+    and both the cumulatively high intensity model and the major
     conflicts only model is effectively one.
@@
 ROC AUC of `r cum_auc`, the overall performance is fairly limited with
-a precision-recall AUC of only `r cum_auc_pr`.
+a precision-recall AUC of only `r cum_auc_pr` out of 1.
 
@@
 suggest that the negative relationship between executive constraints
-and civil conflict onset is fairly robust. However, the same issues of
-model fit continue to plague
+and civil conflict onset is fairly robust. However, these model runs
+similarly suffer from the same issues of relatively poor model fit.
 
@@
 
-How do the results from the presented model relate back to the
-theories potentially linking executive constraints to civil conflict?
-According to the model there is support for the general proposition
-that the risk of civil conflict onset decreases the more constrained
-the executive is by other government branches when controlling for the
-selected covariates. As stated previously, however, which theoretical
-mechanism(s) might be responsible for this relationship lies outside
-the scope of this paper in large part because the mechanisms that were
-discussed are not mutually exclusive. Furthermore, quantitative
-country-year regressions such as this one can only detect general
-patterns and provide evidence towards the associations of particular
-indicators. What they cannot do is examine the microfoundations
-underpinning a particular theory.
-
-That being said, on the basis of the mobilization theory I did argue
-that there may be a potential confounder in the relative threshold for
+How do the presented results from the various model runs relate back
+to the theories potentially linking executive constraints to civil
+conflict? According to the primary model, as well as the robustness
+checks, there is support for the general proposition that the risk of
+civil conflict onset decreases the more constrained the executive is
+by other government branches when controlling for the selected
+covariates. As stated previously, however, an in-depth investigation
+of which theoretical mechanism(s) might be reponsible for this
+relationship lies outside of the scope for this paper in large part
+because the mechanisms that were discussed in the theory section are
+not mutually exclusive.  Furthermore, quantitative country-year
+regressions such as the ones performaned in this thesis can only
+detect general patterns and provide evidence towards the associations
+between particular indicators. What regression models cannot do is
+examine the microfoundations underpinning a particular theory.
+
+That being said, on the basis of the mobilization theory I argued that
+there may be a potential confounder in the relative threshold for
 mobilization. I operationalized this in the model as GDP per capita
-and included an interaction term with the latent construct for
-executive constraints. Accordingly, the negative relationship between
-executive constraints and conflict onset does not hold for low-income
-country-years. In these cases, the effect is practically
-non-existent. According to theory this would mean that where labor
-costs are low and the provisioning of economic selective benefits
-presents less of an obstacle for rebel leaders to mobilize fighters,
-the effect of executive constraints on the risk of civil conflict is
-largely indeterminate. This may be because mobilization in these
-situations is independent of the government's decision to engage in
-repression, assuming that the state does not have the capacity to
-completely eradicate the opposition immediately. More specifically,
-low-income countries may be vulnerable to the types of predatory
-insurgent groups which require significantly less fighters than rebel
-groups seeking regional independence or to topple the central
-regime. This interpretation could also hold even if GDP per capita was
-instead understood as proxying the state's administrative and military
-capacity by roughly capturing the revenue potential available to the
-state [@Fearon2003]. Where the state is particularly weak it simply
-might not matter how constrained the executive is.
-
-Despite these results, there is one significant caveat to the
-presented model. Namely, the model and the estimated predicted
-probabilities were a poor fit to the underlying data. The tradeoff
-between precision and recall was far from satisfactory; the model
-could easily predict non-onset country-years in large part because of
-the class imbalance in the dataset between onsets and non-onsets, but
-had difficultly predicting actual onsets with accuracy. This is
-unsurprising given that civil conflict onsets are relatively rare
-events and studies attempting to examine the impact of structural
-factors, such as political institutions, on onset risk are dealing
-with small effect sizes, noisy data, and relatively few variables of
-actual exogenous change. This is an issue that has plagued much of the
-conflict literature [@Ward2010] so the model presented here is far
-from being an outlier. But it does mean that any results should be
-considered tentative at best, at least until confirmed by additional
-studies.
+and included an interaction term between GDP per capita and the latent
+construct for executive constraints. Accordingly, the results show
+that the negative relationship between executive constraints and civil
+conflict onset does not hold for low-income country-years. In these
+cases, the effect is practically non-existent. According to theory,
+this would mean that where labor costs are low and the provisioning of
+economic selective benefits presents less of an obstacle for rebel
+leaders to mobilize fighters, the effect of executive constraints on
+the risk of civil conflict is largely indeterminate. This may be
+because mobilization in these situations is independent of the
+government's decision to engage in repression, assuming that the state
+does not have the capacity to completely eradicate the opposition
+immediately. More specifically, low-income countries may be vulnerable
+to the types of predatory insurgent groups which require significantly
+less fighters than rebel groups seeking regional independence or to
+topple the central regime. This interpretation could also hold even if
+GDP per capita was instead understood as proxying the state's
+administrative and military capacity by roughly capturing the revenue
+potential available to the state [@Fearon2003]. Where the state is
+particularly weak it simply might not matter how constrained the
+executive is.
+
+These results, however, should be interpreted with caution as there is
+a significant caveat to the presented model. Namely, the model and the
+estimated predicted probabilities were a poor fit to the underlying
+data when examining both low intensity and high intensity civil
+conflicts. The tradeoff between precision and recall was far from
+satisfactory; the model could easily predict non-onset country-years
+in large part because of the class imbalance in the dataset between
+onsets and non-onsets, but had difficultly predicting actual onsets
+with accuracy. This is unsurprising given that civil conflict onsets
+are relatively rare events and studies attempting to examine the
+impact of structural factors, such as political institutions, on onset
+risk are dealing with small effect sizes, noisy data, and relatively
+few variables of actual exogenous change. This is an issue that has
+plagued much of the conflict literature [@Ward2010] so the model
+presented here is far from being an outlier. But, it does mean that any
+results should be considered tentative at best, at least until
+confirmed by additional studies.
 
@@
 unexplored in this paper would be to model the effects of
-repression. A future study could examine the temporal sequence of
-executive constraints on the likelihood of government repression and
-its effect in turn on the risk of civil conflict onset. Such an
-analysis would move beyond the standard country-year logistic
-regression, which is a poor model for the dynamics of civil conflict,
-and attempt to model country-date data, possibly even at the
+repression. A future study could examine the temporal sequence in the
+development of executive constraints on the likelihood of government
+repression and its effect in turn on the risk of civil conflict
+onset. Such an analysis would move beyond the standard country-year
+logistic regression, which is a poor model for the dynamics of civil
+conflict, and attempt to model country-date data, possibly even at the
 subnational level depending on data availability.

### 2020-08-30 paper.Rmd
Manuscript edits

@@
     violence. This can encourage political grievances, lead to a loss of
-    legitimacy for the regime, and provide the opportunity for rebel
+    legitimacy for the regime, and provide the opportunity for would be rebel
     groups to attract new members by offering protection against
@@
 opts_chunk$set(echo = F, message = F, warning = F,
-               fig.width = 5, fig.height = 3, cache = T)
+               fig.width = 5, fig.height = 3, cache = F)
+
+variable_names <- c("rgdpepc" = "GDPpc (ln)",
+                    "rgdpepc_gro" = "GDPpc Growth",
+                    "cinc" = "CINC",
+                    "discrimpop" = "Discriminated Pop.",
+                    "pop_density" = "Pop. Density (ln)",
+                    "meanelev" = "Mean Elevation (ln)",
+                    "rlvt_groups_count" = "Relevant Groups",
+                    "neighbour_conflict" = "Neighbouring Conflict",
+                    "major_ongoing" = "High Int. Conflict",
+                    "independence" = "Independence",
+                    "exec_constraints" = "Exec. Constraints",
+                    "exec * rgdpepc" = "Exec. Constraints * GDPpc")
+
+# Don't include this in thesis.utils since the variable names are hardcoded.
+translate <- function(x) {
+    variable_names[match(x, names(variable_names))] %>% unname
+}
 ```
@@
 cumulative battles deaths from organized violence recorded since 1989
-[@Pettersson2019; @Eck2007; @Sundberg2012], `r
-fmt(civil_deaths_percent, 1)` percent are the result of civil
+[@Pettersson2019; @Eck2007; @Sundberg2012],
+`r fmt(civil_deaths_percent, 1)` percent are the result of civil
 conflict.
@@
 
-## Political Institutions and Civil Conflict
+## Political Institutions
 
@@
     pre/post-processing was done using R 3.6.2 [@Core2019]. The Stan
-    model code is available in [Appendix A](#appendix-a) and the full
+    model code is available in the [appendix](#appendix-a) and the full
     project in the form of a reproducible pipeline, including the
@@
     [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
-    Finally, a simulation analysis assessing the model and it's
-    ability to recover simulated true values is available in [Appendix
-    B](#appendix-b).
 
@@
 
+colnames(df) <- translate(colnames(df))
 q <- summary_stats(df)
@@
 
-```{r results = "hide", fig.width = 8, fig.cap = "\\label{fig:map} (ref:map)"}
+```{r results = "hide", fig.width = 6, fig.cap = "\\label{fig:map} (ref:map)"}
 map_data <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
@@
 
-tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
+rwanda_idx <- filter(final.df, country_name == "Rwanda", year %in% c(2000, 2010)) %$%
     reduced_idx
 
-tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
+rwanda_p <- signif(mean(theta[, rwanda_idx[1]] < theta[, rwanda_idx[2]]) * 100, 2)
 ```
@@
 \ref{fig:tie} plots the $\theta$ estimates for 2010 versus 2000 by
-country. A full list of all countries is available in [Appendix
-C](#appendix-c). Note, the statistical significance of differences
+country. Note, the statistical significance of differences
 between years for a single country cannot strictly be judged by
@@
 probability of approximately $`r chile_p`$ percent that the estimate
-for 2000 is lower than 2010. In the case of Tunisia with even more
-visual overlap, there is a $`r tunisia_p`$ percent posterior
+for 2000 is lower than 2010. In the case of Rwanda with even more
+visual overlap, there is a $`r rwanda_p`$ percent posterior
 probability that the estimate for 2010 is higher than 2000.
@@
 beta <- extract(reg, "beta")
-colnames(beta) <- c("exec_constraints", colnames(X),
+columns <- c("exec_constraints", colnames(X),
              sprintf("exec * %s", colnames(X)[interaction_idx]))
+
+colnames(beta) <- translate(columns)
 
@@
 
-For the $\beta$ parameter for executive constraints there is a `r
-as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
+For the $\beta$ parameter for executive constraints there is a
+`r as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
 parameter is less than zero and a 80 percent posterior probability
@@
 argentina <- filter(merged.df, country_name == "Argentina", year == 1989) %$%
-    prettyNum(rgdpepc, big.mark = ",")
+    fmt(rgdpepc)
 
 slovakia <- filter(merged.df, country_name == "Slovakia", year == 2003) %$%
-    prettyNum(rgdpepc, big.mark = ",")
+    fmt(rgdpepc)
 
@@
     select(pop, area_sqkm) %>%
-    mutate(area_sqkm = signif(area_sqkm, 3),
-           pop = signif(pop, 3))
+    mutate(area_sqkm = fmt(area_sqkm),
+           pop = fmt(pop))
 
@@
     select(pop, area_sqkm, meanelev) %>%
-    mutate(pop = signif(pop, 3),
-           area_sqkm = signif(area_sqkm, 3),
-           meanelev = signif(area_sqkm, 3))
+    mutate(pop = fmt(pop),
+           area_sqkm = fmt(area_sqkm),
+           meanelev = fmt(meanelev))
 
 georgia <- filter(merged.df, country_name == "Georgia", year == 2012) %$%
-    signif(meanelev, 3)
+    fmt(meanelev)
 
 russia <- filter(merged.df, country_name == "Russia", year == 2012) %$%
-    signif(meanelev, 3)
+    fmt(meanelev)
 
 thai <- filter(merged.df, country_name == "Thailand", year == 2002) %$%
-    { discrimpop * 100 }
+    { fmt(discrimpop * 100) }
 
 bhutan <- filter(merged.df, country_name == "Bhutan", year == 1989) %$%
-    { discrimpop * 100 }
+    { fmt(discrimpop * 100) }
 
@@
 civil conflict onset. A one unit increase in population density is
-similar to going from Cambodia in 1987 with a population of $`r
-cmb$pop`$ million and an area of $`r cmb$area_sqkm`$ square kilometers
+similar to going from Cambodia in 1987 with a population of
+$`r cmb$pop`$ million and an area of $`r cmb$area_sqkm`$ square kilometers
 to India in 1976 with a population of $`r india$pop`$ million and an
@@
 unit increase for proportion of the population facing discrimination
-is roughly equivalent to Thailand from 2002 until present with $`r
-thai`$ percent and Bhutan from 1989 until present with $`r bhutan`$
+is roughly equivalent to Thailand from 2002 until present with
+$`r thai`$ percent and Bhutan from 1989 until present with $`r bhutan`$
 percent.
@@
 
-alpha_s <- quantile(alpha, probs = c(0.1, 0.9)) %>% signif(3)
-alpha_interval <- sprintf("[%g,%g]", alpha_s[1], alpha_s[2])
+alpha_s <- quantile(alpha, probs = c(0.1, 0.9)) %>% fmt
+alpha_interval <- sprintf("[%s,%s]", alpha_s[1], alpha_s[2])
 ```
@@
 gdp_rwa <- filter(merged.df, country_name == "Rwanda", year == 2007) %$%
-    rgdpepc %>%
-    prettyNum(big.mark = ",")
+    fmt(rgdpepc)
 
 gdp_phl <- filter(merged.df, country_name == "Philippines", year == 2008) %$%
-    rgdpepc %>%
-    prettyNum(big.mark = ",")
+    fmt(rgdpepc)
 
 gdp_po <- filter(merged.df, country_name == "Portugal", year == 2009) %$%
-    rgdpepc %>%
-    prettyNum(big.mark = ",")
+    fmt(rgdpepc)
 ```
@@
 zero for the high intensity model run, *i.e.* where at least a single
-year crosses the high intensity threshold, is only `r
-fmt(ix_probs["major_only"])` while the probability for the cumulative
+year crosses the high intensity threshold, is only
+`r fmt(ix_probs["major_only"])` while the probability for the cumulative
 intensity model run is only `r fmt(ix_probs["cumulative_only"])`.
@@
 
-## A - Stan Code {#appendix-a}
+## Stan Code {#appendix-a}
 
@@
 
-## B - Prior Simulation {#appendix-b}
-
-In order
-
-```{r eval = F}
-prior_plot <- function(range, fn, title = NULL, ...) {
-    data.frame(x = range) %>%
-        ggplot(aes(x)) +
-        stat_function(fun = fn, ...) +
-        ggtitle(label = title) +
-        theme(axis.title.y = element_blank(),
-              axis.text.y = element_blank(),
-              axis.ticks.y = element_blank())
-}
-
-p1 <- prior_plot(c(-20, 20), dnorm, args = list(mean = 0, sd = 5),
-                 title = "Normal(0, 5)")
-p2 <- prior_plot(c(-10, 10), dnorm, args = list(mean = 0, sd = 2.5),
-                 title = "Normal(0, 2.5)")
-p3 <- prior_plot(c(0, 5), dlnorm, args = list(meanlog = 0, sdlog = 0.5),
-                 title = "Log-Normal(0, 0.5)")
-p4 <- prior_plot(c(0, 3), dweibull, args = list(shape = 5, scale = 1),
-                 title = "Weibull(5, 1)")
-p5 <- prior_plot(c(0, 10), dhcauchy, args = list(sigma = 1),
-                 title = "HalfCauchy(0, 1)")
-p6 <- prior_plot(c(0, 1), dbeta, args = list(shape1 = 5, shape2 = 5),
-                 title = "Beta(5, 5)")
-
-grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3)
-
-```
-
-```{r eval = F, warning = F}
-true_values <- readRDS("posteriors/sim/simulated_parameters.rds")
-
-fa <- read_post("posteriors/sim/fa_posteriors.csv")
-reg <- read_post("posteriors/sim/reg_posteriors.csv")
-
-fit <- bind_cols(fa, reg)
-
-post <- post_summarise(fit, c("lambda", "psi", "gamma", "delta",
-                              "kappa", "rho", "eta", "alpha", "beta",
-                              "sigma")) %>%
-    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
-    mutate(type = "Posterior Estimate")
-
-pars.df <- filter(true_values, !grepl("theta", parameter)) %>%
-    bind_rows(post) %>%
-    mutate(type = as.factor(type))
-
-ggplot(pars.df, aes(parameter, point, color = type)) +
-    geom_point() +
-    geom_errorbar(aes(parameter, ymin = codelow, ymax = codehigh),
-                  width = 0) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-
-```{r eval = F, warning = F}
-latent.df <- post_summarise(fit, pars = "theta", probs = c(0.025, 0.975)) %>%
-    rename(codelow = `2.5%`, codehigh = `97.5`)
-
-x <- mean(theta <= latent.df$codehigh & theta >= latent.df$codelow)
-
-```
-
-```
-
-\newpage
-
 # References {-}

### 2020-08-30 R/thesis.utils/R/plot.R
Update documentation for fmt and remove v2lgotovst from FA

@@ -70,7 +70,7 @@ overlay_pars <- function(pars = NULL, legend_labels = NULL, ...) {
         post_summarise(args[[i]], pars = pars, probs = c(0.10, 0.50, 0.90)) %>%
             dplyr::rename(codelow = `10%`, median = `50%`, codehigh = `90%`) %>%
             dplyr::mutate(parameter = factor(parameter, levels = parameter),
-                   model = i)
+                          model = i)
     })
 
     df <- dplyr::bind_rows(ll)
@@ -169,12 +169,15 @@ tfplot <- function(df) {
     levels <- dplyr::filter(df, year == dplyr::last(year))
     levels <- dplyr::arrange(levels, median) %$% country_name
 
+    df$country_name <- factor(df$country_name, levels = levels)
+
     ggplot(df, aes_string("country_name", "median", colour = "year")) +
         geom_errorbar(aes_string(ymax = "codehigh", ymin = "codelow"),
                       position = position_dodge(width = 0.5)) +
         theme_tufte() +
         scale_color_colorblind() +
-        theme(axis.title.y = element_blank(),
+        theme(axis.title.x = element_blank(),
+              axis.title.y = element_blank(),
               legend.title = element_blank()) +
         coord_flip()
 }

### 2020-08-30 R/thesis.utils/R/utils.R

@@ -271,6 +271,20 @@ summary_stats.data.frame <- function(x, vars = colnames(x)) {
 to_idx <- function(x, ...) factor(x, ...) %>% as.numeric
 
 
+#' Pretty print numerics
+#'
+#' This function formats a numeric value by rounding to the specified
+#' number of digits before then calling `prettyNum`.
+#'
+#' @param x NumericVector
+#' @param digits Number of decimal digits to round to.
+#'
+#' @return CharacterVector
+#'
+#' @examples
+#' fmt(10000.1243)
+#' fmt(9999.999)
+#'
 #' @export
 fmt <- function(x, digits = 2) {
     prettyNum(round(x, digits), big.mark = ",")

### 2020-08-30 R/transform.R

@@ -10,8 +10,7 @@ merged.df <- readRDS("data/merged_data.rds")
 # Manifest variables for exec_constraints
 constraint_vars <- c("v2lginvstp", "v2lgfunds", "v2lgqstexp",
                      "v2lgoppart", "v2juhcind", "v2juncind",
-                     "v2juhccomp", "v2jucomp", "v2exrescon",
-                     "v2lgotovst")
+                     "v2juhccomp", "v2jucomp", "v2exrescon")
 
 ###
 # Start by first replacing NA with Inf for v2lg* variables where

### 2020-08-31 paper.Rmd
Final edits

@@
 abstract: |
-    **Abstract**: Do institutional constraints on a regime's executive decrease the
-    likelihood of civil conflict onset? An unconstrained executive is free
-    to loot state resources, undermining the state's capacity to
-    effectively deal with nascent rebellions, and may be more likely to
-    engage in violent repression, especially in the form of indiscriminate
-    violence. This can encourage political grievances, lead to a loss of
-    legitimacy for the regime, and provide the opportunity for would be rebel
-    groups to attract new members by offering protection against
-    government violence. Furthermore, the lack of guaranteed checks to
-    executive power may incentivise actors to avoid bargained solutions if
-    they fear future extralegal retributions. Taken together, where there
-    is a lack of effective constraints against the executive there may be
-    an increased risk for the onset of civil conflict. To test this
-    proposition I employ a bayesian latent variable model which estimates
-    executive constraint as a latent factor derived from several manifest
-    variables and relates it to a binary measure of civil conflict onset
-    through a logistic regression.  The primary finding is that there is a
-    negative relationship such that the predicted probability for civil
-    conflict is lower where there exists higher levels of executive
-    constraints. However, this is conditional on the level of GDP per
-    capita; for low-income country-years the relationship between
-    executive constraints and civil conflict onset is indeterminate
-    possibly because it is easier to recruit and mobilize fighters
-    regardless of the overall level of executive constraints. The issue of
-    poor model fit means that the presented results should be considered
-    tentative at best.
+    **Abstract**: Do institutional constraints on a regime's executive
+    decrease the likelihood of civil conflict onset? An unconstrained
+    executive is free to loot state resources undermining the state's
+    capacity to effectively deal with nascent rebellions, and may be
+    more likely to engage in violent repression, especially in the
+    form of indiscriminate violence. This can encourage political
+    grievances, lead to a loss of legitimacy for the regime, and
+    provide the opportunity for would be rebel groups to attract new
+    members by offering protection against government
+    violence. Furthermore, the lack of guaranteed checks to executive
+    power may incentivise actors to avoid bargained solutions if they
+    fear future extralegal retributions. Taken together, where there
+    is a lack of effective constraints against the executive there may
+    be an increased risk for the onset of civil conflict. To test this
+    proposition I employ a bayesian latent variable model which
+    estimates executive constraint as a latent factor derived from
+    several manifest variables and relate it to a binary measure of
+    civil conflict onset through a logistic regression.  The primary
+    finding is that there is a negative relationship such that the
+    predicted probability for civil conflict is lower where there
+    exists higher levels of executive constraints. However, this is
+    conditional on the level of GDP per capita; for low-income
+    country-years the relationship between executive constraints and
+    civil conflict onset is indeterminate possibly because it is
+    easier to recruit and mobilize fighters in such settings
+    regardless of the overall level of executive constraints. The
+    model, however, is a poor fit to the data meaning that the
+    presented results should be considered tentative at
+    best. Nonetheless, this study helps to further the work on
+    examining specific political institutions as potential risk
+    factors for the onset of civil conflict.
 ---
@@
                     "neighbour_conflict" = "Neighbouring Conflict",
-                    "major_ongoing" = "High Int. Conflict",
+                    "major_ongoing" = "Major Conflict",
                     "independence" = "Independence",
@@
 
-The principle finding of this paper is that there is tentative
-evidence that executive constraints do matter and that there is a
-negative relationship between the latent level of constraints on the
-executive and the risk of civil conflict onset. However, this
-relationship largely disappears for low-income country-years when an
-interaction term between executive constraints and gross domestic
-product (GDP) per capita is included. If the primary theoretical
-mechanism involves the impact of executive constraints on rebel
-mobilization --- an unchecked executive may be more likely to engage
-in indiscriminate violence which could lead to a loss of legitimacy as
-well as the ability of rebel groups to offer local security guarantees
-to attract recruits --- then the incentives for violence in low-income
+To study the potential link between executive constraints and civil
+conflict onset I use a bayesian latent variable model to estimate
+executive constraints as a latent factor drawn from the basis of
+several observed manifest variables. I then relate this measure to a
+binary indicator for civil conflict onset using a cross-country time
+series logistic regression.
+
+The principle finding is that there is tentative evidence that
+executive constraints do matter and that there is a negative
+relationship between the latent level of constraints on the executive
+and the risk of civil conflict onset. However, this relationship
+largely disappears for low-income country-years when an interaction
+term between executive constraints and gross domestic product (GDP)
+per capita is included. If the primary theoretical mechanism involves
+the impact of executive constraints on rebel mobilization --- an
+unchecked executive may be more likely to engage in indiscriminate
+violence which could lead to a loss of legitimacy as well as the
+ability of rebel groups to offer local security guarantees to attract
+recruits --- then the incentives for violence in low-income
 country-years due to say economic factors may be so great that it does
@@
 mechanisms linking executive constraints to conflict onset will then
-be discussed followed by an empirical analysis using a bayesian latent
-variable model to estimate executive constraints from the basis of
-several observed manifest variables. This latent measure will then be
-used to predict civil conflict onset using a cross-country time series
-logistic regression.
+be discussed followed by a presentation of the paper's research design
+and empirical analysis, including both the latent variable measurement
+model for executive constraints and the cross-country time series
+logistic regression predicting civil conflict onset. The final section
+provides conclusions and possible suggestions for future research.
 
@@
 
-(ref:onsets) Proportions of countries per year experiencing at least
-    one ongoing civil conflict and at least one new onset.
+(ref:onsets) Proportions of countries experiencing at least one
+    ongoing civil conflict and at least one new onset per year.
 
@@
     ylab(target) +
-    xlab("Theta")
+    xlab("Executive Constraints")
 
@@
 
-#### Random Effects
+#### Varying Intercepts
 
@@
 capita, going from the worst to best in executive constraints
-decreases the predicted probability from [1.5, 4.5] down to below
+decreases the predicted probability from [1.0, 4.5] down to below
 0.5. This effect diminishes, however, for lower values of GDP per
@@
     theme_tufte() +
-    xlab("GDP per Capita")
+    xlab("GDP per Capita") +
+    theme(axis.title.y = element_blank())
+
 ```
@@
 
-(ref:confusion) Confusion matrix with a 0.168 cutoff for the median
+(ref:confusion) Confusion matrix with a 0.17 cutoff for the median
     posterior predicted probabilities.
@@
 ```{r confusion}
-m <- table(p_median > .168, final.df$lepisode_onset)
+m <- table(p_median > .17, final.df$lepisode_onset)
 
@@
 compare the predicted probabilities to the actual data, table
-\ref{tab:confusion} presents a confusion matrix with a 0.168 cutoff,
+\ref{tab:confusion} presents a confusion matrix with a 0.17 cutoff,
 which corresponds to the third quartile in the summary table. `r tpr`
@@
 
-(ref:confusion-major) Confusion matrix with a 0.168 cutoff for the
+(ref:confusion-major) Confusion matrix with a 0.17 cutoff for the
     median posterior predicted probabilities from the single year high
@@
 
-m <- table(p_median > .168, major$final.df$lmajor_onset)
+m <- table(p_median > .17, major$final.df$lmajor_onset)
 rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
@@
 single year high intensity model run using the median posterior
-predicted probabilities. A cutoff of 0.168 was reused for comparative
-purposes to the original results. In this case, the model is unable to
-predict even a single major conflict onset with the highest median
-posterior predicted probability for an actual onset being only
+predicted probabilities. A cutoff of 0.17 was reused for comparative
+purposes to the original results. In this case, the model is able to
+only correctly predict a single major conflict onset with the highest
+median posterior predicted probability for an actual onset being only
 `r fmt(max(p_median[major$final.df$lmajor_onset == 1]))`. Furthermore,
@@
 
-(ref:confusion-cumulative) Confusion matrix with a 0.168 cutoff for
+(ref:confusion-cumulative) Confusion matrix with a 0.17 cutoff for
     the median posterior predicted probabilities from the cumulative
@@
 
-m <- table(p_median > .168, cumulative$final.df$lcum_onset)
+m <- table(p_median > .17, cumulative$final.df$lcum_onset)
 rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
